---
title: "MMRF OS/PFS prediction"
author: "Edgar Gonzalez-Kozlova"
date: "Aug 26 2025"
output: 
  html_document:
  toc: true
  toc_depth: 2
  toc_float: true
  ighlight: espresso
number_sections: yes
theme: cerulean
---

#### Start

#### Load Libraries and set working directory

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}
### 
libs<-c('scater','loomR','Seurat','patchwork','SeuratDisk','monocle3','ggpubr','cowplot','ggplot2','ggbeeswarm','clusterProfiler',
        'monocle3','limma','edgeR','fitdistrplus','factoextra','ggrepel','tidyverse','pheatmap','reshape2','ComplexHeatmap','survRM2',
        "survminer","survival","ggalluvial",'gtsummary','variancePartition','DirichletReg','cowsay','emmeans','forestmodel','ROCR','rms')
lapply(libs, require, character.only = TRUE) ; rm(libs)
### 
setwd("/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/")
```

#### Custom Functions

```{r message=FALSE, warning=FALSE}
###
get_auc  <- function (m,df) {
  predictions <- predict(object = m, df[,-which(colnames(df) %in% 'censpfs')],se.fit = TRUE)
  labels <- df$censpfs
  pred <- prediction(predictions, labels)
  perf <- performance(pred,'tpr','fpr')
  auc_ROCR <- performance(pred, measure = "auc")@y.values[[1]]  
  return(list(perf=perf,auc=auc_ROCR,predictions=predictions,labels=labels,pred=pred))
}
###
get_auc2  <- function (m,df) {
  predictions <- predict(object = m, df[which(!is.na(df$d_dx_amm_iss_stage)),-which(colnames(df) %in% 'censpfs')],se.fit = TRUE,type="fitted.ind")
  labels <- df$censpfs[which(!is.na(df$d_dx_amm_iss_stage))]
  pred <- prediction(predictions, labels)
  perf <- performance(pred,'tpr','fpr')
  auc_ROCR <- performance(pred, measure = "auc")@y.values[[1]]  
  return(list(perf=perf,auc=auc_ROCR,predictions=predictions,labels=labels,pred=pred))
}
###
get_lrm_validation <- function(m){
  x <- validate(m,B=1000)
  class(x) <- 'matrix'
  x <- as.data.frame(x)
  x$auc <- x[1,1]/2 +0.5
return(x)
}
```

#### Discovery Cohort (data formating 1)

```{r message=FALSE, warning=FALSE}
### pops
file_717_per_cell_md <- readRDS(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/RData/1110_per_cell_md.rds')
x <- table(file_717_per_cell_md$subcluster_V03072023[file_717_per_cell_md$visit_type %in% 'baseline_diagnosis'],
           file_717_per_cell_md$sample_id[file_717_per_cell_md$visit_type %in% 'baseline_diagnosis'])
class(x) <- 'matrix' ; x[x>0]<-1 ; x <- rowSums(x) ; x <- sort(x)
sample_size_df <- data.frame(subcluster=names(x),samples=as.numeric(x))
sample_size_df$more_than_median <- sample_size_df$samples > median(sample_size_df$samples)
sample_size_df$more_than_q1 <- sample_size_df$samples > quantile(sample_size_df$samples, probs=c(0.25))
```

#### Discovery Cohort (adding clinical data)

```{r message=FALSE, warning=FALSE}
### clinical
load(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/RData/clinical_metadata_IA22_Feb03_release.RData')
###
cmia22 <- clinical_metadata_IA22_Feb03_release[which(clinical_metadata_IA22_Feb03_release$public_id %in% file_717_per_cell_md$public_id),]
cmia22 <- cmia22[,colnames(cmia22) %in% c('censpfs','ttcpfs','censos','ttcos', 'collection_event','sample_id','d_pt_sex','public_id')]
cmia22 <- unique(cmia22)
###
my_vars <- c('public_id','davies_based_risk','visit_type','siteXbatch','progression_group','d_dx_amm_age','d_dx_amm_iss_stage','collection_event','sample_id')
###
x <- file_717_per_cell_md[,which(colnames(file_717_per_cell_md) %in% my_vars)]
rownames(x) <- NULL ; x <- unique(x)
x$site <- tidyr::separate(data.frame(x$siteXbatch),1, sep="_", c("a","b",'c','d'))$a
x$batch <- tidyr::separate(data.frame(x$siteXbatch),1, sep="_", c("a","b",'c','d'))$b
x$siteXbatch <- NULL
meta_df <- x
meta_df <- merge(meta_df,cmia22,by='public_id',all=TRUE) ; rm(cmia22)
###
ix <- which(meta_df$davies_based_risk %in% c('standard_risk','high_risk') & meta_df$collection_event %in% 'Baseline')
meta_baseline_df <- meta_df[ix,] ; rm(meta_df)
```

#### Discovery Cohort (Compartments)

```{r message=FALSE, warning=FALSE}
### COMP
y <- table(file_717_per_cell_md$subcluster_V03072023,file_717_per_cell_md$sample_id)
class(y) <- 'matrix'
my_doublets <- unique(as.character(file_717_per_cell_md$subcluster_V03072023[file_717_per_cell_md$doublet_pred %in% c('dblet_cluster','poss_dblet_cluster')]))
y <- y[!rownames(y) %in% my_doublets,]
y <- y[,which(colnames(y) %in% meta_baseline_df$sample_id)]
Compartment <- c('NkT','BEry','Ery','Myeloid','Plasma') #,'Full'
cell_proportions <- list()
for(iii in Compartment){
  ix <- grep(paste('^',iii,sep=''), rownames(y))
  z <- y[ix,]
  cell_proportions[[iii]] = DR_data(t(z))
}
cell_props_comp <- do.call(cbind,cell_proportions)
###
subcluster_comp_fulldf <- cbind(meta_baseline_df,cell_props_comp)
```

#### Discovery Cohort (Removing flaged doublets)

```{r message=FALSE, warning=FALSE}
### All populations
y <- table(file_717_per_cell_md$subcluster_V03072023,file_717_per_cell_md$sample_id)
class(y) <- 'matrix'
my_doublets <- unique(as.character(file_717_per_cell_md$subcluster_V03072023[file_717_per_cell_md$doublet_pred %in% c('dblet_cluster','poss_dblet_cluster')]))
y <- y[!rownames(y) %in% my_doublets,]
### remove non baseline/relevant
y <- y[,which(colnames(y) %in% meta_baseline_df$sample_id)]
cell_props_all = DR_data(t(y))
```

#### Discovery Cohort (Adjusting ordering of data)

```{r message=FALSE, warning=FALSE}
identical(meta_baseline_df$sample_id,rownames(cell_props_all))
```

```{r message=FALSE, warning=FALSE}
cell_props_all <- cell_props_all[match(meta_baseline_df$sample_id,rownames(cell_props_all)),]
identical(meta_baseline_df$sample_id,rownames(cell_props_all))
```

#### Discovery Cohort (integrating proportions to metadata)

```{r message=FALSE, warning=FALSE}
subcluster_all_fulldf <- cbind(meta_baseline_df,cell_props_all)

### function to label quartiles
get_four_quantiles <- function(p,x){
  y <- x[,p] ; z <- x[,p]
  z[y < quantile(y, probs=c(0.25))] <- 'Q1'
  z[y > quantile(y, probs=c(0.75))] <- 'Q4'
  z[y >= quantile(y, probs=c(0.25)) & y <= quantile(y, probs=c(0.5))] <- 'Q2'
  z[y <= quantile(y, probs=c(0.75)) & y >= quantile(y, probs=c(0.5))] <- 'Q3'
  return(z)  
}

### function to label medians
get_median <- function(p,x){
  y <- x[,p] ; z <- rep('Low',length(y))
  z[y > median(y)] <- 'High'  
  return(z)
}
```

```{r message=FALSE, warning=FALSE}
dim(subcluster_all_fulldf) # non-corrected data for prediction estimated by cell type
```

```{r message=FALSE, warning=FALSE}
dim(subcluster_comp_fulldf) # non-corrected data for prediction estimated by compartments
```

#### clean workspace

```{r message=FALSE, warning=FALSE}
rm(list=setdiff(ls(), c("subcluster_all_fulldf","subcluster_comp_fulldf",
                        'get_four_quantiles','get_median','get_auc2','get_auc','get_lrm_validation') ))
###
#table(duplicated(gsub("_SM","",subcluster_all_fulldf$sample_id)))
subcluster_all_fulldf$sample_id <- gsub("_SM","",subcluster_all_fulldf$sample_id)
```

#### Validation data and metadata update from 2024 

```{r message=FALSE, warning=FALSE}
###
updated_meta_per_cell <- readRDS(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/updateMMRF_yered_william/091524_per_cell_md.rds')
###
#table(updated_meta_per_cell$collection_event,updated_meta_per_cell$cohort)
###
ix <- which(updated_meta_per_cell$collection_event %in% "Baseline")
updated_meta_per_cell <- updated_meta_per_cell[ix,]
#table(updated_meta_per_cell$expanded_risk_definition)
#table(updated_meta_per_cell$IMWG_2024_HR)
#table(updated_meta_per_cell$IMWG_2024_HR, updated_meta_per_cell$expanded_risk_definition)
#table(updated_meta_per_cell$cohort)
updated_meta_per_cell_validation <-  updated_meta_per_cell[which(updated_meta_per_cell$cohort %in% 'validation'),]
#unique(updated_meta_per_cell_validation$aliquot_id)
updated_meta_per_cell <- updated_meta_per_cell[which(updated_meta_per_cell$cohort %in% 'discovery'),]

### Simplify
#"d_amm_tx_asct_1st","siteXbatch"
temp <- updated_meta_per_cell[!duplicated(updated_meta_per_cell$sample_id),]
updated_meta_per_cell <- temp
### non matching IDs
#table(updated_meta_per_cell$sample_id %in% subcluster_all_fulldf$sample_id)
#updated_meta_per_cell$sample_id[!(updated_meta_per_cell$sample_id %in% subcluster_all_fulldf$sample_id)]
#subcluster_all_fulldf#c("sample_id",'d_dx_amm_age','expanded_risk_definition','site','batch','d_dx_amm_iss_stage','censos','censpfs','ttcos','ttcpfs')

#table(updated_meta_per_cell$sample_id %in% subcluster_all_fulldf$sample_id)
updated_meta_per_cell <- updated_meta_per_cell[updated_meta_per_cell$sample_id %in% subcluster_all_fulldf$sample_id,]
#table(updated_meta_per_cell$sample_id %in% subcluster_all_fulldf$sample_id)
###
subcluster_all_fulldf <- subcluster_all_fulldf[subcluster_all_fulldf$sample_id %in% updated_meta_per_cell$sample_id,]
#table(updated_meta_per_cell$sample_id %in% subcluster_all_fulldf$sample_id)
identical(updated_meta_per_cell$sample_id,subcluster_all_fulldf$sample_id)
subcluster_all_fulldf <- subcluster_all_fulldf[match(updated_meta_per_cell$sample_id,subcluster_all_fulldf$sample_id),]
identical(updated_meta_per_cell$sample_id,subcluster_all_fulldf$sample_id)
### add new risk
subcluster_all_fulldf$IMWG_2024_HR <- updated_meta_per_cell$IMWG_2024_HR
### add asct
subcluster_all_fulldf$d_amm_tx_asct_1st <- updated_meta_per_cell$d_amm_tx_asct_1st
#table(subcluster_all_fulldf$IMWG_2024_HR,subcluster_all_fulldf$davies_based_risk)
#table(subcluster_all_fulldf$IMWG_2024_HR,subcluster_all_fulldf$d_amm_tx_asct_1st)
#table(subcluster_all_fulldf$IMWG_2024_HR,subcluster_all_fulldf$d_dx_amm_iss_stage)
#table(subcluster_all_fulldf$IMWG_2024_HR,subcluster_all_fulldf$censos)
#table(subcluster_all_fulldf$IMWG_2024_HR,subcluster_all_fulldf$censpfs)
subcluster_all_fulldf <- subcluster_all_fulldf[subcluster_all_fulldf$IMWG_2024_HR %in% c('high_risk','standard_risk'),]
#table(subcluster_all_fulldf$IMWG_2024_HR,subcluster_all_fulldf$davies_based_risk)
#table(subcluster_all_fulldf$IMWG_2024_HR,subcluster_all_fulldf$censpfs)

### Frequencies for Validation
validation_frequencies <- table(updated_meta_per_cell_validation$aliquot_id,updated_meta_per_cell_validation$seurat_subclusters_label_transferring_Yizhe_v1)
class(validation_frequencies) <- 'matrix'
validation_frequencies <- as.data.frame(validation_frequencies)

z <- DR_data(validation_frequencies)
class(z) <- 'matrix'
z <- as.data.frame(z)
### integrate
table(rownames(z) %in% updated_meta_per_cell_validation$aliquot_id)
#identical(rownames(z),updated_meta_per_cell_validation$aliquot_id)
updated_meta_per_cell_validation<-updated_meta_per_cell_validation[match(rownames(z),updated_meta_per_cell_validation$aliquot_id),]
#identical(rownames(z),updated_meta_per_cell_validation$aliquot_id)
validation_dataset_w_meta <- cbind(z,updated_meta_per_cell_validation)
```


```{r message=FALSE, warning=FALSE}
#store data as input/output
rm(list=setdiff(ls(), c("subcluster_all_fulldf","subcluster_comp_fulldf","yered_prop_data",
                        'get_four_quantiles','get_median','get_auc2','get_auc','get_lrm_validation',
                        'validation_dataset_w_meta') ))
# save(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/RData/temp_re_analysis_sep2024.RData',
#      subcluster_all_fulldf,subcluster_comp_fulldf,get_four_quantiles,get_median,get_auc,get_lrm_validation,
#      validation_frequencies,updated_meta_per_cell_validation,validation_dataset_w_meta)
```

#### LRM modeling: raw proportions (subcluster_all_fulldf)

```{r message=FALSE, warning=FALSE}
set.seed(42)
```

```{r message=FALSE, warning=FALSE}
ddist <- datadist(subcluster_all_fulldf)
```

```{r message=FALSE, warning=FALSE}
options(datadist='ddist')
```

```{r message=FALSE, warning=FALSE}
predictions_list <- list()
pred_validation_list <- list()
```

#### Models with Clinical data (lrms)

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(censos ~ d_dx_amm_iss_stage, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V1 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V1 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ d_amm_tx_asct_1st, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_asct <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_asct <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ d_dx_amm_iss_stage + d_dx_amm_age, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V2 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V2 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ d_dx_amm_iss_stage + d_amm_tx_asct_1st, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V3 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V3 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ d_dx_amm_iss_stage + d_amm_tx_asct_1st + d_dx_amm_age + d_pt_sex, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V4 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V4 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ d_dx_amm_iss_stage + d_amm_tx_asct_1st + d_dx_amm_age + d_pt_sex + batch + site, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V5 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V5 <- get_lrm_validation(mod.bi)
```

#### Models with Davies (lrms)

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(censos ~ davies_based_risk + d_pt_sex, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V1 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V1 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ davies_based_risk + d_dx_amm_age, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V2 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V2 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ davies_based_risk + d_dx_amm_age + d_pt_sex, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V3 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V3 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ davies_based_risk + d_dx_amm_age + d_pt_sex + batch + site, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V4 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V4 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ davies_based_risk + d_amm_tx_asct_1st + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V5 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V5 <- get_lrm_validation(mod.bi)
```

#### Models with IMWG_2024_HR (lrms)

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(censos ~ IMWG_2024_HR + d_pt_sex, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V1 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V1 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ IMWG_2024_HR + d_dx_amm_age, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V2 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V2 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V3 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V3 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V4 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V4 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censos ~ IMWG_2024_HR + d_amm_tx_asct_1st + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V5 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V5 <- get_lrm_validation(mod.bi)
```

#### Models for Cell Populations and overall survival (OS) (lrms)

Sort Subpops present in at least more than 50 patients

```{r message=FALSE, warning=FALSE}
my_subcluster_names <- colnames(subcluster_all_fulldf)[16:104]
x <- list()
for(iii in my_subcluster_names){
  x[[iii]]<- length(table(subcluster_all_fulldf[[iii]]))
}
my_subcluster_names <- names(sort(unlist(x))[sort(unlist(x)) > 50])
###
ix <- which(colnames(subcluster_all_fulldf) %in% my_subcluster_names)
###
my_subcluster_names <- names(colSums(subcluster_all_fulldf[,ix])[colSums(subcluster_all_fulldf[,ix]) >1])
```

#### formulas for interative testing

```{r message=FALSE, warning=FALSE}
my_form_V1 <- list()
for(iii in my_subcluster_names){ my_form_V1[[iii]] <- as.formula(paste("censos ~", iii)) }

my_form_V2 <- list()
for(iii in my_subcluster_names){ my_form_V2[[iii]] <- as.formula(paste("censos ~ d_dx_amm_age +", iii)) }

my_form_V2 <- list()
for(iii in my_subcluster_names){ my_form_V2[[iii]] <- as.formula(paste("censos ~ d_dx_amm_age + d_pt_sex +", iii)) }

my_form_V3 <- list()
for(iii in my_subcluster_names){ my_form_V3[[iii]] <- as.formula(paste("censos ~ d_dx_amm_age + d_pt_sex + batch + site +", iii)) }

my_form_V4 <- list()
for(iii in my_subcluster_names){ my_form_V4[[iii]] <- as.formula(paste("censos ~ d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +", iii)) }

my_form_V5 <- list()
for(iii in my_subcluster_names){ my_form_V5[[iii]] <- as.formula(paste("censos ~ davies_based_risk + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +", iii)) }

my_form_V6 <- list()
for(iii in my_subcluster_names){ my_form_V6[[iii]] <- as.formula(paste("censos ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +", iii)) }

```

#### lrm computation for all models (feature bootstrapping)

This part is computationally expensive, run times can be a couple hours at minimum.

```{r message=FALSE, warning=FALSE}
### V1
for(iii in names(my_form_V1)) {
  mod.bi <- lrm(my_form_V1[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V1",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)
  pred_validation_list[[paste(iii,"V1",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V2
for(iii in names(my_form_V2)) {
  mod.bi <- lrm(my_form_V2[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V2",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)
  pred_validation_list[[paste(iii,"V2",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V3
for(iii in names(my_form_V3)) {
  mod.bi <- lrm(my_form_V3[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V3",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
  pred_validation_list[[paste(iii,"V3",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V4
for(iii in names(my_form_V4)) {
  mod.bi <- lrm(my_form_V4[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V4",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
  pred_validation_list[[paste(iii,"V4",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V5
for(iii in names(my_form_V5)) {
  mod.bi <- lrm(my_form_V5[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V5",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
  pred_validation_list[[paste(iii,"V5",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V6
for(iii in names(my_form_V6)) {
  mod.bi <- lrm(my_form_V6[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V6",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
  pred_validation_list[[paste(iii,"V6",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

#### Iterative Feature Selection Model based on performance vs number of features

```{r message=FALSE, warning=FALSE}
b <- paste( my_subcluster_names[!my_subcluster_names %in% c('')] ,'+') #NkT.8','NkT.7','NkT.11','NkT.9.0
a <- paste("censos ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +") #
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
#subcluster_all_fulldf <- type_convert(subcluster_all_fulldf)
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
dim(subcluster_all_fulldf)
#vif(mod.bi)
predictions_list[["Feat_Sel"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel"]] <- get_lrm_validation(mod.bi)
```

```{r message=FALSE, warning=FALSE}
z <- as.data.frame(as.matrix(anova(mod.bi)))
z <- rownames(z)[z$P<0.1]
z <- z[z %in% my_subcluster_names]
z
Feat_Sel_R1_Features <- z
```

```{r message=FALSE, warning=FALSE}
b <- paste(z,'+')
a <- paste("censos ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list[["Feat_Sel_R1"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel_R1"]] <- get_lrm_validation(mod.bi)
```

```{r message=FALSE, warning=FALSE}
z <- as.data.frame(as.matrix(anova(mod.bi)))
z <- rownames(z)[z$P<0.1]
z <- z[z %in% my_subcluster_names]
z
Feat_Sel_R2_Features <- z
```

```{r message=FALSE, warning=FALSE}
b <- paste(z,'+')
a <- paste("censos ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list[["Feat_Sel_R2"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel_R2"]] <- get_lrm_validation(mod.bi)

```

```{r message=FALSE, warning=FALSE}
z <- as.data.frame(as.matrix(anova(mod.bi)))
z <- rownames(z)[z$P<0.1]
z <- z[z %in% my_subcluster_names]
z
Feat_Sel_R3_Features <- z
```

```{r message=FALSE, warning=FALSE}
b <- paste(z,'+')
a <- paste("censos ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list[["Feat_Sel_R3"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel_R3"]] <- get_lrm_validation(mod.bi)
```

```{r message=FALSE, warning=FALSE}
z <- as.data.frame(as.matrix(anova(mod.bi)))
z <- rownames(z)[z$P<0.1]
z <- z[z %in% my_subcluster_names]
z
Feat_Sel_R4_Features <- z
```

```{r message=FALSE, warning=FALSE}
b <- paste(z,'+')
a <- paste("censos ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list[["Feat_Sel_R4"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel_R4"]] <- get_lrm_validation(mod.bi)
```

### Summary of predictions OS

```{r message=FALSE, warning=FALSE}
predictions_OS <- list() ; labels_OS <- list() ; aucs_OS<-list()
for(iii in names(predictions_list)){ 
  predictions_OS[[iii]] <- predictions_list[[iii]]$predictions
  labels_OS[[iii]] <- predictions_list[[iii]]$labels
  aucs_OS[[iii]] <- predictions_list[[iii]]$auc
}
###
used_features_OS<-list(R1=Feat_Sel_R1_Features,R2=Feat_Sel_R2_Features,
                       R3=Feat_Sel_R3_Features,R4=Feat_Sel_R4_Features)
###
key_predictions_OS_list <- list(predictions=predictions_OS,
                                labels=labels_OS,
                                aucs=aucs_OS,
                                used_features=used_features_OS)
```

```{r message=FALSE, warning=FALSE}
predictions_list_OS <- predictions_list
#saveRDS(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/RData/key_predictions_OS_list.RDS',key_predictions_OS_list)
```

#### Figures for Discovery Cohort OS.

```{r message=FALSE, warning=FALSE}
pred <- prediction(key_predictions_OS_list$predictions, key_predictions_OS_list$labels)
perf <- performance(pred,'tpr','fpr')
#plot(perf,lwd=2,main='ROC curves from 10-fold cross-validation')
x<- perf@x.values
y<- perf@y.values
names(x) <- names(predictions_list_OS)
names(y) <- names(predictions_list_OS)
for(i in 1:length(x)){ x[[i]]<- data.frame(model=names(x)[i],x=x[[i]])}
x <- do.call(rbind,x)
x$y <- unlist(y)
rownames(x)<-NULL
### redo aucs
aucs <- list()
for(i in 1:length(predictions_list_OS)) { aucs[[i]] <- data.frame(auc=predictions_list_OS[[i]]$auc,model=names(predictions_list_OS)[i]) }
aucs <- do.call(rbind,aucs)
x <- merge(x,aucs,by='model') ; rm(aucs)
#y <- unique(x[,c(1,4)])
#y <- y[order(y$auc),]
```

```{r message=FALSE, warning=FALSE}
x$type <- x$model
x$type[grep('cyto',x$model)] <- 'cyto'
x$type[grep('NC',x$model)] <- 'NC'
x$type[grep('V1',x$model)] <- 'Single'
x$type[grep('V2',x$model)] <- '+Age+Sex'
x$type[grep('V3',x$model)] <- '+Age+Sex+B*Site'
x$type[grep('V4',x$model)] <- '+Age+Sex+B*Site+ISS+ASTC'
x$type[grep('V5',x$model)] <- '+Cyto+Age+Sex+B*Site+ISS+ASTC'
x$type[grep('V6',x$model)] <- '+NC+Age+Sex+B*Site+ISS+ASTC'

x$type <- factor(x$type,levels = c('cyto','NC','Single','+Age+Sex','+Age+Sex+B*Site','+Age+Sex+B*Site+ISS+ASTC',
                                   '+Cyto+Age+Sex+B*Site+ISS+ASTC','+NC+Age+Sex+B*Site+ISS+ASTC',
                                   "Feat_Sel","Feat_Sel_R1","Feat_Sel_R2","Feat_Sel_R3","Feat_Sel_R4","mod_asct" ))
```


```{r message=FALSE, warning=FALSE,fig.width=6,fig.height=3}
ggplot(data=x[which(!is.na(x$type)),])+aes(x=x,y=y,color=type) + geom_line(show.legend = F,aes(group=model)) + theme_classic() + geom_smooth() + geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
```

```{r message=FALSE, warning=FALSE,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/ROC_article_boostraping_model_type_OS.pdf',width = 6,height = 3)
ggplot(data=x[which(!is.na(x$type)),])+aes(x=x,y=y,color=type) + geom_line(show.legend = F,aes(group=model),alpha=0.4) + 
  theme_classic() + geom_smooth() + geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
dev.off()
```

```{r message=FALSE, warning=FALSE,fig.width=4,fig.height=3}
ggplot(data=x)+aes(x=x,y=y,color=type) + geom_line(show.legend = F,aes(group=model)) + theme_classic() + #geom_smooth() + 
  geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
```

```{r message=FALSE, warning=FALSE,fig.width=6,fig.height=3}
ggplot(data=x)+aes(x=x,y=y,color=type) + geom_line(show.legend = T,aes(group=model)) + theme_classic() + #geom_smooth() + 
  geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
```

```{r message=FALSE, warning=FALSE,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/ROC_article_boostraping_model_type_full_OS.pdf',width = 4,height = 3)
ggplot(data=x)+aes(x=x,y=y,color=type) + geom_line(show.legend = F,aes(group=model)) + theme_classic() + #geom_smooth() + 
  geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
dev.off()
```

```{r message=FALSE, warning=FALSE,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/ROC_article_boostraping_model_type_full_w_labels_OS.pdf',width = 6,height = 3)
ggplot(data=x)+aes(x=x,y=y,color=type) + geom_line(show.legend = T,aes(group=model)) + theme_classic() + #geom_smooth() + 
  geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
dev.off()
```

```{r message=FALSE, warning=FALSE,fig.width=3,fig.height=3}
ggplot(data=unique(x[,colnames(x) %in% c('model','type','auc')])) + aes(x=type,y=auc)+geom_boxplot()+theme_classic()+rotate_x_text(angle = 45)+
  labs(x='Model Type',y='Area Under the Curve',title = '') + ylim(0,1)
```

```{r message=FALSE, warning=FALSE,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/Boxplot_AUC_article_boostraping_model_type_OS.pdf',width = 3,height = 3)
ggplot(data=unique(x[,colnames(x) %in% c('model','type','auc')])) + aes(x=type,y=auc)+geom_boxplot()+theme_classic()+rotate_x_text(angle = 45) + 
  labs(x='Model Type',y='Area Under the Curve',title = '') + ylim(0,1)
dev.off()
```

### Compare the different populations

```{r message=FALSE, warning=FALSE}
x$class <- 'Integrative'
x$class[grep('^BEry',x$model)] <- 'Bery'
x$class[grep('Plasma',x$model)] <- 'Plasma'
x$class[grep('NkT',x$model,ignore.case = T)] <- 'NkT'
x$class[grep('Myeloid',x$model)] <- 'Myeloid'
x$class[grep('^Ery',x$model)] <- 'Ery'
x$class[grep('^Full',x$model)] <- 'Fibroblasts'
x$class[grep('Cyto',x$model,ignore.case = T)] <- 'Cytogenetic'
x$class[grep('iss',x$model,ignore.case = T)] <- 'ISS'
```

```{r message=FALSE, warning=FALSE,fig.width=3,fig.height=3}
ggplot(data=unique(x[,colnames(x) %in% c('model','class','type','auc')])) + aes(x=reorder(class,auc,median),y=auc)+geom_boxplot()+theme_classic()+rotate_x_text(angle = 45)+
  labs(x='Model Class',y='AUC')+ ylim(0,1)
```

```{r message=FALSE, warning=FALSE,fig.width=3,fig.height=3}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/Boxplot_AUC_article_boostraping_model_class_OS.pdf',width = 3,height = 3)
ggplot(data=unique(x[,colnames(x) %in% c('model','class','type','auc')])) + aes(x=reorder(class,auc,median),y=auc)+geom_boxplot()+theme_classic()+rotate_x_text(angle = 45)+
  labs(x='Model Class',y='AUC')+ ylim(0,1)
dev.off()
```

```{r message=FALSE, warning=FALSE,fig.width=9,fig.height=3}
z<-unique(x[,colnames(x) %in% c('model','class','type','auc')])
iii <- which(!z$class %in% 'Integrative')
z <- z[iii,]
ggplot(data=z) + aes(x=reorder(class,auc,median),y=auc)+geom_boxplot()+
  theme_classic()+rotate_x_text(angle = 45)+facet_wrap(~type,ncol=6) + labs(y='AUC', x='Model Type * Compartment')
```

```{r message=FALSE, warning=FALSE,fig.width=7,fig.height=3,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/Boxplot_AUC_article_boostraping_model_type_compartment_OS.pdf',width = 9,height = 3)
ggplot(data=z) + aes(x=reorder(class,auc,median),y=auc)+geom_boxplot()+
  theme_classic()+rotate_x_text(angle = 45)+facet_wrap(~type,ncol=6) + labs(y='AUC', x='Model Type * Compartment')
dev.off()
```

```{r message=FALSE, warning=FALSE}
x$subcluster <- tidyr::separate(data.frame(x$model),1, sep="_", c("a","b",'c','d'))$a
z$subcluster <- tidyr::separate(data.frame(z$model),1, sep="_", c("a","b",'c','d'))$a
```

```{r message=FALSE, warning=FALSE,fig.width=8,fig.height=4}
ggplot(data=z) + aes(x=reorder(subcluster,auc,median),y=auc,color=type)+
  geom_point()+theme_classic()+
  rotate_x_text(angle = 45)+facet_wrap(~class,ncol=4, scales = 'free_x')
```

```{r message=FALSE, warning=FALSE}
predictions_list_OS <- predictions_list
pred_validation_list_OS <- pred_validation_list
#
predictions_list <- list()
pred_validation_list <- list()
```

#### Models for Cell Populations and progression free survival (PFS) (lrms)

#### clinical lrms

```{r message=FALSE, warning=FALSE}

mod.bi <- lrm(censpfs ~ d_dx_amm_iss_stage, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V1 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V1 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ d_dx_amm_iss_stage + d_dx_amm_age, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V2 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V2 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ d_dx_amm_iss_stage + d_amm_tx_asct_1st, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V3 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V3 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ d_dx_amm_iss_stage + d_amm_tx_asct_1st + d_dx_amm_age + d_pt_sex, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V4 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V4 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ d_dx_amm_iss_stage + d_amm_tx_asct_1st + d_dx_amm_age + d_pt_sex + batch + site, subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list$mod_iss_V5 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_iss_V5 <- get_lrm_validation(mod.bi)
```

#### davies lrms

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(censpfs ~ davies_based_risk + d_pt_sex, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V1 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V1 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ davies_based_risk + d_dx_amm_age, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V2 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V2 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ davies_based_risk + d_dx_amm_age + d_pt_sex, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V3 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V3 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ davies_based_risk + d_dx_amm_age + d_pt_sex + batch + site, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V4 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V4 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ davies_based_risk + d_amm_tx_asct_1st + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_cyto_V5 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_cyto_V5 <- get_lrm_validation(mod.bi)
```

#### IMWG_2024_HR

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(censpfs ~ IMWG_2024_HR + d_pt_sex, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V1 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V1 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ IMWG_2024_HR + d_dx_amm_age, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V2 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V2 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V3 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V3 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V4 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V4 <- get_lrm_validation(mod.bi)

mod.bi <- lrm(censpfs ~ IMWG_2024_HR + d_amm_tx_asct_1st + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage, subcluster_all_fulldf, x=TRUE, y=TRUE)
predictions_list$mod_NC_V5 <- get_auc2(mod.bi,subcluster_all_fulldf)
pred_validation_list$mod_NC_V5 <- get_lrm_validation(mod.bi)
```

#### Sort Subpops present in at least more than 50 patients

```{r message=FALSE, warning=FALSE}
my_subcluster_names <- colnames(subcluster_all_fulldf)[16:104]
x <- list()
for(iii in my_subcluster_names){
  x[[iii]]<- length(table(subcluster_all_fulldf[[iii]]))
}
my_subcluster_names <- names(sort(unlist(x))[sort(unlist(x)) > 50])
###
ix <- which(colnames(subcluster_all_fulldf) %in% my_subcluster_names)
###
my_subcluster_names <- names(colSums(subcluster_all_fulldf[,ix])[colSums(subcluster_all_fulldf[,ix]) >1])
```

#### formulas

```{r message=FALSE, warning=FALSE}
my_form_V1 <- list()
for(iii in my_subcluster_names){ my_form_V1[[iii]] <- as.formula(paste("censpfs ~", iii)) }

my_form_V2 <- list()
for(iii in my_subcluster_names){ my_form_V2[[iii]] <- as.formula(paste("censpfs ~ d_dx_amm_age +", iii)) }

my_form_V2 <- list()
for(iii in my_subcluster_names){ my_form_V2[[iii]] <- as.formula(paste("censpfs ~ d_dx_amm_age + d_pt_sex +", iii)) }

my_form_V3 <- list()
for(iii in my_subcluster_names){ my_form_V3[[iii]] <- as.formula(paste("censpfs ~ d_dx_amm_age + d_pt_sex + batch + site +", iii)) }

my_form_V4 <- list()
for(iii in my_subcluster_names){ my_form_V4[[iii]] <- as.formula(paste("censpfs ~ d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +", iii)) }

my_form_V5 <- list()
for(iii in my_subcluster_names){ my_form_V5[[iii]] <- as.formula(paste("censpfs ~ davies_based_risk + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +", iii)) }

my_form_V6 <- list()
for(iii in my_subcluster_names){ my_form_V6[[iii]] <- as.formula(paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +", iii)) }

```

#### models

```{r message=FALSE, warning=FALSE}
### V1
for(iii in names(my_form_V1)) {
  mod.bi <- lrm(my_form_V1[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V1",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)
  pred_validation_list[[paste(iii,"V1",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V2
for(iii in names(my_form_V2)) {
  mod.bi <- lrm(my_form_V2[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V2",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)
  pred_validation_list[[paste(iii,"V2",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V3
for(iii in names(my_form_V3)) {
  mod.bi <- lrm(my_form_V3[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V3",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
  pred_validation_list[[paste(iii,"V3",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V4
for(iii in names(my_form_V4)) {
  mod.bi <- lrm(my_form_V4[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V4",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
  pred_validation_list[[paste(iii,"V4",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V5
for(iii in names(my_form_V5)) {
  mod.bi <- lrm(my_form_V5[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V5",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
  pred_validation_list[[paste(iii,"V5",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

```{r message=FALSE, warning=FALSE}
### V6
for(iii in names(my_form_V6)) {
  mod.bi <- lrm(my_form_V6[[iii]], subcluster_all_fulldf, x=TRUE,y=TRUE)
  predictions_list[[paste(iii,"V6",sep='_')]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
  pred_validation_list[[paste(iii,"V6",sep='_')]] <- get_lrm_validation(mod.bi)
}
```

#### Feature Selection Model

```{r message=FALSE, warning=FALSE}
b <- paste( my_subcluster_names[!my_subcluster_names %in% c('')] ,'+') #NkT.8','NkT.7','NkT.11','NkT.9.0
a <- paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +") #
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
#subcluster_all_fulldf <- type_convert(subcluster_all_fulldf)
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
dim(subcluster_all_fulldf)
#vif(mod.bi)
predictions_list[["Feat_Sel"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel"]] <- get_lrm_validation(mod.bi)
```

```{r message=FALSE, warning=FALSE}
z <- as.data.frame(as.matrix(anova(mod.bi)))
z <- rownames(z)[z$P<0.1]
z <- z[z %in% my_subcluster_names]
z
Feat_Sel_R1_Features <- z
```

```{r message=FALSE, warning=FALSE}
b <- paste(z,'+')
a <- paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list[["Feat_Sel_R1"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel_R1"]] <- get_lrm_validation(mod.bi)
```

```{r message=FALSE, warning=FALSE}
z <- as.data.frame(as.matrix(anova(mod.bi)))
z <- rownames(z)[z$P<0.1]
z <- z[z %in% my_subcluster_names]
z
Feat_Sel_R2_Features <- z
```

```{r message=FALSE, warning=FALSE}
b <- paste(z,'+')
a <- paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list[["Feat_Sel_R2"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel_R2"]] <- get_lrm_validation(mod.bi)

```

```{r message=FALSE, warning=FALSE}
z <- as.data.frame(as.matrix(anova(mod.bi)))
z <- rownames(z)[z$P<0.1]
z <- z[z %in% my_subcluster_names]
z
Feat_Sel_R3_Features <- z
```

```{r message=FALSE, warning=FALSE}
b <- paste(z,'+')
a <- paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list[["Feat_Sel_R3"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel_R3"]] <- get_lrm_validation(mod.bi)
```

```{r message=FALSE, warning=FALSE}
z <- as.data.frame(as.matrix(anova(mod.bi)))
z <- rownames(z)[z$P<0.1]
z <- z[z %in% my_subcluster_names]
z
Feat_Sel_R4_Features <- z
```

```{r message=FALSE, warning=FALSE}
b <- paste(z,'+')
a <- paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)
```

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(as.formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
predictions_list[["Feat_Sel_R4"]] <- get_auc2(mod.bi,subcluster_all_fulldf)  
pred_validation_list[["Feat_Sel_R4"]] <- get_lrm_validation(mod.bi)
```

### Summary of predictions PFS

```{r message=FALSE, warning=FALSE}
predictions_PFS <- list() ; labels_PFS <- list() ; aucs_PFS<-list()
for(iii in names(predictions_list)){ 
  predictions_PFS[[iii]] <- predictions_list[[iii]]$predictions
  labels_PFS[[iii]] <- predictions_list[[iii]]$labels
  aucs_PFS[[iii]] <- predictions_list[[iii]]$auc
}
###
used_features_PFS<-list(R1=Feat_Sel_R1_Features,R2=Feat_Sel_R2_Features,
                       R3=Feat_Sel_R3_Features,R4=Feat_Sel_R4_Features)
###
key_predictions_PFS_list <- list(predictions=predictions_PFS,
                                 labels=labels_PFS,
                                 aucs=aucs_PFS,
                                 used_features=used_features_PFS)
```

```{r message=FALSE, warning=FALSE}
#saveRDS(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/RData/key_predictions_PFS_list.RDS',key_predictions_PFS_list)
```

```{r message=FALSE, warning=FALSE}
#save(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/RData/predictions_lists_os_pfs.RData',
#     predictions_list_OS,pred_validation_list_OS,predictions_list,pred_validation_list)
```


```{r message=FALSE, warning=FALSE}
pred <- prediction(key_predictions_PFS_list$predictions, key_predictions_PFS_list$labels)
perf <- performance(pred,'tpr','fpr')
#plot(perf,lwd=2,main='ROC curves from 10-fold cross-validation')
x<- perf@x.values
y<- perf@y.values
names(x) <- names(predictions_list)
names(y) <- names(predictions_list)
for(i in 1:length(x)){ x[[i]]<- data.frame(model=names(x)[i],x=x[[i]])}
x <- do.call(rbind,x)
x$y <- unlist(y)
rownames(x)<-NULL
### redo aucs
aucs <- list()
for(i in 1:length(predictions_list)) { aucs[[i]] <- data.frame(auc=predictions_list[[i]]$auc,model=names(predictions_list)[i]) }
aucs <- do.call(rbind,aucs)
x <- merge(x,aucs,by='model') ; rm(aucs)
#y <- unique(x[,c(1,4)])
#y <- y[order(y$auc),]
```


```{r message=FALSE, warning=FALSE}
x$type <- x$model
x$type[grep('cyto',x$model)] <- 'cyto'
x$type[grep('NC',x$model)] <- 'NC'
x$type[grep('V1',x$model)] <- 'Single'
x$type[grep('V2',x$model)] <- '+Age+Sex'
x$type[grep('V3',x$model)] <- '+Age+Sex+B*Site'
x$type[grep('V4',x$model)] <- '+Age+Sex+B*Site+ISS+ASTC'
x$type[grep('V5',x$model)] <- '+Cyto+Age+Sex+B*Site+ISS+ASTC'
x$type[grep('V6',x$model)] <- '+NC+Age+Sex+B*Site+ISS+ASTC'

x$type <- factor(x$type,levels = c('cyto','NC','Single','+Age+Sex','+Age+Sex+B*Site','+Age+Sex+B*Site+ISS+ASTC',
                                   '+Cyto+Age+Sex+B*Site+ISS+ASTC','+NC+Age+Sex+B*Site+ISS+ASTC',
                                   "Feat_Sel","Feat_Sel_R1","Feat_Sel_R2","Feat_Sel_R3","Feat_Sel_R4","mod_asct" ))
```


```{r message=FALSE, warning=FALSE,fig.width=5,fig.height=3}
ggplot(data=x[which(!is.na(x$type)),])+aes(x=x,y=y,color=type) + geom_line(show.legend = F,aes(group=model)) + theme_classic() + geom_smooth() + geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
```

```{r message=FALSE, warning=FALSE,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/ROC_article_boostraping_model_type_PFS.pdf',width = 6,height = 3)
ggplot(data=x[which(!is.na(x$type)),])+aes(x=x,y=y,color=type) + geom_line(show.legend = F,aes(group=model),alpha=0.4) + 
  theme_classic() + geom_smooth() + geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
dev.off()
```


```{r message=FALSE, warning=FALSE,fig.width=4,fig.height=3}
ggplot(data=x)+aes(x=x,y=y,color=type) + geom_line(show.legend = F,aes(group=model)) + theme_classic() + #geom_smooth() + 
  geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
```

```{r message=FALSE, warning=FALSE,fig.width=6,fig.height=3}
ggplot(data=x)+aes(x=x,y=y,color=type) + geom_line(show.legend = T,aes(group=model)) + theme_classic() + #geom_smooth() + 
  geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
```

```{r message=FALSE, warning=FALSE,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/ROC_article_boostraping_model_type_full_PFS.pdf',width = 4,height = 3)
ggplot(data=x)+aes(x=x,y=y,color=type) + geom_line(show.legend = F,aes(group=model)) + theme_classic() + #geom_smooth() + 
  geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
dev.off()
```

```{r message=FALSE, warning=FALSE,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/ROC_article_boostraping_model_type_full_w_labels_PFS.pdf',width = 6,height = 3)
ggplot(data=x)+aes(x=x,y=y,color=type) + geom_line(show.legend = T,aes(group=model)) + theme_classic() + #geom_smooth() + 
  geom_abline(color='black') +
  theme(legend.position = 'right') + labs(x='False Positive Rate',y='True Positive Rate',title = 'Receiver Operating Characteristic Curve')
dev.off()
```

```{r message=FALSE, warning=FALSE,fig.width=3,fig.height=3}
ggplot(data=unique(x[,colnames(x) %in% c('model','type','auc')])) + aes(x=type,y=auc)+geom_boxplot()+theme_classic()+rotate_x_text(angle = 45)+
  labs(x='Model Type',y='Area Under the Curve',title = '') + ylim(0,1)
```

```{r message=FALSE, warning=FALSE,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/Boxplot_AUC_article_boostraping_model_type_PFS.pdf',width = 3,height = 3)
ggplot(data=unique(x[,colnames(x) %in% c('model','type','auc')])) + aes(x=type,y=auc)+geom_boxplot()+theme_classic()+rotate_x_text(angle = 45) + 
  labs(x='Model Type',y='Area Under the Curve',title = '') + ylim(0,1)
dev.off()
```

### Compare the different populations

```{r message=FALSE, warning=FALSE}
x$class <- 'Integrative'
x$class[grep('^BEry',x$model)] <- 'Bery'
x$class[grep('Plasma',x$model)] <- 'Plasma'
x$class[grep('NkT',x$model,ignore.case = T)] <- 'NkT'
x$class[grep('Myeloid',x$model)] <- 'Myeloid'
x$class[grep('^Ery',x$model)] <- 'Ery'
x$class[grep('^Full',x$model)] <- 'Fibroblasts'
x$class[grep('Cyto',x$model,ignore.case = T)] <- 'Cytogenetic'
x$class[grep('iss',x$model,ignore.case = T)] <- 'ISS'
```

```{r message=FALSE, warning=FALSE,fig.width=3,fig.height=3}
ggplot(data=unique(x[,colnames(x) %in% c('model','class','type','auc')])) + aes(x=reorder(class,auc,median),y=auc)+geom_boxplot()+theme_classic()+rotate_x_text(angle = 45)+
  labs(x='Model Class',y='AUC')+ ylim(0,1)
```

```{r message=FALSE, warning=FALSE,fig.width=3,fig.height=3}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/Boxplot_AUC_article_boostraping_model_class_PFS.pdf',width = 3,height = 3)
ggplot(data=unique(x[,colnames(x) %in% c('model','class','type','auc')])) + aes(x=reorder(class,auc,median),y=auc)+geom_boxplot()+theme_classic()+rotate_x_text(angle = 45)+
  labs(x='Model Class',y='AUC')+ ylim(0,1)
dev.off()
```

```{r message=FALSE, warning=FALSE,fig.width=9,fig.height=3}
z<-unique(x[,colnames(x) %in% c('model','class','type','auc')])
iii <- which(!z$class %in% 'Integrative')
z <- z[iii,]
ggplot(data=z) + aes(x=reorder(class,auc,median),y=auc)+geom_boxplot()+
  theme_classic()+rotate_x_text(angle = 45)+facet_wrap(~type,ncol=6) + labs(y='AUC', x='Model Type * Compartment')
```

```{r message=FALSE, warning=FALSE,fig.width=7,fig.height=3,include=FALSE}
pdf(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/Boxplot_AUC_article_boostraping_model_type_compartment_PFS.pdf',width = 9,height = 3)
ggplot(data=z) + aes(x=reorder(class,auc,median),y=auc)+geom_boxplot()+
  theme_classic()+rotate_x_text(angle = 45)+facet_wrap(~type,ncol=6) + labs(y='AUC', x='Model Type * Compartment')
dev.off()
```

```{r message=FALSE, warning=FALSE}
x$subcluster <- tidyr::separate(data.frame(x$model),1, sep="_", c("a","b",'c','d'))$a
z$subcluster <- tidyr::separate(data.frame(z$model),1, sep="_", c("a","b",'c','d'))$a
```

```{r message=FALSE, warning=FALSE,fig.width=8,fig.height=4}
ggplot(data=z) + aes(x=reorder(subcluster,auc,median),y=auc,color=type)+
  geom_point()+theme_classic()+
  rotate_x_text(angle = 45)+facet_wrap(~class,ncol=4, scales = 'free_x')
```

#### Inspect Survival / Progression

```{r message=FALSE, warning=FALSE}
tmp_df <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),] 
forest_model( coxph(Surv(ttcos, censos==1) ~ d_pt_sex + d_dx_amm_age + d_dx_amm_iss_stage + d_amm_tx_asct_1st + site + batch, data = tmp_df) )
```

```{r message=FALSE, warning=FALSE}
ggColorHue(11)
```

#### Top Hits PFS

```{r message=FALSE, warning=FALSE}
x<-list()
for(i in 1:length(key_predictions_PFS_list$aucs)){
  x[[i]] <- data.frame(name=names(key_predictions_PFS_list$aucs)[i],auc=as.numeric(key_predictions_PFS_list$aucs[i])) 
}
x<-do.call(rbind,x)
x <- x[order(x$auc),]
```

```{r message=FALSE, warning=FALSE}
tail(x[grep('V1',x$name),])
```

```{r message=FALSE, warning=FALSE}
tail(x[grep('V6',x$name),])
```

#### Top Hits OS

```{r message=FALSE, warning=FALSE}
x<-list()
for(i in 1:length(key_predictions_OS_list$aucs)){
  x[[i]] <- data.frame(name=names(key_predictions_OS_list$aucs)[i],auc=as.numeric(key_predictions_OS_list$aucs[i])) 
}
x<-do.call(rbind,x)
x <- x[order(x$auc),]
```

```{r message=FALSE, warning=FALSE}
tail(x[grep('V1',x$name),])
```

```{r message=FALSE, warning=FALSE}
tail(x[grep('V6',x$name),])
```

#### Simple Figure with key models (OS)

#### Demographics alone + technical correction (OS)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list_OS$mod_iss_V5$predictions) > median(as.numeric(predictions_list_OS$mod_iss_V5$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_demographics+tech_OS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#64B200')) +labs(title = 'Age+Sex+ISS+ASCT')
dev.off()
```

#### Demographics alone (OS)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list_OS$mod_iss_V4$predictions) > median(as.numeric(predictions_list_OS$mod_iss_V4$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```


```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_demographics_OS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#64B200')) +labs(title = 'Age+Sex+ISS+ASCT')
dev.off()
```

#### Cytogenetics alone (OS)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list_OS$mod_NC_V1$predictions) > median(as.numeric(predictions_list_OS$mod_NC_V1$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_cytogenetics_OS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00BD5C')) +labs(title = 'CYTO')
dev.off()
```

#### Cytogenetics alone LABELS (OS)

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcos, censos==1) ~ IMWG_2024_HR, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
tmp_df_V3$p <- tmp_df_V3$IMWG_2024_HR
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_cytogenetics_labels_OS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00BD5C')) +labs(title = 'CYTO')
dev.off()
```

#### Cytogenetics alone (OS)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list_OS$mod_NC_V5$predictions) > median(as.numeric(predictions_list_OS$mod_NC_V5$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_cytogenetics+clinical_OS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00BD5C')) +labs(title = 'CYTO')
dev.off()
```

#### get best AI auc

```{r message=FALSE, warning=FALSE}
x<-list()
for(i in 1:length(predictions_list_OS)){
  x[[i]] <- data.frame(name=names(predictions_list_OS)[i],auc=predictions_list_OS[[i]]$auc) 
}
x<-do.call(rbind,x)
x <- x[order(x$auc),]
tail(x[grep('V1',x$name),])
```


#### IA alone

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list_OS$BEry.6_V1$predictions) > median(as.numeric(predictions_list_OS$BEry.6_V1$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA_alone_OS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+alone')
dev.off()
```

#### IA+COV

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list_OS$Myeloid.10_V6$predictions) > median(as.numeric(predictions_list_OS$Myeloid.10_V6$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA_+clinical_OS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+clinical')
dev.off()
```

#### IA+CYTO+CLINICAL

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list_OS$Feat_Sel_R2$predictions) > median(as.numeric(predictions_list_OS$Feat_Sel_R2$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA+CYTO+CLIN_OS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+clinical')
dev.off()
```


#### IA+CYTO+CLINICAL

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list_OS$Feat_Sel_R4$predictions) > median(as.numeric(predictions_list_OS$Feat_Sel_R4$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA+CYTO+CLIN_R4_OS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+clinical')
dev.off()
```




### Simple Figure with key models (PFS)

#### Demographics alone + technical correction (PFS)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list$mod_iss_V5$predictions) > median(as.numeric(predictions_list$mod_iss_V5$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_demographics+tech_PFS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#64B200')) +labs(title = 'Age+Sex+ISS+ASCT')
dev.off()
```

#### Demographics alone (PFS)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list$mod_iss_V4$predictions) > median(as.numeric(predictions_list$mod_iss_V4$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```


```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_demographics_PFS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#64B200')) +labs(title = 'Age+Sex+ISS+ASCT')
dev.off()
```

#### Cytogenetics alone (PFS)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list$mod_NC_V1$predictions) > median(as.numeric(predictions_list$mod_NC_V1$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_cytogenetics_PFS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00BD5C')) +labs(title = 'CYTO')
dev.off()
```

#### Cytogenetics alone LABELS (PFS)

```{r message=FALSE, warning=FALSE}
tmp_df_V3$IMWG_2024_HR <- factor(tmp_df_V3$IMWG_2024_HR, levels = c('standard_risk','high_risk') )
coxph(Surv(ttcpfs, censpfs==1) ~ IMWG_2024_HR, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
tmp_df_V3$p <- tmp_df_V3$IMWG_2024_HR
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_cytogenetics_labels_PFS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00BD5C')) +labs(title = 'CYTO')
dev.off()
```

#### Cytogenetics with clinical (PFS)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(predictions_list$mod_NC_V5$predictions) > median(as.numeric(predictions_list$mod_NC_V5$predictions))
tmp_df_V3$p[tmp_df_V3$p == '1'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == '0'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_cytogenetics+clinical_PFS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00BD5C')) +labs(title = 'CYTO')
dev.off()
```

#### IA alone

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(key_predictions_PFS_list$predictions$BEry.6_V1) > median(as.numeric(key_predictions_PFS_list$predictions$BEry.6_V1))
tmp_df_V3$p[tmp_df_V3$p == 'TRUE'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == 'FALSE'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA_alone_PFS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+alone')
dev.off()
```

#### IA+COV

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(key_predictions_PFS_list$predictions$Myeloid.8_V6) > median(as.numeric(key_predictions_PFS_list$predictions$Myeloid.8_V6))
tmp_df_V3$p[tmp_df_V3$p == 'TRUE'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == 'FALSE'] <- 'Low'
```

```{r message=FALSE, warning=FALSE}
tmp_df_V3$p <- factor(tmp_df_V3$p, levels = c('Low','High'))
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA_+clinical_PFS.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+clinical')
dev.off()
```

#### IA+CYTO+CLINICAL (ALL)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(key_predictions_PFS_list$predictions$Feat_Sel) > median(as.numeric(key_predictions_PFS_list$predictions$Feat_Sel))
tmp_df_V3$p[tmp_df_V3$p == 'TRUE'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == 'FALSE'] <- 'Low'
tmp_df_V3$p <- factor(tmp_df_V3$p, levels = c('Low','High'))
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA+CYTO+CLIN_PFS_ALL.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+CYTO+PFS')
dev.off()
```

#### IA+CYTO+CLINICAL (feat selection 1)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(key_predictions_PFS_list$predictions$Feat_Sel_R1) > median(as.numeric(key_predictions_PFS_list$predictions$Feat_Sel_R1))
tmp_df_V3$p[tmp_df_V3$p == 'TRUE'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == 'FALSE'] <- 'Low'
tmp_df_V3$p <- factor(tmp_df_V3$p, levels = c('Low','High'))
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA+CYTO+CLIN_PFS_FS1.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+CYTO+PFS')
dev.off()
```

#### IA+CYTO+CLINICAL (feat selection 2)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(key_predictions_PFS_list$predictions$Feat_Sel_R2) > median(as.numeric(key_predictions_PFS_list$predictions$Feat_Sel_R2))
tmp_df_V3$p[tmp_df_V3$p == 'TRUE'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == 'FALSE'] <- 'Low'
tmp_df_V3$p <- factor(tmp_df_V3$p, levels = c('Low','High'))
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA+CYTO+CLIN_PFS_FS2.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+CYTO+PFS')
dev.off()
```

#### IA+CYTO+CLINICAL (feat selection 3)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(key_predictions_PFS_list$predictions$Feat_Sel_R3) > median(as.numeric(key_predictions_PFS_list$predictions$Feat_Sel_R3))
tmp_df_V3$p[tmp_df_V3$p == 'TRUE'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == 'FALSE'] <- 'Low'
tmp_df_V3$p <- factor(tmp_df_V3$p, levels = c('Low','High'))
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA+CYTO+CLIN_PFS_FS3.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+CYTO+PFS')
dev.off()
```

#### IA+CYTO+CLINICAL (feat selection 4)

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
tmp_df_V3$p <- as.numeric(key_predictions_PFS_list$predictions$Feat_Sel_R4) > median(as.numeric(key_predictions_PFS_list$predictions$Feat_Sel_R4))
tmp_df_V3$p[tmp_df_V3$p == 'TRUE'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == 'FALSE'] <- 'Low'
tmp_df_V3$p <- factor(tmp_df_V3$p, levels = c('Low','High'))
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA+CYTO+CLIN_PFS_FS4.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+CYTO+PFS')
dev.off()
```

```{r message=FALSE, warning=FALSE}
key_predictions_PFS_list$used_features$R4
```



### Validation dataset (PFS)

```{r message=FALSE, warning=FALSE}
ddist <- datadist(subcluster_all_fulldf)
options(datadist='ddist')
```

```{r message=FALSE, warning=FALSE}
#validation_dataset_w_meta
### formula
b <- paste(key_predictions_PFS_list$used_features$R4,'+')
a <- paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage + d_amm_tx_asct_1st +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)

### model
mod.bi <- lrm(formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
```

```{r message=FALSE, warning=FALSE}
predictions <- predict(object = mod.bi, 
                       subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),-which(colnames(subcluster_all_fulldf) %in% 'censpfs')],
                       se.fit = TRUE,type="fitted.ind")

labels <- subcluster_all_fulldf$censpfs[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage))]
pred <- prediction(predictions, labels)
perf <- performance(pred,'tpr','fpr')
performance(pred, measure = "auc")@y.values[[1]]  
```

```{r message=FALSE, warning=FALSE}
#validation_dataset_w_meta
### formula
b <- paste(key_predictions_PFS_list$used_features$R4,'+')
a <- paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)

### model
mod.bi <- lrm(formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
```

```{r message=FALSE, warning=FALSE}
predictions <- predict(object = mod.bi, 
                       subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),-which(colnames(subcluster_all_fulldf) %in% 'censpfs')],
                       se.fit = TRUE,type="fitted.ind")

labels <- subcluster_all_fulldf$censpfs[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage))]
pred <- prediction(predictions, labels)
perf <- performance(pred,'tpr','fpr')
performance(pred, measure = "auc")@y.values[[1]]  
```

```{r message=FALSE, warning=FALSE}
#validation_dataset_w_meta
subcluster_all_fulldf$sitexbatch <- paste(subcluster_all_fulldf$site,subcluster_all_fulldf$batch,sep='_')
### formula
full_form <- "censpfs ~ davies_based_risk + d_dx_amm_age + d_pt_sex + d_dx_amm_iss_stage + 
              NkT.2.0 + Myeloid.1 + BEry.12 + Plasma.4 + NkT.2.1 + Plasma.2 + Myeloid.10 + NkT.2.2 + Myeloid.8 + Myeloid.4 + NkT.14"
### model
mod.bi <- lrm(formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
```

```{r message=FALSE, warning=FALSE}
predictions <- predict(object = mod.bi, 
                       subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),-which(colnames(subcluster_all_fulldf) %in% 'censpfs')],
                       se.fit = TRUE,type="fitted.ind")

labels <- subcluster_all_fulldf$censpfs[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage))]
pred <- prediction(predictions, labels)
perf <- performance(pred,'tpr','fpr')
performance(pred, measure = "auc")@y.values[[1]]  
```

```{r message=FALSE, warning=FALSE}
validation_dataset_w_meta$DISCOVERY.doublet_pred <- NULL
validation_dataset_w_meta$sample_id <- NULL
validation_dataset_w_meta$AT_RESPONSEASSES <- NULL
validation_dataset_w_meta$D_IM_CD81_DETECTED <- NULL
validation_dataset_w_meta$D_IM_CD81_PC_PERCENT <- NULL
validation_dataset_w_meta$batch <- 'B4'
validation_dataset_w_meta$site <- gsub('_B4','',validation_dataset_w_meta$siteXbatch)

validation_dataset_w_meta <- validation_dataset_w_meta %>% select_if(~!all(is.na(.)))
validation_dataset_w_meta <- as.data.frame(validation_dataset_w_meta)
validation_dataset_w_meta$sitexbatch <- paste(validation_dataset_w_meta$site,validation_dataset_w_meta$batch,sep='_')
```

```{r message=FALSE, warning=FALSE}
ddist <- datadist(validation_dataset_w_meta)
options(datadist='ddist')
```

```{r message=FALSE, warning=FALSE}
mod.bi <- lrm(formula(full_form), validation_dataset_w_meta[which(!is.na(validation_dataset_w_meta$d_dx_amm_iss_stage) & 
                                                         !(validation_dataset_w_meta$IMWG_2024_HR %in% 'not_calculable')),], x=TRUE,y=TRUE)
```

```{r message=FALSE, warning=FALSE}
predictions <- predict(object = mod.bi, 
                       validation_dataset_w_meta[which(!is.na(validation_dataset_w_meta$d_dx_amm_iss_stage) & 
                                                         !(validation_dataset_w_meta$IMWG_2024_HR %in% 'not_calculable')),-which(colnames(validation_dataset_w_meta) %in% 'censpfs')],
                       se.fit = TRUE,type="fitted.ind")

labels <- validation_dataset_w_meta$censpfs[which(!is.na(validation_dataset_w_meta$d_dx_amm_iss_stage) & 
                                                         !(validation_dataset_w_meta$IMWG_2024_HR %in% 'not_calculable'))]
pred <- prediction(predictions, labels)
perf <- performance(pred,'tpr','fpr')
performance(pred, measure = "auc")@y.values[[1]]  
```

```{r message=FALSE, warning=FALSE}
ddist <- datadist(subcluster_all_fulldf)
options(datadist='ddist')

predictions <- predict(object = mod.bi, 
                       subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),],
                       se.fit = TRUE,type="fitted.ind")

labels <- subcluster_all_fulldf$censpfs[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage))]
pred <- prediction(predictions, labels)
perf <- performance(pred,'tpr','fpr')
performance(pred, measure = "auc")@y.values[[1]]  
```

```{r message=FALSE, warning=FALSE}
tmp_df_V3 <- validation_dataset_w_meta[which(!is.na(validation_dataset_w_meta$d_dx_amm_iss_stage) & 
                                                         !(validation_dataset_w_meta$IMWG_2024_HR %in% 'not_calculable')),]

tmp_df_V3$p <- as.numeric(predictions) > median(as.numeric(predictions))
tmp_df_V3$p[tmp_df_V3$p == 'TRUE'] <- 'High'
tmp_df_V3$p[tmp_df_V3$p == 'FALSE'] <- 'Low'
tmp_df_V3$p <- factor(tmp_df_V3$p, levels = c('Low','High'))
```

```{r message=FALSE, warning=FALSE}
coxph(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r message=FALSE, warning=FALSE}
#pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA+CYTO+CLIN_PFS_FS4.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ p, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'IA+CYTO+PFS')
dev.off()
```
```{r message=FALSE, warning=FALSE}
#pdf(file = '/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/figures/km_labels_model_IA+CYTO+CLIN_PFS_FS4.pdf',width = 3.85, height = 4.95)
ggsurvplot( fit = survfit(Surv(ttcpfs, censpfs==1) ~ d_amm_tx_asct_1st, data = tmp_df_V3), 
            type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "survdiff",pval.method = TRUE,
            xlab = "Days", ylab = "PFS Prob.",pval = TRUE,risk.table = TRUE,palette = c('black','#00C1A7')) +labs(title = 'ASCT')
#dev.off()
```

```{r message=FALSE, warning=FALSE}
#validation_dataset_w_meta
### formula
b <- paste(key_predictions_PFS_list$used_features$R1,'+')
a <- paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)

### model
mod.bi <- lrm(formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
```

```{r message=FALSE, warning=FALSE}
predictions <- predict(object = mod.bi, 
                       subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),-which(colnames(subcluster_all_fulldf) %in% 'censpfs')],
                       se.fit = TRUE,type="fitted.ind")

labels <- subcluster_all_fulldf$censpfs[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage))]
pred <- prediction(predictions, labels)
perf <- performance(pred,'tpr','fpr')
performance(pred, measure = "auc")@y.values[[1]]  
```

```{r message=FALSE, warning=FALSE}
validation_dataset_w_meta$batch <- 'B4'
validation_dataset_w_meta$site <- gsub('_B4','',validation_dataset_w_meta$siteXbatch)

predictions <- predict(object = mod.bi, 
                       validation_dataset_w_meta[which(!is.na(validation_dataset_w_meta$d_dx_amm_iss_stage) & 
                                                         !(validation_dataset_w_meta$IMWG_2024_HR %in% 'not_calculable')),-which(colnames(validation_dataset_w_meta) %in% 'censpfs')],
                       se.fit = TRUE,type="fitted.ind")

labels <- validation_dataset_w_meta$censpfs[which(!is.na(validation_dataset_w_meta$d_dx_amm_iss_stage) & 
                                                         !(validation_dataset_w_meta$IMWG_2024_HR %in% 'not_calculable'))]
pred <- prediction(predictions, labels)
perf <- performance(pred,'tpr','fpr')
performance(pred, measure = "auc")@y.values[[1]]  

```

```{r message=FALSE, warning=FALSE}
### formula
b <- paste( my_subcluster_names[!my_subcluster_names %in% c('')] ,'+') 
a <- paste("censpfs ~ IMWG_2024_HR + d_dx_amm_age + d_pt_sex + batch + site + d_dx_amm_iss_stage +")
full_form <- capture.output(cat(a,b))
full_form <- gsub(' \\+$','',full_form)

### model
mod.bi <- lrm(formula(full_form), subcluster_all_fulldf, x=TRUE,y=TRUE)
```

```{r message=FALSE, warning=FALSE}
predictions <- predict(object = mod.bi, 
                       subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),-which(colnames(subcluster_all_fulldf) %in% 'censpfs')],
                       se.fit = TRUE,type="fitted.ind")

labels <- subcluster_all_fulldf$censpfs[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage))]
pred <- prediction(predictions, labels)
perf <- performance(pred,'tpr','fpr')
performance(pred, measure = "auc")@y.values[[1]]  
```

```{r message=FALSE, warning=FALSE}
validation_dataset_w_meta$batch <- 'B4'
validation_dataset_w_meta$site <- gsub('_B4','',validation_dataset_w_meta$siteXbatch)

predictions <- predict(object = mod.bi, 
                       validation_dataset_w_meta[which(!is.na(validation_dataset_w_meta$d_dx_amm_iss_stage) & 
                                                         !(validation_dataset_w_meta$IMWG_2024_HR %in% 'not_calculable')),-which(colnames(validation_dataset_w_meta) %in% 'censpfs')],
                       se.fit = TRUE,type="fitted.ind")

labels <- validation_dataset_w_meta$censpfs[which(!is.na(validation_dataset_w_meta$d_dx_amm_iss_stage) & 
                                                         !(validation_dataset_w_meta$IMWG_2024_HR %in% 'not_calculable'))]
pred <- prediction(predictions, labels)
perf <- performance(pred,'tpr','fpr')
performance(pred, measure = "auc")@y.values[[1]]  

```

```{r message=FALSE, warning=FALSE}
final_discovery_cohort <- subcluster_all_fulldf[which(!is.na(subcluster_all_fulldf$d_dx_amm_iss_stage)),]
final_validation_cohort <- validation_dataset_w_meta[which(!is.na(validation_dataset_w_meta$d_dx_amm_iss_stage) & 
                                                         !(validation_dataset_w_meta$IMWG_2024_HR %in% 'not_calculable')),]
###
# saveRDS(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/RData/final_discovery_cohort.RDS',final_discovery_cohort)
# saveRDS(file='/Users/gonzae34/Documents/projects_gnjatic/MMRF_projects/immune_atlas/RData/final_validation_cohort.RDS',final_validation_cohort)
```



```{r message=FALSE, warning=FALSE}
```
