---
title: "Figure 4 - Survival"
author: "William Pilcher"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
---

Computes survival curves for select CD3+ T cell markers (Figures 4L, M, N)

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Load object and selected signatures
```{r init}
library(UCell)
source("source/Validation/validation_load_data.R", local = T)

nkt <- merged |> subset(subset = (lineage_group %in% c("CD8", "CD4") &
    VJ_INTERVAL == "Baseline"))
rm(merged)
gc()
```

```{r}
color_low <- "#2c90c3"
color_high <- "#5a221b"

output_surv <- file.path(output, "Validation", "CytoTox_HeatMap", "ALL_PAT")
dir.create(output_surv, recursive = T, showWarnings = F)

sig <- readxl::read_xlsx(file.path("/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation", "data/resources/pan-cancer_tcell_atlas_chu_et_al.xlsx"), sheet = "Table S4", skip = 1, )
sig_cd4 <- readxl::read_xlsx(file.path("/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation", "data/resources/pan-cancer_tcell_atlas_chu_et_al.xlsx"), sheet = "Table S6", skip = 1, )

combined_naive <- unique(c(sig_cd4[[1]] |> na.omit(), sig[[1]] |> na.omit()))
combined_cyto <- unique(c(sig_cd4[[5]] |> na.omit(), sig[[5]] |> na.omit()))
combined_exh <- unique(c(sig_cd4[[3]] |> na.omit(), sig[[3]] |> na.omit()))

list_of_signatures <- list("Naive" = combined_naive, "Cytotoxic" = combined_cyto, "Exhaustion" = combined_exh)

valid_sig <- make.names(names(list_of_signatures))
```

# Signatures, Survival
UCell computes a signature score which is on a range between 0 and 1. These will be mutlipled by 100 when fitting the CoxPH model, so that the HR can be interpretted as the increase in risk per 1 UCell signature score point increase
```{r}
nkt <- AddModuleScore_UCell(nkt, features = list_of_signatures, name = "") # AddModuleScore(nkt, features = list_of_signatures, name = valid_sig)

averageMod <- nkt@meta.data %>%
    dplyr::group_by(public_id) %>%
    dplyr::summarise(
        Naive = mean(!!sym(paste0(valid_sig[[1]]))),
        Cytotoxicity = mean(!!sym(paste0(valid_sig[[2]]))),
        Exhaustion = mean(!!sym(paste0(valid_sig[[3]]))),
    )

nkt@meta.data <- nkt@meta.data |> dplyr::mutate(maf_combined_call = dplyr::case_when(
    maf_call == "yes" | mafa_call == "yes" | mafb_call == "yes" ~ "yes",
    maf_call == "no" & mafa_call == "no" & mafb_call == "no" ~ "no",
    is.na(maf_call) | is.na(mafa_call) | is.na(mafb_call) ~ NA,
    .default = "ERROR"
))

subset_patMD <- nkt@meta.data |>
    dplyr::filter(d_lot_1_start_day < 2) |>
    dplyr::distinct(public_id, progression_group, censpfs, pfscdy, d_pt_sex, d_dx_amm_age, d_dx_amm_iss_stage, siteXbatch, d_tx_induction_cat, d_amm_tx_asct_1st, cohort, maf_combined_call, nsd2_call, seqwgs_cp_17p13_call, seqwgs_cp_1q21_call)

patMD_averageMod <- dplyr::left_join(subset_patMD, averageMod, by = "public_id")
```


```{r}
gene_avg_expr <- AverageExpression(nkt, assays = "RNA", group.by = c("public_id"), features = c(
    "GNLY", "PRF1", "GZMB", "GZMH", "FGFBP2", "IFNG", "CD74", "CST7", "B3GAT1", "ZEB2", "KLRG1", "PDCD1", "TIGIT", "TOX", "LAG3", "CCR7", "SELL", "TCF7", "LTB", "IL7R", "LEF1", "CD27", "CD28"
))$RNA
# colnames(avg_expr) <- gsub("-","_", colnames(avg_expr))
```

```{r}
gene_avg_expr_df <- as.data.frame(gene_avg_expr)
gene_avg_expr_df$gene <- rownames(gene_avg_expr_df)

gene_avg_expr_df <- gene_avg_expr_df |> tidyr::pivot_longer(cols = colnames(gene_avg_expr), names_to = "name", values_to = "Expression")

gene_avg_expr_df <- gene_avg_expr_df |> tidyr::separate(name, sep = "_", c("public_id")) # dplyr::mutate(public_id = str_split(name, "_")[[1]][1], lineage_group = str_split(name, "_")[[1]][2])
gene_avg_expr_df$public_id <- gsub("-", "_", gene_avg_expr_df$public_id)
gene_avg_expr_df$Module <- gene_avg_expr_df$gene
gene_avg_expr_df$Score <- gene_avg_expr_df$Expression

diff_expr_avg <- dplyr::left_join(gene_avg_expr_df, subset_patMD, by = "public_id")

gene_pat_mean <- diff_expr_avg |>
    dplyr::group_by(Module) |>
    dplyr::summarize(mean_module = mean(Score), var_module = var(Score))
```

```{r}
# pat_mean <- patMD_averageMod |>
#     dplyr::group_by(Module) |>
#     dplyr::summarize(mean_module = mean(Score), var_module = var(Score))
```


```{r}
tmp <- diff_expr_avg |>
    dplyr::left_join(gene_pat_mean, by = c("Module")) |>
    group_by(Module) |>
    dplyr::mutate(z = (Score - mean_module) / sd(Score), min_max = 2 * ((Score - min(Score)) / max(Score - min(Score)) - 0.5))


tmp_b <- patMD_averageMod |>
    dplyr::mutate(z_cyto = (Cytotoxicity - mean(Cytotoxicity)) / sd(Cytotoxicity), z_naive = (Naive - mean(Naive)) / sd(Naive))



global_mean <- tmp |>
    group_by(public_id)

tmp_b$progression_group <- factor(tmp_b$progression_group, levels = c("NP", "P", "Inc", "RP"))

tmp_b <- tmp_b |>
    dplyr::arrange(progression_group, Cytotoxicity)

global_mean$public_id <- factor(global_mean$public_id, levels = tmp_b$public_id)
global_mean$Module <- factor(global_mean$Module, levels = rev(c("GNLY", "PRF1", "GZMB", "GZMH", "FGFBP2", "CST7", "IFNG", "CD74", "B3GAT1", "ZEB2", "KLRG1", "PDCD1", "TIGIT", "TOX", "LAG3", "CCR7", "SELL", "TCF7", "LTB", "IL7R", "LEF1", "CD27", "CD28")))
global_mean <- global_mean |> dplyr::filter(Module %in% c("GNLY", "PRF1", "GZMH", "CST7", "CCR7", "SELL", "TCF7", "LTB", "IL7R", "LEF1", "KLRG1", "CD27", "CD28"))
tmp_b$public_id <- factor(tmp_b$public_id, levels = tmp_b$public_id)

global_mean <- global_mean |> dplyr::filter(progression_group %in% c("NP", "RP"), cohort == "validation")
tmp_b <- tmp_b |> dplyr::filter(progression_group %in% c("NP", "RP"), cohort == "validation")

all_order <- global_mean |>
    reshape2::acast(public_id ~ Module, value.var = "z")

np_order <- global_mean |>
    dplyr::filter(progression_group == "NP") |>
    reshape2::acast(public_id ~ Module, value.var = "z")
rp_order <- global_mean |>
    dplyr::filter(progression_group == "RP") |>
    reshape2::acast(public_id ~ Module, value.var = "z")

d_uni_factor <- dist(t(all_order) / sqrt(rowSums(t(all_order) * t(all_order))), method = "euclidean")
uni_dendrogram_factor <- hclust(d_uni_factor)
gene_ordering <- uni_dendrogram_factor$labels[uni_dendrogram_factor$order]

d_uni_cluster <- dist(np_order / sqrt(rowSums(np_order * np_order)), method = "euclidean")
uni_dendrogram_clusters <- hclust(d_uni_cluster)
np_patient_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]


d_uni_cluster <- dist(rp_order / sqrt(rowSums(rp_order * rp_order)), method = "euclidean")
uni_dendrogram_clusters <- hclust(d_uni_cluster)
rp_patient_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]

patient_ordering <- c(np_patient_ordering, rp_patient_ordering)


global_mean$Module <- factor(global_mean$Module, levels = rev(gene_ordering))
global_mean$public_id <- factor(global_mean$public_id, levels = patient_ordering)
tmp_b$public_id <- factor(tmp_b$public_id, levels = patient_ordering)



RATIO_PAT_INFO <- 1
p <- ggplot(global_mean, x = public_id, y = Module, fill = z) +
    geom_tile(aes(x = public_id, y = Module, fill = z), color = "#484848", lwd = 0.1, linetype = 1) +
    scale_fill_gradient2(
        low = "#1d9ae3",
        high = "#f63575",
        limits = range(c(-2, 2)),
        oob = squish
    ) +
    theme_prism() +
    xlab("Patient") +
    ylab("Gene") +
    theme(
        plot.margin = margin(
            14 / 2, 14 / 2,
            0, 14 / 2
        ),
        axis.text.x = element_blank(),
        plot.title = element_blank(),
        axis.ticks.x = element_blank()
    ) +
    coord_fixed(ratio = 2) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    theme(legend.position = "bottom") +
    scale_y_discrete(expand = c(0, 0))

p_cyto <- ggplot(tmp_b, x = public_id, y = "Cytotoxicity", fill = z_cyto) +
    geom_tile(aes(x = public_id, y = "Cytotoxicity", fill = z_cyto), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_gradient2(
        low = "#1d9ae3",
        high = "#f63575",
        limits = range(c(-2, 2)),
        oob = squish
    ) +
    theme_prism() +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank()
        # axis.text.y = element_blank(),
        # axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +
    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    theme(legend.position = "bottom") +
    scale_y_discrete(expand = c(0, 0))


p_naive <- ggplot(tmp_b, x = public_id, y = "Naive", fill = z_naive, show.legend = FALSE) +
    geom_tile(aes(x = public_id, y = "Naive", fill = z_naive), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_gradient2(
        low = "#1d9ae3",
        high = "#f63575",
        limits = range(c(-3, 3)),
        oob = squish
    ) +
    theme_prism() +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0.5,
            unit = "lines"
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        # axis.text.y = element_blank(),
        # axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +
    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    theme(legend.position = "bottom") +
    scale_y_discrete(expand = c(0, 0))


p_sig <- ggplot(global_mean, x = public_id, y = "0", fill = progression_group, show.legend = F) +
    geom_tile(aes(x = public_id, y = "0", fill = progression_group), color = "#484848", lwd = 0.1, linetype = 1, show.legend = F) +
    scale_fill_manual(
        values = c("NP" = fill_NP, "RP" = fill_RP, "P" = "white", "Inc" = "grey")
    ) +
    theme_prism() +
    theme(
        plot.margin = margin(
            0.5, 0,
            0.5, 0,
            unit = "lines"
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))


p_tile_1q21 <- ggplot(global_mean, x = public_id, y = "Gain 1q21", fill = seqwgs_cp_1q21_call) +
    geom_tile(aes(x = public_id, y = "Gain 1q21", fill = seqwgs_cp_1q21_call), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_manual(
        values = c("no" = "white", "yes" = "#e48a04")
    ) +
    theme_prism() +
    xlab("Patient") +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))


p_tile_17p13 <- ggplot(global_mean, x = public_id, y = "Del 17p13", fill = seqwgs_cp_17p13_call) +
    geom_tile(aes(x = public_id, y = "Del 17p13", fill = seqwgs_cp_17p13_call), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_manual(
        values = c("no" = "white", "yes" = "#009bcf")
    ) +
    theme_prism() +
    xlab("Patient") +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))

p_tile_nsd2 <- ggplot(global_mean, x = public_id, y = "NSD2 t(4;14)", fill = nsd2_call) +
    geom_tile(aes(x = public_id, y = "NSD2 t(4;14)", fill = nsd2_call), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_manual(
        values = c("no" = "white", "yes" = "#c32400")
    ) +
    theme_prism() +
    xlab("Patient") +
    ggtitle("CD3+ T Cells") +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))

p_tile_maf <- ggplot(global_mean, x = public_id, y = "MAF, MAFA, MAFB", fill = maf_combined_call) +
    geom_tile(aes(x = public_id, y = "MAF, MAFA, MAFB", fill = maf_combined_call), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_manual(
        values = c("no" = "white", "yes" = "#00b02c")
    ) +
    theme_prism() +
    xlab("Patient") +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        plot.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +
    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))

TEST <- patchwork::wrap_plots(list(p_tile_nsd2, p_tile_1q21, p_tile_maf, p_tile_17p13, p_sig, p_cyto, p_naive, p), ncol = 1, nrow = 8)
pdf_and_png(TEST, output_surv, "val_cytotox_heatmap", pdfWidth = 9, pdfHeight = 5, scale = 2)
```

```{r}
test <- tmp_b
test$progression_group <- droplevels(test$progression_group)
pvals_wilcox <- test |>
    rstatix::t_test(Cytotoxicity ~ progression_group) |>
    rstatix::add_xy_position()


p <- ggplot(tmp_b, aes(x = progression_group, fill = progression_group, color = progression_group, y = Cytotoxicity)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    scale_fill_manual(values = c("NP" = fill_NP, "RP" = fill_RP)) +
    scale_color_manual(values = c("NP" = fill_NP, "RP" = fill_RP)) +
    theme_prism() +
    xlab("Progression Group") +
    ylab("Cytotoxicity") +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
```

```{r}
test <- tmp_b
test$progression_group <- droplevels(test$progression_group)
pvals_wilcox <- test |>
    rstatix::t_test(Naive ~ progression_group) |>
    rstatix::add_xy_position()


p <- ggplot(tmp_b, aes(x = progression_group, fill = progression_group, color = progression_group, y = Naive)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    scale_fill_manual(values = c("NP" = fill_NP, "RP" = fill_RP)) +
    scale_color_manual(values = c("NP" = fill_NP, "RP" = fill_RP)) +
    theme_prism() +
    xlab("Progression Group") +
    ylab("Naive") +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
```

```{r}
pid_tc <- nkt@meta.data |> dplyr::distinct(public_id, .keep_all = T)
pid_tc$public_id <- gsub("_", "-", pid_tc$public_id)
pseudo_tc <- AggregateExpression(nkt |> subset(lineage_group == "CD8"), assays = "RNA", return.seurat = T, group.by = c("public_id"))
pseudo_tc@meta.data$sample_name <- rownames(pseudo_tc@meta.data)
pseudo_tc@meta.data <- dplyr::left_join(pseudo_tc@meta.data, pid_tc, by = "public_id")
rownames(pseudo_tc@meta.data) <- pseudo_tc@meta.data$sample_name

counts <- GetAssayData(pseudo_tc, "RNA", "counts")

d0 <- DGEList(counts)
d0 <- calcNormFactors(d0)

cutoff <- 2
drop <- which(apply(cpm(d0), 1, max) < cutoff)
d <- d0[-drop, ]
dim(d) # number of genes left

mkrs <- list()
risk <- "progression_group"

md_bulk_tmp <- pseudo_tc@meta.data |> dplyr::filter(progression_group %in% c("NP", "RP"), VJ_INTERVAL == "Baseline", cohort == "validation")
md_bulk_tmp$public_id <- gsub("_", "-", md_bulk_tmp$public_id)
d_select <- d[, md_bulk_tmp$public_id]


mm <- model.matrix(as.formula(paste0("~ 0 + ", risk, "+ siteXbatch")), data = md_bulk_tmp)
y <- voom(d_select, mm, plot = T)

fit <- lmFit(y, mm)

contr <- makeContrasts(paste0(risk, "RP", " - ", risk, "NP"), levels = colnames(coef(fit)))
tmp <- contrasts.fit(fit, contr)
tmp <- eBayes(tmp)
top.table <- topTable(tmp, sort.by = "logFC", n = Inf) |> dplyr::arrange(desc(logFC))
top.table$gene <- rownames(top.table)
top.table <- top.table |> dplyr::relocate(gene)
mkrs[[risk]] <- top.table

write.csv(mkrs[[risk]], file.path(output_surv, "PB_val_rp_v_np.csv"))
```


```{r}
cd8_averageMod <- nkt@meta.data %>%
    dplyr::filter(lineage_group == "CD8") %>%
    dplyr::group_by(public_id) %>%
    dplyr::summarise(
        Naive = mean(!!sym(paste0(valid_sig[[1]]))),
        Cytotoxicity = mean(!!sym(paste0(valid_sig[[2]]))),
        Exhaustion = mean(!!sym(paste0(valid_sig[[3]]))),
    )

subset_patMD <- nkt@meta.data |>
    dplyr::filter(d_lot_1_start_day < 2) |>
    dplyr::distinct(public_id, progression_group, censpfs, pfscdy, d_pt_sex, d_dx_amm_age, d_dx_amm_iss_stage, siteXbatch, d_tx_induction_cat, d_amm_tx_asct_1st, cohort, maf_combined_call, nsd2_call, seqwgs_cp_17p13_call, seqwgs_cp_1q21_call)

patMD_averageMod <- dplyr::left_join(subset_patMD, averageMod, by = "public_id")

gene_avg_expr <- AverageExpression(nkt |> subset(lineage_group == "CD8"), assays = "RNA", group.by = c("public_id"), features = c(
    "GNLY", "PRF1", "GZMB", "GZMH", "FGFBP2", "IFNG", "CD74", "B3GAT1", "ZEB2", "KLRG1", "PDCD1", "TIGIT", "TOX", "LAG3", "CCR7", "SELL", "TCF7", "LTB", "IL7R", "LEF1"
))$RNA
# colnames(avg_expr) <- gsub("-","_", colnames(avg_expr))
```

```{r}
gene_avg_expr_df <- as.data.frame(gene_avg_expr)
gene_avg_expr_df$gene <- rownames(gene_avg_expr_df)

gene_avg_expr_df <- gene_avg_expr_df |> tidyr::pivot_longer(cols = colnames(gene_avg_expr), names_to = "name", values_to = "Expression")

gene_avg_expr_df <- gene_avg_expr_df |> tidyr::separate(name, sep = "_", c("public_id")) # dplyr::mutate(public_id = str_split(name, "_")[[1]][1], lineage_group = str_split(name, "_")[[1]][2])
gene_avg_expr_df$public_id <- gsub("-", "_", gene_avg_expr_df$public_id)
gene_avg_expr_df$Module <- gene_avg_expr_df$gene
gene_avg_expr_df$Score <- gene_avg_expr_df$Expression

diff_expr_avg <- dplyr::left_join(gene_avg_expr_df, subset_patMD, by = "public_id")

gene_pat_mean <- diff_expr_avg |>
    dplyr::group_by(Module) |>
    dplyr::summarize(mean_module = mean(Score), var_module = var(Score))
```

```{r}
# pat_mean <- patMD_averageMod |>
#     dplyr::group_by(Module) |>
#     dplyr::summarize(mean_module = mean(Score), var_module = var(Score))
```


```{r}
tmp <- diff_expr_avg |>
    dplyr::left_join(gene_pat_mean, by = c("Module")) |>
    group_by(Module) |>
    dplyr::mutate(z = (Score - mean_module) / sd(Score), min_max = 2 * ((Score - min(Score)) / max(Score - min(Score)) - 0.5))


tmp_b <- patMD_averageMod |>
    dplyr::mutate(z_cyto = (Cytotoxicity - mean(Cytotoxicity)) / sd(Cytotoxicity), z_naive = (Naive - mean(Naive)) / sd(Naive))



global_mean <- tmp |>
    group_by(public_id)

tmp_b$progression_group <- factor(tmp_b$progression_group, levels = c("NP", "P", "Inc", "RP"))

tmp_b <- tmp_b |>
    dplyr::arrange(progression_group, Cytotoxicity)

global_mean$public_id <- factor(global_mean$public_id, levels = tmp_b$public_id)
global_mean$Module <- factor(global_mean$Module, levels = rev(c("GNLY", "PRF1", "GZMB", "GZMH", "FGFBP2", "IFNG", "CD74", "B3GAT1", "ZEB2", "KLRG1", "PDCD1", "TIGIT", "TOX", "LAG3", "CCR7", "SELL", "TCF7", "LTB", "IL7R", "LEF1")))
global_mean <- global_mean |> dplyr::filter(Module %in% c("GNLY", "PRF1", "GZMH", "FGFBP2", "IFNG", "CCR7", "SELL", "TCF7", "LTB", "IL7R", "LEF1", "KLRG1"))
tmp_b$public_id <- factor(tmp_b$public_id, levels = tmp_b$public_id)

global_mean <- global_mean |> dplyr::filter(progression_group %in% c("NP", "RP"))
tmp_b <- tmp_b |> dplyr::filter(progression_group %in% c("NP", "RP"))

RATIO_PAT_INFO <- 1
p <- ggplot(global_mean, x = public_id, y = Module, fill = z) +
    geom_tile(aes(x = public_id, y = Module, fill = z), color = "#484848", lwd = 0.1, linetype = 1) +
    scale_fill_gradient2(
        low = "#1d9ae3",
        high = "#f63575",
        limits = range(c(-3, 3)),
        oob = squish
    ) +
    theme_prism() +
    xlab("Patient") +
    ylab("Gene") +
    theme(
        plot.margin = margin(
            14 / 2, 14 / 2,
            0, 14 / 2
        ),
        axis.text.x = element_blank(),
        plot.title = element_blank(),
        axis.ticks.x = element_blank()
    ) +
    coord_fixed(ratio = 2) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    theme(legend.position = "bottom") +
    scale_y_discrete(expand = c(0, 0))

p_cyto <- ggplot(tmp_b, x = public_id, y = "Cytotoxicity", fill = z_cyto) +
    geom_tile(aes(x = public_id, y = "Cytotoxicity", fill = z_cyto), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_gradient2(
        low = "#1d9ae3",
        high = "#f63575",
        limits = range(c(-3, 3)),
        oob = squish
    ) +
    theme_prism() +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank()
        # axis.text.y = element_blank(),
        # axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +
    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    theme(legend.position = "bottom") +
    scale_y_discrete(expand = c(0, 0))


p_naive <- ggplot(tmp_b, x = public_id, y = "Naive", fill = z_naive, show.legend = FALSE) +
    geom_tile(aes(x = public_id, y = "Naive", fill = z_naive), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_gradient2(
        low = "#1d9ae3",
        high = "#f63575",
        limits = range(c(-3, 3)),
        oob = squish
    ) +
    theme_prism() +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0.5,
            unit = "lines"
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        # axis.text.y = element_blank(),
        # axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +
    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    theme(legend.position = "bottom") +
    scale_y_discrete(expand = c(0, 0))


p_sig <- ggplot(global_mean, x = public_id, y = "0", fill = progression_group, show.legend = F) +
    geom_tile(aes(x = public_id, y = "0", fill = progression_group), color = "#484848", lwd = 0.1, linetype = 1, show.legend = F) +
    scale_fill_manual(
        values = c("NP" = fill_NP, "RP" = fill_RP, "P" = "white", "Inc" = "grey")
    ) +
    theme_prism() +
    theme(
        plot.margin = margin(
            0.5, 0,
            0.5, 0,
            unit = "lines"
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))


p_tile_1q21 <- ggplot(global_mean, x = public_id, y = "Gain 1q21", fill = seqwgs_cp_1q21_call) +
    geom_tile(aes(x = public_id, y = "Gain 1q21", fill = seqwgs_cp_1q21_call), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_manual(
        values = c("no" = "white", "yes" = "#e48a04")
    ) +
    theme_prism() +
    xlab("Patient") +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))


p_tile_17p13 <- ggplot(global_mean, x = public_id, y = "Del 17p13", fill = seqwgs_cp_17p13_call) +
    geom_tile(aes(x = public_id, y = "Del 17p13", fill = seqwgs_cp_17p13_call), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_manual(
        values = c("no" = "white", "yes" = "#009bcf")
    ) +
    theme_prism() +
    xlab("Patient") +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))

p_tile_nsd2 <- ggplot(global_mean, x = public_id, y = "NSD2 t(4;14)", fill = nsd2_call) +
    geom_tile(aes(x = public_id, y = "NSD2 t(4;14)", fill = nsd2_call), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_manual(
        values = c("no" = "white", "yes" = "#c32400")
    ) +
    theme_prism() +
    xlab("Patient") +
    ggtitle("CD3+ T Cells") +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +

    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))

p_tile_maf <- ggplot(global_mean, x = public_id, y = "MAF, MAFA, MAFB", fill = maf_combined_call) +
    geom_tile(aes(x = public_id, y = "MAF, MAFA, MAFB", fill = maf_combined_call), color = "#484848", lwd = 0.1, linetype = 1, show.legend = FALSE) +
    scale_fill_manual(
        values = c("no" = "white", "yes" = "#00b02c")
    ) +
    theme_prism() +
    xlab("Patient") +
    theme(
        plot.margin = margin(
            0, 0,
            0, 0
        ),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.length.x = unit(0, "lines"),
        plot.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
    ) +
    coord_fixed(ratio = RATIO_PAT_INFO) +
    # element_text(
    #     angle = 90,
    #     vjust = 0.5, hjust = 1,
    # )) +
    guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
    scale_y_discrete(expand = c(0, 0))

TEST <- patchwork::wrap_plots(list(p_tile_nsd2, p_tile_1q21, p_tile_maf, p_tile_17p13, p_sig, p_cyto, p_naive, p), ncol = 1, nrow = 8)
pdf_and_png(TEST, output_surv, "cd8_val_cytotox_heatmap", pdfWidth = 9, pdfHeight = 5, scale = 2)
```



# Figure 4M - Naive Signature
```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)
library(survminer)

feat <- "Naive"
tmp <- patMD_averageMod |> dplyr::filter(cohort == "validation")

tmp[, feat] <- tmp[, feat] * 100
tmp$survVar <- tmp[, feat]
med_val <- median(tmp$survVar)

med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_naive <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = "Progression Free Survival"
    ) +
    add_censor_mark(size = 3) +
    add_risktable() +
    theme_prism() +
    ggtitle("Chu et al.,\nNaive Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]

pval <- round(pval, digits = 4)
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)

annotations <- data.frame(
    xpos = c(-Inf),
    ypos = c(-Inf),
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_naive_cox <- p_naive + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)
p_naive_cox <- ggsurvfit_build(p_naive_cox)

pdf_and_png(p_naive_cox, output_surv, "NaiveSig_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```

# Figure 4L - Cytotoxicity Signature
```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)

feat <- "Cytotoxicity"
tmp <- patMD_averageMod |> dplyr::filter(cohort == "validation")

tmp[, feat] <- tmp[, feat] * 100

tmp$survVar <- tmp[, feat]
# med_val <- median(tmp$survVar)
med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_cyto <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = ""
    ) +
    add_censor_mark(size = 3) +
    add_risktable() +
    theme_prism() +
    ggtitle("Chu et al.,\nCytotoxicity Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)
pval <- round(pval, digits = 4)

annotations <- data.frame(
    xpos = c(-Inf),
    ypos = c(-Inf),
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_cyto_cox <- p_cyto + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)
p_cyto_cox <- ggsurvfit_build(p_cyto_cox)

pdf_and_png(p_cyto_cox, output_surv, "CytoSig_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```


# Figure 4N - Exhaustion Signature
```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)

feat <- "Exhaustion"
tmp <- patMD_averageMod |> dplyr::filter(cohort == "validation")

tmp[, feat] <- tmp[, feat] * 100

tmp$survVar <- tmp[, feat]
# med_val <- median(tmp$survVar)
med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_exh <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = " "
    ) +
    add_censor_mark(size = 3) +
    add_risktable() +
    theme_prism() +
    ggtitle("Chu et al.,\nExhaustion Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)

pval <- round(pval, digits = 4)

annotations <- data.frame(
    xpos = c(-Inf),
    ypos = c(-Inf),
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_exh_cox <- p_exh + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)

p_exh_cox <- ggsurvfit_build(p_exh_cox)


pdf_and_png(p_exh_cox, output_surv, "ExhSig_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```

# Combine figures 4L, 4M, and 4N, into one figure (to ensure consistent rendering sizes)
```{r}
legend_surv <- cowplot::get_legend(p_naive_cox)

p_all_sig <- cowplot::plot_grid(p_cyto_cox + NoLegend(), p_naive_cox + NoLegend(), p_exh_cox + NoLegend(), legend_surv, nrow = 1, rel_widths = c(1, 1, 1, 0.25))
pdf_and_png(p_all_sig, output_surv, "All_CD3_Sig", pdfWidth = W_FULL * 3, pdfHeight = H_FULL, scale = 3)
```

# Cut Figures

Average expression of select exhaustion markers
```{r}
tmp2 <- AverageExpression(nkt, assays = "RNA", group.by = "public_id", features = c(list_of_signatures$Exhaustion, "LAG3"))$RNA |> # c("TIGIT", "TOX", "PDCD1", "LAG3", "HAVCR2", "CTLA4"))$RNA |>
    t() |>
    as.data.frame()

tmp2$public_id <- rownames(tmp2)

tmp <- dplyr::left_join(subset_patMD, tmp2, by = "public_id")

tmp <- tmp |> dplyr::filter(progression_group %in% c("NP", "RP"))

feat <- "LAG3"

pvals_wilcox <- tmp |>
    dplyr::filter(progression_group %in% c("NP", "RP")) |>
    rstatix::wilcox_test(as.formula(paste0(feat, " ~ progression_group"))) |>
    rstatix::add_xy_position()

p_box_lag3 <- ggplot(tmp, aes(x = progression_group, y = !!sym(feat), fill = progression_group, color = progression_group)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_jitter(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    ylab(paste0("Per Patient Average Expression")) +
    ggtitle(feat) +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)

write.csv(tmp, file.path(output_surv, "per_pat_avg_LAG3"))
```

```{r}
tmp2 <- AverageExpression(nkt, assays = "RNA", group.by = "public_id", features = c(list_of_signatures$Exhaustion, "LAG3"))$RNA |> # c("TIGIT", "TOX", "PDCD1", "LAG3", "HAVCR2", "CTLA4"))$RNA |>
    t() |>
    as.data.frame()

tmp2$public_id <- rownames(tmp2)

tmp <- dplyr::left_join(subset_patMD, tmp2, by = "public_id")

tmp <- tmp |> dplyr::filter(progression_group %in% c("NP", "RP"))

feat <- "TOX"

pvals_wilcox <- tmp |>
    dplyr::filter(progression_group %in% c("NP", "RP")) |>
    rstatix::wilcox_test(as.formula(paste0(feat, " ~ progression_group"))) |>
    rstatix::add_xy_position()
p_box_tox <- ggplot(tmp, aes(x = progression_group, y = !!sym(feat), fill = progression_group, color = progression_group)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_jitter(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    ylab(bquote("Average Log-Normalized Expression Across" ~ CD3^"+" ~ "T Cells")) +
    ggtitle(feat) +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "{scales::pvalue(p)}", inherit.aes = F, hide.ns = T, size = 5)

write.csv(tmp, file.path(output_surv, "exh_per_pat_average_lognorm_expr.csv"))
```

```{r}
p_exh_comb <- p_box_tox + p_box_lag3

pdf_and_png(p_exh_comb, output_surv, "lognorm_avg_cd3_tox_lag3_violin", pdfWidth = W_FULL, pdfHeight = H_FULL, scale = 3)
```

```{r}
test <- nkt |> subset(subset = (subcluster_V03072023 == "NkT.2.1" & davies_based_risk %in% c("high_risk", "standard_risk")))

test <- AddModuleScore_UCell(test, features = list("Dysfunctional" = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2")))

testb <- DotPlot(test, features = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2"), group.by = "davies_based_risk")


tmp <- test@meta.data
pvals_wilcox <- tmp |>
    rstatix::wilcox_test(as.formula(paste0("Dysfunctional_UCell", " ~ davies_based_risk"))) |>
    rstatix::add_xy_position()

p <- ggplot(tmp, aes(x = davies_based_risk, y = Dysfunctional_UCell, fill = davies_based_risk, color = davies_based_risk)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.1) +
    ylab(paste0("Per Cell Dysfunctional Score")) +
    ggtitle("Dysfunctional Score Within CD8+ HLA T Cells") +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
```

```{r}
averageMod_dys <- test@meta.data %>%
    dplyr::group_by(public_id) %>%
    dplyr::summarise(
        DysfunctionalScore = mean(Dysfunctional_UCell)
    )

subset_patMD_dys <- test@meta.data |>
    dplyr::distinct(public_id, davies_based_risk)

patMD_averageMod_dys <- dplyr::left_join(subset_patMD_dys, averageMod_dys, by = "public_id")
```

```{r}
tmp <- patMD_averageMod_dys
pvals_wilcox <- tmp |>
    rstatix::wilcox_test(as.formula(paste0("DysfunctionalScore", " ~ davies_based_risk"))) |>
    rstatix::add_xy_position()

p <- ggplot(tmp, aes(x = davies_based_risk, y = DysfunctionalScore, fill = davies_based_risk, color = davies_based_risk)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    ylab(paste0("Per Cell Dysfunctional Score")) +
    ggtitle("Dysfunctional Score Within CD8+ HLA T Cells") +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
```

```{r}
nkt <- AddModuleScore_UCell(nkt, features = list("Dysfunctional" = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2")))

testb <- DotPlot(nkt, features = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2", "Dysfunctional_UCell"), group.by = "cellID_short")
```