---
title: "Figure 4 - Survival"
author: "William Pilcher"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
---

Computes survival curves for select CD3+ T cell markers (Figures 4L, M, N)

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Load object and selected signatures
```{r init}
library(UCell)
source("source/Validation/validation_load_data.R", local = T)

nkt <- merged |> subset(subset = (lineage_group %in% c("CD8", "CD4") &
    VJ_INTERVAL == "Baseline"))

color_low <- "#2c90c3"
color_high <- "#5a221b"

output_surv <- file.path(output, "Validation", "Survival")
dir.create(output_surv, recursive = T, showWarnings = F)

sig <- readxl::read_xlsx(file.path("/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation", "data/resources/pan-cancer_tcell_atlas_chu_et_al.xlsx"), sheet = "Table S4", skip = 1, )
sig_cd4 <- readxl::read_xlsx(file.path("/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation", "data/resources/pan-cancer_tcell_atlas_chu_et_al.xlsx"), sheet = "Table S6", skip = 1, )

combined_naive <- unique(c(sig_cd4[[1]] |> na.omit(), sig[[1]] |> na.omit()))
combined_cyto <- unique(c(sig_cd4[[5]] |> na.omit(), sig[[5]] |> na.omit()))
combined_exh <- unique(c(sig_cd4[[3]] |> na.omit(), sig[[3]] |> na.omit()))
dysfunctional_custom <- c("B3GAT1+", "ZEB2+", "KLRG1+", "KLRK1+", "TIGIT+", "PDCD1+", "LAG3+", "CTLA4+", "HAVCR2+", "TOX+", "CD160+", "CD27-", "CD28-")

list_of_signatures <- list("Naive" = combined_naive, "Cytotoxic" = combined_cyto, "Exhaustion" = combined_exh, "Dysfunctional" = dysfunctional_custom)

valid_sig <- make.names(names(list_of_signatures))
```

```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)
library(survminer)

color_low <- "#2c90c3"
color_high <- "#5a221b"

bl_md <- merged@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::distinct(public_id, .keep_all = T)

tmp <- bl_md

tmp$survVar <- tmp[, "cohort"]

# ggsurvplot( fit = survfit(Surv(ttcos, censos==1) ~ davies_based_risk, data = tmp_df),
#             type=c("kaplan-meier"),surv.median.line = 'hv',log.rank.weights= "n",pval.method = TRUE,
#             xlab = "Days", ylab = "OS Prob.",pval = TRUE,risk.table = TRUE,palette = c('orange','skyblue3'))

p_naive <- survfit2(Surv(ttcpfs, censpfs) ~ survVar, data = tmp) %>% ggsurvplot(
    size = 1, type = c("kaplan-meier"), surv.median.line = "hv", log.rank.weights = "n", pval.method = FALSE,
    xlab = "Days", ylab = "PFS Prob.", pval = TRUE, risk.table = TRUE, palette = c("#aab5bd", "#6D1E46"),
    legend = c(0.8, 0.85), legend.title = "Cohort", legend.labs = c("Discovery", "Validation")
)
just_surv <- p_naive[[1]] + theme(plot.background = element_rect(fill = "transparent", color = NA), rect = element_rect(fill = "transparent", color = NA), panel.border = element_blank(), plot.border = element_blank())
# labs(
#     x = "Days",
#     y = "Progression Free Survival"
# ) +
# add_censor_mark(size = 3) +
# add_risktable() +
# theme_prism() +
# add_pvalue() +
# ggtitle("Survival Outcomes in Discovery and Validation Cohorts") +
# scale_color_manual(values = c("discovery" = color_low, "validation" = color_high))

test <- patchwork::wrap_plots(list(p_naive[[1]], p_naive[[2]]), nrow = 2, heights = c(3.5, 1))

pdf_and_png(just_surv, outputDir = output_surv, filename = "surv_plot_cleaned", pdfWidth = W_FULL, pdfHeight = H_FULL, scale = 1.5)
```

```{r}
valMD <- bl_md |> dplyr::filter(cohort == "validation")
valMD$progression_group <- factor(valMD$progression_group, levels = rev(c("RP", "P", "Inc", "NP")))
valMD$Group <- "Progression Group"
p <- ggplot(valMD, aes(x = Group, fill = progression_group)) +
    geom_bar(stat = "count", position = "fill") +
    theme_survminer() +
    ylab("Patient Count") +
    theme(
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA),
        legend.position = "right",
        legend.background = element_rect(fill = "transparent", color = NA),
        legend.title = element_blank()
    ) +
    scale_fill_manual(values = c(
        "NP" = fill_NP,
        "RP" = fill_RP,
        "P" = "#f1c0b5",
        "Inc" = "#dce7f3"
    )) +
    scale_y_continuous(position = "right", expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0))

pdf_and_png(p, output_surv, "stacked_chart", pdfHeight = H_FULL, pdfWidth = 1.6, scale = 1.5)
```

```{r}
valMD <- bl_md |> dplyr::filter(cohort == "validation")
valMD$progression_group <- factor(valMD$progression_group, levels = rev(c("RP", "P", "Inc", "NP")))
valMD$Group <- "Progression Group"
p <- ggplot(valMD, aes(x = Group, fill = progression_group)) +
    geom_bar(stat = "count", position = "fill") +
    theme_survminer() +
    ylab("Patient Count") +
    theme(
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA),
        legend.position = "right",
        legend.background = element_rect(fill = "transparent", color = NA),
        legend.title = element_blank()
    ) +
    scale_fill_manual(values = c(
        "NP" = fill_NP,
        "RP" = fill_RP,
        "P" = "#f1c0b5",
        "Inc" = "#dce7f3"
    )) +
    scale_y_continuous(position = "right", expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0))

pdf_and_png(p, output_surv, "stacked_chart", pdfHeight = H_FULL, pdfWidth = 1.6, scale = 1.5)
```

```{r}
valMD <- bl_md |> dplyr::filter(cohort == "validation")
counts_1q <- table(valMD$seqwgs_cp_1q21_call == "yes")
counts_nsd2_call <- table(valMD$nsd2_call == "yes")
counts_17p13_del <- table(valMD$seqwgs_cp_17p13_call == "yes")
counts_maf_call <- table(valMD$maf_call == "yes")
counts_mafb_call <- table(valMD$mafb_call == "yes")
counts_mafa_call <- table(valMD$mafa_call == "yes")

barchart_risk_simp_df <- data.frame(risk_names = c(
    "chr1q_gain",
    "chr4_14_WHSC1",
    "chr17p_del",
    "chr14_16_MAF",
    "chr14_20_MAFB",
    "chr8_14_MAFA"
), risk_counts = c(
    counts_1q[["TRUE"]],
    counts_nsd2_call[["TRUE"]],
    counts_17p13_del[["TRUE"]],
    counts_maf_call[["TRUE"]],
    0,
    0
))

barchart_risk_simp_df$risk_names <- factor(barchart_risk_simp_df$risk_names, levels = barchart_risk_simp_df$risk_names)


p <- ggplot(barchart_risk_simp_df, aes(x = risk_names, y = risk_counts)) +
    geom_bar(stat = "identity", fill = "#5682b1", color = NA, show.legend = F) +
    geom_text(aes(label = risk_counts, y = risk_counts), vjust = 0) +
    theme_survminer() +
    ylab("Number of Patients") +
    xlab("Genetic Events") +
    theme(
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 10),
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA)
    ) +
    ylim(c(0, 30))

pdf_and_png(p, output_surv, "risk_chart_fig1_style", pdfHeight = H_FULL, pdfWidth = 2, scale = 1.5)
```


```{r}
risk_copy <- valMD
risk_copy[, c("nsd2_call", "mafa_call", "maf_call", "mafb_call", "seqwgs_cp_17p13_call", "seqwgs_cp_1q21_call")] <- risk_copy[, c("nsd2_call", "mafa_call", "maf_call", "mafb_call", "seqwgs_cp_17p13_call", "seqwgs_cp_1q21_call")] == "yes"

risk_copy <- risk_copy |>
    dplyr::select(nsd2_call, mafa_call, maf_call, mafb_call, seqwgs_cp_17p13_call, seqwgs_cp_1q21_call) |>
    na.omit()

risk_copy[risk_copy == FALSE] <- 0
risk_copy[risk_copy == TRUE] <- 1

library(UpSetR)
p <- UpSetR::upset(as.data.frame(risk_copy), nsets = 9)
pdf(file = file.path(output_surv, "upset_val.pdf"), height = 4, width = 6)
print(p)
dev.off()
```

```{r}
bl_md <- merged@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::distinct(public_id, .keep_all = T)

for (risk_type in c("skerget_based_risk", "davies_based_risk", "skerget_with_amp1q", "expanded_risk_definition", "expanded_risk_definition_NO_CHROMOTHRIPSIS", "expanded_risk_definition_NO_APOBEC", "expanded_risk_definition_NO_TP53_MUT", "expanded_risk_definition_NO_1q_AMP")) {
    tmp <- bl_md |> dplyr::filter(!!sym(risk_type) %in% c("high_risk", "standard_risk"))
    tmp$survVar <- tmp[, risk_type]
    p_naive <- survfit2(Surv(pfscdy, censpfs) ~ survVar, data = tmp) %>% ggsurvfit(size = 2) +
        labs(
            x = "Days",
            y = "Progression Free Survival"
        ) +
        add_censor_mark(size = 3) +
        add_risktable() +
        theme_prism() +
        add_pvalue() +
        ggtitle(risk_type) +
        scale_color_manual(values = c("standard_risk" = color_low, "high_risk" = color_high))

    p_bar <- ggplot(tmp, aes(x = !!sym(risk_type), fill = !!sym(risk_type), color = !!sym(risk_type))) +
        geom_bar() +
        geom_text(stat = "count", aes(label = ..count..), vjust = -1) +
        ylab(paste0("Count of, ", risk_type)) +
        ggtitle(risk_type) +
        theme_prism() +
        theme(axis.title.x = element_blank()) +
        scale_fill_prism("floral") +
        scale_color_prism("floral") +
        NoLegend()

    p_combined <- ggsurvfit_build(p_naive)
    pdf_and_png(p_combined, output_surv, paste0(risk_type, "_survival"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
    pdf_and_png(p_bar, output_surv, paste0(risk_type, "_frequency"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
}

for (risk_element in c("nsd2_call", "mafa_call", "maf_call", "mafb_call", "seqwgs_cp_1q21_call", "seqwgs_cp_17p13_call", "seqwgs_cp_derived_amp_1q", "tp53_mut_call", "chromothripsis", "APOBEC_Enriched")) {
    tmp <- bl_md |> dplyr::filter(!!sym(risk_element) %in% c("yes", "no"))
    tmp$survVar <- tmp[, risk_element]

    p_naive <- survfit2(Surv(pfscdy, censpfs) ~ survVar, data = tmp) %>% ggsurvfit(size = 2) +
        labs(
            x = "Days",
            y = "Progression Free Survival"
        ) +
        add_censor_mark(size = 3) +
        add_pvalue() +
        add_risktable() +
        theme_prism() +
        ggtitle(risk_element) +
        scale_color_manual(values = c("no" = color_low, "yes" = color_high))

    p_bar <- ggplot(tmp, aes(x = !!sym(risk_element), fill = !!sym(risk_element), color = !!sym(risk_element))) +
        geom_bar() +
        geom_text(stat = "count", aes(label = ..count..), vjust = -1) +
        ylab(paste0("Count of, ", risk_element)) +
        ggtitle(risk_type) +
        theme_prism() +
        theme(axis.title.x = element_blank()) +
        scale_fill_prism("floral") +
        scale_color_prism("floral") +
        NoLegend()

    p_combined <- ggsurvfit_build(p_naive)
    pdf_and_png(p_combined, output_surv, paste0(risk_element, "_survival"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
    pdf_and_png(p_bar, output_surv, paste0(risk_element, "_frequency"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
}
```

```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)
library(survminer)
output_val <- file.path(output_surv, "validation")
dir.create(output_val)
bl_md <- merged@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline", cohort == "validation") |>
    dplyr::distinct(public_id, .keep_all = T)

for (risk_type in c("skerget_based_risk", "davies_based_risk", "skerget_with_amp1q", "expanded_risk_definition", "expanded_risk_definition_NO_CHROMOTHRIPSIS", "expanded_risk_definition_NO_APOBEC", "expanded_risk_definition_NO_TP53_MUT", "expanded_risk_definition_NO_1q_AMP")) {
    tmp <- bl_md |> dplyr::filter(!!sym(risk_type) %in% c("high_risk", "standard_risk"))
    tmp$survVar <- tmp[, risk_type]
    tryCatch(expr = {
        p_naive <- survfit2(Surv(pfscdy, censpfs) ~ survVar, data = tmp) %>% ggsurvfit(size = 2) +
            labs(
                x = "Days",
                y = "Progression Free Survival"
            ) +
            add_censor_mark(size = 3) +
            add_risktable() +
            theme_prism() +
            add_pvalue() +
            ggtitle(risk_type) +
            scale_color_manual(values = c("standard_risk" = color_low, "high_risk" = color_high))

        p_combined <- ggsurvfit_build(p_naive)
        pdf_and_png(p_combined, output_val, paste0(risk_type, "_survival"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
    })
    p_bar <- ggplot(tmp, aes(x = !!sym(risk_type), fill = !!sym(risk_type), color = !!sym(risk_type))) +
        geom_bar() +
        geom_text(stat = "count", aes(label = ..count..), vjust = -1) +
        ylab(paste0("Count of, ", risk_type)) +
        ggtitle(risk_type) +
        theme_prism() +
        theme(axis.title.x = element_blank()) +
        scale_fill_prism("floral") +
        scale_color_prism("floral") +
        NoLegend()

    pdf_and_png(p_bar, output_val, paste0(risk_type, "_frequency"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
}

for (risk_element in c("nsd2_call", "mafa_call", "maf_call", "mafb_call", "seqwgs_cp_1q21_call", "seqwgs_cp_17p13_call", "seqwgs_cp_derived_amp_1q", "tp53_mut_call", "chromothripsis", "APOBEC_Enriched")) {
    tmp <- bl_md |> dplyr::filter(!!sym(risk_element) %in% c("yes", "no"))
    tmp$survVar <- tmp[, risk_element]
    if (risk_element != "mafa_call" && risk_element != "mafb_call") {
        p_naive <- survfit2(Surv(pfscdy, censpfs) ~ survVar, data = tmp) %>% ggsurvfit(size = 2) +
            labs(
                x = "Days",
                y = "Progression Free Survival"
            ) +
            add_censor_mark(size = 3) +
            add_pvalue() +
            add_risktable() +
            theme_prism() +
            ggtitle(risk_element) +
            scale_color_manual(values = c("no" = color_low, "yes" = color_high))

        p_combined <- ggsurvfit_build(p_naive)
        pdf_and_png(p_combined, output_val, paste0(risk_element, "_survival"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
    }
    p_bar <- ggplot(tmp, aes(x = !!sym(risk_element), fill = !!sym(risk_element), color = !!sym(risk_element))) +
        geom_bar() +
        geom_text(stat = "count", aes(label = ..count..), vjust = -1) +
        ylab(paste0("Count of, ", risk_element)) +
        ggtitle(risk_type) +
        theme_prism() +
        theme(axis.title.x = element_blank()) +
        scale_fill_prism("floral") +
        scale_color_prism("floral") +
        NoLegend()


    pdf_and_png(p_bar, output_val, paste0(risk_element, "_frequency"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
}
```

```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)
library(survminer)
output_val <- file.path(output_surv, "discovery")
dir.create(output_val)
bl_md <- merged@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline", cohort == "discovery") |>
    dplyr::distinct(public_id, .keep_all = T)

for (risk_type in c("skerget_based_risk", "davies_based_risk", "skerget_with_amp1q", "expanded_risk_definition", "expanded_risk_definition_NO_CHROMOTHRIPSIS", "expanded_risk_definition_NO_APOBEC", "expanded_risk_definition_NO_TP53_MUT", "expanded_risk_definition_NO_1q_AMP")) {
    tmp <- bl_md |> dplyr::filter(!!sym(risk_type) %in% c("high_risk", "standard_risk"))
    tmp$survVar <- tmp[, risk_type]
    p_naive <- survfit2(Surv(pfscdy, censpfs) ~ survVar, data = tmp) %>% ggsurvfit(size = 2) +
        labs(
            x = "Days",
            y = "Progression Free Survival"
        ) +
        add_censor_mark(size = 3) +
        add_risktable() +
        theme_prism() +
        add_pvalue() +
        ggtitle(risk_type) +
        scale_color_manual(values = c("standard_risk" = color_low, "high_risk" = color_high))

    p_bar <- ggplot(tmp, aes(x = !!sym(risk_type), fill = !!sym(risk_type), color = !!sym(risk_type))) +
        geom_bar() +
        geom_text(stat = "count", aes(label = ..count..), vjust = -1) +
        ylab(paste0("Count of, ", risk_type)) +
        ggtitle(risk_type) +
        theme_prism() +
        theme(axis.title.x = element_blank()) +
        scale_fill_prism("floral") +
        scale_color_prism("floral") +
        NoLegend()

    p_combined <- ggsurvfit_build(p_naive)
    pdf_and_png(p_combined, output_val, paste0(risk_type, "_survival"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
    pdf_and_png(p_bar, output_val, paste0(risk_type, "_frequency"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
}

for (risk_element in c("nsd2_call", "mafa_call", "maf_call", "mafb_call", "seqwgs_cp_1q21_call", "seqwgs_cp_17p13_call", "seqwgs_cp_derived_amp_1q", "tp53_mut_call", "chromothripsis", "APOBEC_Enriched")) {
    tmp <- bl_md |> dplyr::filter(!!sym(risk_element) %in% c("yes", "no"))
    tmp$survVar <- tmp[, risk_element]

    p_naive <- survfit2(Surv(pfscdy, censpfs) ~ survVar, data = tmp) %>% ggsurvfit(size = 2) +
        labs(
            x = "Days",
            y = "Progression Free Survival"
        ) +
        add_censor_mark(size = 3) +
        add_pvalue() +
        add_risktable() +
        theme_prism() +
        ggtitle(risk_element) +
        scale_color_manual(values = c("no" = color_low, "yes" = color_high))

    p_bar <- ggplot(tmp, aes(x = !!sym(risk_element), fill = !!sym(risk_element), color = !!sym(risk_element))) +
        geom_bar() +
        geom_text(stat = "count", aes(label = ..count..), vjust = -1) +
        ylab(paste0("Count of, ", risk_element)) +
        ggtitle(risk_type) +
        theme_prism() +
        theme(axis.title.x = element_blank()) +
        scale_fill_prism("floral") +
        scale_color_prism("floral") +
        NoLegend()

    p_combined <- ggsurvfit_build(p_naive)
    pdf_and_png(p_combined, output_val, paste0(risk_element, "_survival"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
    pdf_and_png(p_bar, output_val, paste0(risk_element, "_frequency"), pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
}
```

# Signatures, Survival
UCell computes a signature score which is on a range between 0 and 1. These will be mutlipled by 100 when fitting the CoxPH model, so that the HR can be interpretted as the increase in risk per 1 UCell signature score point increase
```{r}
nkt <- AddModuleScore_UCell(nkt, features = list_of_signatures, name = "") # AddModuleScore(nkt, features = list_of_signatures, name = valid_sig)

averageMod <- nkt@meta.data %>%
    dplyr::group_by(public_id) %>%
    dplyr::summarise(
        Naive = mean(!!sym(paste0(valid_sig[[1]]))),
        Cytotoxicity = mean(!!sym(paste0(valid_sig[[2]]))),
        Exhaustion = mean(!!sym(paste0(valid_sig[[3]]))),
        Dysfunctional = mean(!!sym(paste0(valid_sig[[4]])))
    )

subset_patMD <- nkt@meta.data |>
    dplyr::filter(d_lot_1_start_day < 2) |>
    dplyr::distinct(public_id, progression_group, censpfs, pfscdy, d_pt_sex, d_dx_amm_age, d_dx_amm_iss_stage, siteXbatch, d_tx_induction_cat, d_amm_tx_asct_1st, cohort)

patMD_averageMod <- dplyr::left_join(subset_patMD, averageMod, by = "public_id")
```

# Figure 4M - Naive Signature
```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)
library(survminer)

feat <- "Naive"
tmp <- patMD_averageMod |> dplyr::filter(cohort == "validation")

tmp[, feat] <- tmp[, feat] * 100
tmp$survVar <- tmp[, feat]
med_val <- median(tmp$survVar)

med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_naive <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = "Progression Free Survival"
    ) +
    add_censor_mark(size = 3) +
    add_risktable() +
    theme_prism() +
    ggtitle("Chu et al.,\nNaive Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]

pval <- round(pval, digits = 4)
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)

annotations <- data.frame(
    xpos = c(-Inf),
    ypos = c(-Inf),
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_naive_cox <- p_naive + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)
p_naive_cox <- ggsurvfit_build(p_naive_cox)

pdf_and_png(p_naive_cox, output_surv, "NaiveSig_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```

# Figure 4L - Cytotoxicity Signature
```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)

feat <- "Cytotoxicity"
tmp <- patMD_averageMod |> dplyr::filter(cohort == "validation")

tmp[, feat] <- tmp[, feat] * 100

tmp$survVar <- tmp[, feat]
# med_val <- median(tmp$survVar)
med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_cyto <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = ""
    ) +
    add_censor_mark(size = 3) +
    # add_risktable() +
    theme_prism() +
    ggtitle("Chu et al.,\nCytotoxicity Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)
pval <- round(pval, digits = 4)

annotations <- data.frame(
    xpos = 0,
    ypos = 0,
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_cyto_cox <- p_cyto + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)
p_cyto_cox <- ggsurvfit_build(p_cyto_cox)

pdf_and_png(p_cyto_cox, output_surv, "CytoSig_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```


# Figure 4N - Exhaustion Signature
```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)

feat <- "Exhaustion"
tmp <- patMD_averageMod |> dplyr::filter(cohort == "validation")

tmp[, feat] <- tmp[, feat] * 100

tmp$survVar <- tmp[, feat]
# med_val <- median(tmp$survVar)
med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_exh <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = " "
    ) +
    add_censor_mark(size = 3) +
    add_risktable() +
    theme_prism() +
    ggtitle("Chu et al.,\nExhaustion Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)

pval <- round(pval, digits = 4)

annotations <- data.frame(
    xpos = c(-Inf),
    ypos = c(-Inf),
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_exh_cox <- p_exh + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)

p_exh_cox <- ggsurvfit_build(p_exh_cox)


pdf_and_png(p_exh_cox, output_surv, "ExhSig_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```


# Dysfunctional

```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)

feat <- "Dysfunctional"
tmp <- patMD_averageMod |> dplyr::filter(cohort == "validation")

tmp[, feat] <- tmp[, feat] * 100

tmp$survVar <- tmp[, feat]
# med_val <- median(tmp$survVar)
med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_dys <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = ""
    ) +
    add_censor_mark(size = 3) +
    # add_risktable() +
    theme_prism() +
    ggtitle("Dysfunctional Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)
pval <- round(pval, digits = 4)

annotations <- data.frame(
    xpos = 0,
    ypos = 0,
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_dys_cox <- p_dys + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)
p_dys_cox <- ggsurvfit_build(p_dys_cox)

pdf_and_png(p_dys_cox, output_surv, "dysfunctional_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```


# Combine figures 4L, 4M, and 4N, into one figure (to ensure consistent rendering sizes)
```{r}
legend_surv <- cowplot::get_legend(p_naive_cox)

p_all_sig <- cowplot::plot_grid(p_cyto_cox, p_dys_cox, nrow = 2)
pdf_and_png(p_all_sig, output_surv, "All_CD3_Sig", pdfWidth = W_FULL * 1, pdfHeight = H_FULL * 1.75, scale = 2.5)
```

# Cut Figures

Average expression of select exhaustion markers
```{r}
tmp2 <- AverageExpression(nkt, assays = "RNA", group.by = "public_id", features = c(list_of_signatures$Exhaustion, "LAG3"))$RNA |> # c("TIGIT", "TOX", "PDCD1", "LAG3", "HAVCR2", "CTLA4"))$RNA |>
    t() |>
    as.data.frame()

tmp2$public_id <- rownames(tmp2)

tmp <- dplyr::left_join(subset_patMD, tmp2, by = "public_id")

tmp <- tmp |> dplyr::filter(progression_group %in% c("NP", "RP"))

feat <- "LAG3"

pvals_wilcox <- tmp |>
    dplyr::filter(progression_group %in% c("NP", "RP")) |>
    rstatix::wilcox_test(as.formula(paste0(feat, " ~ progression_group"))) |>
    rstatix::add_xy_position()

p_box_lag3 <- ggplot(tmp, aes(x = progression_group, y = !!sym(feat), fill = progression_group, color = progression_group)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_jitter(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    ylab(paste0("Per Patient Average Expression")) +
    ggtitle(feat) +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)

write.csv(tmp, file.path(output_surv, "per_pat_avg_LAG3"))
```

```{r}
tmp2 <- AverageExpression(nkt, assays = "RNA", group.by = "public_id", features = c(list_of_signatures$Exhaustion, "LAG3"))$RNA |> # c("TIGIT", "TOX", "PDCD1", "LAG3", "HAVCR2", "CTLA4"))$RNA |>
    t() |>
    as.data.frame()

tmp2$public_id <- rownames(tmp2)

tmp <- dplyr::left_join(subset_patMD, tmp2, by = "public_id")

tmp <- tmp |> dplyr::filter(progression_group %in% c("NP", "RP"))

feat <- "TOX"

pvals_wilcox <- tmp |>
    dplyr::filter(progression_group %in% c("NP", "RP")) |>
    rstatix::wilcox_test(as.formula(paste0(feat, " ~ progression_group"))) |>
    rstatix::add_xy_position()
p_box_tox <- ggplot(tmp, aes(x = progression_group, y = !!sym(feat), fill = progression_group, color = progression_group)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_jitter(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    ylab(bquote("Average Log-Normalized Expression Across" ~ CD3^"+" ~ "T Cells")) +
    ggtitle(feat) +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "{scales::pvalue(p)}", inherit.aes = F, hide.ns = T, size = 5)

write.csv(tmp, file.path(output_surv, "exh_per_pat_average_lognorm_expr.csv"))
```

```{r}
p_exh_comb <- p_box_tox + p_box_lag3

pdf_and_png(p_exh_comb, output_surv, "lognorm_avg_cd3_tox_lag3_violin", pdfWidth = W_FULL, pdfHeight = H_FULL, scale = 3)
```

```{r}
test <- nkt |> subset(subset = (subcluster_V03072023 == "NkT.2.1" & davies_based_risk %in% c("high_risk", "standard_risk")))

test <- AddModuleScore_UCell(test, features = list("Dysfunctional" = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2")))

testb <- DotPlot(test, features = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2"), group.by = "davies_based_risk")


tmp <- test@meta.data
pvals_wilcox <- tmp |>
    rstatix::wilcox_test(as.formula(paste0("Dysfunctional_UCell", " ~ davies_based_risk"))) |>
    rstatix::add_xy_position()

p <- ggplot(tmp, aes(x = davies_based_risk, y = Dysfunctional_UCell, fill = davies_based_risk, color = davies_based_risk)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.1) +
    ylab(paste0("Per Cell Dysfunctional Score")) +
    ggtitle("Dysfunctional Score Within CD8+ HLA T Cells") +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
```

```{r}
averageMod_dys <- test@meta.data %>%
    dplyr::group_by(public_id) %>%
    dplyr::summarise(
        DysfunctionalScore = mean(Dysfunctional_UCell)
    )

subset_patMD_dys <- test@meta.data |>
    dplyr::distinct(public_id, davies_based_risk)

patMD_averageMod_dys <- dplyr::left_join(subset_patMD_dys, averageMod_dys, by = "public_id")
```

```{r}
tmp <- patMD_averageMod_dys
pvals_wilcox <- tmp |>
    rstatix::wilcox_test(as.formula(paste0("DysfunctionalScore", " ~ davies_based_risk"))) |>
    rstatix::add_xy_position()

p <- ggplot(tmp, aes(x = davies_based_risk, y = DysfunctionalScore, fill = davies_based_risk, color = davies_based_risk)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    ylab(paste0("Per Cell Dysfunctional Score")) +
    ggtitle("Dysfunctional Score Within CD8+ HLA T Cells") +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
```

```{r}
nkt <- AddModuleScore_UCell(nkt, features = list("Dysfunctional" = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2")))

testb <- DotPlot(nkt, features = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2", "Dysfunctional_UCell"), group.by = "cellID_short")
```