---
title: "Fig 3-5: All Differential Abundance Tests"
author: "William Pilcher"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
---

# This code is a wrapper to perform differential abundance tests for a variety of variables.
# All differetial abundances analyses perform binary comparions between a group with or without an event, adjusting for siteXbatch as a covariate.
# All differential abundances use limma with a robust linear model to fit to the logit-transformed proportions, in a mechanism similar to propeller.
# Logit is preferred for its interpretability as opposed to arcsin-transformations.


```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Setup
```{r init}
library(ggpubr)
source("source/Validation/validation_load_data.R", local = T)
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)

cluster_group <- "subcluster_V03072023"
output_expr <- file.path(output, "Validation", "Individual_Risk", "Diff_Abundance", "LINEAR_LOGIT", "RISK_CATEGORIES_STUDY_SITE")

output_extra <- file.path(output_F3, "Diff_Abundance")
output_box <- file.path(output_extra, "Diff_Abundance_Box")
output_raw_csv <- file.path(output_extra, "CSV")
output_abun_volcano <- file.path(output_extra, "Diff_Abun_Volcano")
dir.create(output_extra, showWarnings = F, recursive = T)
dir.create(output_box, showWarnings = F, recursive = T)
dir.create(output_raw_csv, showWarnings = F, recursive = T)
dir.create(output_abun_volcano, showWarnings = F, recursive = T)
dir.create(output_expr, showWarnings = F, recursive = T)


floral_color_theme <- list(
    "SIG (+)" = fill_RP,
    "+" = color_RP,
    "SIG (-)" = fill_NP,
    "-" = color_NP
)

floral_fill_theme <- floral_color_theme
```

```{r}
var_to_test <- "nsd2_call"
groupA <- "yes"
groupB <- "no"

# function_string <- "d_pt_race_1"
method <- "LINEAR"

# Note - 'include_multivariate' not used for this analysis. Figures show 'univariate' model
variable_info_structure <- rev(list(
    "davies_based_risk" = list(
        isFactor = TRUE,
        factorLevels = c("standard_risk", "high_risk"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "High Risk"
    ),

    # Core davies based risk covariates
    "nsd2_call" = list(
        isFactor = TRUE,
        factorLevels = c("no", "yes"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = TRUE,
        variable_name = "NSD2 t(4;14)"
    ),
    "seqwgs_cp_1q21_call" = list(
        isFactor = TRUE,
        factorLevels = c("no", "yes"),
        perform_univariate = TRUE,
        univariate_batch = "siteXbatch",
        include_multivariate = TRUE,
        variable_name = "Gain 1q21"
    ),
    "maf_combined_call" = list(
        isFactor = TRUE,
        factorLevels = c("no", "yes"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = TRUE,
        variable_name = "MAF, MAFA, MAFB"
    ),
    "seqwgs_cp_17p13_call" = list(
        isFactor = TRUE,
        factorLevels = c("no", "yes"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = TRUE,
        variable_name = "Del 17p13"
    ),
    # Mutation information
    "tp53_mut_call" = list(
        isFactor = TRUE,
        factorLevels = c("no", "yes"),
        perform_univariate = TRUE,
        univariate_batch = "siteXbatch",
        include_multivariate = FALSE,
        variable_name = "TP53 Mut"
    ),
    "seqwgs_cp_cdkn2c_call" = list(
        isFactor = TRUE,
        factorLevels = c("no", "yes"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Del 1p32"
    ),
    # 1Q Amp
    "seqwgs_cp_derived_amp_1q" = list(
        isFactor = TRUE,
        factorLevels = c("no", "yes"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Del 1p32"
    ),

    # Multi-level factors for nsd2+1q21 and tp53
    "tp53_inactivation_state" = list(
        isFactor = TRUE,
        factorLevels = c("NONE", "PARTIAL", "COMPLETE"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        perform_contrasts = TRUE,
        variable_name = "TP53 Inactivation State"
    ),
    "mixed_nsd2_1q21_class" = list(
        isFactor = TRUE,
        factorLevels = c("Normal", "NSD2", "Gain1q21", "NSD2.Gain1q21"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        perform_contrasts = TRUE,
        variable_name = "TP53 Inactivation State"
    ),
    # Alternative covariates for IMWG 2024 HR Definition (not included in core analyses)
    "IMWG_FLAG_tp53" = list(
        isFactor = TRUE,
        factorLevels = c(FALSE, TRUE),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "TP53 Inactivation"
    ),
    "IMWG_FLAG_1P32_AND_1Q21" = list(
        isFactor = TRUE,
        factorLevels = c(FALSE, TRUE),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Del 1p32 + Gain 1q21"
    ),
    "IMWG_FLAG_BIALLELIC_1P32_DELETION" = list(
        isFactor = TRUE,
        factorLevels = c(FALSE, TRUE),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Del 1p32 + Gain 1q21"
    ),
    "IMWG_FLAG_NSD2_AND_1Q21" = list(
        isFactor = TRUE,
        factorLevels = c(FALSE, TRUE),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "NSD2 t(4;14) + Gain 1q21"
    ),
    "IMWG_FLAG_NSD2_AND_1P32" = list(
        isFactor = TRUE,
        factorLevels = c(FALSE, TRUE),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "NSD2 t(4;14) + Gain 1q21"
    ),
    "IMWG_FLAG_MAF_1Q21" = list(
        isFactor = TRUE,
        factorLevels = c(FALSE, TRUE),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "NSD2 t(4;14) + Gain 1q21"
    ),
    "IMWG_FLAG_MAF_1P32" = list(
        isFactor = TRUE,
        factorLevels = c(FALSE, TRUE),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "NSD2 t(4;14) + Gain 1q21"
    ),

    # Interaction with 1q21 and other covariates
    "mixed_1p32_1q21_class" = list(
        isFactor = TRUE,
        factorLevels = c("Normal", "1p32", "Gain1q21", "1p32.Gain1q21"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        perform_contrasts = TRUE,
        variable_name = "TP53 Inactivation State"
    ),
    "mixed_maf_1q21_class" = list(
        isFactor = TRUE,
        factorLevels = c("Normal", "MAF", "Gain1q21", "MAF.Gain1q21"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        perform_contrasts = TRUE,
        variable_name = "TP53 Inactivation State"
    ),
    "mixed_HR_1q21_class" = list(
        isFactor = TRUE,
        factorLevels = c("Normal", "HR", "Gain1q21", "HR.Gain1q21"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        perform_contrasts = TRUE,
        variable_name = "TP53 Inactivation State"
    ),
    "mixed_cn_1q21" = list(
        isFactor = TRUE,
        factorLevels = c("Normal", "Gain1q21", "Amp1q21"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        perform_contrasts = TRUE,
        variable_name = "TP53 Inactivation State"
    ),
    # External chromothripsis/APOBEC categories
    "chromothripsis" = list(
        isFactor = TRUE,
        factorLevels = c("no", "yes"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Chromothripsis"
    ),
    "APOBEC_Enriched" = list(
        isFactor = TRUE,
        factorLevels = c("no", "yes"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "APOBEC Signature"
    ),

    # BM PC PCT
    "scaled_PC_PCT" = list(
        isFactor = FALSE,
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Plasma Cell Percentage"
    ),
    # logit transformed PC PCT
    "logit_pc_proportion" = list(
        isFactor = FALSE,
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Plasma Cell Proportion (logit)"
    ),

    # Binned based on the upper quartile
    "binned_PC_PCT" = list(
        isFactor = TRUE,
        factorLevels = c("<20", ">=20"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Plasma Cell Percent >= 20"
    ),
    "Study_Site" = list(
        isFactor = TRUE,
        perform_univariate = FALSE,
        include_multivariate = FALSE
    )
))
```

# Differential Abundance - Setup Counts Data

```{r}
baseline <- merged |> subset(subset = (VJ_INTERVAL == "Baseline" & cohort == "discovery"))

Idents(baseline) <- baseline@meta.data[, cluster_group]
```

```{r}
counts <- table(baseline$public_id, baseline@meta.data[, cluster_group])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(maf_combined_call = dplyr::case_when(
    maf_call == "yes" | mafa_call == "yes" | mafb_call == "yes" ~ "yes",
    maf_call == "no" & mafa_call == "no" & mafb_call == "no" ~ "no",
    is.na(maf_call) | is.na(mafa_call) | is.na(mafb_call) ~ NA,
    .default = "ERROR"
))

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(age_binned = dplyr::case_when(
    d_dx_amm_age < 65 ~ "<65",
    d_dx_amm_age >= 65 ~ ">=65",
    .default = "ERROR"
))

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(age_binned_60 = dplyr::case_when(
    d_dx_amm_age < 60 ~ "<60",
    d_dx_amm_age >= 60 ~ ">=60",
    .default = "ERROR"
))
baseline@meta.data <- baseline@meta.data |> dplyr::mutate(scaled_age = (d_dx_amm_age - min(d_dx_amm_age)) / max(d_dx_amm_age))
baseline@meta.data <- baseline@meta.data |> dplyr::mutate(scaled_PC_PCT = (D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH - min(D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH)) / max(D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH))
baseline@meta.data <- baseline@meta.data |> dplyr::mutate(pc_proportion = (D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH) / 100)
baseline@meta.data <- baseline@meta.data |> dplyr::mutate(logit_pc_proportion = car::logit(pc_proportion, adjust = 0.025))

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(binned_PC_PCT = dplyr::case_when(
    D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH < 20 ~ "<20",
    D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH >= 20 ~ ">=20",
    .default = "ERROR"
))
```




```{r}
per_pat <- read.csv(file = "/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation/data/metadata_1110/mmrf_commpass_IA22_active_standardized_per_patient_20230505.csv", header = T)
per_pat <- per_pat |> dplyr::select(public_id, IMWG_RISK_CLASS)

baseline@meta.data <- baseline@meta.data |> dplyr::left_join(per_pat, by = "public_id")
# rownames(merged@meta.data) <- merged$cellname
```

```{r}
baseline@meta.data$IMWG_RISK_CLASS <- factor(baseline@meta.data$IMWG_RISK_CLASS, levels = c(0, 1, 2))
baseline@meta.data <- baseline@meta.data |> dplyr::mutate(IMWG_int_or_high = dplyr::case_when(
    IMWG_RISK_CLASS == 0 ~ "standard_risk",
    IMWG_RISK_CLASS %in% c(1, 2) ~ "high_risk",
    .default = NA
))
```

```{r}
baseline@meta.data$SR_davies.progression_group <- paste0(baseline@meta.data$davies_based_risk, ".", baseline@meta.data$progression_group)
baseline@meta.data$HR_davies.progression_group <- paste0(baseline@meta.data$davies_based_risk, ".", baseline@meta.data$progression_group)

baseline@meta.data$SR_IMWG2024.progression_group <- paste0(baseline@meta.data$IMWG_2024_HR, ".", baseline@meta.data$progression_group)
baseline@meta.data$HR_IMWG2024.progression_group <- paste0(baseline@meta.data$IMWG_2024_HR, ".", baseline@meta.data$progression_group)

baseline@meta.data$SR_EXPANDED.progression_group <- paste0(baseline@meta.data$expanded_risk_definition, ".", baseline@meta.data$progression_group)
baseline@meta.data$HR_EXPANDED.progression_group <- paste0(baseline@meta.data$expanded_risk_definition, ".", baseline@meta.data$progression_group)


baseline@meta.data <- baseline@meta.data |> dplyr::mutate(triplet_therapy = dplyr::case_when(
    d_tx_induction_cat %in% c("imid_pi_steroid", "chemo_imid_pi_steroid") ~ "Triplet",
    d_tx_induction_cat %in% c("pi_steroid", "chemo_pi_steroid", "imid_steroid", "chemo_imid_steroid") ~ "Doublet",
    .default = "ERROR MISSED CASE"
))

baseline@meta.data$Triplet_therapy.progression_group <- paste0(baseline@meta.data$triplet_therapy, ".", baseline@meta.data$progression_group)
baseline@meta.data$Doublet_therapy.progression_group <- paste0(baseline@meta.data$triplet_therapy, ".", baseline@meta.data$progression_group)
```

```{r}
baseline@meta.data <- baseline@meta.data |> dplyr::mutate(tp53_inactivation_state = dplyr::case_when(
    is.na(seqwgs_cp_17p13_call) | is.na(tp53_mut_call) ~ NA,
    as.numeric(seqwgs_cp_17p13) < log2(0.4) ~ "COMPLETE",
    seqwgs_cp_17p13_call == "yes" & tp53_mut_call == "yes" ~ "COMPLETE",
    seqwgs_cp_17p13_call == "yes" | tp53_mut_call == "yes" ~ "PARTIAL", # NA for one would mean partial, probably okay though
    seqwgs_cp_17p13_call == "no" & tp53_mut_call == "no" ~ "NONE",
    .default = "UNHANDLED CASE"
))

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(tp53_inactivation_state_alt = dplyr::case_when(
    is.na(seqwgs_cp_17p13_call) | is.na(tp53_mut_call) ~ NA,
    as.numeric(seqwgs_cp_17p13) < log2(0.4) ~ "COMPLETE",
    seqwgs_cp_17p13_call == "yes" & tp53_mut_call == "yes" ~ "COMPLETE",
    seqwgs_cp_17p13_call == "yes" & tp53_mut_call == "no" ~ "MonoDeletion", # NA for one would mean partial, probably okay though
    seqwgs_cp_17p13_call == "no" & tp53_mut_call == "yes" ~ "MonoMutation", # NA for one would mean partial, probably okay though
    seqwgs_cp_17p13_call == "no" & tp53_mut_call == "no" ~ "NONE",
    .default = "UNHANDLED CASE"
))

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(mixed_nsd2_1q21_class = dplyr::case_when(
    is.na(seqwgs_cp_1q21_call) | is.na(nsd2_call) ~ NA,
    seqwgs_cp_1q21_call == "yes" & nsd2_call == "yes" ~ "NSD2.Gain1q21",
    seqwgs_cp_1q21_call == "yes" & nsd2_call == "no" ~ "Gain1q21", # no yes is not possible
    seqwgs_cp_1q21_call == "no" & nsd2_call == "yes" ~ "NSD2", # no yes is not possible
    seqwgs_cp_1q21_call == "no" & nsd2_call == "no" ~ "Normal",
    .default = "UNHANDLED CASE"
))

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(mixed_maf_1q21_class = dplyr::case_when(
    is.na(seqwgs_cp_1q21_call) | is.na(maf_combined_call) ~ NA,
    seqwgs_cp_1q21_call == "yes" & maf_combined_call == "yes" ~ "MAF.Gain1q21",
    seqwgs_cp_1q21_call == "yes" & maf_combined_call == "no" ~ "Gain1q21", # no yes is not possible
    seqwgs_cp_1q21_call == "no" & maf_combined_call == "yes" ~ "MAF", # no yes is not possible
    seqwgs_cp_1q21_call == "no" & maf_combined_call == "no" ~ "Normal",
    .default = "UNHANDLED CASE"
))


baseline@meta.data <- baseline@meta.data |> dplyr::mutate(mixed_1p32_1q21_class = dplyr::case_when(
    is.na(seqwgs_cp_1q21_call) | is.na(seqwgs_cp_cdkn2c_call) ~ NA,
    seqwgs_cp_1q21_call == "yes" & seqwgs_cp_cdkn2c_call == "yes" ~ "1p32.Gain1q21",
    seqwgs_cp_1q21_call == "yes" & seqwgs_cp_cdkn2c_call == "no" ~ "Gain1q21", # no yes is not possible
    seqwgs_cp_1q21_call == "no" & seqwgs_cp_cdkn2c_call == "yes" ~ "1p32", # no yes is not possible
    seqwgs_cp_1q21_call == "no" & seqwgs_cp_cdkn2c_call == "no" ~ "Normal",
    .default = "UNHANDLED CASE"
))

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(any_high_risk = dplyr::case_when(
    is.na(seqwgs_cp_cdkn2c_call) | is.na(nsd2_call) | is.na(seqwgs_cp_17p13_call) | is.na(maf_combined_call) ~ NA,
    seqwgs_cp_cdkn2c_call == "yes" | nsd2_call == "yes" | seqwgs_cp_17p13_call == "yes" | maf_combined_call == "yes" ~ "yes",
    .default = "no"
))

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(mixed_HR_1q21_class = dplyr::case_when(
    is.na(seqwgs_cp_1q21_call) | is.na(any_high_risk) ~ NA,
    seqwgs_cp_1q21_call == "yes" & any_high_risk == "yes" ~ "HR.Gain1q21",
    seqwgs_cp_1q21_call == "yes" & any_high_risk == "no" ~ "Gain1q21", # no yes is not possible
    seqwgs_cp_1q21_call == "no" & any_high_risk == "yes" ~ "HR", # no yes is not possible
    seqwgs_cp_1q21_call == "no" & any_high_risk == "no" ~ "Normal",
    .default = "UNHANDLED CASE"
))

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(mixed_cn_1q21 = dplyr::case_when(
    is.na(seqwgs_cp_1q21_call) ~ NA,
    seqwgs_cp_1q21_call == "yes" & seqwgs_cp_derived_amp_1q == "yes" ~ "Amp1q21",
    seqwgs_cp_1q21_call == "yes" & seqwgs_cp_derived_amp_1q == "no" ~ "Gain1q21", # no yes is not possible
    seqwgs_cp_1q21_call == "no" & seqwgs_cp_derived_amp_1q == "no" ~ "Normal",
    .default = "UNHANDLED CASE"
))
```




```{r}
md_bl <- baseline@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::select(public_id, siteXbatch, dplyr::all_of(names(variable_info_structure)))

md_bl <- md_bl |> dplyr::distinct(public_id, .keep_all = T)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(!!sym(cluster_group), isDoublet, subcluster_V03072023_compartment, lineage_group, lineage_order, cellID_short, cellID_long)
colnames(cluster_md)[1] <- "cluster"
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")
```

```{r}
factor_names <- c()
for (i in names(variable_info_structure)) {
    if (variable_info_structure[[i]]$isFactor) {
        factor_names <- c(factor_names, paste0(i, variable_info_structure[[i]]$factorLevels[2]))
    } else {
        factor_names <- c(factor_names, i)
    }
}
```

# Core differential abundance analysis for risk - summary figures generated automatically. Final figures rendered out separately below.
For risk and BM PC PCT, 'All', 'Non-Malignant', 'Immune', and 'CD8+' are utilized.

```{r}
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)

counts <- table(baseline$public_id, baseline@meta.data[, cluster_group])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl <- baseline@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::select(public_id, siteXbatch, dplyr::all_of(names(variable_info_structure)))

md_bl <- md_bl |> dplyr::distinct(public_id, .keep_all = T)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(!!sym(cluster_group), isDoublet, subcluster_V03072023_compartment, lineage_group, lineage_order, cellID_short, cellID_long)
colnames(cluster_md)[1] <- "cluster"
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")

types_list <- list(
    "CD3+ Cells" = c("CD8", "CD4"),
    "Myeloid Cells" = "M",
    "B Cells" = "B",
    "CD8+" = "CD8",
    "CD4+" = "CD4",
    "Nk" = "Nk",
    "Immune" = c("CD8", "CD4", "M", "B", "Nk", "Other"),
    "All_NoPlasma" = c("CD8", "CD4", "M", "B", "Nk", "Other", "E"),
    "All" = c("CD8", "CD4", "M", "B", "Other", "Nk", "P", "E")
)
file_name_list <- list(
    "CD3+ Cells" = "cd3_only",
    "Myeloid Cells" = "myeloid_only",
    "B Cells" = "b_only",
    "CD8+" = "cd8_only",
    "CD4+" = "cd4_only",
    "Nk" = "nk_only",
    "Immune" = "immune_only",
    "All_NoPlasma" = "all_noplasma",
    "All" = "all_cells"
)


env_list <- list()
for (idx_name in names(types_list)) {
    all_cells <- new.env()

    all_cells$counts <- counts |> dplyr::filter(lineage_group %in% types_list[[idx_name]])
    all_cells$outs <- DiffAbun_Wrapper(all_cells$counts, variable_info_structure = variable_info_structure, DOMULTIVAR = F, method = "LINEAR", transform = "logit")

    # all_cells$outs <- quickLinearAsinAndPlots(all_cells$counts, variable_info_structure)
    write.csv(all_cells$outs$wilcox_out |> dplyr::arrange(factor), file.path(output_expr, paste0("linear_reg_", file_name_list[[idx_name]], ".csv")))
    write.csv(all_cells$outs$wilcox_out_uni |> dplyr::arrange(factor), file.path(output_expr, paste0("linear_reg_UNI", file_name_list[[idx_name]], ".csv")))
    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::filter(variable != "siteXbatch")
    all_cells$cluster_md <- cluster_md
    all_cells$cluster_md$subcluster_V03072023 <- all_cells$cluster_md$cluster

    all_cells$outs$wilcox_out <- dplyr::left_join(all_cells$outs$wilcox_out, all_cells$cluster_md, by = "cluster")
    all_cells$outs$wilcox_out_uni <- dplyr::left_join(all_cells$outs$wilcox_out_uni, all_cells$cluster_md, by = "cluster")

    all_cells$outs$wilcox_out$p.val <- as.numeric(all_cells$outs$wilcox_out$p.val)
    all_cells$outs$wilcox_out$coef <- as.numeric(all_cells$outs$wilcox_out$coef)

    all_cells$outs$wilcox_out_uni$p.val <- as.numeric(all_cells$outs$wilcox_out_uni$p.val)
    all_cells$outs$wilcox_out_uni$coef <- as.numeric(all_cells$outs$wilcox_out_uni$coef)


    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out$binned_pval <- factor(all_cells$outs$wilcox_out$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))

    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$binned_pval <- factor(all_cells$outs$wilcox_out_uni$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))



    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(sig_levels = dplyr::case_when(
        p.val < 0.01 ~ "Significant (p<0.01)",
        p.val < 0.05 ~ "Significant (p<0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$sig_levels <- factor(all_cells$outs$wilcox_out_uni$sig_levels, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))


    custom_fill_theme <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "grey"
    )

    custom_fill_theme_continuous <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "#f5f5f5"
    )
    library(reshape2)

    mat_multi <- reshape2::acast(all_cells$outs$wilcox_out, cellID_short ~ factor, value.var = "coef")
    d_multi_cluster <- dist(mat_multi / sqrt(rowSums(mat_multi * mat_multi)), method = "euclidean")
    multi_dendrogram_clusters <- hclust(d_multi_cluster)
    d_multi_factor <- dist(t(mat_multi) / sqrt(rowSums(t(mat_multi) * t(mat_multi))), method = "euclidean")
    multi_dendrogram_factor <- hclust(d_multi_factor)

    test <- all_cells$outs$wilcox_out_uni |> dplyr::filter(cluster == "Myeloid.0")
    mat_uni <- reshape2::acast(all_cells$outs$wilcox_out_uni, cellID_short ~ factor, value.var = "coef")
    d_uni_cluster <- dist(mat_uni / sqrt(rowSums(mat_uni * mat_uni)), method = "euclidean")
    uni_dendrogram_clusters <- hclust(d_uni_cluster)
    d_uni_factor <- dist(t(mat_uni) / sqrt(rowSums(t(mat_uni) * t(mat_uni))), method = "euclidean")
    uni_dendrogram_factor <- hclust(d_uni_factor)

    uni_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    uni_factor_ordering <- unique(all_cells$outs$wilcox_out_uni$factor) # factor_names # uni_dendrogram_factor$labels[uni_dendrogram_factor$order]


    all_cells$outs$wilcox_out_uni$factor <- factor(all_cells$outs$wilcox_out_uni$factor, levels = uni_factor_ordering)
    all_cells$outs$wilcox_out_uni$cellID_short <- factor(all_cells$outs$wilcox_out_uni$cellID_short, levels = uni_cluster_ordering)

    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4
    scale_factor <- 12

    p_uni_indiv <- ggplot(all_cells$outs$wilcox_out_uni) +
        geom_point(aes(x = cellID_short, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686") +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("Factor") +
        ggtitle("Univariate Linear Model (Corr sXb)") +
        coord_fixed() +
        guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
        theme(axis.text.x = element_text(
            size = 14 * 0.95,
            family = "sans", face = "bold",
            margin = margin(t = 0.8 * 14 / 4),
            angle = 90,
            vjust = 0.5,
            hjust = 1,
        ))

    # p_uni_comb <- (p_uni + theme(axis.text.x = element_blank(), axis.title.x = element_blank())) / (p_uni_2 + theme(plot.title = element_blank()))
    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4

    pdf_and_png(p_uni_indiv, output_expr, paste0("linear_reg_uni_ALT_COMBINED_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    plot_list <- list()
    for (i in types_list[[idx_name]]) {
        all_cells$sub_wilcox_out <- all_cells$outs$wilcox_out_uni |> dplyr::filter(lineage_group %in% i)

        sub_mat_uni <- reshape2::acast(all_cells$sub_wilcox_out, cellID_short ~ factor, value.var = "coef")
        sub_d_uni_cluster <- dist(sub_mat_uni / sqrt(rowSums(sub_mat_uni * sub_mat_uni)), method = "euclidean")
        sub_uni_dendrogram_clusters <- hclust(sub_d_uni_cluster)


        sub_uni_cluster_ordering <- sub_uni_dendrogram_clusters$labels[sub_uni_dendrogram_clusters$order]


        all_cells$sub_wilcox_out$sig_binned_value <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c(
            "NS", "Significant (p<0.05)", "Significant (p<0.01)"
        ))
        all_cells$sub_wilcox_out$binned_pval <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c("Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"))
        #                low = "#290AD8", mid = "grey", high = "#A50021", limits = c(-0.31, 0.31),

        plot_list[[i]] <- ggplot(all_cells$sub_wilcox_out) +
            geom_point(aes(x = cellID_short, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
            scale_fill_gradient2(
                low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
                oob = squish
            ) +
            scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
            scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
            theme_prism() +
            xlab("Cluster") +
            ylab("Factor") +
            ggtitle(i) +
            coord_fixed() +
            guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
            theme(
                axis.text.x = element_text(
                    size = 14 * 0.95,
                    family = "sans", face = "bold",
                    margin = margin(t = 0.8 * 14 / 4),
                    angle = 90,
                    vjust = 0.5,
                    hjust = 1,
                ),
                axis.title.x = element_blank(),
                plot.margin = unit(c(0, 1, 0, 0), "lines")
            )
        if (i != types_list[[idx_name]][[1]]) {
            plot_list[[i]] <- plot_list[[i]] + theme(
                axis.text.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.title.y = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.length.y = unit(0, "lines")
            )
        }
    }

    for (i in unique(all_cells$outs$wilcox_out_uni$cluster)) {
        plot <- patchwork::wrap_plots(plot_list, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")
    }
    pdf_and_png(plot, output_expr, paste0("linear_reg_uni_ALT_COMBINED_FACETTED_GROUPS_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    env_list[[idx_name]] <- all_cells
}
```



# Figure 3A - Summary of differnetial abundance results for davies based risk
Non-malignant shown for risk summary figures.

```{r}
idx_name <- "All_NoPlasma"
all_cells <- env_list[[idx_name]]

wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("maf_combined_call", "seqwgs_cp_17p13_call", "seqwgs_cp_1q21_call", "nsd2_call", "davies_based_risk"))
tmp_VIS <- variable_info_structure[c("seqwgs_cp_17p13_call", "maf_combined_call", "seqwgs_cp_1q21_call", "nsd2_call", "davies_based_risk")]

# regretting past design decisions but its okay
variable_names <- c(
    tmp_VIS[[names(tmp_VIS)[[1]]]][["variable_name"]],
    tmp_VIS[[names(tmp_VIS)[[2]]]][["variable_name"]],
    tmp_VIS[[names(tmp_VIS)[[3]]]][["variable_name"]],
    tmp_VIS[[names(tmp_VIS)[[4]]]][["variable_name"]],
    tmp_VIS[[names(tmp_VIS)[[5]]]][["variable_name"]]
)

lineage_group_names_df <- data.frame(
    lineage_group = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    lineage_group_name = c("CD8+ T Cell", "CD4+ T Cell", "Myeloid", "B Cell", "Other", "Nk Cell", "Erythroid")
)
rownames(lineage_group_names_df) <- lineage_group_names_df$lineage_group
vis_df <- data.frame(variable = names(tmp_VIS), variable_names = variable_names)
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, vis_df, by = "variable")
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, lineage_group_names_df, by = "lineage_group")

wilcox_out_uni$variable_names <- factor(wilcox_out_uni$variable_names, levels = variable_names)

heights <- length(unique(wilcox_out_uni$factor))
char_len_uni <- max(nchar(levels(wilcox_out_uni$variable_names)))
widths <- length(unique(wilcox_out_uni$cluster)) + char_len_uni / 4
scale_factor <- 12

top_plot_info <- wilcox_out_uni |> dplyr::filter(variable %in% c("davies_based_risk"))

bottom_plot_info <- wilcox_out_uni |> dplyr::filter(variable %in% c("maf_combined_call", "seqwgs_cp_17p13_call", "seqwgs_cp_1q21_call", "nsd2_call"))

top_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
    sub_wilcox_out <- top_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    top_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        ggtitle(lineage_group_name) +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.x = element_blank(),
            axis.line.x = element_blank(),
            plot.margin = unit(c(0, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        top_plot_list[[i]] <- top_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    top_plot_list[[i]] <- top_plot_list[[i]] + coord_fixed()
}

bottom_plot_list <- list()
for (i in types_list[[idx_name]]) {
    sub_wilcox_out <- bottom_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    bottom_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_text(
                size = 14 * 0.95,
                family = "sans", face = "bold",
                margin = margin(t = 0.8 * 14 / 4),
                angle = 90,
                vjust = 0.5,
                hjust = 1,
            ),
            axis.title.x = element_blank(),
            plot.margin = unit(c(0.25, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        bottom_plot_list[[i]] <- bottom_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    bottom_plot_list[[i]] <- bottom_plot_list[[i]] + coord_fixed()
}

plot <- patchwork::wrap_plots(c(top_plot_list, bottom_plot_list), nrow = 2, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")

pdf_and_png(plot, output_expr, paste0("Figure_3_Diff_Abundance_Summary_Figure_noplasma"), pdfWidth = W_FULL * 3, pdfHeight = 2, scale = 3)
```




# NSD2+1q21 and all CD8+ T cells
```{r}
idx_name <- "CD8+"
all_cells <- env_list[[idx_name]]

wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("mixed_nsd2_1q21_class", "davies_based_risk", "nsd2_call", "seqwgs_cp_1q21_call"))

# regretting past design decisions but its okay
vis_df <- data.frame(
    factor = c(
        "davies_based_riskhigh_risk",
        "nsd2_callyes",
        "seqwgs_cp_1q21_callyes",
        "mixed_nsd2_1q21_class: NSD2 - Normal",
        "mixed_nsd2_1q21_class: Gain1q21 - Normal",
        "mixed_nsd2_1q21_class: NSD2.Gain1q21 - Normal",
        "mixed_nsd2_1q21_class: Gain1q21 - NSD2",
        "mixed_nsd2_1q21_class: NSD2.Gain1q21 - Gain1q21",
        "mixed_nsd2_1q21_class: NSD2.Gain1q21 - NSD2"
    ),
    variable_names = c(
        "High Risk",
        "NSD2 t(4;14)",
        "Gain 1q21",
        "NSD2 - Neither",
        "1q21 - Neither",
        "NSD2+1q21 - Neither",
        "1q21 - NSD2",
        "NSD2+1q21 - 1q21",
        "NSD2+1q21 - NSD2"
    )
)


lineage_group_names_df <- data.frame(
    lineage_group = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    lineage_group_name = c("CD8+ T Cell", "CD4+ T Cell", "Myeloid", "B Cell", "Other", "Nk Cell", "Erythroid")
)

rownames(lineage_group_names_df) <- lineage_group_names_df$lineage_group
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, vis_df, by = "factor")
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, lineage_group_names_df, by = "lineage_group")


wilcox_out_uni$variable_names <- factor(wilcox_out_uni$variable_names, levels = rev(c(
    "High Risk",
    "NSD2 t(4;14)",
    "Gain 1q21",
    "NSD2 - Neither",
    "1q21 - Neither",
    "NSD2+1q21 - Neither",
    "1q21 - NSD2",
    "NSD2+1q21 - 1q21",
    "NSD2+1q21 - NSD2"
)))

heights <- length(unique(wilcox_out_uni$factor))
char_len_uni <- max(nchar(levels(wilcox_out_uni$variable_names)))
widths <- length(unique(wilcox_out_uni$cluster)) + char_len_uni / 4
scale_factor <- 12

top_plot_info <- wilcox_out_uni |> dplyr::filter(variable %in% c("davies_based_risk", "nsd2_call", "seqwgs_cp_1q21_call"))

middle_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "NSD2 - Neither",
    "1q21 - Neither",
    "NSD2+1q21 - Neither"
))


bottom_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "1q21 - NSD2",
    "NSD2+1q21 - 1q21",
    "NSD2+1q21 - NSD2"
))

top_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
    sub_wilcox_out <- top_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    top_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        ggtitle(lineage_group_name) +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.x = element_blank(),
            axis.line.x = element_blank(),
            plot.margin = unit(c(0, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        top_plot_list[[i]] <- top_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    top_plot_list[[i]] <- top_plot_list[[i]] + coord_fixed()
}


middle_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
    sub_wilcox_out <- middle_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    middle_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.x = element_blank(),
            axis.line.x = element_blank(),
            plot.margin = unit(c(0, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        middle_plot_list[[i]] <- middle_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    middle_plot_list[[i]] <- middle_plot_list[[i]] + coord_fixed()
}

bottom_plot_list <- list()
for (i in types_list[[idx_name]]) {
    sub_wilcox_out <- bottom_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    bottom_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_text(
                size = 14 * 0.95,
                family = "sans", face = "bold",
                margin = margin(t = 0.8 * 14 / 4),
                angle = 90,
                vjust = 0.5,
                hjust = 1,
            ),
            axis.title.x = element_blank(),
            plot.margin = unit(c(0.25, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        bottom_plot_list[[i]] <- bottom_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    bottom_plot_list[[i]] <- bottom_plot_list[[i]] + coord_fixed()
}

plot <- patchwork::wrap_plots(c(top_plot_list, middle_plot_list, bottom_plot_list), nrow = 3, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")

pdf_and_png(plot, output_expr, paste0("Figure_3_Diff_Abundance_T_Cell_NSD2_1q21"), pdfWidth = W_FULL * 3, pdfHeight = 2, scale = 3)
```



# Fig 3, Chromothripsis + APOBEC 
Non-Malignant summary dot plot

```{r}
# sapply(test$factor, FUN = function(x) {
#     nchar(x)
# })

idx_name <- "All_NoPlasma"
all_cells <- env_list[[idx_name]]

wilcox_out_uni <- all_cells$outs$wilcox_out_uni #|> dplyr::filter(variable %in% c("mixed_nsd2_1q21_class", "davies_based_risk", "nsd2_call", "seqwgs_cp_1q21_call"))

# regretting past design decisions but its okay
vis_df <- data.frame(
    factor = c(
        "davies_based_riskhigh_risk",
        "chromothripsisyes",
        "APOBEC_Enrichedyes"
    ),
    variable_names = c(
        "High Risk (Davies)",
        "Chromothripsis",
        "APOBEC"
    )
)


lineage_group_names_df <- data.frame(
    lineage_group = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    lineage_group_name = c("CD8+ T Cell", "CD4+ T Cell", "Myeloid", "B Cell", "Other", "Nk Cell", "Erythroid")
)

rownames(lineage_group_names_df) <- lineage_group_names_df$lineage_group
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, vis_df, by = "factor")
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, lineage_group_names_df, by = "lineage_group")


wilcox_out_uni$variable_names <- factor(wilcox_out_uni$variable_names, levels = rev(c(
    "High Risk (Davies)",
    "Chromothripsis",
    "APOBEC",
    "Amp 1q21",
    "Del 1p32",
    "TP53 Mutation"
)))

heights <- length(unique(wilcox_out_uni$factor))
char_len_uni <- max(nchar(levels(wilcox_out_uni$variable_names)))
widths <- length(unique(wilcox_out_uni$cluster)) + char_len_uni / 4
scale_factor <- 12

top_plot_info <- wilcox_out_uni |> dplyr::filter(variable %in% c("davies_based_risk"))

# middle_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
#     "NSD2 - Neither",
#     "1q21 - Neither",
#     "NSD2+1q21 - Neither"
# ))


bottom_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "Chromothripsis",
    "APOBEC",
    "Amp 1q21",
    "Del 1p32",
    "TP53 Mutation"
))

top_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
    sub_wilcox_out <- top_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    top_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        ggtitle(lineage_group_name) +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.x = element_blank(),
            axis.line.x = element_blank(),
            plot.margin = unit(c(0, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        top_plot_list[[i]] <- top_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    top_plot_list[[i]] <- top_plot_list[[i]] + coord_fixed()
}


# middle_plot_list <- list()
# for (i in types_list[[idx_name]]) {
#     lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
#     sub_wilcox_out <- middle_plot_info |> dplyr::filter(lineage_group %in% i)

#     sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
#         "NS", "Significant (p<0.05)", "Significant (p<0.01)"
#     ))

#     middle_plot_list[[i]] <- ggplot(sub_wilcox_out) +
#         geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
#         scale_fill_gradient2(
#             low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
#             oob = squish
#         ) +
#         scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
#         scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
#         theme_prism() +
#         xlab("Cluster") +
#         ylab("") +
#         guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
#         theme(
#             axis.text.x = element_blank(),
#             axis.ticks.x = element_blank(),
#             axis.title.x = element_blank(),
#             axis.line.x = element_blank(),
#             plot.margin = unit(c(0, 1, 0, 0), "lines")
#         )
#     if (i != types_list[[idx_name]][[1]]) {
#         middle_plot_list[[i]] <- middle_plot_list[[i]] + theme(
#             axis.text.y = element_blank(),
#             axis.ticks.y = element_blank(),
#             axis.title.y = element_blank(),
#             axis.line.y = element_blank(),
#             axis.ticks.length.y = unit(0, "lines")
#         )
#     }

#     middle_plot_list[[i]] <- middle_plot_list[[i]] + coord_fixed()
# }

bottom_plot_list <- list()
for (i in types_list[[idx_name]]) {
    sub_wilcox_out <- bottom_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    bottom_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_text(
                size = 14 * 0.95,
                family = "sans", face = "bold",
                margin = margin(t = 0.8 * 14 / 4),
                angle = 90,
                vjust = 0.5,
                hjust = 1,
            ),
            axis.title.x = element_blank(),
            plot.margin = unit(c(0.25, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        bottom_plot_list[[i]] <- bottom_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    bottom_plot_list[[i]] <- bottom_plot_list[[i]] + coord_fixed()
}

plot <- patchwork::wrap_plots(c(top_plot_list, bottom_plot_list), nrow = 2, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")

pdf_and_png(plot, output_expr, paste0("Figure_3_Diff_Abundance_Other_Covariates"), pdfWidth = W_FULL * 3, pdfHeight = 2, scale = 3)
```

# Fig 3, 1Q Gain, Amp

```{r}
# sapply(test$factor, FUN = function(x) {
#     nchar(x)
# })

idx_name <- "All_NoPlasma"
all_cells <- env_list[[idx_name]]

wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("mixed_cn_1q21", "davies_based_risk", "seqwgs_cp_derived_amp_1q", "seqwgs_cp_1q21_call"))

# regretting past design decisions but its okay
vis_df <- data.frame(
    factor = c(
        "davies_based_riskhigh_risk",
        "seqwgs_cp_1q21_callyes",
        "seqwgs_cp_derived_amp_1q",
        "mixed_cn_1q21: Amp1q21 - Normal",
        "mixed_cn_1q21: Gain1q21 - Normal",
        "mixed_cn_1q21: Amp1q21 - Gain1q21"
    ),
    variable_names = c(
        "High Risk",
        "Gain 1q21",
        "Amp 1q21",
        "Amp - Normal",
        "Gain - Normal",
        "AMP - Gain"
    )
)


lineage_group_names_df <- data.frame(
    lineage_group = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    lineage_group_name = c("CD8+ T Cell", "CD4+ T Cell", "Myeloid", "B Cell", "Other", "Nk Cell", "Erythroid")
)

rownames(lineage_group_names_df) <- lineage_group_names_df$lineage_group
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, vis_df, by = "factor")
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, lineage_group_names_df, by = "lineage_group")


wilcox_out_uni$variable_names <- factor(wilcox_out_uni$variable_names, levels = rev(c(
    "High Risk",
    "Gain 1q21",
    "Amp 1q21",
    "Amp - Normal",
    "Gain - Normal",
    "AMP - Gain"
)))

heights <- length(unique(wilcox_out_uni$factor))
char_len_uni <- max(nchar(levels(wilcox_out_uni$variable_names)))
widths <- length(unique(wilcox_out_uni$cluster)) + char_len_uni / 4
scale_factor <- 12

top_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "High Risk",
    "Gain 1q21",
    "Amp 1q21"
))

middle_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "Amp - Normal",
    "Gain - Normal"
))


bottom_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "AMP - Gain"
))

top_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
    sub_wilcox_out <- top_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    top_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        ggtitle(lineage_group_name) +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.x = element_blank(),
            axis.line.x = element_blank(),
            plot.margin = unit(c(0, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        top_plot_list[[i]] <- top_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    top_plot_list[[i]] <- top_plot_list[[i]] + coord_fixed()
}


middle_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
    sub_wilcox_out <- middle_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    middle_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.x = element_blank(),
            axis.line.x = element_blank(),
            plot.margin = unit(c(0, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        middle_plot_list[[i]] <- middle_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    middle_plot_list[[i]] <- middle_plot_list[[i]] + coord_fixed()
}

bottom_plot_list <- list()
for (i in types_list[[idx_name]]) {
    sub_wilcox_out <- bottom_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    bottom_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_text(
                size = 14 * 0.95,
                family = "sans", face = "bold",
                margin = margin(t = 0.8 * 14 / 4),
                angle = 90,
                vjust = 0.5,
                hjust = 1,
            ),
            axis.title.x = element_blank(),
            plot.margin = unit(c(0.25, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        bottom_plot_list[[i]] <- bottom_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    bottom_plot_list[[i]] <- bottom_plot_list[[i]] + coord_fixed()
}

plot <- patchwork::wrap_plots(c(top_plot_list, middle_plot_list, bottom_plot_list), nrow = 3, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")

pdf_and_png(plot, output_expr, paste0("Figure_3_Diff_Abundance_AllNoPlasma_1q21_gain_amp"), pdfWidth = W_FULL * 3, pdfHeight = 2, scale = 3)
```


# Supplemental Figure 18B

TP53 loss, non-malignant included in paper
```{r}
# sapply(test$factor, FUN = function(x) {
#     nchar(x)
# })
for (idx_name in c("CD3+ Cells", "CD8+", "CD4+", "Myeloid Cells", "All_NoPlasma")) {
    # idx_name <- "All_NoPlasma"
    all_cells <- env_list[[idx_name]]

    wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("tp53_inactivation_state", "davies_based_risk", "tp53_mut_call", "seqwgs_cp_17p13_call"))

    # regretting past design decisions but its okay
    vis_df <- data.frame(
        factor = c(
            "davies_based_riskhigh_risk",
            "seqwgs_cp_17p13_callyes",
            "tp53_mut_callyes",
            "tp53_inactivation_state: PARTIAL - NONE",
            "tp53_inactivation_state: COMPLETE - NONE",
            "tp53_inactivation_state: COMPLETE - PARTIAL"
        ),
        variable_names = c(
            "High Risk",
            "Del 17p13",
            "Mut TP53",
            "Partial Loss - WT",
            "Complete Loss - WT",
            "Complete - Partial Loss"
        )
    )


    lineage_group_names_df <- data.frame(
        lineage_group = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
        lineage_group_name = c("CD8+ T Cell", "CD4+ T Cell", "Myeloid", "B Cell", "Other", "Nk Cell", "Erythroid")
    )

    rownames(lineage_group_names_df) <- lineage_group_names_df$lineage_group
    wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, vis_df, by = "factor")
    wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, lineage_group_names_df, by = "lineage_group")


    wilcox_out_uni$variable_names <- factor(wilcox_out_uni$variable_names, levels = rev(c(
        "High Risk",
        "Del 17p13",
        "Mut TP53",
        "Partial Loss - WT",
        "Complete Loss - WT",
        "Complete - Partial Loss"
    )))

    heights <- length(unique(wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(wilcox_out_uni$variable_names)))
    widths <- length(unique(wilcox_out_uni$cluster)) + char_len_uni / 4
    scale_factor <- 12

    top_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
        "High Risk",
        "Del 17p13",
        "Mut TP53"
    ))

    middle_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
        "Partial Loss - WT",
        "Complete Loss - WT"
    ))


    bottom_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
        "Complete - Partial Loss"
    ))

    top_plot_list <- list()
    for (i in types_list[[idx_name]]) {
        lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
        sub_wilcox_out <- top_plot_info |> dplyr::filter(lineage_group %in% i)

        sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
            "NS", "Significant (p<0.05)", "Significant (p<0.01)"
        ))

        top_plot_list[[i]] <- ggplot(sub_wilcox_out) +
            geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
            scale_fill_gradient2(
                low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
                oob = squish
            ) +
            scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
            scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
            theme_prism() +
            xlab("Cluster") +
            ylab("") +
            ggtitle(lineage_group_name) +
            guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
            theme(
                axis.text.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.title.x = element_blank(),
                axis.line.x = element_blank(),
                plot.margin = unit(c(0, 1, 0, 0), "lines")
            )
        if (i != types_list[[idx_name]][[1]]) {
            top_plot_list[[i]] <- top_plot_list[[i]] + theme(
                axis.text.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.title.y = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.length.y = unit(0, "lines")
            )
        }

        top_plot_list[[i]] <- top_plot_list[[i]] + coord_fixed()
    }


    middle_plot_list <- list()
    for (i in types_list[[idx_name]]) {
        lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
        sub_wilcox_out <- middle_plot_info |> dplyr::filter(lineage_group %in% i)

        sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
            "NS", "Significant (p<0.05)", "Significant (p<0.01)"
        ))

        middle_plot_list[[i]] <- ggplot(sub_wilcox_out) +
            geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
            scale_fill_gradient2(
                low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
                oob = squish
            ) +
            scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
            scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
            theme_prism() +
            xlab("Cluster") +
            ylab("") +
            guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
            theme(
                axis.text.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.title.x = element_blank(),
                axis.line.x = element_blank(),
                plot.margin = unit(c(0, 1, 0, 0), "lines")
            )
        if (i != types_list[[idx_name]][[1]]) {
            middle_plot_list[[i]] <- middle_plot_list[[i]] + theme(
                axis.text.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.title.y = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.length.y = unit(0, "lines")
            )
        }

        middle_plot_list[[i]] <- middle_plot_list[[i]] + coord_fixed()
    }

    bottom_plot_list <- list()
    for (i in types_list[[idx_name]]) {
        sub_wilcox_out <- bottom_plot_info |> dplyr::filter(lineage_group %in% i)

        sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
            "NS", "Significant (p<0.05)", "Significant (p<0.01)"
        ))

        bottom_plot_list[[i]] <- ggplot(sub_wilcox_out) +
            geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
            scale_fill_gradient2(
                low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
                oob = squish
            ) +
            scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
            scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
            theme_prism() +
            xlab("Cluster") +
            ylab("") +
            guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
            theme(
                axis.text.x = element_text(
                    size = 14 * 0.95,
                    family = "sans", face = "bold",
                    margin = margin(t = 0.8 * 14 / 4),
                    angle = 90,
                    vjust = 0.5,
                    hjust = 1,
                ),
                axis.title.x = element_blank(),
                plot.margin = unit(c(0.25, 1, 0, 0), "lines")
            )
        if (i != types_list[[idx_name]][[1]]) {
            bottom_plot_list[[i]] <- bottom_plot_list[[i]] + theme(
                axis.text.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.title.y = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.length.y = unit(0, "lines")
            )
        }

        bottom_plot_list[[i]] <- bottom_plot_list[[i]] + coord_fixed()
    }

    plot <- patchwork::wrap_plots(c(top_plot_list, middle_plot_list, bottom_plot_list), nrow = 3, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")

    pdf_and_png(plot, output_expr, paste0("Figure_3_Diff_Abundance_", idx_name, "_tp53_state_amp"), pdfWidth = W_FULL * 3, pdfHeight = 2, scale = 3)
}
```

# Supplemental Figure 18A

```{r}
# sapply(test$factor, FUN = function(x) {
#     nchar(x)
# })

idx_name <- "All_NoPlasma"
all_cells <- env_list[[idx_name]]

wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("mixed_HR_1q21_class", "seqwgs_cp_1q21_call", "davies_based_risk"))

vis_df <- data.frame(
    factor = c(
        "davies_based_riskhigh_risk",
        "seqwgs_cp_1q21_callyes",
        "mixed_HR_1q21_class: HR - Normal",
        "mixed_HR_1q21_class: Gain1q21 - Normal",
        "mixed_HR_1q21_class: HR.Gain1q21 - Normal",
        "mixed_HR_1q21_class: Gain1q21 - HR",
        "mixed_HR_1q21_class: HR.Gain1q21 - Gain1q21",
        "mixed_HR_1q21_class: HR.Gain1q21 - HR"
    ),
    variable_names = c(
        "High Risk",
        "Gain 1q21",
        "HR_No1q - Normal",
        "1q21 - Normal",
        "HR+1q21 - Normal",
        "1q21 - HR_No1q",
        "HR+1q21 - 1q21",
        "HR+1q21 - HR_No1q"
    )
)


lineage_group_names_df <- data.frame(
    lineage_group = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    lineage_group_name = c("CD8+ T Cell", "CD4+ T Cell", "Myeloid", "B Cell", "Other", "Nk Cell", "Erythroid")
)


rownames(lineage_group_names_df) <- lineage_group_names_df$lineage_group
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, vis_df, by = "factor")
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, lineage_group_names_df, by = "lineage_group")


wilcox_out_uni$variable_names <- factor(wilcox_out_uni$variable_names, levels = rev(c(
    "High Risk",
    "Gain 1q21",
    "HR_No1q - Normal",
    "1q21 - Normal",
    "HR+1q21 - Normal",
    "1q21 - HR_No1q",
    "HR+1q21 - 1q21",
    "HR+1q21 - HR_No1q"
)))

heights <- length(unique(wilcox_out_uni$factor))
char_len_uni <- max(nchar(levels(wilcox_out_uni$variable_names)))
widths <- length(unique(wilcox_out_uni$cluster)) + char_len_uni / 4
scale_factor <- 12

top_plot_info <- wilcox_out_uni |> dplyr::filter(variable %in% c("davies_based_risk", "seqwgs_cp_1q21_call"))

middle_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "HR_No1q - Normal",
    "1q21 - Normal",
    "HR+1q21 - Normal"
))


bottom_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "1q21 - HR_No1q",
    "HR+1q21 - 1q21",
    "HR+1q21 - HR_No1q"
))

top_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
    sub_wilcox_out <- top_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    top_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        ggtitle(lineage_group_name) +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.x = element_blank(),
            axis.line.x = element_blank(),
            plot.margin = unit(c(0, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        top_plot_list[[i]] <- top_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    top_plot_list[[i]] <- top_plot_list[[i]] + coord_fixed()
}


middle_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]
    sub_wilcox_out <- middle_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    middle_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.x = element_blank(),
            axis.line.x = element_blank(),
            plot.margin = unit(c(0, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        middle_plot_list[[i]] <- middle_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    middle_plot_list[[i]] <- middle_plot_list[[i]] + coord_fixed()
}

bottom_plot_list <- list()
for (i in types_list[[idx_name]]) {
    sub_wilcox_out <- bottom_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    bottom_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_text(
                size = 14 * 0.95,
                family = "sans", face = "bold",
                margin = margin(t = 0.8 * 14 / 4),
                angle = 90,
                vjust = 0.5,
                hjust = 1,
            ),
            axis.title.x = element_blank(),
            plot.margin = unit(c(0.25, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        bottom_plot_list[[i]] <- bottom_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    bottom_plot_list[[i]] <- bottom_plot_list[[i]] + coord_fixed()
}

plot <- patchwork::wrap_plots(c(top_plot_list, middle_plot_list, bottom_plot_list), nrow = 3, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")

pdf_and_png(plot, output_expr, paste0("Figure_3_Diff_Abundance_HR_1q21"), pdfWidth = W_FULL * 3, pdfHeight = 2, scale = 3)
```


# Figure 3C - CD8+ HLA Diff Abundance
CD8+ Teff HLA as a fraction of CD8+ Cells
Shown for davies High and Standard Risk, and for various combinations of NSD2 and 1q21 gain

```{r}
idx_name <- "CD8+"
all_cells <- env_list[[idx_name]]
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("mixed_nsd2_1q21_class", "seqwgs_cp_1q21_call", "davies_based_risk", "nsd2_call"))


counts_hr_v_sr <- all_cells$outs$counts_uni[["davies_based_risk"]] |> dplyr::filter(cluster == "NkT.2.1")

hr_only_pvalue <- wilcox_out_uni |> dplyr::filter(cluster == "NkT.2.1", variable == "davies_based_risk")


counts_hr_v_sr <- counts_hr_v_sr |> dplyr::mutate(nice_names = dplyr::case_when(
    davies_based_risk == "high_risk" ~ "High\nRisk",
    davies_based_risk == "standard_risk" ~ "Standard\nRisk"
))

counts_hr_v_sr$nice_names <- factor(counts_hr_v_sr$nice_names, levels = c("Standard\nRisk", "High\nRisk"))

hr_stat_df <- data.frame(p.val = hr_only_pvalue$p.val, group1 = "High\nRisk", group2 = "Standard\nRisk")


p1 <- counts_hr_v_sr |>
    ggplot() +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = nice_names, fill = davies_based_risk, color = davies_based_risk, group = davies_based_risk, y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = nice_names, fill = davies_based_risk, color = davies_based_risk, group = davies_based_risk, y = prop)) +
    stat_pvalue_manual(hr_stat_df, label = "{scales::pvalue(p.val)}", size = 5, y.position = .3, step_increase = 0.025, hide.ns = T) +
    theme_prism() +
    # ylim(c(0, box_plots$tmp_distinct$y.position)) +
    scale_fill_manual(values = c(
        "high_risk" = fill_RP,
        "standard_risk" = fill_NP
    )) +
    scale_color_manual(values = c(
        "high_risk" = color_RP,
        "standard_risk" = color_NP
    )) +
    theme(axis.title.x = element_blank()) +
    ylim(0, .35) +
    ylab(bquote(atop("CD8+ Teff HLA", "Proportion of CD8+ Cells"))) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    theme(plot.title = element_text(size = 14, face = "bold"))

counts_nsd2_mixed <- all_cells$outs$counts_uni[["mixed_nsd2_1q21_class"]] |> dplyr::filter(cluster == "NkT.2.1")
nsd2_pvalues <- wilcox_out_uni |> dplyr::filter(cluster == "NkT.2.1", variable == "mixed_nsd2_1q21_class")
nsd2_stat_df <- data.frame(p.val = nsd2_pvalues$p.val, group1 = c("NSD2", "Gain1q21", "NSD2.Gain1q21", "Gain1q21", "NSD2.Gain1q21", "NSD2.Gain1q21"), group2 = c("Normal", "Normal", "Normal", "NSD2", "NSD2", "Gain1q21"))

nsd2_stat_df <- nsd2_stat_df |> dplyr::filter(p.val < 0.1)
nsd2_stat_df$y.position <- .3
for (idx in 1:length(nsd2_stat_df$y.position)) {
    nsd2_stat_df[idx, "y.position"] <- nsd2_stat_df[idx, "y.position"] + 0.35 * (idx - 1) * .05
}

counts_nsd2_mixed <- counts_nsd2_mixed |> dplyr::mutate(nice_names = dplyr::case_when(
    mixed_nsd2_1q21_class == "Normal" ~ "Neither",
    mixed_nsd2_1q21_class == "NSD2" ~ "NSD2\nt(4;14)",
    mixed_nsd2_1q21_class == "Gain1q21" ~ "Gain 1q21",
    mixed_nsd2_1q21_class == "NSD2.Gain1q21" ~ "NSD2+\nGain1q21"
))

counts_nsd2_mixed$nice_names <- factor(counts_nsd2_mixed$nice_names, levels = c("Neither", "NSD2\nt(4;14)", "Gain 1q21", "NSD2+\nGain1q21"))

nsd2_stat_df <- nsd2_stat_df |> dplyr::mutate(
    group1 = dplyr::case_when(
        group1 == "Normal" ~ "Neither",
        group1 == "NSD2" ~ "NSD2\nt(4;14)",
        group1 == "Gain1q21" ~ "Gain 1q21",
        group1 == "NSD2.Gain1q21" ~ "NSD2+\nGain1q21"
    ),
    group2 = dplyr::case_when(
        group2 == "Normal" ~ "Neither",
        group2 == "NSD2" ~ "NSD2\nt(4;14)",
        group2 == "Gain1q21" ~ "Gain\n1q21",
        group2 == "NSD2.Gain1q21" ~ "NSD2+\nGain1q21"
    )
)


p2 <- counts_nsd2_mixed |>
    ggplot() +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = nice_names, fill = mixed_nsd2_1q21_class, color = mixed_nsd2_1q21_class, group = mixed_nsd2_1q21_class, y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = nice_names, fill = mixed_nsd2_1q21_class, color = mixed_nsd2_1q21_class, group = mixed_nsd2_1q21_class, y = prop)) +
    stat_pvalue_manual(nsd2_stat_df, label = "{scales::pvalue(p.val)}", size = 5, y.position = "y.position", step_increase = 1) +
    theme_prism() +
    # ylim(c(0, box_plots$tmp_distinct$y.position)) +
    scale_fill_manual(values = c(
        "Normal" = fill_NP,
        "NSD2" = "#d9a630",
        "Gain1q21" = fill_RP,
        "NSD2.Gain1q21" = "#d93063"
    )) +
    scale_color_manual(values = c(
        "Normal" = color_NP,
        "NSD2" = "#b58821",
        "Gain1q21" = color_RP,
        "NSD2.Gain1q21" = "#9f1d44"
    )) +
    theme(axis.title.x = element_blank()) +
    ylim(0, .35) +
    ylab("") +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    theme(plot.title = element_text(size = 14, face = "bold"))

test <- patchwork::wrap_plots(list(p1, p2), ncol = 2, nrow = 1, widths = c(1, 2))

pdf_and_png(test, output_expr, "boxplots_HLA_cd3_pct", pdfWidth = W_FULL * 1.25, pdfHeight = H_FULL, scale = 2.5)
```




# Diff Abundance 


### Auto do the rest

```{r}
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)

# baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
# Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline[[var_to_test]], "_", baseline[[var_to_test]])

counts <- table(baseline$public_id, baseline@meta.data[, "lineage_group"])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl_all_var <- baseline@meta.data |> dplyr::distinct(public_id, .keep_all = T)


md_bl <- baseline@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::select(public_id, siteXbatch, dplyr::all_of(names(variable_info_structure)))

md_bl <- md_bl |> dplyr::distinct(public_id, .keep_all = T)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(lineage_group)
cluster_md$cluster <- cluster_md$lineage_group
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")

types_list <- list(
    "LG CD3+ Cells" = c("CD8", "CD4"),
    "LG Immune" = c("CD8", "CD4", "M", "B", "Other", "Nk"),
    "LG All NoPlasma" = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    "LG All" = c("CD8", "CD4", "M", "B", "Other", "Nk", "P", "E")
)
file_name_list <- list(
    "LG CD3+ Cells" = "LINEAGE_GROUP_cd3_only",
    "LG Immune" = "LINEAGE_GROUP_immune_only",
    "LG All NoPlasma" = "LINEAGE_GROUP_All_NoPlasma",
    "LG All" = "LINEAGE_GROUP_all_cells"
)

for (idx_name in names(types_list)) {
    all_cells <- new.env()

    all_cells$counts <- counts |> dplyr::filter(lineage_group %in% types_list[[idx_name]])
    all_cells$outs <- DiffAbun_Wrapper(all_cells$counts, variable_info_structure = variable_info_structure, DOMULTIVAR = F, method = "LINEAR", transform = "logit")

    # all_cells$outs <- quickLinearAsinAndPlots(all_cells$counts, variable_info_structure)
    write.csv(all_cells$outs$wilcox_out |> dplyr::arrange(factor), file.path(output_expr, paste0("linear_reg_", file_name_list[[idx_name]], ".csv")))
    write.csv(all_cells$outs$wilcox_out_uni |> dplyr::arrange(factor), file.path(output_expr, paste0("linear_reg_UNI_", file_name_list[[idx_name]], ".csv")))
    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::filter(variable != "siteXbatch")
    all_cells$cluster_md <- cluster_md
    all_cells$cluster_md$subcluster_V03072023 <- all_cells$cluster_md$cluster

    all_cells$outs$wilcox_out <- dplyr::left_join(all_cells$outs$wilcox_out, all_cells$cluster_md, by = "cluster")
    all_cells$outs$wilcox_out_uni <- dplyr::left_join(all_cells$outs$wilcox_out_uni, all_cells$cluster_md, by = "cluster")

    all_cells$outs$wilcox_out$p.val <- as.numeric(all_cells$outs$wilcox_out$p.val)
    all_cells$outs$wilcox_out$coef <- as.numeric(all_cells$outs$wilcox_out$coef)

    all_cells$outs$wilcox_out_uni$p.val <- as.numeric(all_cells$outs$wilcox_out_uni$p.val)
    all_cells$outs$wilcox_out_uni$coef <- as.numeric(all_cells$outs$wilcox_out_uni$coef)


    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out$binned_pval <- factor(all_cells$outs$wilcox_out$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))

    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$binned_pval <- factor(all_cells$outs$wilcox_out_uni$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))



    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(sig_levels = dplyr::case_when(
        p.val < 0.01 ~ "Significant (p<0.01)",
        p.val < 0.05 ~ "Significant (p<0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$sig_levels <- factor(all_cells$outs$wilcox_out_uni$sig_levels, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))


    custom_fill_theme <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "grey"
    )

    custom_fill_theme_continuous <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "#f5f5f5"
    )
    library(reshape2)

    mat_multi <- reshape2::acast(all_cells$outs$wilcox_out, lineage_group ~ factor, value.var = "coef")
    d_multi_cluster <- dist(mat_multi / sqrt(rowSums(mat_multi * mat_multi)), method = "euclidean")
    multi_dendrogram_clusters <- hclust(d_multi_cluster)
    d_multi_factor <- dist(t(mat_multi) / sqrt(rowSums(t(mat_multi) * t(mat_multi))), method = "euclidean")
    multi_dendrogram_factor <- hclust(d_multi_factor)

    mat_uni <- reshape2::acast(all_cells$outs$wilcox_out_uni, lineage_group ~ factor, value.var = "coef")
    d_uni_cluster <- dist(mat_uni / sqrt(rowSums(mat_uni * mat_uni)), method = "euclidean")
    uni_dendrogram_clusters <- hclust(d_uni_cluster)
    d_uni_factor <- dist(t(mat_uni) / sqrt(rowSums(t(mat_uni) * t(mat_uni))), method = "euclidean")
    uni_dendrogram_factor <- hclust(d_uni_factor)

    uni_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    uni_factor_ordering <- unique(all_cells$outs$wilcox_out_uni$factor) # factor_names # uni_dendrogram_factor$labels[uni_dendrogram_factor$order]

    # multi_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    # multi_factor_ordering <- factor_names # multi_dendrogram_factor$labels[multi_dendrogram_factor$order]


    # all_cells$outs$wilcox_out$factor <- factor(all_cells$outs$wilcox_out$factor, levels = multi_factor_ordering)
    # all_cells$outs$wilcox_out$cellID_short <- factor(all_cells$outs$wilcox_out$cellID_short, levels = multi_cluster_ordering)

    all_cells$outs$wilcox_out_uni$factor <- factor(all_cells$outs$wilcox_out_uni$factor, levels = uni_factor_ordering)
    all_cells$outs$wilcox_out_uni$lineage_group <- factor(all_cells$outs$wilcox_out_uni$lineage_group, levels = uni_cluster_ordering)

    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4
    scale_factor <- 12
    p_uni_indiv <- ggplot(all_cells$outs$wilcox_out_uni) +
        geom_point(aes(x = lineage_group, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686") +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("Factor") +
        ggtitle("Univariate Linear Model (Corr sXb)") +
        coord_fixed() +
        guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
        theme(axis.text.x = element_text(
            size = 14 * 0.95,
            family = "sans", face = "bold",
            margin = margin(t = 0.8 * 14 / 4),
            angle = 90,
            vjust = 0.5,
            hjust = 1,
        ))

    # p_uni_comb <- (p_uni + theme(axis.text.x = element_blank(), axis.title.x = element_blank())) / (p_uni_2 + theme(plot.title = element_blank()))
    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4

    # pdf_and_png(p_uni, output_expr, paste0("linear_reg_uni_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)
    # pdf_and_png(p_uni_2, output_expr, paste0("linear_reg_uni_LOGFC_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)
    # pdf_and_png(p_uni_comb, output_expr, paste0("linear_reg_uni_COMBINED_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor * 2, scale = 3)
    pdf_and_png(p_uni_indiv, output_expr, paste0("linear_reg_uni_ALT_COMBINED_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    plot_list <- list()
    for (i in list("AllComp")) {
        all_cells$sub_wilcox_out <- all_cells$outs$wilcox_out_uni #|> dplyr::filter(lineage_group %in% i)

        sub_mat_uni <- reshape2::acast(all_cells$sub_wilcox_out, lineage_group ~ factor, value.var = "coef")
        sub_d_uni_cluster <- dist(sub_mat_uni / sqrt(rowSums(sub_mat_uni * sub_mat_uni)), method = "euclidean")
        sub_uni_dendrogram_clusters <- hclust(sub_d_uni_cluster)


        sub_uni_cluster_ordering <- sub_uni_dendrogram_clusters$labels[sub_uni_dendrogram_clusters$order]


        # all_cells$sub_wilcox_out$factor <- factor(all_cells$sub_wilcox_out$factor, levels = uni_factor_ordering)
        # all_cells$sub_wilcox_out$cellID_short <- factor(all_cells$sub_wilcox_out$cellID_short, levels = sub_uni_cluster_ordering)


        all_cells$sub_wilcox_out$sig_binned_value <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c(
            "NS", "Significant (p<0.05)", "Significant (p<0.01)"
        ))
        all_cells$sub_wilcox_out$binned_pval <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c("Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"))
        #                low = "#290AD8", mid = "grey", high = "#A50021", limits = c(-0.31, 0.31),

        plot_list[[i]] <- ggplot(all_cells$sub_wilcox_out) +
            geom_point(aes(x = lineage_group, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
            scale_fill_gradient2(
                low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
                oob = squish
            ) +
            scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
            scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
            theme_prism() +
            xlab("Cluster") +
            ylab("Factor") +
            ggtitle(i) +
            coord_fixed() +
            guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
            theme(
                axis.text.x = element_text(
                    size = 14 * 0.95,
                    family = "sans", face = "bold",
                    margin = margin(t = 0.8 * 14 / 4),
                    angle = 90,
                    vjust = 0.5,
                    hjust = 1,
                ),
                axis.title.x = element_blank(),
                plot.margin = unit(c(0, 1, 0, 0), "lines")
            )
    }

    for (i in unique(all_cells$outs$wilcox_out_uni$cluster)) {
        plot <- patchwork::wrap_plots(plot_list, ncol = length(list("AllComp")), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")
    }
    pdf_and_png(plot, output_expr, paste0("linear_reg_uni_ALT_COMBINED_FACETTED_GROUPS_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    env_list[[idx_name]] <- all_cells
}
```



# Diff Abundance - Lineage Group with all CD3+ Cells combined into one category

```{r}
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(lineage_group_combined_T = dplyr::case_when(
    lineage_group %in% c("CD8", "CD4") ~ "T",
    .default = lineage_group
))

counts <- table(baseline$public_id, baseline@meta.data[, "lineage_group_combined_T"])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl_all_var <- baseline@meta.data |> dplyr::distinct(public_id, .keep_all = T)


md_bl <- baseline@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::select(public_id, siteXbatch, dplyr::all_of(names(variable_info_structure)))

md_bl <- md_bl |> dplyr::distinct(public_id, .keep_all = T)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(lineage_group_combined_T)
cluster_md$cluster <- cluster_md$lineage_group_combined_T
cluster_md$lineage_group <- cluster_md$lineage_group_combined_T

counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")
types_list <- list(
    "LG-T Immune" = c("T", "M", "B", "Other", "Nk"),
    "LG-T All NoPlasma" = c("T", "M", "B", "Other", "Nk", "E"),
    "LG-T All" = c("T", "M", "B", "Other", "Nk", "P", "E")
)
file_name_list <- list(
    "LG-T Immune" = "LINEAGE_GROUP_MERGED_T_immune_only",
    "LG-T All NoPlasma" = "LINEAGE_GROUP_MERGED_TAll_NoPlasma",
    "LG-T All" = "LINEAGE_GROUP_MERGED_T_all_cells"
)

for (idx_name in names(types_list)) {
    all_cells <- new.env()

    all_cells$counts <- counts |> dplyr::filter(lineage_group %in% types_list[[idx_name]])
    all_cells$outs <- DiffAbun_Wrapper(all_cells$counts, variable_info_structure = variable_info_structure, DOMULTIVAR = F, method = "LINEAR", transform = "logit")


    # all_cells$outs <- quickLinearAsinAndPlots(all_cells$counts, variable_info_structure)
    write.csv(all_cells$outs$wilcox_out, file.path(output_expr, paste0("linear_reg_", file_name_list[[idx_name]], ".csv")))
    write.csv(all_cells$outs$wilcox_out_uni, file.path(output_expr, paste0("linear_reg_UNI_", file_name_list[[idx_name]], ".csv")))

    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::filter(variable != "siteXbatch")
    all_cells$cluster_md <- cluster_md
    all_cells$cluster_md$subcluster_V03072023 <- all_cells$cluster_md$cluster

    all_cells$outs$wilcox_out <- dplyr::left_join(all_cells$outs$wilcox_out, all_cells$cluster_md, by = "cluster")
    all_cells$outs$wilcox_out_uni <- dplyr::left_join(all_cells$outs$wilcox_out_uni, all_cells$cluster_md, by = "cluster")

    all_cells$outs$wilcox_out$p.val <- as.numeric(all_cells$outs$wilcox_out$p.val)
    all_cells$outs$wilcox_out$coef <- as.numeric(all_cells$outs$wilcox_out$coef)

    all_cells$outs$wilcox_out_uni$p.val <- as.numeric(all_cells$outs$wilcox_out_uni$p.val)
    all_cells$outs$wilcox_out_uni$coef <- as.numeric(all_cells$outs$wilcox_out_uni$coef)


    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out$binned_pval <- factor(all_cells$outs$wilcox_out$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))

    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$binned_pval <- factor(all_cells$outs$wilcox_out_uni$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))



    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(sig_levels = dplyr::case_when(
        p.val < 0.01 ~ "Significant (p<0.01)",
        p.val < 0.05 ~ "Significant (p<0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$sig_levels <- factor(all_cells$outs$wilcox_out_uni$sig_levels, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))


    custom_fill_theme <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "grey"
    )

    custom_fill_theme_continuous <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "#f5f5f5"
    )
    library(reshape2)

    mat_multi <- reshape2::acast(all_cells$outs$wilcox_out, lineage_group ~ factor, value.var = "coef")
    d_multi_cluster <- dist(mat_multi / sqrt(rowSums(mat_multi * mat_multi)), method = "euclidean")
    multi_dendrogram_clusters <- hclust(d_multi_cluster)
    d_multi_factor <- dist(t(mat_multi) / sqrt(rowSums(t(mat_multi) * t(mat_multi))), method = "euclidean")
    multi_dendrogram_factor <- hclust(d_multi_factor)

    mat_uni <- reshape2::acast(all_cells$outs$wilcox_out_uni, lineage_group ~ factor, value.var = "coef")
    d_uni_cluster <- dist(mat_uni / sqrt(rowSums(mat_uni * mat_uni)), method = "euclidean")
    uni_dendrogram_clusters <- hclust(d_uni_cluster)
    d_uni_factor <- dist(t(mat_uni) / sqrt(rowSums(t(mat_uni) * t(mat_uni))), method = "euclidean")
    uni_dendrogram_factor <- hclust(d_uni_factor)

    uni_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    uni_factor_ordering <- unique(all_cells$outs$wilcox_out_uni$factor) # factor_names # uni_dendrogram_factor$labels[uni_dendrogram_factor$order]

    # multi_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    # multi_factor_ordering <- factor_names # multi_dendrogram_factor$labels[multi_dendrogram_factor$order]


    # all_cells$outs$wilcox_out$factor <- factor(all_cells$outs$wilcox_out$factor, levels = multi_factor_ordering)
    # all_cells$outs$wilcox_out$cellID_short <- factor(all_cells$outs$wilcox_out$cellID_short, levels = multi_cluster_ordering)

    all_cells$outs$wilcox_out_uni$factor <- factor(all_cells$outs$wilcox_out_uni$factor, levels = uni_factor_ordering)
    all_cells$outs$wilcox_out_uni$lineage_group <- factor(all_cells$outs$wilcox_out_uni$lineage_group, levels = uni_cluster_ordering)

    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4
    scale_factor <- 12

    p_uni_indiv <- ggplot(all_cells$outs$wilcox_out_uni) +
        geom_point(aes(x = lineage_group, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686") +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("Factor") +
        ggtitle("Univariate Linear Model (Corr sXb)") +
        coord_fixed() +
        guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
        theme(axis.text.x = element_text(
            size = 14 * 0.95,
            family = "sans", face = "bold",
            margin = margin(t = 0.8 * 14 / 4),
            angle = 90,
            vjust = 0.5,
            hjust = 1,
        ))

    # p_uni_comb <- (p_uni + theme(axis.text.x = element_blank(), axis.title.x = element_blank())) / (p_uni_2 + theme(plot.title = element_blank()))
    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4

    pdf_and_png(p_uni_indiv, output_expr, paste0("linear_reg_uni_ALT_COMBINED_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    plot_list <- list()
    for (i in list("AllComp")) {
        all_cells$sub_wilcox_out <- all_cells$outs$wilcox_out_uni #|> dplyr::filter(lineage_group %in% i)

        sub_mat_uni <- reshape2::acast(all_cells$sub_wilcox_out, lineage_group ~ factor, value.var = "coef")
        sub_d_uni_cluster <- dist(sub_mat_uni / sqrt(rowSums(sub_mat_uni * sub_mat_uni)), method = "euclidean")
        sub_uni_dendrogram_clusters <- hclust(sub_d_uni_cluster)


        sub_uni_cluster_ordering <- sub_uni_dendrogram_clusters$labels[sub_uni_dendrogram_clusters$order]


        # all_cells$sub_wilcox_out$factor <- factor(all_cells$sub_wilcox_out$factor, levels = uni_factor_ordering)
        # all_cells$sub_wilcox_out$cellID_short <- factor(all_cells$sub_wilcox_out$cellID_short, levels = sub_uni_cluster_ordering)


        all_cells$sub_wilcox_out$sig_binned_value <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c(
            "NS", "Significant (p<0.05)", "Significant (p<0.01)"
        ))
        all_cells$sub_wilcox_out$binned_pval <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c("Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"))
        #                low = "#290AD8", mid = "grey", high = "#A50021", limits = c(-0.31, 0.31),

        plot_list[[i]] <- ggplot(all_cells$sub_wilcox_out) +
            geom_point(aes(x = lineage_group, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
            scale_fill_gradient2(
                low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
                oob = squish
            ) +
            scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
            scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
            theme_prism() +
            xlab("Cluster") +
            ylab("Factor") +
            ggtitle(i) +
            coord_fixed() +
            guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
            theme(
                axis.text.x = element_text(
                    size = 14 * 0.95,
                    family = "sans", face = "bold",
                    margin = margin(t = 0.8 * 14 / 4),
                    angle = 90,
                    vjust = 0.5,
                    hjust = 1,
                ),
                axis.title.x = element_blank(),
                plot.margin = unit(c(0, 1, 0, 0), "lines")
            )
    }

    for (i in unique(all_cells$outs$wilcox_out_uni$cluster)) {
        plot <- patchwork::wrap_plots(plot_list, ncol = length(list("AllComp")), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")
    }
    pdf_and_png(plot, output_expr, paste0("linear_reg_uni_ALT_COMBINED_FACETTED_GROUPS_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    env_list[[idx_name]] <- all_cells
}
```

```{r}
types_list <- list(
    "CD3+ Cells" = c("CD8", "CD4"),
    "Myeloid Cells" = "M",
    "B Cells" = "B",
    "CD8+" = "CD8",
    "CD4+" = "CD4",
    "Nk" = "Nk",
    "Immune" = c("CD8", "CD4", "M", "B", "Other", "Nk"),
    "All_NoPlasma" = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    "All" = c("CD8", "CD4", "M", "B", "Nk", "Other", "E", "P"),
    "LG CD3+ Cells" = c("CD8", "CD4"),
    "LG Immune" = c("CD8", "CD4", "M", "B", "Other", "Nk"),
    "LG All NoPlasma" = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    "LG All" = c("CD8", "CD4", "M", "B", "Other", "Nk", "P", "E"),
    "LG-T Immune" = c("T", "M", "B", "Other", "Nk"),
    "LG-T All NoPlasma" = c("T", "M", "B", "Other", "Nk", "E"),
    "LG-T All" = c("T", "M", "B", "Other", "Nk", "P", "E")
)
```

# BM PC PCT (Flow) vs Plasma Fraction

```{r}
# alternative

idx_name <- "LG-T All"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["logit_pc_proportion"]] |> dplyr::filter(cluster == "P")
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("logit_pc_proportion"))
model_uni <- all_cells$outs$model_uni[["lmfit"]][["P"]][["logit_pc_proportion"]]
ebayes_uni <- all_cells$outs$model_uni[["ebayes"]][["P"]][["logit_pc_proportion"]]



grouping_var <- "logit_pc_proportion"

tmp_out <- wilcox_out_uni #|> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out <- tmp_out |> dplyr::filter(cluster == "P")
p_val <- as.numeric(tmp_out$p.val)
adj_r_squared <- as.numeric(tmp_out$adj_r_sqrd)

p_val_clean <- pixiedust::pval_string(p_val, format = c("scientific"), digits = 3)
adj_r_squared <- adj_r_squared |> round(digits = 2)

fit_coef <- model_uni$coefficients
x_est <- seq(from = min(tmp$logit_pc_proportion), to = max(tmp$logit_pc_proportion), length.out = 100)

avg_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.25, 100), rep(0.25, 100), rep(0.25, 100)), nrow = 100, ncol = 5)
em_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(0.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
mayo_fit <- matrix(data = c(rep(1, 100), x_est, rep(1.0, 100), rep(0.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
mssm_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(1.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
washu_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(0.0, 100), rep(1.0, 100)), nrow = 100, ncol = 5)

fitted_line <- avg_fit %*% t(fit_coef)

em_fitted_line <- em_fit %*% t(fit_coef)
mayo_fitted_line <- mayo_fit %*% t(fit_coef)
mssm_fitted_line <- mssm_fit %*% t(fit_coef)
washu_fitted_line <- washu_fit %*% t(fit_coef)

all_fits <- data.frame(logit_pc_proportion = x_est, avg = fitted_line, EMORY = em_fitted_line, MAYO = mayo_fitted_line, MSSM = mssm_fitted_line, WUSTL = washu_fitted_line) |> tidyr::pivot_longer(-logit_pc_proportion, names_to = "Study_Site", values_to = "logit_prop")

EMORY_palettes <- paletteer_d("ggsci::blue_material")
WASHU_palettes <- paletteer_d("ggsci::purple_material")
MSSM_palettes <- paletteer_d("ggsci::green_material")
MAYO_palettes <- paletteer_d("ggsci::deep_orange_material")

colors_site <- c("EMORY" = EMORY_palettes[[8]], "WUSTL" = WASHU_palettes[[9]], "MSSM" = MSSM_palettes[[7]], "MAYO" = MAYO_palettes[[7]])
# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)

p_t <- ggplot(tmp, aes(x = !!sym(grouping_var), y = logit_prop)) +
    geom_point(outlier.shape = NA, alpha = 0.8, width = 0.4, aes(fill = Study_Site, color = Study_Site)) +
    geom_abline(linetype = "dashed", color = "grey") +
    geom_line(data = all_fits |> dplyr::filter(Study_Site != "avg"), aes(color = Study_Site, fill = Study_Site), linetype = "dashed", alpha = 1) +
    geom_line(data = all_fits |> dplyr::filter(Study_Site == "avg"), linewidth = 2) +
    # geom_ribbon(data = all_fits |> dplyr::filter(line_grouping=="avg"), aes(ymin=lower, ymax=upper), linetype=2, alpha=0.1) +
    # geom_smooth(method = "lm", formula = y ~ x, aes(color=Study_Site)) +
    scale_color_manual(values = colors_site) +
    theme_prism() +
    coord_cartesian(clip = "off") +
    ggtitle("Plasma Cell Fraction") +
    xlab("Logit-Transformed Plasma Cell Proportion\n(Flow-Cytometry)") +
    ylab("Logit-Transformed Plasma Cell Proportion\n(scRNA-seq)") +
    theme(plot.title = element_text(size = 14, face = "bold")) +
    annotate("text", x = -3.5, y = 1.8, label = sprintf("R\U00B2: %s\nP.Val:%s", adj_r_squared, p_val_clean), hjust = 0, size = 5.5)

pdf_and_png(p_t, output_expr, "logit_transform_plasma_cell_pct", pdfWidth = 2.5, pdfHeight = 2.5, scale = 3)
```


```{r}
# alternative

idx_name <- "LG-T All"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["binned_PC_PCT"]]
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("binned_PC_PCT"))

grouping_var <- "binned_PC_PCT"

per_clust_max <- tmp |>
    dplyr::group_by(cluster, binned_PC_PCT) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.1, y.position.alt = upper_max + 0.1 * IQR_max)

tmp_out <- wilcox_out_uni |> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out$group1 <- "<20"
tmp_out$group2 <- ">=20"

tmp_t <- tmp |> dplyr::filter(cluster == "P")

# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)



tmp_out_t <- tmp_out |> dplyr::filter(cluster == "P")

fill_theme_binary_PCPCT <- c(
    "<20" = fill_NP,
    ">=20" = fill_RP
)

color_theme_binary_PCPCT <- c(
    "<20" = color_NP,
    ">=20" = color_RP
)


p_t <- ggplot(tmp_t) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_t, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, 1)) +
    scale_fill_manual(values = fill_theme_binary_PCPCT) +
    scale_color_manual(values = color_theme_binary_PCPCT) +
    ylab("Plasma Cell Proportion (scRNA-seq)") +
    xlab("Binned Plasma Cell Percentage\n(Flow-Cytometry)") +
    NoLegend() +
    coord_cartesian(clip = "off") +
    ggtitle("Plasma Cells") +
    theme(plot.title = element_text(size = 14, face = "bold"))


pdf_and_png(p_t, output_expr, "binned_plasma_cell_percentage", pdfWidth = 1.8, pdfHeight = 2.5, scale = 3)
```


# Supplemental Figure, Plasma PCT - All cells (including plasma)
```{r}
# Fig 3, IMWG_AND_COMPONENTS

# sapply(test$factor, FUN = function(x) {
#     nchar(x)
# })

idx_name <- "All"
all_cells <- env_list[[idx_name]]

wilcox_out_uni <- all_cells$outs$wilcox_out_uni #|> dplyr::filter(variable %in% c("mixed_nsd2_1q21_class", "davies_based_risk", "nsd2_call", "seqwgs_cp_1q21_call"))

# regretting past design decisions but its okay
vis_df <- data.frame(
    factor = c(
        "logit_pc_proportion",
        "binned_PC_PCT>=20"
    ),
    variable_names = c(
        "Flow-Cyt PC Percentage (Continuous)",
        "Flow-Cyt PC Percentage (Binned)"
    )
)


lineage_group_names_df <- data.frame(
    lineage_group = c("CD8", "CD4", "M", "B", "Other", "Nk", "E", "P"),
    lineage_group_name = c("CD8+ T Cell", "CD4+ T Cell", "Myeloid", "B Cell", "Other", "Nk Cell", "Erythroid", "Plasma Cell")
)

cluster_ordering <- readRDS("data/cellID_short_figure_ordering.RDS")

rownames(lineage_group_names_df) <- lineage_group_names_df$lineage_group
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, vis_df, by = "factor")
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, lineage_group_names_df, by = "lineage_group")


wilcox_out_uni$variable_names <- factor(wilcox_out_uni$variable_names, levels = rev(c(
    "Flow-Cyt PC Percentage (Continuous)",
    "Flow-Cyt PC Percentage (Binned)"
)))

heights <- length(unique(wilcox_out_uni$factor))
char_len_uni <- max(nchar(levels(wilcox_out_uni$variable_names)))
widths <- length(unique(wilcox_out_uni$cluster)) + char_len_uni / 4
scale_factor <- 12


# middle_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
#     "NSD2 - Neither",
#     "1q21 - Neither",
#     "NSD2+1q21 - Neither"
# ))


bottom_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "Flow-Cyt PC Percentage (Continuous)",
    "Flow-Cyt PC Percentage (Binned)"
))

wilcox_out_uni$cellID_short <- factor(wilcox_out_uni$cluster, levels = cluster_ordering)


bottom_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]

    sub_wilcox_out <- bottom_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    bottom_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        ggtitle(lineage_group_name) +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_text(
                size = 14 * 0.95,
                family = "sans", face = "bold",
                margin = margin(t = 0.8 * 14 / 4),
                angle = 90,
                vjust = 0.5,
                hjust = 1,
            ),
            axis.title.x = element_blank(),
            plot.margin = unit(c(0.25, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        bottom_plot_list[[i]] <- bottom_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    bottom_plot_list[[i]] <- bottom_plot_list[[i]] + coord_fixed()
}

plot <- patchwork::wrap_plots(c(bottom_plot_list), nrow = 1, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")

pdf_and_png(plot, output_expr, paste0("PC_PCT_All_Cells"), pdfWidth = W_FULL * 3.5, pdfHeight = 2, scale = 3)
```

# Supplemental Figure Plasma PCT - Immune cells (including plasma)
Summary figure not included

```{r}
idx_name <- "Immune"
all_cells <- env_list[[idx_name]]

wilcox_out_uni <- all_cells$outs$wilcox_out_uni #|> dplyr::filter(variable %in% c("mixed_nsd2_1q21_class", "davies_based_risk", "nsd2_call", "seqwgs_cp_1q21_call"))

# regretting past design decisions but its okay
vis_df <- data.frame(
    factor = c(
        "logit_pc_proportion",
        "binned_PC_PCT>=20"
    ),
    variable_names = c(
        "Flow-Cyt PC Percentage (Continuous)",
        "Flow-Cyt PC Percentage (Binned)"
    )
)


lineage_group_names_df <- data.frame(
    lineage_group = c("CD8", "CD4", "M", "B", "Other", "Nk", "E", "P"),
    lineage_group_name = c("CD8+ T Cell", "CD4+ T Cell", "Myeloid", "B Cell", "Other", "Nk Cell", "Erythroid", "Plasma Cell")
)

cluster_ordering <- readRDS("data/cellID_short_figure_ordering.RDS")

rownames(lineage_group_names_df) <- lineage_group_names_df$lineage_group
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, vis_df, by = "factor")
wilcox_out_uni <- dplyr::left_join(wilcox_out_uni, lineage_group_names_df, by = "lineage_group")


wilcox_out_uni$variable_names <- factor(wilcox_out_uni$variable_names, levels = rev(c(
    "Flow-Cyt PC Percentage (Continuous)",
    "Flow-Cyt PC Percentage (Binned)"
)))

heights <- length(unique(wilcox_out_uni$factor))
char_len_uni <- max(nchar(levels(wilcox_out_uni$variable_names)))
widths <- length(unique(wilcox_out_uni$cluster)) + char_len_uni / 4
scale_factor <- 12


# middle_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
#     "NSD2 - Neither",
#     "1q21 - Neither",
#     "NSD2+1q21 - Neither"
# ))


bottom_plot_info <- wilcox_out_uni |> dplyr::filter(variable_names %in% c(
    "Flow-Cyt PC Percentage (Continuous)",
    "Flow-Cyt PC Percentage (Binned)"
))

wilcox_out_uni$cellID_short <- factor(wilcox_out_uni$cluster, levels = cluster_ordering)


bottom_plot_list <- list()
for (i in types_list[[idx_name]]) {
    lineage_group_name <- lineage_group_names_df[i, "lineage_group_name"]

    sub_wilcox_out <- bottom_plot_info |> dplyr::filter(lineage_group %in% i)

    sub_wilcox_out$sig_binned_value <- factor(sub_wilcox_out$binned_pval, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))

    bottom_plot_list[[i]] <- ggplot(sub_wilcox_out) +
        geom_point(aes(x = cellID_short, y = variable_names, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        ggtitle(lineage_group_name) +
        xlab("Cluster") +
        ylab("") +
        guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
        theme(
            axis.text.x = element_text(
                size = 14 * 0.95,
                family = "sans", face = "bold",
                margin = margin(t = 0.8 * 14 / 4),
                angle = 90,
                vjust = 0.5,
                hjust = 1,
            ),
            axis.title.x = element_blank(),
            plot.margin = unit(c(0.25, 1, 0, 0), "lines")
        )
    if (i != types_list[[idx_name]][[1]]) {
        bottom_plot_list[[i]] <- bottom_plot_list[[i]] + theme(
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks.length.y = unit(0, "lines")
        )
    }

    bottom_plot_list[[i]] <- bottom_plot_list[[i]] + coord_fixed()
}

plot <- patchwork::wrap_plots(c(bottom_plot_list), nrow = 1, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")

pdf_and_png(plot, output_expr, paste0("PC_PCT_Immune_Cells"), pdfWidth = W_FULL * 3, pdfHeight = 2, scale = 3)
```

### Scatter plots for specific immune clusters

#### CD8 Tem
```{r}
idx_name <- "Immune"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["logit_pc_proportion"]] |> dplyr::filter(cluster == "NkT.3.0")
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("logit_pc_proportion"))
model_uni <- all_cells$outs$model_uni[["lmfit"]][["NkT.3.0"]][["logit_pc_proportion"]]
ebayes_uni <- all_cells$outs$model_uni[["ebayes"]][["NkT.3.0"]][["logit_pc_proportion"]]



grouping_var <- "logit_pc_proportion"

tmp_out <- wilcox_out_uni #|> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out <- tmp_out |> dplyr::filter(cluster == "NkT.3.0")
p_val <- as.numeric(tmp_out$p.val)
adj_r_squared <- as.numeric(tmp_out$adj_r_sqrd)

p_val_clean <- pixiedust::pval_string(p_val, format = c("scientific"), digits = 3)
adj_r_squared <- adj_r_squared |> round(digits = 2)

fit_coef <- model_uni$coefficients
x_est <- seq(from = min(tmp$logit_pc_proportion), to = max(tmp$logit_pc_proportion), length.out = 100)

avg_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.25, 100), rep(0.25, 100), rep(0.25, 100)), nrow = 100, ncol = 5)
em_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(0.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
mayo_fit <- matrix(data = c(rep(1, 100), x_est, rep(1.0, 100), rep(0.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
mssm_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(1.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
washu_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(0.0, 100), rep(1.0, 100)), nrow = 100, ncol = 5)

fitted_line <- avg_fit %*% t(fit_coef)

em_fitted_line <- em_fit %*% t(fit_coef)
mayo_fitted_line <- mayo_fit %*% t(fit_coef)
mssm_fitted_line <- mssm_fit %*% t(fit_coef)
washu_fitted_line <- washu_fit %*% t(fit_coef)

all_fits <- data.frame(logit_pc_proportion = x_est, avg = fitted_line, EMORY = em_fitted_line, MAYO = mayo_fitted_line, MSSM = mssm_fitted_line, WUSTL = washu_fitted_line) |> tidyr::pivot_longer(-logit_pc_proportion, names_to = "Study_Site", values_to = "logit_prop")

EMORY_palettes <- paletteer_d("ggsci::blue_material")
WASHU_palettes <- paletteer_d("ggsci::purple_material")
MSSM_palettes <- paletteer_d("ggsci::green_material")
MAYO_palettes <- paletteer_d("ggsci::deep_orange_material")

colors_site <- c("EMORY" = EMORY_palettes[[8]], "WUSTL" = WASHU_palettes[[9]], "MSSM" = MSSM_palettes[[7]], "MAYO" = MAYO_palettes[[7]])
# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)

p_t <- ggplot(tmp, aes(x = !!sym(grouping_var), y = logit_prop)) +
    geom_point(outlier.shape = NA, alpha = 0.8, width = 0.4, aes(fill = Study_Site, color = Study_Site)) +
    geom_abline(linetype = "dashed", color = "grey") +
    geom_line(data = all_fits |> dplyr::filter(Study_Site != "avg"), aes(color = Study_Site, fill = Study_Site), linetype = "dashed", alpha = 1) +
    geom_line(data = all_fits |> dplyr::filter(Study_Site == "avg"), linewidth = 2) +
    # geom_ribbon(data = all_fits |> dplyr::filter(line_grouping=="avg"), aes(ymin=lower, ymax=upper), linetype=2, alpha=0.1) +
    # geom_smooth(method = "lm", formula = y ~ x, aes(color=Study_Site)) +
    scale_color_manual(values = colors_site) +
    theme_prism() +
    coord_cartesian(clip = "off") +
    ggtitle("CD8 T Effector Memory") +
    xlab("Logit-Transformed Plasma Cell Proportion\n(Flow-Cytometry)") +
    ylab("Logit-Transformed Proportion of Immune Cells") +
    theme(plot.title = element_text(size = 14, face = "bold")) +
    annotate("text", x = -3.25, y = -1, label = sprintf("R\U00B2: %s\nP.Val:%s", adj_r_squared, p_val_clean), hjust = 0, size = 5.5)

pdf_and_png(p_t, output_expr, paste0("CD8_Tem_Immune_Fraction"), pdfWidth = 2.5, pdfHeight = 2.5, scale = 3)
```


```{r}
# alternative

idx_name <- "Immune"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["binned_PC_PCT"]]
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("binned_PC_PCT"))

grouping_var <- "binned_PC_PCT"

per_clust_max <- tmp |>
    dplyr::group_by(cluster, binned_PC_PCT) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.1, y.position.alt = upper_max + 0.1 * IQR_max)

tmp_out <- wilcox_out_uni |> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out$group1 <- "<20"
tmp_out$group2 <- ">=20"

tmp_t <- tmp |> dplyr::filter(cluster == "NkT.3.0")

# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)



tmp_out_t <- tmp_out |> dplyr::filter(cluster == "NkT.3.0")

fill_theme_binary_PCPCT <- c(
    "<20" = fill_NP,
    ">=20" = fill_RP
)

color_theme_binary_PCPCT <- c(
    "<20" = color_NP,
    ">=20" = color_RP
)


p_t <- ggplot(tmp_t) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_t, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, 0.3)) +
    scale_fill_manual(values = fill_theme_binary_PCPCT) +
    scale_color_manual(values = color_theme_binary_PCPCT) +
    ylab("Proportion of Immune Cells") +
    xlab("Binned Plasma Cell Percentage\n(Flow-Cytometry)") +
    NoLegend() +
    coord_cartesian(clip = "off") +
    ggtitle("CD8+ T Effector Memory") +
    theme(plot.title = element_text(size = 14, face = "bold"))


pdf_and_png(p_t, output_expr, "binned_CD8_Tem_Immune_percentage", pdfWidth = 1.8, pdfHeight = 2.5, scale = 3)
```



#### Nk BM Res
```{r}
idx_name <- "Immune"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["logit_pc_proportion"]] |> dplyr::filter(cluster == "NkT.9.1")
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("logit_pc_proportion"))
model_uni <- all_cells$outs$model_uni[["lmfit"]][["NkT.9.1"]][["logit_pc_proportion"]]
ebayes_uni <- all_cells$outs$model_uni[["ebayes"]][["NkT.9.1"]][["logit_pc_proportion"]]



grouping_var <- "logit_pc_proportion"

tmp_out <- wilcox_out_uni #|> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out <- tmp_out |> dplyr::filter(cluster == "NkT.9.1")
p_val <- as.numeric(tmp_out$p.val)
adj_r_squared <- as.numeric(tmp_out$adj_r_sqrd)

p_val_clean <- pixiedust::pval_string(p_val, format = c("scientific"), digits = 3)
adj_r_squared <- adj_r_squared |> round(digits = 2)

fit_coef <- model_uni$coefficients
x_est <- seq(from = min(tmp$logit_pc_proportion), to = max(tmp$logit_pc_proportion), length.out = 100)

avg_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.25, 100), rep(0.25, 100), rep(0.25, 100)), nrow = 100, ncol = 5)
em_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(0.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
mayo_fit <- matrix(data = c(rep(1, 100), x_est, rep(1.0, 100), rep(0.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
mssm_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(1.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
washu_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(0.0, 100), rep(1.0, 100)), nrow = 100, ncol = 5)

fitted_line <- avg_fit %*% t(fit_coef)

em_fitted_line <- em_fit %*% t(fit_coef)
mayo_fitted_line <- mayo_fit %*% t(fit_coef)
mssm_fitted_line <- mssm_fit %*% t(fit_coef)
washu_fitted_line <- washu_fit %*% t(fit_coef)

all_fits <- data.frame(logit_pc_proportion = x_est, avg = fitted_line, EMORY = em_fitted_line, MAYO = mayo_fitted_line, MSSM = mssm_fitted_line, WUSTL = washu_fitted_line) |> tidyr::pivot_longer(-logit_pc_proportion, names_to = "Study_Site", values_to = "logit_prop")

EMORY_palettes <- paletteer_d("ggsci::blue_material")
WASHU_palettes <- paletteer_d("ggsci::purple_material")
MSSM_palettes <- paletteer_d("ggsci::green_material")
MAYO_palettes <- paletteer_d("ggsci::deep_orange_material")

colors_site <- c("EMORY" = EMORY_palettes[[8]], "WUSTL" = WASHU_palettes[[9]], "MSSM" = MSSM_palettes[[7]], "MAYO" = MAYO_palettes[[7]])
# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)

p_t <- ggplot(tmp, aes(x = !!sym(grouping_var), y = logit_prop)) +
    geom_point(outlier.shape = NA, alpha = 0.8, width = 0.4, aes(fill = Study_Site, color = Study_Site)) +
    geom_abline(linetype = "dashed", color = "grey") +
    geom_line(data = all_fits |> dplyr::filter(Study_Site != "avg"), aes(color = Study_Site, fill = Study_Site), linetype = "dashed", alpha = 1) +
    geom_line(data = all_fits |> dplyr::filter(Study_Site == "avg"), linewidth = 2) +
    # geom_ribbon(data = all_fits |> dplyr::filter(line_grouping=="avg"), aes(ymin=lower, ymax=upper), linetype=2, alpha=0.1) +
    # geom_smooth(method = "lm", formula = y ~ x) +
    scale_color_manual(values = colors_site) +
    theme_prism() +
    coord_cartesian(clip = "off") +
    ggtitle("Nk Resident Cells") +
    xlab("Logit-Transformed Plasma Cell Proportion\n(Flow-Cytometry)") +
    ylab("Logit-Transformed Proportion of Immune Cells") +
    theme(plot.title = element_text(size = 14, face = "bold")) +
    annotate("text", x = -3.25, y = -1, label = sprintf("R\U00B2: %s\nP.Val:%s", adj_r_squared, p_val_clean), hjust = 0, size = 5.5)

pdf_and_png(p_t, output_expr, paste0("NkRes_Immune_Fraction"), pdfWidth = 2.5, pdfHeight = 2.5, scale = 3)
```


```{r}
# alternative

idx_name <- "Immune"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["binned_PC_PCT"]]
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("binned_PC_PCT"))

grouping_var <- "binned_PC_PCT"

per_clust_max <- tmp |>
    dplyr::group_by(cluster, binned_PC_PCT) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.1, y.position.alt = upper_max + 0.1 * IQR_max)

tmp_out <- wilcox_out_uni |> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out$group1 <- "<20"
tmp_out$group2 <- ">=20"

tmp_t <- tmp |> dplyr::filter(cluster == "NkT.9.1")

# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)



tmp_out_t <- tmp_out |> dplyr::filter(cluster == "NkT.9.1")

fill_theme_binary_PCPCT <- c(
    "<20" = fill_NP,
    ">=20" = fill_RP
)

color_theme_binary_PCPCT <- c(
    "<20" = color_NP,
    ">=20" = color_RP
)


p_t <- ggplot(tmp_t) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_t, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, 0.25)) +
    scale_fill_manual(values = fill_theme_binary_PCPCT) +
    scale_color_manual(values = color_theme_binary_PCPCT) +
    ylab("Proportion of Immune Cells") +
    xlab("Binned Plasma Cell Percentage\n(Flow-Cytometry)") +
    NoLegend() +
    coord_cartesian(clip = "off") +
    ggtitle("Nk Resident Cells") +
    theme(plot.title = element_text(size = 14, face = "bold"))


pdf_and_png(p_t, output_expr, "binned_NkRes_Tem_Immune_percentage", pdfWidth = 1.8, pdfHeight = 2.5, scale = 3)
```

#### Fibro
```{r}
idx_name <- "Immune"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["logit_pc_proportion"]] |> dplyr::filter(cluster == "Full.23")
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("logit_pc_proportion"))
model_uni <- all_cells$outs$model_uni[["lmfit"]][["Full.23"]][["logit_pc_proportion"]]
ebayes_uni <- all_cells$outs$model_uni[["ebayes"]][["Full.23"]][["logit_pc_proportion"]]



grouping_var <- "logit_pc_proportion"

tmp_out <- wilcox_out_uni #|> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out <- tmp_out |> dplyr::filter(cluster == "Full.23")
p_val <- as.numeric(tmp_out$p.val)
adj_r_squared <- as.numeric(tmp_out$adj_r_sqrd)

p_val_clean <- pixiedust::pval_string(p_val, format = c("scientific"), digits = 3)
adj_r_squared <- adj_r_squared |> round(digits = 2)

fit_coef <- model_uni$coefficients
x_est <- seq(from = min(tmp$logit_pc_proportion), to = max(tmp$logit_pc_proportion), length.out = 100)

avg_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.25, 100), rep(0.25, 100), rep(0.25, 100)), nrow = 100, ncol = 5)
em_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(0.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
mayo_fit <- matrix(data = c(rep(1, 100), x_est, rep(1.0, 100), rep(0.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
mssm_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(1.0, 100), rep(0.0, 100)), nrow = 100, ncol = 5)
washu_fit <- matrix(data = c(rep(1, 100), x_est, rep(0.0, 100), rep(0.0, 100), rep(1.0, 100)), nrow = 100, ncol = 5)

fitted_line <- avg_fit %*% t(fit_coef)

em_fitted_line <- em_fit %*% t(fit_coef)
mayo_fitted_line <- mayo_fit %*% t(fit_coef)
mssm_fitted_line <- mssm_fit %*% t(fit_coef)
washu_fitted_line <- washu_fit %*% t(fit_coef)

all_fits <- data.frame(logit_pc_proportion = x_est, avg = fitted_line, EMORY = em_fitted_line, MAYO = mayo_fitted_line, MSSM = mssm_fitted_line, WUSTL = washu_fitted_line) |> tidyr::pivot_longer(-logit_pc_proportion, names_to = "Study_Site", values_to = "logit_prop")

EMORY_palettes <- paletteer_d("ggsci::blue_material")
WASHU_palettes <- paletteer_d("ggsci::purple_material")
MSSM_palettes <- paletteer_d("ggsci::green_material")
MAYO_palettes <- paletteer_d("ggsci::deep_orange_material")

colors_site <- c("EMORY" = EMORY_palettes[[8]], "WUSTL" = WASHU_palettes[[9]], "MSSM" = MSSM_palettes[[7]], "MAYO" = MAYO_palettes[[7]])
# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)

p_t <- ggplot(tmp, aes(x = !!sym(grouping_var), y = logit_prop)) +
    geom_point(outlier.shape = NA, alpha = 0.8, width = 0.4, aes(fill = Study_Site, color = Study_Site)) +
    geom_abline(linetype = "dashed", color = "grey") +
    geom_line(data = all_fits |> dplyr::filter(Study_Site != "avg"), aes(color = Study_Site, fill = Study_Site), linetype = "dashed", alpha = 1) +
    geom_line(data = all_fits |> dplyr::filter(Study_Site == "avg"), linewidth = 2) +
    # geom_ribbon(data = all_fits |> dplyr::filter(line_grouping=="avg"), aes(ymin=lower, ymax=upper), linetype=2, alpha=0.1) +
    # geom_smooth(method = "lm", formula = y ~ x) +
    scale_color_manual(values = colors_site) +
    theme_prism() +
    coord_cartesian(clip = "off") +
    ggtitle("Fibroblasts") +
    xlab("Logit-Transformed Plasma Cell Proportion\n(Flow-Cytometry)") +
    ylab("Logit-Transformed Proportion of Immune Cells") +
    theme(plot.title = element_text(size = 14, face = "bold")) +
    annotate("text", x = -3.25, y = -1, label = sprintf("R\U00B2: %s\nP.Val:%s", adj_r_squared, p_val_clean), hjust = 0, size = 5.5)

pdf_and_png(p_t, output_expr, paste0("Fibroblast_Immune_Fraction"), pdfWidth = 2.5, pdfHeight = 2.5, scale = 3)
```


```{r}
# alternative

idx_name <- "Immune"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["binned_PC_PCT"]]
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("binned_PC_PCT"))

grouping_var <- "binned_PC_PCT"

per_clust_max <- tmp |>
    dplyr::group_by(cluster, binned_PC_PCT) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.1, y.position.alt = upper_max + 0.1 * IQR_max)

tmp_out <- wilcox_out_uni |> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out$group1 <- "<20"
tmp_out$group2 <- ">=20"

tmp_t <- tmp |> dplyr::filter(cluster == "Full.23")

# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)



tmp_out_t <- tmp_out |> dplyr::filter(cluster == "Full.23")

fill_theme_binary_PCPCT <- c(
    "<20" = fill_NP,
    ">=20" = fill_RP
)

color_theme_binary_PCPCT <- c(
    "<20" = color_NP,
    ">=20" = color_RP
)


p_t <- ggplot(tmp_t) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_t, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, 0.025)) +
    scale_fill_manual(values = fill_theme_binary_PCPCT) +
    scale_color_manual(values = color_theme_binary_PCPCT) +
    ylab("Proportion of Immune Cells") +
    xlab("Binned Plasma Cell Percentage\n(Flow-Cytometry)") +
    NoLegend() +
    coord_cartesian(clip = "off") +
    ggtitle("Fibroblasts Cells") +
    theme(plot.title = element_text(size = 14, face = "bold"))


pdf_and_png(p_t, output_expr, "binned_Fibroblasts_Immune_percentage", pdfWidth = 1.8, pdfHeight = 2.5, scale = 3)
```


# Supplemental - BM PC PCT

```{r}
all_cells <- env_list[["All_NoPlasma"]]

tmp_bl <- baseline@meta.data |> dplyr::distinct(public_id, D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH, D_IM_CD138_PC_PERCENT)

library(ggstatsplot)

tmp_counts_with_props <- all_cells$outs$counts_with_props |> dplyr::left_join(tmp_bl, by = "public_id")
tmp_counts_with_props <- tmp_counts_with_props |> dplyr::filter(cluster == "NkT.3.0")

p <- ggplot(tmp_counts_with_props, aes(x = car::logit(D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH / 100), y = logit_prop), alpha = 0.75) +
    geom_point() +
    geom_abline(linetype = "dashed", color = "grey") +
    geom_smooth(method = "lm", formula = y ~ x) +
    theme_prism() +
    xlab("") +
    ylab("Asin-Sqrt CD8 T Effector (Fraction of Immune Cells)") #+
coord_cartesian(xlim = c(-3, 0), ylim = c(-3, 3))


tmp_counts_with_props <- all_cells$outs$counts_with_props |> dplyr::left_join(tmp_bl, by = "public_id")
tmp_counts_with_props <- tmp_counts_with_props |> dplyr::filter(cluster == "Full.23")

p2 <- ggplot(tmp_counts_with_props, aes(x = car::logit(D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH / 100), y = logit_prop), alpha = 0.75) +
    geom_point() +
    geom_abline(linetype = "dashed", color = "grey") +
    geom_smooth(method = "lm", formula = y ~ x) +
    theme_prism() +
    xlab("Asin-Sqrt BM Plasma Cell Proportion (Flow Cytometry)") +
    ylab("Asin-Sqrt Fibroblast (Fraction of Immune Cells)") +
    coord_cartesian(xlim = c(-0.02, 1.25), ylim = c(-0.02, 0.5))


tmp_counts_with_props <- all_cells$outs$counts_with_props |> dplyr::left_join(tmp_bl, by = "public_id")
tmp_counts_with_props <- tmp_counts_with_props |> dplyr::filter(cluster == "BEry.11")

p3 <- ggplot(tmp_counts_with_props, aes(x = car::logit(D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH / 100), y = logit_prop), alpha = 0.75) +
    geom_point() +
    geom_abline(linetype = "dashed", color = "grey") +
    geom_smooth(method = "lm", formula = y ~ x) +
    theme_prism() +
    xlab("") +
    ylab("Asin-Sqrt pDC Proportion (Fraction of Immune Cells)") +
    coord_cartesian(xlim = c(-0.02, 1.25), ylim = c(-0.02, 0.5))

tmp_counts_with_props <- all_cells$outs$counts_with_props |> dplyr::left_join(tmp_bl, by = "public_id")
tmp_counts_with_props <- tmp_counts_with_props |> dplyr::filter(cluster == "Myeloid.15")

p4 <- ggplot(tmp_counts_with_props, aes(x = car::logit(D_IM_FLOWCYT_PCT_PC_IN_BM_HIGH / 100), y = logit_prop), alpha = 0.75) +
    geom_point() +
    geom_abline(linetype = "dashed", color = "grey") +
    geom_smooth(method = "lm", formula = y ~ x) +
    theme_prism() +
    xlab("") +
    ylab("Asin-Sqrt M2 Macro Proportion (Fraction of Immune Cells)") +
    coord_cartesian(xlim = c(-0.02, 1.25), ylim = c(-0.02, 0.5))


combined <- patchwork::wrap_plots(list(p, p2, p3, p4), ncol = 4)
pdf_and_png(combined, output_expr, "immune_fractions_continuous_withm2", pdfWidth = 5, pdfHeight = 2.5, scale = 3)
```

# Supplemental - BM PC PCT and Exhaustion

```{r}
cd8 <- merged |> subset(lineage_group == "CD8")
tem <- merged |> subset(subcluster_V03072023 == "NkT.3.0")


fill_theme_binary_PCPCT <- c(
    "<20" = fill_NP,
    ">=20" = fill_RP
)

color_theme_binary_PCPCT <- c(
    "<20" = color_NP,
    ">=20" = color_RP
)



FEATURES <- c("TIGIT", "TOX")

tmp2_cd8 <- AverageExpression(cd8, assays = "RNA", group.by = "public_id", features = FEATURES)$RNA |>
    t() |>
    as.data.frame()

tmp2_tem <- AverageExpression(tem, assays = "RNA", group.by = "public_id", features = FEATURES)$RNA |>
    t() |>
    as.data.frame()

rownames(tmp2_tem) <- gsub("-", "_", rownames(tmp2_tem))
rownames(tmp2_cd8) <- gsub("-", "_", rownames(tmp2_cd8))


tmp2_cd8$public_id <- rownames(tmp2_cd8)
md_bl <- baseline@meta.data |> dplyr::distinct(binned_PC_PCT, public_id)
tmp <- dplyr::left_join(tmp2_cd8, md_bl, by = "public_id")

plot_list <- list()

for (feat in FEATURES) {
    tmp_nprp <- tmp |> dplyr::filter(binned_PC_PCT %in% c("<20", ">=20"))

    pvals_wilcox <- tmp_nprp |>
        rstatix::wilcox_test(as.formula(paste0(feat, " ~ binned_PC_PCT"))) |>
        rstatix::add_xy_position()

    plot_list[[feat]] <- ggplot(tmp_nprp, aes(x = binned_PC_PCT, y = !!sym(feat), fill = binned_PC_PCT, color = binned_PC_PCT)) +
        geom_violin(alpha = 0.75) +
        geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
        geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
        ggtitle(feat) +
        ylab(sprintf("Average Per Patient Expression in CD8+ Cells")) +
        ggtitle(feat) +
        theme_prism() +
        theme(axis.title.x = element_blank()) +
        scale_fill_manual(values = fill_theme_binary_PCPCT) +
        scale_color_manual(values = color_theme_binary_PCPCT) +
        NoLegend() +
        ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
}

FEATURES <- c("TNFRSF13B", "IGHGP", "SDC1", "FCRL5")

tmp2_cd8 <- AverageExpression(cd8, assays = "RNA", group.by = "public_id", features = FEATURES)$RNA |>
    t() |>
    as.data.frame()
rownames(tmp2_cd8) <- gsub("-", "_", rownames(tmp2_cd8))

tmp2_cd8$public_id <- rownames(tmp2_cd8)
md_bl <- baseline@meta.data |> dplyr::distinct(binned_PC_PCT, public_id)
tmp <- dplyr::left_join(tmp2_cd8, md_bl, by = "public_id")

plot_list <- list()

for (feat in FEATURES) {
    tmp_nprp <- tmp |> dplyr::filter(binned_PC_PCT %in% c("<20", ">=20"))

    pvals_wilcox <- tmp_nprp |>
        rstatix::wilcox_test(as.formula(paste0(feat, " ~ binned_PC_PCT"))) |>
        rstatix::add_xy_position()

    plot_list[[feat]] <- ggplot(tmp_nprp, aes(x = binned_PC_PCT, y = !!sym(feat), fill = binned_PC_PCT, color = binned_PC_PCT)) +
        geom_violin(alpha = 0.75) +
        geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
        geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
        ggtitle(feat) +
        ylab(sprintf("Average Per Patient Expression in CD8+ Cells")) +
        ggtitle(feat) +
        theme_prism() +
        theme(axis.title.x = element_blank()) +
        scale_fill_manual(values = fill_theme_binary_PCPCT) +
        scale_color_manual(values = color_theme_binary_PCPCT) +
        NoLegend() +
        ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
}

ptest <- patchwork::wrap_plots(plot_list, ncol = 4, guides = "collect")


pdf_and_png(ptest, output_expr, "FCRL5_in_tcells", pdfWidth = 5, pdfHeight = 2.5, scale = 3)

ptest <- patchwork::wrap_plots(plot_list, ncol = 2, guides = "collect")

pdf_and_png(ptest, output_expr, "tigit_tox_pcpct", pdfWidth = 2.5, pdfHeight = 2.5, scale = 3)

tmp2_tem$public_id <- rownames(tmp2_tem)
md_bl <- baseline@meta.data |> dplyr::distinct(binned_PC_PCT, public_id)
tmp <- dplyr::left_join(tmp2_tem, md_bl, by = "public_id")

plot_listb <- list()

for (feat in FEATURES) {
    tmp_nprp <- tmp |> dplyr::filter(binned_PC_PCT %in% c("<20", ">=20"))

    pvals_wilcox <- tmp_nprp |>
        rstatix::wilcox_test(as.formula(paste0(feat, " ~ binned_PC_PCT"))) |>
        rstatix::add_xy_position()

    plot_listb[[feat]] <- ggplot(tmp_nprp, aes(x = binned_PC_PCT, y = !!sym(feat), fill = binned_PC_PCT, color = binned_PC_PCT)) +
        geom_violin(alpha = 0.75) +
        geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
        geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
        ggtitle(feat) +
        ylab(sprintf("Average Per Patient Expression in CD8+ Cells")) +
        ggtitle(feat) +
        theme_prism() +
        theme(axis.title.x = element_blank()) +
        scale_fill_manual(values = fill_theme_binary_PCPCT) +
        scale_color_manual(values = color_theme_binary_PCPCT) +
        NoLegend() +
        ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
}

ptestb <- patchwork::wrap_plots(plot_listb, ncol = 2, guides = "collect")
```
