---
title: "Figure 4 - Differential Abundance Plots"
author: "William Pilcher"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
---



Generates the select differential abundance box plots for main figures

Includes: 

Figure 4b (Differential Abundance of Major Cell Types)

Supplemental Figure 8 (CD8% of CD3+ Cells)

Figure 4f (Box plots of select CD3+ cells)

Revisions: Differential abundance methodology for entire paper standardized to logit transformation

The original file performs all calculations for all parts of the paper in one location. This version of the file simplifies down to sections relevant to Figure 4.


# This code is a wrapper to perform differential abundance tests for a variety of variables.
# All differetial abundances analyses perform binary comparions between a group with or without an event, adjusting for siteXbatch as a covariate.
# All differential abundances use limma with a robust linear model to fit to the logit-transformed proportions, in a mechanism similar to propeller.
# Logit is preferred for its interpretability as opposed to arcsin-transformations.


```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Setup
```{r init}
library(ggpubr)
source("source/Validation/validation_load_data.R", local = T)
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)

cluster_group <- "subcluster_V03072023"
output_expr <- file.path(output, "Diff_Abundance", "LINEAR_LOGIT", "RISK_CATEGORIES_STUDY_SITE")

output_extra <- file.path(output_F3, "Diff_Abundance")
output_box <- file.path(output_extra, "Diff_Abundance_Box")
output_raw_csv <- file.path(output_extra, "CSV")
output_abun_volcano <- file.path(output_extra, "Diff_Abun_Volcano")
dir.create(output_extra, showWarnings = F, recursive = T)
dir.create(output_box, showWarnings = F, recursive = T)
dir.create(output_raw_csv, showWarnings = F, recursive = T)
dir.create(output_abun_volcano, showWarnings = F, recursive = T)
dir.create(output_expr, showWarnings = F, recursive = T)

# floral_fill_theme <- list(
#     "groupA (SIG)" = fill_RP,
#     "groupA" = "#a9bad0",
#     "groupB (SIG)" = "#A48AD3",
#     "groupB" = "#baafcf"
# )

floral_color_theme <- list(
    "SIG (+)" = fill_RP,
    "+" = color_RP,
    "SIG (-)" = fill_NP,
    "-" = color_NP
)

floral_fill_theme <- floral_color_theme
```

```{r}
var_to_test <- "nsd2_call"
groupA <- "yes"
groupB <- "no"

# function_string <- "d_pt_race_1"
method <- "LINEAR"

variable_info_structure <- rev(list(
    "progression_group" = list(
        isFactor = TRUE,
        factorLevels = c("NP", "RP"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Progression Group",
        perform_contrasts = TRUE
    ),
    "SR_davies.progression_group" = list(
        isFactor = TRUE,
        factorLevels = c("standard_risk.NP", "standard_risk.RP", "high_risk.NP", "high_risk.RP"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Progression Group (Standard Risk)",
        perform_contrasts = TRUE
    ),
    "Triplet_therapy.progression_group" = list(
        isFactor = TRUE,
        factorLevels = c("Triplet.NP", "Triplet.RP"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Progression Group (Triplet)"
    ),
    "Doublet_therapy.progression_group" = list(
        isFactor = TRUE,
        factorLevels = c("Doublet.NP", "Doublet.RP"),
        perform_univariate = TRUE,
        univariate_batch = "Study_Site",
        include_multivariate = FALSE,
        variable_name = "Progression Group (Doublet)"
    ),
    "Study_Site" = list(
        isFactor = TRUE,
        perform_univariate = FALSE,
        include_multivariate = FALSE
    )
))
```

# Differential Abundance - Setup Counts Data

```{r}
# Only considering the 'discovery' samples at 'baseline'
baseline <- merged |> subset(subset = (VJ_INTERVAL == "Baseline" & cohort == "discovery"))

Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline$progression_group, "_", baseline[[var_to_test]])
```


# Setting up other covariates explored as part of separate analyses
```{r}
counts <- table(baseline$public_id, baseline@meta.data[, cluster_group])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")


baseline@meta.data$IMWG_RISK_CLASS <- factor(baseline@meta.data$IMWG_RISK_CLASS, levels = c(0, 1, 2))
baseline@meta.data <- baseline@meta.data |> dplyr::mutate(IMWG_int_or_high = dplyr::case_when(
    IMWG_RISK_CLASS == 0 ~ "standard_risk",
    IMWG_RISK_CLASS %in% c(1, 2) ~ "high_risk",
    .default = NA
))

baseline@meta.data$SR_davies.progression_group <- paste0(baseline@meta.data$davies_based_risk, ".", baseline@meta.data$progression_group)
baseline@meta.data$HR_davies.progression_group <- paste0(baseline@meta.data$davies_based_risk, ".", baseline@meta.data$progression_group)

baseline@meta.data$SR_IMWG2024.progression_group <- paste0(baseline@meta.data$IMWG_2024_HR, ".", baseline@meta.data$progression_group)
baseline@meta.data$HR_IMWG2024.progression_group <- paste0(baseline@meta.data$IMWG_2024_HR, ".", baseline@meta.data$progression_group)

baseline@meta.data$SR_EXPANDED.progression_group <- paste0(baseline@meta.data$expanded_risk_definition, ".", baseline@meta.data$progression_group)
baseline@meta.data$HR_EXPANDED.progression_group <- paste0(baseline@meta.data$expanded_risk_definition, ".", baseline@meta.data$progression_group)


baseline@meta.data <- baseline@meta.data |> dplyr::mutate(triplet_therapy = dplyr::case_when(
    d_tx_induction_cat %in% c("imid_pi_steroid", "chemo_imid_pi_steroid") ~ "Triplet",
    d_tx_induction_cat %in% c("pi_steroid", "chemo_pi_steroid", "imid_steroid", "chemo_imid_steroid") ~ "Doublet",
    .default = "ERROR MISSED CASE"
))

baseline@meta.data$Triplet_therapy.progression_group <- paste0(baseline@meta.data$triplet_therapy, ".", baseline@meta.data$progression_group)
baseline@meta.data$Doublet_therapy.progression_group <- paste0(baseline@meta.data$triplet_therapy, ".", baseline@meta.data$progression_group)
```



```{r}
md_bl <- baseline@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::select(public_id, dplyr::all_of(names(variable_info_structure)))

md_bl <- md_bl |> dplyr::distinct(public_id, .keep_all = T)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(!!sym(cluster_group), isDoublet, subcluster_V03072023_compartment, lineage_group, lineage_order, cellID_short, cellID_long)
colnames(cluster_md)[1] <- "cluster"
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")
```

```{r}
factor_names <- c()
for (i in names(variable_info_structure)) {
    if (variable_info_structure[[i]]$isFactor) {
        factor_names <- c(factor_names, paste0(i, variable_info_structure[[i]]$factorLevels[2]))
    } else {
        factor_names <- c(factor_names, i)
    }
}
```

# Diff Abundance 


### Perform all DAA tests

```{r}
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)

# baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
# Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline[[var_to_test]], "_", baseline[[var_to_test]])

counts <- table(baseline$public_id, baseline@meta.data[, cluster_group])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl <- baseline@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::select(public_id, siteXbatch, dplyr::all_of(names(variable_info_structure)))

md_bl <- md_bl |> dplyr::distinct(public_id, .keep_all = T)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(!!sym(cluster_group), isDoublet, subcluster_V03072023_compartment, lineage_group, lineage_order, cellID_short, cellID_long)
colnames(cluster_md)[1] <- "cluster"
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")

# Subsets focused on in the paper for 'progression' based analyses
types_list <- list(
    "CD3+ Cells" = c("CD8", "CD4"),
    "Myeloid Cells" = "M",
    "B Cells" = "B",
    "Immune" = c("CD8", "CD4", "M", "B", "Nk", "Other")
)
file_name_list <- list(
    "CD3+ Cells" = "cd3_only",
    "Myeloid Cells" = "myeloid_only",
    "B Cells" = "b_only",
    "Immune" = "immune_only"
)

# Automatic summary plots are generated within this script - separate code blocks generate figures directly for the paper.
env_list <- list()
for (idx_name in names(types_list)) {
    all_cells <- new.env()

    all_cells$counts <- counts |> dplyr::filter(lineage_group %in% types_list[[idx_name]])
    all_cells$outs <- DiffAbun_Wrapper(all_cells$counts, variable_info_structure = variable_info_structure, DOMULTIVAR = F, method = "LINEAR", transform = "logit")

    # all_cells$outs <- quickLinearAsinAndPlots(all_cells$counts, variable_info_structure)
    write.csv(all_cells$outs$wilcox_out |> dplyr::arrange(factor), file.path(output_expr, paste0("linear_reg_", file_name_list[[idx_name]], ".csv")))
    write.csv(all_cells$outs$wilcox_out_uni |> dplyr::arrange(factor), file.path(output_expr, paste0("linear_reg_UNI", file_name_list[[idx_name]], ".csv")))
    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::filter(variable != "siteXbatch")
    all_cells$cluster_md <- cluster_md
    all_cells$cluster_md$subcluster_V03072023 <- all_cells$cluster_md$cluster

    all_cells$outs$wilcox_out <- dplyr::left_join(all_cells$outs$wilcox_out, all_cells$cluster_md, by = "cluster")
    all_cells$outs$wilcox_out_uni <- dplyr::left_join(all_cells$outs$wilcox_out_uni, all_cells$cluster_md, by = "cluster")

    all_cells$outs$wilcox_out$p.val <- as.numeric(all_cells$outs$wilcox_out$p.val)
    all_cells$outs$wilcox_out$coef <- as.numeric(all_cells$outs$wilcox_out$coef)

    all_cells$outs$wilcox_out_uni$p.val <- as.numeric(all_cells$outs$wilcox_out_uni$p.val)
    all_cells$outs$wilcox_out_uni$coef <- as.numeric(all_cells$outs$wilcox_out_uni$coef)


    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out$binned_pval <- factor(all_cells$outs$wilcox_out$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))

    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$binned_pval <- factor(all_cells$outs$wilcox_out_uni$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))



    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(sig_levels = dplyr::case_when(
        p.val < 0.01 ~ "Significant (p<0.01)",
        p.val < 0.05 ~ "Significant (p<0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$sig_levels <- factor(all_cells$outs$wilcox_out_uni$sig_levels, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))


    custom_fill_theme <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "grey"
    )

    custom_fill_theme_continuous <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "#f5f5f5"
    )
    library(reshape2)

    mat_multi <- reshape2::acast(all_cells$outs$wilcox_out, cellID_short ~ factor, value.var = "coef")
    d_multi_cluster <- dist(mat_multi / sqrt(rowSums(mat_multi * mat_multi)), method = "euclidean")
    multi_dendrogram_clusters <- hclust(d_multi_cluster)
    d_multi_factor <- dist(t(mat_multi) / sqrt(rowSums(t(mat_multi) * t(mat_multi))), method = "euclidean")
    multi_dendrogram_factor <- hclust(d_multi_factor)

    test <- all_cells$outs$wilcox_out_uni |> dplyr::filter(cluster == "Myeloid.0")
    mat_uni <- reshape2::acast(all_cells$outs$wilcox_out_uni, cellID_short ~ factor, value.var = "coef")
    d_uni_cluster <- dist(mat_uni / sqrt(rowSums(mat_uni * mat_uni)), method = "euclidean")
    uni_dendrogram_clusters <- hclust(d_uni_cluster)
    d_uni_factor <- dist(t(mat_uni) / sqrt(rowSums(t(mat_uni) * t(mat_uni))), method = "euclidean")
    uni_dendrogram_factor <- hclust(d_uni_factor)

    uni_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    uni_factor_ordering <- unique(all_cells$outs$wilcox_out_uni$factor) # factor_names # uni_dendrogram_factor$labels[uni_dendrogram_factor$order]

    # multi_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    # multi_factor_ordering <- factor_names # multi_dendrogram_factor$labels[multi_dendrogram_factor$order]


    # all_cells$outs$wilcox_out$factor <- factor(all_cells$outs$wilcox_out$factor, levels = multi_factor_ordering)
    # all_cells$outs$wilcox_out$cellID_short <- factor(all_cells$outs$wilcox_out$cellID_short, levels = multi_cluster_ordering)

    all_cells$outs$wilcox_out_uni$factor <- factor(all_cells$outs$wilcox_out_uni$factor, levels = uni_factor_ordering)
    all_cells$outs$wilcox_out_uni$cellID_short <- factor(all_cells$outs$wilcox_out_uni$cellID_short, levels = uni_cluster_ordering)

    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4
    scale_factor <- 12

    p_uni_indiv <- ggplot(all_cells$outs$wilcox_out_uni) +
        geom_point(aes(x = cellID_short, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686") +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("Factor") +
        ggtitle("Univariate Linear Model (Corr sXb)") +
        coord_fixed() +
        guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
        theme(axis.text.x = element_text(
            size = 14 * 0.95,
            family = "sans", face = "bold",
            margin = margin(t = 0.8 * 14 / 4),
            angle = 90,
            vjust = 0.5,
            hjust = 1,
        ))

    # p_uni_comb <- (p_uni + theme(axis.text.x = element_blank(), axis.title.x = element_blank())) / (p_uni_2 + theme(plot.title = element_blank()))
    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4

    # pdf_and_png(p_uni, output_expr, paste0("linear_reg_uni_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)
    # pdf_and_png(p_uni_2, output_expr, paste0("linear_reg_uni_LOGFC_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)
    # pdf_and_png(p_uni_comb, output_expr, paste0("linear_reg_uni_COMBINED_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor * 2, scale = 3)
    pdf_and_png(p_uni_indiv, output_expr, paste0("linear_reg_uni_ALT_COMBINED_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    plot_list <- list()
    for (i in types_list[[idx_name]]) {
        all_cells$sub_wilcox_out <- all_cells$outs$wilcox_out_uni |> dplyr::filter(lineage_group %in% i)

        sub_mat_uni <- reshape2::acast(all_cells$sub_wilcox_out, cellID_short ~ factor, value.var = "coef")
        sub_d_uni_cluster <- dist(sub_mat_uni / sqrt(rowSums(sub_mat_uni * sub_mat_uni)), method = "euclidean")
        sub_uni_dendrogram_clusters <- hclust(sub_d_uni_cluster)


        sub_uni_cluster_ordering <- sub_uni_dendrogram_clusters$labels[sub_uni_dendrogram_clusters$order]


        # all_cells$sub_wilcox_out$factor <- factor(all_cells$sub_wilcox_out$factor, levels = uni_factor_ordering)
        # all_cells$sub_wilcox_out$cellID_short <- factor(all_cells$sub_wilcox_out$cellID_short, levels = sub_uni_cluster_ordering)


        all_cells$sub_wilcox_out$sig_binned_value <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c(
            "NS", "Significant (p<0.05)", "Significant (p<0.01)"
        ))
        all_cells$sub_wilcox_out$binned_pval <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c("Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"))
        #                low = "#290AD8", mid = "grey", high = "#A50021", limits = c(-0.31, 0.31),

        plot_list[[i]] <- ggplot(all_cells$sub_wilcox_out) +
            geom_point(aes(x = cellID_short, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
            scale_fill_gradient2(
                low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
                oob = squish
            ) +
            scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
            scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
            theme_prism() +
            xlab("Cluster") +
            ylab("Factor") +
            ggtitle(i) +
            coord_fixed() +
            guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
            theme(
                axis.text.x = element_text(
                    size = 14 * 0.95,
                    family = "sans", face = "bold",
                    margin = margin(t = 0.8 * 14 / 4),
                    angle = 90,
                    vjust = 0.5,
                    hjust = 1,
                ),
                axis.title.x = element_blank(),
                plot.margin = unit(c(0, 1, 0, 0), "lines")
            )
        if (i != types_list[[idx_name]][[1]]) {
            plot_list[[i]] <- plot_list[[i]] + theme(
                axis.text.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.title.y = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.length.y = unit(0, "lines")
            )
        }
    }

    for (i in unique(all_cells$outs$wilcox_out_uni$cluster)) {
        plot <- patchwork::wrap_plots(plot_list, ncol = length(types_list[[idx_name]]), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")
    }
    pdf_and_png(plot, output_expr, paste0("linear_reg_uni_ALT_COMBINED_FACETTED_GROUPS_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    env_list[[idx_name]] <- all_cells
}
```

# Supplementary Figures - All Box plots for RP vs NP (Not automatically constructed for all covariates)

```{r}
types_list <- list(
    "CD3+ Cells" = c("CD8", "CD4"),
    "Myeloid Cells" = "M",
    "B Cells" = "B",
    "Immune" = c("CD8", "CD4", "M", "B", "Nk", "Other")
)

yaxis_str <- c(
    "CD3+ Cells" = "CD3+ Cells",
    "Myeloid Cells" = "Myeloid Cells",
    "B Cells" = "B Cells",
    "Immune" = "Immune Cells"
)

cov <- "Progression Group"
for (idx_name in names(types_list)) {
    all_cells <- env_list[[idx_name]]

    wilcox_out_uni <- all_cells$outs$wilcox_out_uni

    factor_list <- unique(wilcox_out_uni$factor)

    # for (cov in factor_list) {
    wilcox_plot_env <- new.env()
    wilcox_plot_env$wilcox_out <- wilcox_out_uni |> dplyr::filter(variable == "progression_group")

    variable <- unique(wilcox_plot_env$wilcox_out$variable)

    variable_info <- variable_info_structure[[variable]]

    # if (variable_info$isFactor) {
    groupA <- "RP" # unique(wilcox_plot_env$wilcox_out$groupA)
    groupB <- "NP" # unique(wilcox_plot_env$wilcox_out$groupB)


    x_max_range <- max(abs(wilcox_plot_env$wilcox_out$coef)) * 1.1

    wilcox_plot_env$p <- ggplot(wilcox_plot_env$wilcox_out) +
        geom_point(aes(x = coef, y = -log10(p.val + 1e-200), fill = direction, color = direction, alpha = 0.5), show.legend = F) +
        geom_vline(xintercept = 0, color = "black", size = 1) +
        geom_hline(yintercept = -log10(0.05 + 1e-200), color = "black", size = 1) +
        geom_label_repel(aes(x = coef, y = -log10(p.val + 1e-200), label = cellID_short, fill = direction), color = "white", max.overlaps = 15, segment.colour = "black") +
        xlim(-x_max_range, x_max_range) +
        ggtitle(paste0("Differential Abundance, ", cov)) +
        theme_prism()

    wilcox_plot_env$annotations <- data.frame(
        xpos = c(-Inf, Inf),
        ypos = c(Inf, Inf),
        annotateText = c(
            sprintf("%s Enriched", groupB),
            sprintf("%s Enriched", groupA)
        ),
        hjustvar = c(-.05, 1),
        vjustvar = c(1, 1)
    )

    wilcox_plot_env$p <- wilcox_plot_env$p +
        scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c(sprintf("%s*", groupA), sprintf("%s", groupA), sprintf("%s*", groupB), sprintf("%s", groupB))) +
        scale_color_manual(values = floral_color_theme) +
        xlab("Adjusted Log-Odds Ratio") +
        ylab("-log10 P Value") +
        ggtitle(paste0("Differential Abundance, ", yaxis_str[[idx_name]])) +
        geom_text(data = wilcox_plot_env$annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

    # browser()
    count_plot <- all_cells$outs$counts_uni[[variable]]

    box_plots <- new.env()
    box_plots$data <- count_plot |> dplyr::filter(progression_group %in% c("NP", "RP"))


    # Add cluster size
    box_plots$data <- box_plots$data |> dplyr::arrange(lineage_group) # , desc(sum))

    box_plots$panels <- unique(box_plots$data$cluster)

    box_plots$plot_list <- list()
    box_plots$plot_list_bar <- list()

    box_plots$per_clust_max <- box_plots$data |>
        dplyr::group_by(cluster, progression_group) |>
        dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
        dplyr::group_by(cluster) |>
        dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.5 * IQR_max, y.position.alt = upper_max + 0.2 * IQR_max)

    box_plots$per_clust_max <- box_plots$per_clust_max |> dplyr::mutate(
        y.position = dplyr::case_when(
            y.position == 0 ~ 0.001,
            .default = y.position
        ),
        y.position.alt = dplyr::case_when(
            y.position.alt == 0 ~ 0.001,
            .default = y.position.alt
        )
    )

    box_plots$data <- box_plots$data |> dplyr::left_join(box_plots$per_clust_max, by = "cluster")
    pval_info <- wilcox_out_uni |> dplyr::filter(variable == "progression_group")
    pval_info <- pval_info |> dplyr::left_join(box_plots$per_clust_max, by = "cluster")
    pval_info$group1 <- pval_info$groupB
    pval_info$group2 <- pval_info$groupA # swapping for positioning
    grouping_var <- "progression_group"
    wrap_point <- 6

    custom_lin_group_order <- types_list[[idx_name]] # unique(all_cells$counts$lineage_group)

    for (k in custom_lin_group_order) {
        box_plots$panels_select <- box_plots$data |>
            dplyr::filter(lineage_group == k) |>
            dplyr::distinct(cluster)

        box_plots$panels_select <- box_plots$panels_select[["cluster"]]

        box_plots$plot_list[[k]] <- list()

        for (i in box_plots$panels_select) {
            box_plots$tmp <- box_plots$data |> dplyr::filter(cluster == i)
            # browser()

            box_plots$tmp_distinct <- pval_info |> dplyr::filter(cluster == i) # box_plots$tmp |> dplyr::distinct(cluster, p, y.position, y.position.alt, cluster_size, cellID_short, group1, group2)
            box_plots$tmp_distinct$p.val <- as.numeric(box_plots$tmp_distinct$p.val)
            box_plots$tmp_distinct_pval_filt <- box_plots$tmp_distinct |> dplyr::filter(p.val <= 0.05)
            box_plots$plot_list[[k]][[i]] <- ggplot(box_plots$tmp) +
                geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
                geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
                stat_pvalue_manual(box_plots$tmp_distinct_pval_filt, label = "{scales::pvalue(p.val)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
                theme_prism() +
                ylim(c(0, box_plots$tmp_distinct$y.position)) +
                scale_fill_manual(values = fill_theme_binary) +
                scale_color_manual(values = color_theme_binary) +
                theme(axis.title.x = element_blank()) +
                ylab((paste0("Proportion\n", "of ", yaxis_str[[idx_name]]))) +
                NoLegend() +
                coord_cartesian(clip = "off") +
                theme(axis.title.x = element_blank()) +
                ggtitle(box_plots$tmp_distinct$cellID_short) +
                theme(plot.title = element_text(size = 14, face = "bold"))
        }
    }

    test <- list()
    test_b <- list()
    heights <- c()
    heights_b <- list()
    widths_b <- list()

    heights_b_con <- c()
    widths_b_con <- c()

    for (i in 1:(4 * wrap_point)) {
        test_b[[i]] <- plot_spacer()
    }
    # Use the Design matrix to add gaps in the plot
    for (k in custom_lin_group_order) {
        for (i in seq_along(box_plots$plot_list[[k]])) {
            if (mod(i - 1, wrap_point) != 0) {
                box_plots$plot_list[[k]][[i]] <- box_plots$plot_list[[k]][[i]] + theme(axis.title.y = element_blank())
            }
        }
        numPlots <- length(box_plots$plot_list[[k]])
        if (numPlots > 0) {
            if (mod(numPlots, wrap_point) != 0) {
                idx <- 1
                for (i in (mod(numPlots, wrap_point) + 1):wrap_point) {
                    box_plots$plot_list[[k]][[numPlots + idx]] <- plot_spacer()
                    idx <- idx + 1
                }
            }
            updated_size <- length(box_plots$plot_list[[k]])
            for (i in 1:6) {
                box_plots$plot_list[[k]][[updated_size + i]] <- plot_spacer()
            }
            widths_b[[k]] <- rep(1.0, length(box_plots$plot_list[[k]]))
            heights_b[[k]] <- rep(1.0, length(box_plots$plot_list[[k]]))

            widths_b_con <- c(widths_b_con, widths_b[[k]])
            heights_b_con <- c(heights_b_con, heights_b[[k]])

            test_b <- c(test_b, box_plots$plot_list[[k]])
            test[[k]] <- wrap_plots(box_plots$plot_list[[k]], ncol = wrap_point)
            heights[k] <- ceiling(length(box_plots$plot_list[[k]]) / wrap_point)
        }
    }

    patchwork_nrows <- ceiling(length(test_b) / wrap_point)
    patchwork_area <- ceiling(sqrt(length(test_b))) # ?????????????????

    # dark magic
    design_mat <- patchwork::area()
    base <- 0
    for (i in seq_along(test_b)) {
        col <- mod(i - 1, wrap_point) + 1
        row <- ceiling(i / wrap_point)

        design_mat <- c(design_mat, area(t = base + row, b = base + row, l = col, r = col))
    }

    test_multi <- wrap_plots(test_b, ncol = wrap_point, heights = heights_b_con)

    patchwork_nrows <- ceiling(length(test_b) / wrap_point)
    test_multi_with_wilcox <- test_multi + wilcox_plot_env$p + plot_layout(design = c(
        design_mat,
        area(t = 1, b = 3, l = 2, r = 5)
    ))

    test_multi_with_wilcox_free <- test_multi + free(wilcox_plot_env$p) + plot_layout(design = c(
        design_mat,
        area(t = 1, b = 3, l = 2, r = 5)
    ))

    pdf_and_png(test_multi_with_wilcox_free, output_expr, paste0("wrapped_free_", file_name_list[[idx_name]], ".pdf"), pdfWidth = 3 * W_FULL, pdfHeight = H_FULL * row / 2, scale = 3)
    # }
    # }
}
```

# Triplet Therapy vs Normal Scatter Plot Supplementary Figure

```{r}
all_cells <- env_list[["Immune"]]

wilcox_out_uni <- all_cells$outs$wilcox_out_uni

all_prog <- wilcox_out_uni |>
    dplyr::filter(variable == "progression_group") |>
    dplyr::select(cellID_short, coef) |>
    dplyr::rename(RP_v_NP = "coef")
triplet_prog <- wilcox_out_uni |>
    dplyr::filter(variable == "Triplet_therapy.progression_group") |>
    dplyr::select(cellID_short, coef) |>
    dplyr::rename(RP.T_v_NP.T = "coef")

diff_abun_test <- dplyr::left_join(all_prog, triplet_prog, by = "cellID_short")


diff_abun_test <- diff_abun_test |> dplyr::mutate(direction = dplyr::case_when(
    RP.T_v_NP.T > 0 & RP_v_NP > 0 ~ "RP Enriched Consistent",
    RP.T_v_NP.T < 0 & RP_v_NP < 0 ~ "NP Enriched Consistent",
    RP.T_v_NP.T > 0 & RP_v_NP < 0 ~ "NP Overall, RP Triplet",
    RP.T_v_NP.T < 0 & RP_v_NP > 0 ~ "NP Triplet, RP Overall",
    RP.T_v_NP.T == 0 & RP_v_NP > 0 ~ "RP Overall, NC Triplet",
    RP.T_v_NP.T == 0 & RP_v_NP < 0 ~ "NP Overall, NC Triplet",
    RP.T_v_NP.T > 0 & RP_v_NP == 0 ~ "RP Triplet, NC Overall",
    RP.T_v_NP.T < 0 & RP_v_NP == 0 ~ "NP Triplet, NC Overall",
    RP.T_v_NP.T == 0 & RP_v_NP == 0 ~ "NC",
    .default = "UNKNOWN CASE"
))

p_scatter <- ggplot(diff_abun_test) +
    geom_point(aes(x = RP.T_v_NP.T, y = RP_v_NP, fill = direction, color = direction)) +
    geom_label_repel(aes(x = RP.T_v_NP.T, y = RP_v_NP, label = cellID_short, fill = direction), max.overlaps = 20) +
    geom_vline(xintercept = 0, color = "black", size = 1) +
    geom_hline(yintercept = 0, color = "black", size = 1) +
    theme_prism() +
    coord_equal() +
    xlab("RP-T v NP-T Log-Odds Ratio") +
    ylab("RP v NP Log-Odds Ratio") +
    xlim(c(-0.35, 0.35)) +
    ylim(c(-0.35, 0.35))

pdf_and_png(p_scatter, output_expr, "Triplet_v_overall_scatter", pdfWidth = 6, pdfHeight = 6, scale = 3)
```

# Performing the analysis on a 'lineage group' basis versus subcluster_V03072023 basis

```{r}
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)

# baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
# Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline[[var_to_test]], "_", baseline[[var_to_test]])

counts <- table(baseline$public_id, baseline@meta.data[, "lineage_group"])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl_all_var <- baseline@meta.data |> dplyr::distinct(public_id, .keep_all = T)


md_bl <- baseline@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::select(public_id, siteXbatch, dplyr::all_of(names(variable_info_structure)))

md_bl <- md_bl |> dplyr::distinct(public_id, .keep_all = T)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(lineage_group)
cluster_md$cluster <- cluster_md$lineage_group
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")

types_list <- list(
    "LG CD3+ Cells" = c("CD8", "CD4"),
    "LG Immune" = c("CD8", "CD4", "M", "B", "Other", "Nk"),
    "LG All NoPlasma" = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    "LG All" = c("CD8", "CD4", "M", "B", "Other", "Nk", "P", "E")
)
file_name_list <- list(
    "LG CD3+ Cells" = "LINEAGE_GROUP_cd3_only",
    "LG Immune" = "LINEAGE_GROUP_immune_only",
    "LG All NoPlasma" = "LINEAGE_GROUP_All_NoPlasma",
    "LG All" = "LINEAGE_GROUP_all_cells"
)

for (idx_name in names(types_list)) {
    all_cells <- new.env()

    all_cells$counts <- counts |> dplyr::filter(lineage_group %in% types_list[[idx_name]])
    all_cells$outs <- DiffAbun_Wrapper(all_cells$counts, variable_info_structure = variable_info_structure, DOMULTIVAR = F, method = "LINEAR", transform = "logit")

    # all_cells$outs <- quickLinearAsinAndPlots(all_cells$counts, variable_info_structure)
    write.csv(all_cells$outs$wilcox_out |> dplyr::arrange(factor), file.path(output_expr, paste0("linear_reg_", file_name_list[[idx_name]], ".csv")))
    write.csv(all_cells$outs$wilcox_out_uni |> dplyr::arrange(factor), file.path(output_expr, paste0("linear_reg_UNI_", file_name_list[[idx_name]], ".csv")))
    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::filter(variable != "siteXbatch")
    all_cells$cluster_md <- cluster_md
    all_cells$cluster_md$subcluster_V03072023 <- all_cells$cluster_md$cluster

    all_cells$outs$wilcox_out <- dplyr::left_join(all_cells$outs$wilcox_out, all_cells$cluster_md, by = "cluster")
    all_cells$outs$wilcox_out_uni <- dplyr::left_join(all_cells$outs$wilcox_out_uni, all_cells$cluster_md, by = "cluster")

    all_cells$outs$wilcox_out$p.val <- as.numeric(all_cells$outs$wilcox_out$p.val)
    all_cells$outs$wilcox_out$coef <- as.numeric(all_cells$outs$wilcox_out$coef)

    all_cells$outs$wilcox_out_uni$p.val <- as.numeric(all_cells$outs$wilcox_out_uni$p.val)
    all_cells$outs$wilcox_out_uni$coef <- as.numeric(all_cells$outs$wilcox_out_uni$coef)


    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out$binned_pval <- factor(all_cells$outs$wilcox_out$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))

    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$binned_pval <- factor(all_cells$outs$wilcox_out_uni$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))



    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(sig_levels = dplyr::case_when(
        p.val < 0.01 ~ "Significant (p<0.01)",
        p.val < 0.05 ~ "Significant (p<0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$sig_levels <- factor(all_cells$outs$wilcox_out_uni$sig_levels, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))


    custom_fill_theme <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "grey"
    )

    custom_fill_theme_continuous <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "#f5f5f5"
    )
    library(reshape2)

    mat_multi <- reshape2::acast(all_cells$outs$wilcox_out, lineage_group ~ factor, value.var = "coef")
    d_multi_cluster <- dist(mat_multi / sqrt(rowSums(mat_multi * mat_multi)), method = "euclidean")
    multi_dendrogram_clusters <- hclust(d_multi_cluster)
    d_multi_factor <- dist(t(mat_multi) / sqrt(rowSums(t(mat_multi) * t(mat_multi))), method = "euclidean")
    multi_dendrogram_factor <- hclust(d_multi_factor)

    mat_uni <- reshape2::acast(all_cells$outs$wilcox_out_uni, lineage_group ~ factor, value.var = "coef")
    d_uni_cluster <- dist(mat_uni / sqrt(rowSums(mat_uni * mat_uni)), method = "euclidean")
    uni_dendrogram_clusters <- hclust(d_uni_cluster)
    d_uni_factor <- dist(t(mat_uni) / sqrt(rowSums(t(mat_uni) * t(mat_uni))), method = "euclidean")
    uni_dendrogram_factor <- hclust(d_uni_factor)

    uni_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    uni_factor_ordering <- unique(all_cells$outs$wilcox_out_uni$factor) # factor_names # uni_dendrogram_factor$labels[uni_dendrogram_factor$order]

    # multi_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    # multi_factor_ordering <- factor_names # multi_dendrogram_factor$labels[multi_dendrogram_factor$order]


    # all_cells$outs$wilcox_out$factor <- factor(all_cells$outs$wilcox_out$factor, levels = multi_factor_ordering)
    # all_cells$outs$wilcox_out$cellID_short <- factor(all_cells$outs$wilcox_out$cellID_short, levels = multi_cluster_ordering)

    all_cells$outs$wilcox_out_uni$factor <- factor(all_cells$outs$wilcox_out_uni$factor, levels = uni_factor_ordering)
    all_cells$outs$wilcox_out_uni$lineage_group <- factor(all_cells$outs$wilcox_out_uni$lineage_group, levels = uni_cluster_ordering)

    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4
    scale_factor <- 12

    p_uni_indiv <- ggplot(all_cells$outs$wilcox_out_uni) +
        geom_point(aes(x = lineage_group, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686") +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("Factor") +
        ggtitle("Univariate Linear Model (Corr sXb)") +
        coord_fixed() +
        guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
        theme(axis.text.x = element_text(
            size = 14 * 0.95,
            family = "sans", face = "bold",
            margin = margin(t = 0.8 * 14 / 4),
            angle = 90,
            vjust = 0.5,
            hjust = 1,
        ))

    # p_uni_comb <- (p_uni + theme(axis.text.x = element_blank(), axis.title.x = element_blank())) / (p_uni_2 + theme(plot.title = element_blank()))
    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4
    pdf_and_png(p_uni_indiv, output_expr, paste0("linear_reg_uni_ALT_COMBINED_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    plot_list <- list()
    for (i in list("AllComp")) {
        all_cells$sub_wilcox_out <- all_cells$outs$wilcox_out_uni #|> dplyr::filter(lineage_group %in% i)

        sub_mat_uni <- reshape2::acast(all_cells$sub_wilcox_out, lineage_group ~ factor, value.var = "coef")
        sub_d_uni_cluster <- dist(sub_mat_uni / sqrt(rowSums(sub_mat_uni * sub_mat_uni)), method = "euclidean")
        sub_uni_dendrogram_clusters <- hclust(sub_d_uni_cluster)


        sub_uni_cluster_ordering <- sub_uni_dendrogram_clusters$labels[sub_uni_dendrogram_clusters$order]


        all_cells$sub_wilcox_out$sig_binned_value <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c(
            "NS", "Significant (p<0.05)", "Significant (p<0.01)"
        ))
        all_cells$sub_wilcox_out$binned_pval <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c("Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"))
        #                low = "#290AD8", mid = "grey", high = "#A50021", limits = c(-0.31, 0.31),

        plot_list[[i]] <- ggplot(all_cells$sub_wilcox_out) +
            geom_point(aes(x = lineage_group, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
            scale_fill_gradient2(
                low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
                oob = squish
            ) +
            scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
            scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
            theme_prism() +
            xlab("Cluster") +
            ylab("Factor") +
            ggtitle(i) +
            coord_fixed() +
            guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
            theme(
                axis.text.x = element_text(
                    size = 14 * 0.95,
                    family = "sans", face = "bold",
                    margin = margin(t = 0.8 * 14 / 4),
                    angle = 90,
                    vjust = 0.5,
                    hjust = 1,
                ),
                axis.title.x = element_blank(),
                plot.margin = unit(c(0, 1, 0, 0), "lines")
            )
    }

    for (i in unique(all_cells$outs$wilcox_out_uni$cluster)) {
        plot <- patchwork::wrap_plots(plot_list, ncol = length(list("AllComp")), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")
    }
    pdf_and_png(plot, output_expr, paste0("linear_reg_uni_ALT_COMBINED_FACETTED_GROUPS_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    env_list[[idx_name]] <- all_cells
}
```



# Alternative where CD8+ and CD4+ populations are combined together as just 'T' cells. 

```{r}
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)

# baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
# Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline[[var_to_test]], "_", baseline[[var_to_test]])

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(lineage_group_combined_T = dplyr::case_when(
    lineage_group %in% c("CD8", "CD4") ~ "T",
    .default = lineage_group
))

counts <- table(baseline$public_id, baseline@meta.data[, "lineage_group_combined_T"])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl_all_var <- baseline@meta.data |> dplyr::distinct(public_id, .keep_all = T)


md_bl <- baseline@meta.data |>
    dplyr::filter(VJ_INTERVAL == "Baseline") |>
    dplyr::select(public_id, siteXbatch, dplyr::all_of(names(variable_info_structure)))

md_bl <- md_bl |> dplyr::distinct(public_id, .keep_all = T)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(lineage_group_combined_T)
cluster_md$cluster <- cluster_md$lineage_group_combined_T
cluster_md$lineage_group <- cluster_md$lineage_group_combined_T

counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")
types_list <- list(
    "LG-T Immune" = c("T", "M", "B", "Other", "Nk"),
    "LG-T All NoPlasma" = c("T", "M", "B", "Other", "Nk", "E"),
    "LG-T All" = c("T", "M", "B", "Other", "Nk", "P", "E")
)
file_name_list <- list(
    "LG-T Immune" = "LINEAGE_GROUP_MERGED_T_immune_only",
    "LG-T All NoPlasma" = "LINEAGE_GROUP_MERGED_TAll_NoPlasma",
    "LG-T All" = "LINEAGE_GROUP_MERGED_T_all_cells"
)

for (idx_name in names(types_list)) {
    all_cells <- new.env()

    all_cells$counts <- counts |> dplyr::filter(lineage_group %in% types_list[[idx_name]])
    all_cells$outs <- DiffAbun_Wrapper(all_cells$counts, variable_info_structure = variable_info_structure, DOMULTIVAR = F, method = "LINEAR", transform = "logit")


    # all_cells$outs <- quickLinearAsinAndPlots(all_cells$counts, variable_info_structure)
    write.csv(all_cells$outs$wilcox_out, file.path(output_expr, paste0("linear_reg_", file_name_list[[idx_name]], ".csv")))
    write.csv(all_cells$outs$wilcox_out_uni, file.path(output_expr, paste0("linear_reg_UNI_", file_name_list[[idx_name]], ".csv")))

    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::filter(variable != "siteXbatch")
    all_cells$cluster_md <- cluster_md
    all_cells$cluster_md$subcluster_V03072023 <- all_cells$cluster_md$cluster

    all_cells$outs$wilcox_out <- dplyr::left_join(all_cells$outs$wilcox_out, all_cells$cluster_md, by = "cluster")
    all_cells$outs$wilcox_out_uni <- dplyr::left_join(all_cells$outs$wilcox_out_uni, all_cells$cluster_md, by = "cluster")

    all_cells$outs$wilcox_out$p.val <- as.numeric(all_cells$outs$wilcox_out$p.val)
    all_cells$outs$wilcox_out$coef <- as.numeric(all_cells$outs$wilcox_out$coef)

    all_cells$outs$wilcox_out_uni$p.val <- as.numeric(all_cells$outs$wilcox_out_uni$p.val)
    all_cells$outs$wilcox_out_uni$coef <- as.numeric(all_cells$outs$wilcox_out_uni$coef)


    all_cells$outs$wilcox_out <- all_cells$outs$wilcox_out |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out$binned_pval <- factor(all_cells$outs$wilcox_out$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))

    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(binned_pval = dplyr::case_when(
        p.val < 0.001 & coef > 0 ~ "Enriched (<0.001)",
        p.val < 0.01 & coef > 0 ~ "Enriched (<0.01)",
        p.val < 0.05 & coef > 0 ~ "Enriched (0.05)",
        p.val < 0.001 & coef < 0 ~ "Depleted (<0.001)",
        p.val < 0.01 & coef < 0 ~ "Depleted (<0.01)",
        p.val < 0.05 & coef < 0 ~ "Depleted (0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$binned_pval <- factor(all_cells$outs$wilcox_out_uni$binned_pval, levels = c(
        "Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"
    ))



    all_cells$outs$wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::mutate(sig_levels = dplyr::case_when(
        p.val < 0.01 ~ "Significant (p<0.01)",
        p.val < 0.05 ~ "Significant (p<0.05)",
        .default = "NS"
    ))

    all_cells$outs$wilcox_out_uni$sig_levels <- factor(all_cells$outs$wilcox_out_uni$sig_levels, levels = c(
        "NS", "Significant (p<0.05)", "Significant (p<0.01)"
    ))


    custom_fill_theme <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "grey"
    )

    custom_fill_theme_continuous <- list(
        "Enriched (<0.001)" = "#A50021",
        "Enriched (<0.01)" = "#F76D5E",
        "Enriched (0.05)" = "#FFAD72",
        "Depleted (<0.001)" = "#290AD8",
        "Depleted (<0.01)" = "#3FA0FF",
        "Depleted (0.05)" = "#AAF7FF",
        "NS" = "#f5f5f5"
    )
    library(reshape2)

    mat_multi <- reshape2::acast(all_cells$outs$wilcox_out, lineage_group ~ factor, value.var = "coef")
    d_multi_cluster <- dist(mat_multi / sqrt(rowSums(mat_multi * mat_multi)), method = "euclidean")
    multi_dendrogram_clusters <- hclust(d_multi_cluster)
    d_multi_factor <- dist(t(mat_multi) / sqrt(rowSums(t(mat_multi) * t(mat_multi))), method = "euclidean")
    multi_dendrogram_factor <- hclust(d_multi_factor)

    mat_uni <- reshape2::acast(all_cells$outs$wilcox_out_uni, lineage_group ~ factor, value.var = "coef")
    d_uni_cluster <- dist(mat_uni / sqrt(rowSums(mat_uni * mat_uni)), method = "euclidean")
    uni_dendrogram_clusters <- hclust(d_uni_cluster)
    d_uni_factor <- dist(t(mat_uni) / sqrt(rowSums(t(mat_uni) * t(mat_uni))), method = "euclidean")
    uni_dendrogram_factor <- hclust(d_uni_factor)

    uni_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    uni_factor_ordering <- unique(all_cells$outs$wilcox_out_uni$factor) # factor_names # uni_dendrogram_factor$labels[uni_dendrogram_factor$order]

    # multi_cluster_ordering <- uni_dendrogram_clusters$labels[uni_dendrogram_clusters$order]
    # multi_factor_ordering <- factor_names # multi_dendrogram_factor$labels[multi_dendrogram_factor$order]


    # all_cells$outs$wilcox_out$factor <- factor(all_cells$outs$wilcox_out$factor, levels = multi_factor_ordering)
    # all_cells$outs$wilcox_out$cellID_short <- factor(all_cells$outs$wilcox_out$cellID_short, levels = multi_cluster_ordering)

    all_cells$outs$wilcox_out_uni$factor <- factor(all_cells$outs$wilcox_out_uni$factor, levels = uni_factor_ordering)
    all_cells$outs$wilcox_out_uni$lineage_group <- factor(all_cells$outs$wilcox_out_uni$lineage_group, levels = uni_cluster_ordering)

    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4
    scale_factor <- 12

    p_uni_indiv <- ggplot(all_cells$outs$wilcox_out_uni) +
        geom_point(aes(x = lineage_group, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686") +
        scale_fill_gradient2(
            low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
            oob = squish
        ) +
        scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
        scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
        theme_prism() +
        xlab("Cluster") +
        ylab("Factor") +
        ggtitle("Univariate Linear Model (Corr sXb)") +
        coord_fixed() +
        guides(size = guide_legend(override.aes = list(shape = 21, colour = "grey", fill = "white"))) +
        theme(axis.text.x = element_text(
            size = 14 * 0.95,
            family = "sans", face = "bold",
            margin = margin(t = 0.8 * 14 / 4),
            angle = 90,
            vjust = 0.5,
            hjust = 1,
        ))

    # p_uni_comb <- (p_uni + theme(axis.text.x = element_blank(), axis.title.x = element_blank())) / (p_uni_2 + theme(plot.title = element_blank()))
    heights <- length(unique(all_cells$outs$wilcox_out_uni$factor))
    char_len_uni <- max(nchar(levels(all_cells$outs$wilcox_out_uni$factor)))

    widths <- length(unique(all_cells$outs$wilcox_out_uni$cluster)) + char_len_uni / 4

    # pdf_and_png(p_uni, output_expr, paste0("linear_reg_uni_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)
    # pdf_and_png(p_uni_2, output_expr, paste0("linear_reg_uni_LOGFC_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)
    # pdf_and_png(p_uni_comb, output_expr, paste0("linear_reg_uni_COMBINED_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor * 2, scale = 3)
    pdf_and_png(p_uni_indiv, output_expr, paste0("linear_reg_uni_ALT_COMBINED_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    plot_list <- list()
    for (i in list("AllComp")) {
        all_cells$sub_wilcox_out <- all_cells$outs$wilcox_out_uni #|> dplyr::filter(lineage_group %in% i)

        sub_mat_uni <- reshape2::acast(all_cells$sub_wilcox_out, lineage_group ~ factor, value.var = "coef")
        sub_d_uni_cluster <- dist(sub_mat_uni / sqrt(rowSums(sub_mat_uni * sub_mat_uni)), method = "euclidean")
        sub_uni_dendrogram_clusters <- hclust(sub_d_uni_cluster)


        sub_uni_cluster_ordering <- sub_uni_dendrogram_clusters$labels[sub_uni_dendrogram_clusters$order]


        # all_cells$sub_wilcox_out$factor <- factor(all_cells$sub_wilcox_out$factor, levels = uni_factor_ordering)
        # all_cells$sub_wilcox_out$cellID_short <- factor(all_cells$sub_wilcox_out$cellID_short, levels = sub_uni_cluster_ordering)


        all_cells$sub_wilcox_out$sig_binned_value <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c(
            "NS", "Significant (p<0.05)", "Significant (p<0.01)"
        ))
        all_cells$sub_wilcox_out$binned_pval <- factor(all_cells$sub_wilcox_out$binned_pval, levels = c("Depleted (<0.001)", "Depleted (<0.01)", "Depleted (0.05)", "NS", "Enriched (0.05)", "Enriched (<0.01)", "Enriched (<0.001)"))
        #                low = "#290AD8", mid = "grey", high = "#A50021", limits = c(-0.31, 0.31),

        plot_list[[i]] <- ggplot(all_cells$sub_wilcox_out) +
            geom_point(aes(x = lineage_group, y = factor, fill = coef, size = sig_levels, shape = sig_levels), color = "#868686", show.legend = TRUE) +
            scale_fill_gradient2(
                low = fill_NP, mid = "grey", high = fill_RP, limits = c(-0.31, 0.31),
                oob = squish
            ) +
            scale_size_manual(values = c("Significant (p<0.01)" = 7.5, "Significant (p<0.05)" = 6, "NS" = 4.25), drop = F) +
            scale_shape_manual(values = c("Significant (p<0.01)" = 22, "Significant (p<0.05)" = 23, "NS" = 21), drop = F) +
            theme_prism() +
            xlab("Cluster") +
            ylab("Factor") +
            ggtitle(i) +
            coord_fixed() +
            guides(size = guide_legend(override.aes = list(colour = "grey", fill = "white"))) +
            theme(
                axis.text.x = element_text(
                    size = 14 * 0.95,
                    family = "sans", face = "bold",
                    margin = margin(t = 0.8 * 14 / 4),
                    angle = 90,
                    vjust = 0.5,
                    hjust = 1,
                ),
                axis.title.x = element_blank(),
                plot.margin = unit(c(0, 1, 0, 0), "lines")
            )
    }

    for (i in unique(all_cells$outs$wilcox_out_uni$cluster)) {
        plot <- patchwork::wrap_plots(plot_list, ncol = length(list("AllComp")), byrow = T, guides = "collect", axes = "collect", axes_titles = "collect")
    }
    pdf_and_png(plot, output_expr, paste0("linear_reg_uni_ALT_COMBINED_FACETTED_GROUPS_", file_name_list[[idx_name]]), pdfWidth = 2 + widths / scale_factor, pdfHeight = 3 + heights / scale_factor, scale = 3)

    env_list[[idx_name]] <- all_cells
}
```

```{r}
types_list <- list(
    "CD3+ Cells" = c("CD8", "CD4"),
    "Myeloid Cells" = "M",
    "B Cells" = "B",
    "Nk" = "Nk",
    "Immune" = c("CD8", "CD4", "M", "B", "Other", "Nk"),
    "LG CD3+ Cells" = c("CD8", "CD4"),
    "LG Immune" = c("CD8", "CD4", "M", "B", "Other", "Nk"),
    "LG All NoPlasma" = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    "LG All" = c("CD8", "CD4", "M", "B", "Other", "Nk", "P", "E"),
    "LG-T Immune" = c("T", "M", "B", "Other", "Nk"),
    "LG-T All NoPlasma" = c("T", "M", "B", "Other", "Nk", "E"),
    "LG-T All" = c("T", "M", "B", "Other", "Nk", "P", "E")
)
```

### Figure 4b, panel formatting
Figure 4b, final output, with revised differential abundance methodology
Immune Only Fraction

```{r}
# alternative

idx_name <- "LG-T Immune"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["progression_group"]]
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("progression_group"))

grouping_var <- "progression_group"

per_clust_max <- tmp |>
    dplyr::group_by(cluster, progression_group) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.1, y.position.alt = upper_max + 0.1 * IQR_max)

tmp_out <- wilcox_out_uni |> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out$group1 <- "NP"
tmp_out$group2 <- "RP"

tmp_t <- tmp |> dplyr::filter(cluster == "T")
tmp_b <- tmp |> dplyr::filter(cluster == "B")
tmp_m <- tmp |> dplyr::filter(cluster == "M")

# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)



tmp_out_t <- tmp_out |> dplyr::filter(cluster == "T")
tmp_out_b <- tmp_out |> dplyr::filter(cluster == "B")
tmp_out_m <- tmp_out |> dplyr::filter(cluster == "M")


p_t <- ggplot(tmp_t) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_t, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, 1)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    ylab("Immune Cell Fraction") +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle("T Cell") +
    theme(plot.title = element_text(size = 14, face = "bold"))


p_b <- ggplot(tmp_b) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    stat_pvalue_manual(tmp_out_b, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, tmp_out_b$y.position + 0.05)) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank()) +
    ggtitle("B Cell") +
    theme(plot.title = element_text(size = 14, face = "bold"))


p_m <- ggplot(tmp_m) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_m, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, tmp_out_m$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank()) +
    ggtitle("Myeloid") +
    xlab("Progression Group") +
    theme(plot.title = element_text(size = 14, face = "bold"))

p_indp <- p_t + p_m + p_b

pdf_and_png(p_indp, output_expr, "FIGURE4B_major_cell_types_t_b_m_ind", pdfHeight = H_HALF, pdfWidth = W_FULL, scale = 3, SAVE_RDS = F)
```

### Supplementary Figure - CD8 T cell percentage of CD3+ Cells
Supplemental figure, final output, with revised differential abundance methodology
CD3+ Fraction

```{r}
# alternative

idx_name <- "LG CD3+ Cells"
all_cells <- env_list[[idx_name]]
tmp <- all_cells$outs$counts_uni[["progression_group"]]
wilcox_out_uni <- all_cells$outs$wilcox_out_uni |> dplyr::filter(variable %in% c("progression_group"))

grouping_var <- "progression_group"

per_clust_max <- tmp |>
    dplyr::group_by(cluster, progression_group) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.1, y.position.alt = upper_max + 0.1 * IQR_max)

tmp_out <- wilcox_out_uni |> dplyr::left_join(per_clust_max, by = "cluster")
tmp_out$p <- tmp_out$p.val

tmp_out$group1 <- "NP"
tmp_out$group2 <- "RP"

tmp_t <- tmp |> dplyr::filter(cluster == "CD8")


# df_p_val <- tmp |>
#     rstatix::group_by(cluster) %>%
#     rstatix::wilcox_test(as.formula(sprintf("prop ~ %s", "progression_group")), detailed = T)



tmp_out_t <- tmp_out |> dplyr::filter(cluster == "CD8")


p_t <- ggplot(tmp_t) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_t, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, 1)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    ylab("Fraction of CD3+ Cells") +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle("CD8+ T Cells") +
    theme(plot.title = element_text(size = 14, face = "bold"))

pdf_and_png(p_t, output_expr, "Supplemental_CD8_Positive_Fraction", pdfHeight = H_FULL, pdfWidth = W_FULL, scale = 3, SAVE_RDS = F)
```






### Figure 4F Rendering (Select CD3+ clusters enriched in NP or RP)
Percentage of CD3+ Cells
```{r}
# alternative

tmp_info <- tmp |> dplyr::distinct(cluster, cellID_short, cellID_long)
tmp_out <- tmp_out |> dplyr::left_join(tmp_info, by = "cluster")

tmp_20 <- tmp |> dplyr::filter(cluster == "NkT.2.0")
tmp_21 <- tmp |> dplyr::filter(cluster == "NkT.2.1")
tmp_6 <- tmp |> dplyr::filter(cluster == "NkT.6")
tmp_31 <- tmp |> dplyr::filter(cluster == "NkT.3.1")

tmp_out_20 <- tmp_out |> dplyr::filter(cluster == "NkT.2.0")
tmp_out_21 <- tmp_out |> dplyr::filter(cluster == "NkT.2.1")
tmp_out_6 <- tmp_out |> dplyr::filter(cluster == "NkT.6")
tmp_out_31 <- tmp_out |> dplyr::filter(cluster == "NkT.3.1")

p_t_6 <- ggplot(tmp_6) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_6, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, tmp_out_6$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    theme(axis.title.x = element_blank()) +
    ylab(bquote(atop("Proportion", "of" ~ CD3^"+" ~ "T Cells"))) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle(tmp_6$cellID_short) +
    theme(plot.title = element_text(size = 14, face = "bold"))

p_t_31 <- ggplot(tmp_31) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_31, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, tmp_out_31$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    theme(axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank()) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle(tmp_31$cellID_short) +
    theme(plot.title = element_text(size = 14, face = "bold"))

p_t_20 <- ggplot(tmp_20) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_20, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, tmp_out_20$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    theme(axis.title.x = element_blank()) +
    ylab(bquote(atop("Proportion", "of" ~ CD3^"+" ~ "T Cells"))) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle(tmp_20$cellID_short) +
    theme(plot.title = element_text(size = 14, face = "bold"))

p_t_21 <- ggplot(tmp_21) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_21, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    scale_y_continuous(limits = c(0, tmp_out_21$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    theme(axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank()) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle(tmp_21$cellID_short) +
    theme(plot.title = element_text(size = 14, face = "bold"))


p_naive <- p_t_6 + p_t_31
p_cyto <- p_t_20 + p_t_21

# The NULLS just add some vertical padding without messing with plot margins
p_grid <- cowplot::plot_grid(NULL, p_naive, NULL, p_cyto, nrow = 4, ncol = 1, labels = c("", "Less Differentiated", "", "More Differentiated"), label_size = 18, label_x = 0.5, hjust = 0, vjust = 0.5, rel_heights = c(0.025, 1, 0.025, 1))
pdf_and_png(p_grid, output_expr, "FIGURE4F_select_cd8cd4_clusters", pdfHeight = H_FULL, pdfWidth = W_FULL, scale = 3, SAVE_RDS = F)
```
