---
title: "Figure 4 - Differential Abundance Plots"
author: "William Pilcher"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
---

Generates the select differential abundance box plots for main figures

Includes: 

Figure 4b (Differential Abundance of Major Cell Types)

Supplemental Figure 8 (CD8% of CD3+ Cells)

Figure 4f (Box plots of select CD3+ cells)

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Setup
```{r init}
library(ggpubr)
source("source/Figure_3/Fig3_standard_load_data.R", local = T)
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)



output_extra <- file.path(output_F3, "Diff_Abundance")
output_box <- file.path(output_extra, "Diff_Abundance_Box")
output_raw_csv <- file.path(output_extra, "CSV")
output_abun_volcano <- file.path(output_extra, "Diff_Abun_Volcano")
dir.create(output_extra, showWarnings = F, recursive = T)
dir.create(output_box, showWarnings = F, recursive = T)
dir.create(output_raw_csv, showWarnings = F, recursive = T)
dir.create(output_abun_volcano, showWarnings = F, recursive = T)
```


# Diff Abundance 

## Setup Categories 

# Flag CD45 Positive/Immune Populations
```{r}
cd45neg_notplasma <- c("BEry.0", "BEry.4", "BEry.13", "BEry.17", "Ery.0", "Ery.1", "Ery.2", "Ery.3", "Ery.4", "Ery.5", "Ery.6", "Ery.8", "Ery.9", "Full.23")

merged@meta.data <- merged@meta.data %>% mutate(isImmune = dplyr::case_when(
    .data[[cluster_group]] %in% cd45neg_notplasma ~ "No",
    .default = "Yes"
))
```

# Flag Non-Malginant, Plasma Populations
```{r}
merged@meta.data <- merged@meta.data %>% mutate(isNonMalignantImmune = dplyr::case_when(
    subcluster_V03072023_compartment == "Plasma" ~ "No",
    .default = isImmune
))

merged$pfs_risk <- paste0(merged$progression_group, "_", merged$davies_based_risk)
```


# Differential Abundance - Major Cell Populations (4B)
## Setup Count Data

```{r}
baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline$progression_group, "_", baseline$davies_based_risk)

baseline@meta.data <- baseline@meta.data |> dplyr::mutate(lineage_group_T = dplyr::case_when(
    lineage_group %in% c("CD4", "CD8") ~ "T",
    .default = lineage_group
))

counts <- table(baseline$public_id, baseline@meta.data[, "lineage_group_T"])
counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl <- baseline@meta.data |> dplyr::distinct(public_id, progression_group, Batch, siteXbatch, Study_Site, Batch, davies_based_risk, d_pt_race_1, pfs_risk)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")
```


## Perform wilcoxon tests
```{r}
counts_majfil <- counts |> dplyr::filter(cluster != "LQ" & cluster != "P")

outs <- quickWilcoxAndPlots(counts_majfil, "progression_group", "NP", "RP")
```

Scatter plot of all cell types
```{r}
floral_fill_theme <- list(
    "NP (SIG)" = "#7CA0D4",
    "NP" = "#a9bad0",
    "RP (SIG)" = "#A48AD3",
    "RP" = "#baafcf"
)

floral_color_theme <- list(
    "NP (SIG)" = "#285291",
    "NP" = "#43618e",
    "RP (SIG)" = "#4f2b8e",
    "RP" = "#5f4886"
)

annotations <- data.frame(
    xpos = c(-Inf, Inf),
    ypos = c(Inf, Inf),
    annotateText = c(
        "RP Enriched",
        "NP Enriched"
    ),
    hjustvar = c(-.05, 1),
    vjustvar = c(1, 1)
)

p_wilcoxplot <- outs$plotWilcoxOut +
    scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP*", "NP", "RP*", "RP")) +
    scale_color_manual(values = floral_color_theme) +
    xlab("Average Log2 Fold Change") +
    ylab("-log10 P Value") +
    ggtitle("Differential Abundance, All Clusters") +
    geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

pdf_and_png(p_wilcoxplot, output_abun_volcano, "major_cell_types", pdfWidth = 10 * 1.3, pdfHeight = 8 * 1.3)
```

```{r}
write.csv(outs$wilcox_out, file.path(output_raw_csv, "major_cell_types.csv"))
counts_tmp <- counts
```

Formatting for Figure 4B

```{r}
grouping_var <- "progression_group"
counts <- counts_majfil
counts <- counts |> dplyr::filter(progression_group %in% c("NP", "RP"))

cluster_size <- counts |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(cluster_size = sum(Counts))

counts <- dplyr::left_join(counts, cluster_size)


counts <- counts |>
    dplyr::group_by(public_id) |>
    dplyr::summarise(pat_total = sum(Counts)) |>
    dplyr::right_join(counts) |>
    dplyr::filter(pat_total != 0) |> # remove patients with zero cells in group
    dplyr::mutate(prop = Counts / pat_total)

tmp <- counts |> dplyr::filter(cluster %in% c("T", "M", "B"))
tmp <- tmp |> dplyr::mutate(cluster_label = dplyr::case_when(
    cluster == "T" ~ "T Cell",
    cluster == "M" ~ "Myeloid",
    cluster == "B" ~ "B Cell"
))

tmp$cluster_label <- factor(tmp$cluster_label, levels = c("T Cell", "Myeloid", "B Cell"))


tmp_out <- outs$wilcox_out
tmp_out <- tmp_out |> dplyr::filter(cluster %in% c("T", "M", "B"))

tmp_out <- tmp_out |> dplyr::mutate(cluster_label = dplyr::case_when(
    cluster == "T" ~ "T Cell",
    cluster == "M" ~ "Myeloid",
    cluster == "B" ~ "B Cell"
))
tmp_out$cluster_label <- factor(tmp_out$cluster_label, levels = c("T Cell", "Myeloid", "B Cell"))
tmp_out <- tmp_out |> arrange(cluster_label)

per_clust_max <- tmp |>
    dplyr::group_by(cluster_label, progression_group) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster_label) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.1, y.position.alt = upper_max + 0.1 * IQR_max)

tmp_out <- tmp_out |> dplyr::left_join(per_clust_max, by = "cluster_label")

p <- ggplot(tmp) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out, label = "{scales::pvalue(p)}", size = 5) +
    theme_prism() +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    facet_grid(~cluster_label, scales = "free") +
    xlab("Progression Group") +
    ylab("Proportion") +
    NoLegend() +
    coord_cartesian(clip = "off")

pdf_and_png(p, output_box, "major_cell_types_t_b_m", pdfHeight = 1.22 * 2.5, pdfWidth = 2.966 * 2.5, SAVE_RDS = F)
```


### Figure 4b, panel formatting
Figure 4B, final output


```{r}
# alternative

tmp_t <- tmp |> dplyr::filter(cluster == "T")
tmp_b <- tmp |> dplyr::filter(cluster == "B")
tmp_m <- tmp |> dplyr::filter(cluster == "M")

tmp_out_t <- tmp_out |> dplyr::filter(cluster == "T")
tmp_out_b <- tmp_out |> dplyr::filter(cluster == "B")
tmp_out_m <- tmp_out |> dplyr::filter(cluster == "M")


p_t <- ggplot(tmp_t) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_t, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    ylim(c(0, tmp_out_t$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    ylab("Proportion\nof Non-Malignant Cells") +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle("T Cell") +
    theme(plot.title = element_text(size = 14, face = "bold"))


p_b <- ggplot(tmp_b) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    stat_pvalue_manual(tmp_out_b, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    ylim(c(0, tmp_out_b$y.position + 0.05)) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank()) +
    ggtitle("B Cell") +
    theme(plot.title = element_text(size = 14, face = "bold"))


p_m <- ggplot(tmp_m) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_m, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    ylim(c(0, tmp_out_m$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank()) +
    ggtitle("Myeloid") +
    xlab("Progression Group") +
    theme(plot.title = element_text(size = 14, face = "bold"))

p_indp <- p_t + p_m + p_b

pdf_and_png(p_indp, output_box, "major_cell_types_t_b_m_ind", pdfHeight = H_HALF, pdfWidth = W_FULL, scale = 3, SAVE_RDS = F)
```

# Supplemental Figure 8: CD8% of CD3+ Cells

## Setup counts
```{r}
baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline$progression_group, "_", baseline$davies_based_risk)

counts <- table(baseline$public_id, baseline@meta.data[, "lineage_group"])
counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl <- baseline@meta.data |> dplyr::distinct(public_id, progression_group, Batch, siteXbatch, Study_Site, Batch, davies_based_risk, d_pt_race_1, pfs_risk)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")
```

## Run Wilcoxon

```{r}
counts_majfil <- counts |> dplyr::filter(cluster %in% c("CD8", "CD4"))

outs <- quickWilcoxAndPlots(counts_majfil, "progression_group", "NP", "RP")
```

```{r}
write.csv(outs$wilcox_out, file.path(output_raw_csv, "CD8_v_CD4.csv"))
counts_tmp <- counts_majfil
```

Setup for Supplemental Figure 8 Rendering
```{r}
grouping_var <- "progression_group"
counts <- counts_majfil
counts <- counts |> dplyr::filter(progression_group %in% c("NP", "RP"))

cluster_size <- counts |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(cluster_size = sum(Counts))

counts <- dplyr::left_join(counts, cluster_size)


counts <- counts |>
    dplyr::group_by(public_id) |>
    dplyr::summarise(pat_total = sum(Counts)) |>
    dplyr::right_join(counts) |>
    dplyr::filter(pat_total != 0) |> # remove patients with zero cells in group
    dplyr::mutate(prop = Counts / pat_total)
tmp <- counts
tmp <- tmp |> dplyr::mutate(cluster_label = dplyr::case_when(
    cluster == "CD8" ~ "CD8+ T Cell",
    cluster == "CD4" ~ "CD4+ T Cell"
))

tmp_out <- outs$wilcox_out
tmp_out <- tmp_out |> dplyr::mutate(cluster_label = dplyr::case_when(
    cluster == "CD8" ~ "CD8+ T Cell",
    cluster == "CD4" ~ "CD4+ T Cell"
))


per_clust_max <- tmp |>
    dplyr::group_by(cluster_label, progression_group) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster_label) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.1, y.position.alt = upper_max + 0.1 * IQR_max)

tmp_out <- tmp_out |> dplyr::left_join(per_clust_max, by = "cluster_label")

p <- ggplot(tmp) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out, label = "{scales::pvalue(p)}", size = 5) +
    theme_prism() +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    facet_grid(~cluster_label, scales = "free") +
    xlab("Progression Group") +
    ylab("Proportion") +
    NoLegend() +
    coord_cartesian(clip = "off")

pdf_and_png(p, output_box, "cd8_and_cd4", pdfHeight = H_FULL, pdfWidth = W_FULL, scale = 3)
```

```{r}
# alternative
tmp_t <- tmp |> dplyr::filter(cluster == "CD8")
tmp_out_t <- tmp_out |> dplyr::filter(cluster == "CD8")

p_t <- ggplot(tmp_t) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_t, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    ylim(c(0, tmp_out_t$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    ylab(bquote(atop("Proportion", "of" ~ CD3^"+" ~ "T Cells"))) +
    xlab("Progression Group") +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle("CD8+ T Cell Fraction") +
    theme(plot.title = element_text(size = 14, face = "bold"))


p_indp <- p_t

pdf_and_png(p_indp, output_box, "cd8_percent_only", pdfHeight = H_FULL, pdfWidth = W_FULL, scale = 3, SAVE_RDS = F)
```

### Supplemental Figure 8 Panel Layout

```{r}
tmp_t <- tmp |> dplyr::filter(cluster == "CD8")
tmp_out_t <- tmp_out |> dplyr::filter(cluster == "CD8")

write.csv(tmp_t, file.path(output_box, "proportion_data_plotting_cd8.csv"))
write.csv(tmp_out_t, file.path(output_box, "cluster_proportion_data_plotting_cd8.csv"))

p_t <- ggplot(tmp_t) +
    geom_violin(alpha = 0.75, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.5, width = 0.2, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.75, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_t, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    ylim(c(0, tmp_out_t$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    ylab(bquote(atop("Proportion", "of" ~ CD3^"+" ~ "T Cells"))) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle(bquote(CD8^"+" ~ "T Cell Fraction")) +
    theme(plot.title = element_text(size = 14, face = "bold"))


p_indp <- p_t

pdf_and_png(p_indp, output_box, "cd8_percent_only_violin", pdfHeight = H_FULL, pdfWidth = W_FULL, scale = 3, SAVE_RDS = F)
```



# Main Figure 4F - Differential abundance across CD3+ T Cells

## Setup counts object
```{r}
baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline$progression_group, "_", baseline$davies_based_risk)

counts <- table(baseline$public_id, baseline@meta.data[, cluster_group])
counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl <- baseline@meta.data |> dplyr::distinct(public_id, progression_group, Batch, siteXbatch, Study_Site, Batch, davies_based_risk, d_pt_race_1, pfs_risk)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(!!sym(cluster_group), isDoublet, isImmune, isNonMalignantImmune, subcluster_V03072023_compartment, lineage_group, lineage_order, cellID_short, cellID_long)
colnames(cluster_md)[1] <- "cluster"
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")
```

## Perform Wilcoxon rank-sum 
```{r}
counts_cd8cd4 <- counts |> dplyr::filter(lineage_group %in% c("CD4", "CD8"))
outs <- quickWilcoxAndPlots(counts_cd8cd4, "progression_group", "NP", "RP")
```

```{r}
# floral_fill_theme <- list(
#     "NP (SIG)" = "#7CA0D4",
#     "NP" = "#a9bad0",
#     "RP (SIG)" = "#A48AD3",
#     "RP" = "#baafcf"
# )

# floral_color_theme <- list(
#     "NP (SIG)" = "#285291",
#     "NP" = "#43618e",
#     "RP (SIG)" = "#4f2b8e",
#     "RP" = "#5f4886"
# )

# annotations <- data.frame(
#     xpos = c(-Inf, Inf),
#     ypos = c(Inf, Inf),
#     annotateText = c(
#         "RP Enriched",
#         "NP Enriched"
#     ),
#     hjustvar = c(-.05, 1),
#     vjustvar = c(1, 1)
# )

# p_wilcoxplot <- outs$plotWilcoxOut +
#     scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP*", "NP", "RP*", "RP")) +
#     scale_color_manual(values = floral_color_theme) +
#     xlab("Average Log2 Fold Change") +
#     ylab("-log10 P Value") +
#     ggtitle("Differential Abundance, All Clusters") +
#     geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

# pdf_and_png(p_wilcoxplot, output_abun_volcano, "cd8_cd4_wilcox", pdfWidth = 10 * 1.3, pdfHeight = 8 * 1.3)
```

```{r}
write.csv(outs$wilcox_out, file.path(output_raw_csv, "cd8_cd4_only.csv"))
```

Formatting for Figure 4F

```{r}
grouping_var <- "progression_group"
counts <- counts_cd8cd4
counts <- counts |> dplyr::filter(progression_group %in% c("NP", "RP"))

cluster_size <- counts |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(cluster_size = sum(Counts))

counts <- dplyr::left_join(counts, cluster_size)


counts <- counts |>
    dplyr::group_by(public_id) |>
    dplyr::summarise(pat_total = sum(Counts)) |>
    dplyr::right_join(counts) |>
    dplyr::filter(pat_total != 0) |> # remove patients with zero cells in group
    dplyr::mutate(prop = Counts / pat_total)
tmp <- counts |> dplyr::filter(cluster %in% c("NkT.6", "NkT.3.1", "NkT.2.0", "NkT.2.1"))
tmp <- tmp |> dplyr::mutate(cluster_label = cellID_short)


tmp_out <- outs$wilcox_out |> dplyr::filter(cluster %in% c("NkT.6", "NkT.3.1", "NkT.2.0", "NkT.2.1"))

per_clust_max <- tmp |>
    dplyr::group_by(cluster, progression_group) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.1, y.position.alt = upper_max + 0.1 * IQR_max)

tmp_out <- tmp_out |> dplyr::left_join(per_clust_max, by = "cluster")

p <- ggplot(tmp) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out, label = "{scales::pvalue(p)}", size = 5) +
    theme_prism() +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    facet_grid(~cluster, scales = "free") +
    xlab("Progression Group") +
    ylab("Proportion") +
    NoLegend() +
    coord_cartesian(clip = "off")

pdf_and_png(p, output_box, "various", pdfHeight = H_FULL, pdfWidth = W_FULL, scale = 3)
```

### Figure 4F Rendering (Select CD3+ clusters enriched in NP or RP)

```{r}
# alternative

tmp_info <- tmp |> dplyr::distinct(cluster, cellID_short, cellID_long)
tmp_out <- tmp_out |> dplyr::left_join(tmp_info, by = "cluster")

tmp_20 <- tmp |> dplyr::filter(cluster == "NkT.2.0")
tmp_21 <- tmp |> dplyr::filter(cluster == "NkT.2.1")
tmp_6 <- tmp |> dplyr::filter(cluster == "NkT.6")
tmp_31 <- tmp |> dplyr::filter(cluster == "NkT.3.1")

tmp_out_20 <- tmp_out |> dplyr::filter(cluster == "NkT.2.0")
tmp_out_21 <- tmp_out |> dplyr::filter(cluster == "NkT.2.1")
tmp_out_6 <- tmp_out |> dplyr::filter(cluster == "NkT.6")
tmp_out_31 <- tmp_out |> dplyr::filter(cluster == "NkT.3.1")

p_t_6 <- ggplot(tmp_6) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_6, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    ylim(c(0, tmp_out_6$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    theme(axis.title.x = element_blank()) +
    ylab(bquote(atop("Proportion", "of" ~ CD3^"+" ~ "T Cells"))) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle(tmp_6$cellID_short) +
    theme(plot.title = element_text(size = 14, face = "bold"))

p_t_31 <- ggplot(tmp_31) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_31, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    ylim(c(0, tmp_out_31$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    theme(axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank()) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle(tmp_31$cellID_short) +
    theme(plot.title = element_text(size = 14, face = "bold"))

p_t_20 <- ggplot(tmp_20) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_20, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    ylim(c(0, tmp_out_20$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    theme(axis.title.x = element_blank()) +
    ylab(bquote(atop("Proportion", "of" ~ CD3^"+" ~ "T Cells"))) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle(tmp_20$cellID_short) +
    theme(plot.title = element_text(size = 14, face = "bold"))

p_t_21 <- ggplot(tmp_21) +
    geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
    stat_pvalue_manual(tmp_out_21, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
    theme_prism() +
    ylim(c(0, tmp_out_21$y.position + 0.05)) +
    scale_fill_manual(values = fill_theme_binary) +
    scale_color_manual(values = color_theme_binary) +
    theme(axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank()) +
    NoLegend() +
    coord_cartesian(clip = "off") +
    theme(axis.title.x = element_blank()) +
    ggtitle(tmp_21$cellID_short) +
    theme(plot.title = element_text(size = 14, face = "bold"))


p_naive <- p_t_6 + p_t_31
p_cyto <- p_t_20 + p_t_21

# The NULLS just add some vertical padding without messing with plot margins
p_grid <- cowplot::plot_grid(NULL, p_naive, NULL, p_cyto, nrow = 4, ncol = 1, labels = c("", "Less Differentiated", "", "More Differentiated"), label_size = 18, label_x = 0.5, hjust = 0, vjust = 0.5, rel_heights = c(0.025, 1, 0.025, 1))
pdf_and_png(p_grid, output_box, "select_cd8cd4_clusters", pdfHeight = H_FULL, pdfWidth = W_FULL, scale = 3, SAVE_RDS = F)
```



# Other differential abundance testing (not used for main figures)
```{r}
baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline$progression_group, "_", baseline$davies_based_risk)

counts <- table(baseline$public_id, baseline@meta.data[, cluster_group])
counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl <- baseline@meta.data |> dplyr::distinct(public_id, progression_group, Batch, siteXbatch, Study_Site, Batch, davies_based_risk, d_pt_race_1, pfs_risk)
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(!!sym(cluster_group), isDoublet, isImmune, isNonMalignantImmune, subcluster_V03072023_compartment, lineage_group, lineage_order, cellID_short, cellID_long)
colnames(cluster_md)[1] <- "cluster"
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")
```

```{r}
# outs <- quickWilcoxAndPlots(counts |> dplyr::filter(lineage_group != "LQ"), "progression_group", "NP", "RP")
```

```{r}
# floral_fill_theme <- list(
#     "NP (SIG)" = "#7CA0D4",
#     "NP" = "#a9bad0",
#     "RP (SIG)" = "#A48AD3",
#     "RP" = "#baafcf"
# )

# floral_color_theme <- list(
#     "NP (SIG)" = "#285291",
#     "NP" = "#43618e",
#     "RP (SIG)" = "#4f2b8e",
#     "RP" = "#5f4886"
# )

# annotations <- data.frame(
#     xpos = c(-Inf, Inf),
#     ypos = c(Inf, Inf),
#     annotateText = c(
#         "RP Enriched",
#         "NP Enriched"
#     ),
#     hjustvar = c(-.05, 1),
#     vjustvar = c(1, 1)
# )

# p_wilcoxplot <- outs$plotWilcoxOut +
#     scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP*", "NP", "RP*", "RP")) +
#     scale_color_manual(values = floral_color_theme) +
#     xlab("Average Log2 Fold Change") +
#     ylab("-log10 P Value") +
#     ggtitle("Differential Abundance, All Clusters") +
#     geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

# pdf_and_png(p_wilcoxplot, output_abun_volcano, "all_cells", pdfWidth = 10 * 1.3, pdfHeight = 8 * 1.3)
```

```{r}
# write.csv(outs$wilcox_out, file.path(output_raw_csv, "all_cells.csv"))
```

# CD3+ DiffAbun
```{r}
# outs <- quickWilcoxAndPlots(counts |> dplyr::filter(lineage_group %in% c("CD8", "CD4")), "progression_group", "NP", "RP")
```

```{r}
# floral_fill_theme <- list(
#     "NP (SIG)" = "#7CA0D4",
#     "NP" = "#a9bad0",
#     "RP (SIG)" = "#A48AD3",
#     "RP" = "#baafcf"
# )

# floral_color_theme <- list(
#     "NP (SIG)" = "#285291",
#     "NP" = "#43618e",
#     "RP (SIG)" = "#4f2b8e",
#     "RP" = "#5f4886"
# )

# annotations <- data.frame(
#     xpos = c(-Inf, Inf),
#     ypos = c(Inf, Inf),
#     annotateText = c(
#         "RP Enriched",
#         "NP Enriched"
#     ),
#     hjustvar = c(-.05, 1),
#     vjustvar = c(1, 1)
# )

# p_wilcoxplot <- outs$plotWilcoxOut +
#     scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP*", "NP", "RP*", "RP")) +
#     scale_color_manual(values = floral_color_theme) +
#     xlab("Average Log2 Fold Change") +
#     ylab("-log10 P Value") +
#     ggtitle("Differential Abundance, CD3+ Cells") +
#     geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

# pdf_and_png(p_wilcoxplot, output_abun_volcano, "cd3_positive_cells", pdfWidth = 10 * 1.3, pdfHeight = 8 * 1.3)
```

# Myeloid DiffAbun
```{r}
# outs <- quickWilcoxAndPlots(counts |> dplyr::filter(lineage_group %in% c("M")), "progression_group", "NP", "RP")
```

```{r}
# floral_fill_theme <- list(
#     "NP (SIG)" = "#7CA0D4",
#     "NP" = "#a9bad0",
#     "RP (SIG)" = "#A48AD3",
#     "RP" = "#baafcf"
# )

# floral_color_theme <- list(
#     "NP (SIG)" = "#285291",
#     "NP" = "#43618e",
#     "RP (SIG)" = "#4f2b8e",
#     "RP" = "#5f4886"
# )

# annotations <- data.frame(
#     xpos = c(-Inf, Inf),
#     ypos = c(Inf, Inf),
#     annotateText = c(
#         "RP Enriched",
#         "NP Enriched"
#     ),
#     hjustvar = c(-.05, 1),
#     vjustvar = c(1, 1)
# )

# p_wilcoxplot <- outs$plotWilcoxOut +
#     scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP*", "NP", "RP*", "RP")) +
#     scale_color_manual(values = floral_color_theme) +
#     xlab("Average Log2 Fold Change") +
#     ylab("-log10 P Value") +
#     ggtitle("Differential Abundance, Myeloid") +
#     geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

# pdf_and_png(p_wilcoxplot, output_abun_volcano, "myeloid_cells", pdfWidth = 10 * 1.3, pdfHeight = 8 * 1.3)
```


# B DiffAbun
```{r}
# outs <- quickWilcoxAndPlots(counts |> dplyr::filter(lineage_group %in% c("B")), "progression_group", "NP", "RP")
```

```{r}
# floral_fill_theme <- list(
#     "NP (SIG)" = "#7CA0D4",
#     "NP" = "#a9bad0",
#     "RP (SIG)" = "#A48AD3",
#     "RP" = "#baafcf"
# )

# floral_color_theme <- list(
#     "NP (SIG)" = "#285291",
#     "NP" = "#43618e",
#     "RP (SIG)" = "#4f2b8e",
#     "RP" = "#5f4886"
# )

# annotations <- data.frame(
#     xpos = c(-Inf, Inf),
#     ypos = c(Inf, Inf),
#     annotateText = c(
#         "RP Enriched",
#         "NP Enriched"
#     ),
#     hjustvar = c(-.05, 1),
#     vjustvar = c(1, 1)
# )

# p_wilcoxplot <- outs$plotWilcoxOut +
#     scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP*", "NP", "RP*", "RP")) +
#     scale_color_manual(values = floral_color_theme) +
#     xlab("Average Log2 Fold Change") +
#     ylab("-log10 P Value") +
#     ggtitle("Differential Abundance, B Cells") +
#     geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

# pdf_and_png(p_wilcoxplot, output_abun_volcano, "B_cells", pdfWidth = 10 * 1.3, pdfHeight = 8 * 1.3)
```

# Immune
```{r}
# outs <- quickWilcoxAndPlots(counts |> dplyr::filter(lineage_group %in% c("CD8", "CD4", "M", "B", "Nk", "Other")), "progression_group", "NP", "RP")
```

```{r}
# floral_fill_theme <- list(
#     "NP (SIG)" = "#7CA0D4",
#     "NP" = "#a9bad0",
#     "RP (SIG)" = "#A48AD3",
#     "RP" = "#baafcf"
# )

# floral_color_theme <- list(
#     "NP (SIG)" = "#285291",
#     "NP" = "#43618e",
#     "RP (SIG)" = "#4f2b8e",
#     "RP" = "#5f4886"
# )

# annotations <- data.frame(
#     xpos = c(-Inf, Inf),
#     ypos = c(Inf, Inf),
#     annotateText = c(
#         "RP Enriched",
#         "NP Enriched"
#     ),
#     hjustvar = c(-.05, 1),
#     vjustvar = c(1, 1)
# )

# p_wilcoxplot <- outs$plotWilcoxOut +
#     scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP*", "NP", "RP*", "RP")) +
#     scale_color_manual(values = floral_color_theme) +
#     xlab("Average Log2 Fold Change") +
#     ylab("-log10 P Value") +
#     ggtitle("Differential Abundance, Immune Cells") +
#     geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

# pdf_and_png(p_wilcoxplot, output_abun_volcano, "immune_cells", pdfWidth = 10 * 1.3, pdfHeight = 8 * 1.3)
```

```{r}
```