---
title: "Figure 4 - Survival"
author: "William Pilcher"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
---

Computes survival curves for select CD3+ T cell markers (Figures 4L, M, N)

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Load object and selected signatures
```{r init}
library(UCell)
source("source/Figure_3/Fig3_standard_load_data.R", local = T)

nkt <- merged |> subset(subset = (lineage_group %in% c("CD8", "CD4") &
    VJ_INTERVAL == "Baseline"))

color_low <- "#2c90c3"
color_high <- "#5a221b"

output_surv <- file.path(output_F3, "Survival")
dir.create(output_surv, recursive = T, showWarnings = F)

sig <- readxl::read_xlsx(file.path("/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation", "data/resources/pan-cancer_tcell_atlas_chu_et_al.xlsx"), sheet = "Table S4", skip = 1, )
sig_cd4 <- readxl::read_xlsx(file.path("/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation", "data/resources/pan-cancer_tcell_atlas_chu_et_al.xlsx"), sheet = "Table S6", skip = 1, )

combined_naive <- unique(c(sig_cd4[[1]] |> na.omit(), sig[[1]] |> na.omit()))
combined_cyto <- unique(c(sig_cd4[[5]] |> na.omit(), sig[[5]] |> na.omit()))
combined_exh <- unique(c(sig_cd4[[3]] |> na.omit(), sig[[3]] |> na.omit()))

list_of_signatures <- list("Naive" = combined_naive, "Cytotoxic" = combined_cyto, "Exhaustion" = combined_exh)

valid_sig <- make.names(names(list_of_signatures))
```

# Signatures, Survival
UCell computes a signature score which is on a range between 0 and 1. These will be mutlipled by 100 when fitting the CoxPH model, so that the HR can be interpretted as the increase in risk per 1 UCell signature score point increase
```{r}
nkt <- AddModuleScore_UCell(nkt, features = list_of_signatures, name = "") # AddModuleScore(nkt, features = list_of_signatures, name = valid_sig)

averageMod <- nkt@meta.data %>%
    dplyr::group_by(public_id) %>%
    dplyr::summarise(
        Naive = mean(!!sym(paste0(valid_sig[[1]]))),
        Cytotoxicity = mean(!!sym(paste0(valid_sig[[2]]))),
        Exhaustion = mean(!!sym(paste0(valid_sig[[3]]))),
    )

subset_patMD <- nkt@meta.data |>
    dplyr::filter(d_lot_1_start_day < 2) |>
    dplyr::distinct(public_id, progression_group, censpfs, pfscdy, d_pt_sex, d_dx_amm_age, d_dx_amm_iss_stage, siteXbatch, d_tx_induction_cat, d_amm_tx_asct_1st)

patMD_averageMod <- dplyr::left_join(subset_patMD, averageMod, by = "public_id")
```

# Figure 4M - Naive Signature
```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)
library(survminer)

feat <- "Naive"
tmp <- patMD_averageMod

tmp[, feat] <- tmp[, feat] * 100
tmp$survVar <- tmp[, feat]
med_val <- median(tmp$survVar)

med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_naive <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = "Progression Free Survival"
    ) +
    add_censor_mark(size = 3) +
    # add_risktable() +
    theme_prism() +
    ggtitle("Chu et al.,\nNaive Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]

pval <- round(pval, digits = 4)
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)

annotations <- data.frame(
    xpos = c(-Inf),
    ypos = c(-Inf),
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_naive_cox <- p_naive + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)
p_naive_cox <- ggsurvfit_build(p_naive_cox)

pdf_and_png(p_naive_cox, output_surv, "NaiveSig_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```

# Figure 4L - Cytotoxicity Signature
```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)

feat <- "Cytotoxicity"
tmp <- patMD_averageMod

tmp[, feat] <- tmp[, feat] * 100

tmp$survVar <- tmp[, feat]
# med_val <- median(tmp$survVar)
med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_cyto <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = ""
    ) +
    add_censor_mark(size = 3) +
    # add_risktable() +
    theme_prism() +
    ggtitle("Chu et al.,\nCytotoxicity Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)
pval <- round(pval, digits = 4)

annotations <- data.frame(
    xpos = c(-Inf),
    ypos = c(-Inf),
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_cyto_cox <- p_cyto + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)
p_cyto_cox <- ggsurvfit_build(p_cyto_cox)

pdf_and_png(p_cyto_cox, output_surv, "CytoSig_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```


# Figure 4N - Exhaustion Signature
```{r}
library(ggstatsplot)
library(ggpubr)
library(forestmodel)
library(gtsummary)
library(survival)
library(ggsurvfit)

feat <- "Exhaustion"
tmp <- patMD_averageMod

tmp[, feat] <- tmp[, feat] * 100

tmp$survVar <- tmp[, feat]
# med_val <- median(tmp$survVar)
med_val <- surv_cutpoint(tmp, time = "pfscdy", event = "censpfs", variables = "survVar", minprop = 0.2)$cutpoint$cutpoint

tmp <- tmp %>% mutate(median_cutoff = case_when(
    .data[[feat]] < med_val ~ "Low",
    .data[[feat]] >= med_val ~ "High"
))




tmp$time <- tmp$pfscdy
p_exh <- survfit2(Surv(pfscdy, censpfs) ~ median_cutoff, data = tmp) %>% ggsurvfit(size = 2) +
    labs(
        x = "Days",
        y = " "
    ) +
    add_censor_mark(size = 3) +
    # add_risktable() +
    theme_prism() +
    ggtitle("Chu et al.,\nExhaustion Signature") +
    scale_color_manual(values = c("Low" = color_low, "High" = color_high))


cox_cont_sxb <- coxph(Surv(pfscdy, censpfs) ~ survVar + siteXbatch, data = tmp)

pval <- summary(cox_cont_sxb)$coefficients["survVar", "Pr(>|z|)"]
HR <- summary(cox_cont_sxb)$coefficients["survVar", "exp(coef)"]
HR <- round(HR, digits = 4)

pval <- round(pval, digits = 4)

annotations <- data.frame(
    xpos = c(-Inf),
    ypos = c(-Inf),
    annotateText = c(
        paste0("\n ", "CoxPH P.Value: ", scales::pvalue(pval), "\n ", "HR: ", HR)
    ),
    hjustvar = c(0.0),
    vjustvar = c(0)
)
p_exh_cox <- p_exh + geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("black"), size = 6)

p_exh_cox <- ggsurvfit_build(p_exh_cox)


pdf_and_png(p_exh_cox, output_surv, "ExhSig_All_CD3", pdfWidth = W_DBL / 3, pdfHeight = H_FULL, scale = 3)
```

# Combine figures 4L, 4M, and 4N, into one figure (to ensure consistent rendering sizes)
```{r}
legend_surv <- cowplot::get_legend(p_naive_cox)

p_all_sig <- cowplot::plot_grid(p_cyto_cox + NoLegend(), p_naive_cox + NoLegend(), p_exh_cox + NoLegend(), legend_surv, nrow = 1, rel_widths = c(1, 1, 1, 0.25))
pdf_and_png(p_all_sig, output_surv, "All_CD3_Sig", pdfWidth = W_FULL * 3, pdfHeight = H_FULL, scale = 3)
```

# Cut Figures

Average expression of select exhaustion markers
```{r}
tmp2 <- AverageExpression(nkt, assays = "RNA", group.by = "public_id", features = c(list_of_signatures$Exhaustion, "LAG3"))$RNA |> # c("TIGIT", "TOX", "PDCD1", "LAG3", "HAVCR2", "CTLA4"))$RNA |>
    t() |>
    as.data.frame()

tmp2$public_id <- rownames(tmp2)

tmp <- dplyr::left_join(subset_patMD, tmp2, by = "public_id")

tmp <- tmp |> dplyr::filter(progression_group %in% c("NP", "RP"))

feat <- "LAG3"

pvals_wilcox <- tmp |>
    dplyr::filter(progression_group %in% c("NP", "RP")) |>
    rstatix::wilcox_test(as.formula(paste0(feat, " ~ progression_group"))) |>
    rstatix::add_xy_position()

p_box_lag3 <- ggplot(tmp, aes(x = progression_group, y = !!sym(feat), fill = progression_group, color = progression_group)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_jitter(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    ylab(paste0("Per Patient Average Expression")) +
    ggtitle(feat) +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)

write.csv(tmp, file.path(output_surv, "per_pat_avg_LAG3"))
```

```{r}
tmp2 <- AverageExpression(nkt, assays = "RNA", group.by = "public_id", features = c(list_of_signatures$Exhaustion, "LAG3"))$RNA |> # c("TIGIT", "TOX", "PDCD1", "LAG3", "HAVCR2", "CTLA4"))$RNA |>
    t() |>
    as.data.frame()

tmp2$public_id <- rownames(tmp2)

tmp <- dplyr::left_join(subset_patMD, tmp2, by = "public_id")

tmp <- tmp |> dplyr::filter(progression_group %in% c("NP", "RP"))

feat <- "TOX"

pvals_wilcox <- tmp |>
    dplyr::filter(progression_group %in% c("NP", "RP")) |>
    rstatix::wilcox_test(as.formula(paste0(feat, " ~ progression_group"))) |>
    rstatix::add_xy_position()
p_box_tox <- ggplot(tmp, aes(x = progression_group, y = !!sym(feat), fill = progression_group, color = progression_group)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_jitter(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    ylab(bquote("Average Log-Normalized Expression Across" ~ CD3^"+" ~ "T Cells")) +
    ggtitle(feat) +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "{scales::pvalue(p)}", inherit.aes = F, hide.ns = T, size = 5)

write.csv(tmp, file.path(output_surv, "exh_per_pat_average_lognorm_expr.csv"))
```

```{r}
p_exh_comb <- p_box_tox + p_box_lag3

pdf_and_png(p_exh_comb, output_surv, "lognorm_avg_cd3_tox_lag3_violin", pdfWidth = W_FULL, pdfHeight = H_FULL, scale = 3)
```

```{r}
test <- nkt |> subset(subset = (subcluster_V03072023 == "NkT.2.1" & davies_based_risk %in% c("high_risk", "standard_risk")))

test <- AddModuleScore_UCell(test, features = list("Dysfunctional" = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2")))

testb <- DotPlot(test, features = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2"), group.by = "davies_based_risk")


tmp <- test@meta.data
pvals_wilcox <- tmp |>
    rstatix::wilcox_test(as.formula(paste0("Dysfunctional_UCell", " ~ davies_based_risk"))) |>
    rstatix::add_xy_position()

p <- ggplot(tmp, aes(x = davies_based_risk, y = Dysfunctional_UCell, fill = davies_based_risk, color = davies_based_risk)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.1) +
    ylab(paste0("Per Cell Dysfunctional Score")) +
    ggtitle("Dysfunctional Score Within CD8+ HLA T Cells") +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
```

```{r}
averageMod_dys <- test@meta.data %>%
    dplyr::group_by(public_id) %>%
    dplyr::summarise(
        DysfunctionalScore = mean(Dysfunctional_UCell)
    )

subset_patMD_dys <- test@meta.data |>
    dplyr::distinct(public_id, davies_based_risk)

patMD_averageMod_dys <- dplyr::left_join(subset_patMD_dys, averageMod_dys, by = "public_id")
```

```{r}
tmp <- patMD_averageMod_dys
pvals_wilcox <- tmp |>
    rstatix::wilcox_test(as.formula(paste0("DysfunctionalScore", " ~ davies_based_risk"))) |>
    rstatix::add_xy_position()

p <- ggplot(tmp, aes(x = davies_based_risk, y = DysfunctionalScore, fill = davies_based_risk, color = davies_based_risk)) +
    geom_violin(alpha = 0.75) +
    geom_boxplot(outlier.shape = NA, width = 0.2, alpha = 0.5) +
    geom_quasirandom(shape = 1, show.legend = F, width = 0.3, alpha = 0.75) +
    ylab(paste0("Per Cell Dysfunctional Score")) +
    ggtitle("Dysfunctional Score Within CD8+ HLA T Cells") +
    theme_prism() +
    theme(axis.title.x = element_blank()) +
    scale_fill_prism("floral") +
    scale_color_prism("floral") +
    NoLegend() +
    ggpubr::stat_pvalue_manual(pvals_wilcox, xmin = "xmin", xmax = "xmax", label = "p", inherit.aes = F, size = 5)
```

```{r}
nkt <- AddModuleScore_UCell(nkt, features = list("Dysfunctional" = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2")))

testb <- DotPlot(nkt, features = c("B3GAT1", "ZEB2", "KLRG1", "KLRK1", "TIGIT", "PDCD1", "LAG3", "CTLA4", "HAVCR2", "Dysfunctional_UCell"), group.by = "cellID_short")
```