---
title: "Figure 4 - Differential Expression, Myeloid Compartment"
author: "William Pilcher"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
---

Generates both the Volcano plot seen in Figure 4D (Myeloid CD14+ Mono Diff Expression) and the Pathway Analysis in Figure 4E (Myeloid CD14+ Mono Pathways)

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Setup
```{r init}
source("source/Figure_3/Fig3_standard_load_data.R", local = T)
output_volcano <- file.path(output_F3, "Volcano")
dir.create(output_volcano, showWarnings = F, recursive = T)
```

## Figure 4D - Differential Expression across CD14+ Monocytes
Using Limma-trend, correcting for siteXbatch
Highlighting select markers
```{r}
library(paletteer)
library(limma)

# Subset cluster of interest, get expression matrix
cluster_of_interest <- merged |> subset(subset = (lineage_group %in% c("M") & progression_group %in% c("NP", "RP") & !!sym(cluster_group) %in% c("Myeloid.0", "Myeloid.12", "Myeloid.2", "Myeloid.1", "Myeloid.8", "Myeloid.3")) & VJ_INTERVAL == "Baseline")
expr <- as.matrix(GetAssayData(cluster_of_interest, slot = "data", assay = "RNA"))

# Fit a model for NP and RP, correcting for siteXbatch
design <- model.matrix(~ 0 + progression_group + siteXbatch, data = cluster_of_interest@meta.data)
fit <- lmFit(expr, design)

# Make a contrast for NP - RP (positive - NP enriched, negative - RP enriched), and fit with limma trend = TrUE
x <- c("progression_groupRP - progression_groupNP")
contrast <- makeContrasts(contrasts = x, levels = design) # NP is positive, RP is negative
fit <- contrasts.fit(fit, contrast)
fit <- eBayes(fit, trend = TRUE, robust = TRUE)

# Get the table of fitted (adjusted) p-values, logFCs. Sort by the t statistic (effectively signed p-value). Break ties with logFC
forVolcano <- topTable(fit, coef = 1, adjust = "BH", n = Inf, sort.by = "t", resort.by = "logFC") # Top by p.value - set 'sort.by' to 'logFC' for top by logFC

# Select the top and bottom ten by t statistic (top up, top down) for labeling
nGenes <- length(rownames(forVolcano))
genelist <- rownames(forVolcano)[c(1:15, (nGenes - 14):nGenes)] # Top and bottom ten by t statistic (ties broken by logFC)
geneList <- c("CXCL8", "IL1B", "CCL5", "CCL3", "CCL4", "MMP1", "MMP3", "CCL2", "CCL8", "TGFB1", "IL6", "CXCL1")
# Volcano plot code (or heatmap)


ISG15_lab <- c("IFI6", "IFIT3", "IFI44L", "MX1", "OAS1", "ISG15")
SASP_lab <- c("CXCL8", "IL1B", "CCL5", "CCL3", "CCL4", "MMP1", "MMP3", "CCL2", "CCL8", "TGFB1", "IL6", "CXCL1")
STRESS <- c("FKBP5", "TSC22D3", "KLF6", "TXNIP")

colors_volcano <- paletteer_d("suffrager::CarolMan")
# create custom key-value pairs for different cell-types
# this can be achieved with nested ifelse statements
keyvals.color <- ifelse(
    rownames(forVolcano) %in% ISG15_lab, colors_volcano[[4]],
    ifelse(rownames(forVolcano) %in% SASP_lab, colors_volcano[[1]],
        ifelse(rownames(forVolcano) %in% STRESS, colors_volcano[[2]], "grey")
    )
)

keyvals.alpha <- ifelse(
    rownames(forVolcano) %in% ISG15_lab, 1,
    ifelse(rownames(forVolcano) %in% SASP_lab, 1,
        ifelse(rownames(forVolcano) %in% STRESS, 1, 0.5)
    )
)

keyvals.color[is.na(keyvals.color)] <- "grey"
names(keyvals.color)[keyvals.color == "grey"] <- "Other"
names(keyvals.color)[keyvals.color == colors_volcano[[1]]] <- "Inflammatory"
names(keyvals.color)[keyvals.color == colors_volcano[[4]]] <- "ISG15"
names(keyvals.color)[keyvals.color == colors_volcano[[2]]] <- "Stress"

selectLab <- rownames(forVolcano |> dplyr::filter(adj.P.Val < 10e-40 & abs(logFC) > 0.1))


labels.color <- ifelse(
    selectLab %in% ISG15_lab, colors_volcano[[4]],
    ifelse(selectLab %in% SASP_lab, colors_volcano[[1]],
        ifelse(selectLab %in% STRESS, colors_volcano[[2]], "grey")
    )
)


annotations <- data.frame(
    xpos = c(-Inf, Inf),
    ypos = c(Inf, Inf),
    annotateText = c(
        "NP Enriched",
        "RP Enriched"
    ),
    hjustvar = c(-.05, 1),
    vjustvar = c(0.9, 0.9)
)

p <- forVolcano |> EnhancedVolcano(
    lab = rownames(forVolcano),
    x = "logFC",
    y = "adj.P.Val",
    title = bquote(CD14^"+" ~ "Classical Monocytes"),
    selectLab = selectLab,
    labCol = labels.color,
    colCustom = keyvals.color,
    colAlpha = keyvals.alpha,
    caption = "",
    xlim = c(-0.5, 0.5),
    ylim = c(-5, 315),
    pCutoff = 10e-50,
    FCcutoff = 0.11,
    pointSize = 3.0,
    drawConnectors = TRUE,
    widthConnectors = 0.75,
    max.overlaps = 20,
    lengthConnectors = unit(0.007, "npc"),
    labSize = 4.0
) +
    theme_prism() +
    theme(plot.caption = element_blank(), plot.subtitle = element_blank()) +
    geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c(color_theme_binary[["NP"]], color_theme_binary[["RP"]]), size = 6)

# Alternative rendering with no select highlights
p2 <- forVolcano |> EnhancedVolcano(
    lab = rownames(forVolcano),
    x = "logFC",
    y = "adj.P.Val",
    title = "CD14+ Classical Monocytes",
    selectLab = selectLab,
    xlim = c(-0.5, 0.5),
    ylim = c(-5, 315),
    pCutoff = 10e-50,
    FCcutoff = 0.1,
    pointSize = 3.0,
    drawConnectors = TRUE,
    widthConnectors = 0.75,
    max.overlaps = 20,
    lengthConnectors = unit(0.007, "npc"),
    labSize = 4.0
) +
    theme_prism() +
    theme(plot.caption = element_blank(), plot.subtitle = element_blank()) +
    geom_text(data = annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), color = c("#4f2b8e", "#285291"), size = 6)

write.csv(forVolcano, file.path(output_volcano, "CD14_limma_markers_toptable.csv"))
pdf_and_png(p, output_volcano, "limma_trend_np_rp_baseline_AUC_ClassicalMyeloid_select_SASP", pdfWidth = W_FULL, pdfHeight = H_FULL, scale = 3)
pdf_and_png(p2, output_volcano, "limma_trend_np_rp_baseline_AUC_ClassicalMyeloid_select_SASP_DEFAULT_COLORS", pdfWidth = W_FULL, pdfHeight = H_FULL, scale = 3)
```

## Figure 4E - Computed pathways with ReactomePA

```{r}
library("org.Hs.eg.db")
library(clusterProfiler)
library(ReactomePA)

forVolcano <- forVolcano |> arrange(logFC)
gene_list <- rownames(forVolcano)
symbols <- mapIds(org.Hs.eg.db, keys = gene_list, column = "ENTREZID", keytype = "SYMBOL")

m_markers <- forVolcano |> arrange(logFC, "descending")
m_markers$gene <- rownames(m_markers)
gene_list <- rownames(m_markers)
symbols <- bitr(gene_list, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")
colnames(symbols) <- c("gene", "ENTREZID")
m_markers <- dplyr::left_join(symbols, m_markers, by = "gene") # Drops unmapped
rownames(m_markers) <- m_markers$ENTREZID

logfc_m <- m_markers[, "logFC"]
names(logfc_m) <- rownames(m_markers)
logfc_m <- logfc_m |> sort(decreasing = T)

cd8_gsea <- gsePathway(logfc_m,
    pvalueCutoff = 0.2,
    pAdjustMethod = "BH",
    verbose = FALSE
)
write.csv(cd8_gsea, file.path(output_volcano, "cd14_mono_reactome.csv"))
filt <- cd8_gsea |> dplyr::filter(p.adjust < 0.1)

write.csv(symbols, file.path(output_volcano, "gene2id.csv"))
results <- filt@result
results <- cd8_gsea@result |> arrange(NES)
results$Description_wrap <- stringr::str_wrap(results$Description, width = 45)

results$Description_wrap <- factor(results$Description_wrap, levels = results$Description_wrap)

results <- results |> dplyr::mutate(signed.p.adjust = -log10(p.adjust) * sign(NES))

p2 <- results |>
    arrange(NES) |>
    ggplot() +
    geom_bar(aes(y = Description_wrap, x = signed.p.adjust, fill = NES), stat = "identity") +
    scale_fill_gradient2(
        low =
            fill_theme_binary[["NP"]],
        mid = "white",
        high = fill_theme_binary[["RP"]],
        limits = c(-4, 4),
        oob = squish
    ) +
    geom_vline(xintercept = 0, color = "black") +
    theme_prism() +
    theme(axis.text.y = element_text(size = 10)) +
    xlab("Signed FDR") +
    ylab("Significantly Enriched Pathways") +
    labs(fill = "Normalized\nEnrichment\nScore") +
    theme(legend.title = element_text())

pdf_and_png(p2, output_volcano, "myeloid_reactomePA_pathways_test", pdfWidth = W_FULL, pdfHeight = H_FULL, scale = 3)

p2 <- results |>
    arrange(NES) |>
    ggplot() +
    geom_bar(aes(y = Description_wrap, x = NES, fill = signed.p.adjust), stat = "identity") +
    scale_fill_gradient2(
        low =
            fill_theme_binary[["NP"]],
        mid = "white",
        high = fill_theme_binary[["RP"]],
        limits = c(-4, 4),
        oob = squish
    ) +
    geom_vline(xintercept = 0, color = "black") +
    theme_prism() +
    theme(axis.text.y = element_text(size = 10)) +
    xlab("Normalized Enrichment Score") +
    ylab("Significantly Enriched Pathways") +
    labs(fill = "Sign(NES)*\nLog10(FDR)") +
    theme(legend.title = element_text())

pdf_and_png(p2, output_volcano, "myeloid_reactomePA_pathways_test_alt", pdfWidth = W_FULL, pdfHeight = H_FULL, scale = 3)
```

### Also save the leadning edge genes for each pathway
```{r}
output_leadingedge <- file.path(output_volcano, "M_LE")
dir.create(output_leadingedge, recursive = T, showWarnings = F)
for (idx in 1:nrow(results)) {
    id <- results[idx, "ID"]
    signaling <- results |> dplyr::filter(ID == id)
    LE <- signaling$core_enrichment |> stringr::str_split("/")
    filt_markers <- m_markers
    rownames(filt_markers) <- filt_markers$ENTREZID
    filt_markers <- filt_markers[LE[[1]], ]
    filename <- paste0(gsub(" ", "_", signaling$Description, fixed = T), "_LE.csv")
    filename <- paste0(gsub("/", "-slash-", filename, fixed = T), "_LE.csv")
    filename <- paste0(gsub(":", "-COLON-", filename, fixed = T), "_LE.csv")

    write.csv(filt_markers, file.path(output_leadingedge, filename))
}
```


```{r}
```