---
title: "Supplemental Figure 13/Figure 5 - Standard Risk NP v RP Differential Abundance Plots"
author: "William Pilcher"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
---

Renders the Standard Risk Only NP v RP Differential Abundance Supplemental Figure (Supplemental Figure 13)


```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Setup
```{r init}
library(ggpubr)
source("source/Figure_3/Fig3_standard_load_data.R", local = T)
source("source/Helper_Functions/COMMON_diff_abundance_functions.R", local = T)


output_expr <- file.path(output_F3, "Supplemental", "Experimental", "standard_risk_only")

output_extra <- file.path(output_F3, "Diff_Abundance")
output_box <- file.path(output_extra, "Diff_Abundance_Box")
output_raw_csv <- file.path(output_extra, "CSV")
output_abun_volcano <- file.path(output_extra, "Diff_Abun_Volcano")
dir.create(output_extra, showWarnings = F, recursive = T)
dir.create(output_box, showWarnings = F, recursive = T)
dir.create(output_raw_csv, showWarnings = F, recursive = T)
dir.create(output_abun_volcano, showWarnings = F, recursive = T)
dir.create(output_expr, showWarnings = F, recursive = T)

fill_theme_binary <- c(
    "NP.SR" = fill_NP,
    "RP.SR" = fill_RP
)

color_theme_binary <- c(
    "NP.SR" = color_NP,
    "RP.SR" = color_RP
)

fill_theme_emphasis <- c(
    "NP.SR" = fill_NP,
    "NP.SR*" = color_NP,
    "RP.SR" = fill_RP,
    "RP.SR*" = color_RP,
    "No Difference" = "white"
)

floral_fill_theme <- list(
    "NP.SR (SIG)" = fill_NP,
    "NP.SR" = color_NP,
    "RP.SR (SIG)" = fill_RP,
    "RP.SR" = color_RP
)

floral_color_theme <- list(
    "NP.SR (SIG)" = fill_NP,
    "NP.SR" = color_NP,
    "RP.SR (SIG)" = fill_RP,
    "RP.SR" = color_RP
)
```

# Differential Abundance - Setup Counts Data

```{r}
baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline$progression_group, "_", baseline$davies_based_risk)

counts <- table(baseline$public_id, baseline@meta.data[, cluster_group])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl <- baseline@meta.data |> dplyr::distinct(public_id, progression_group, Batch, siteXbatch, Study_Site, Batch, davies_based_risk, d_pt_race_1)

md_bl$pfs_risk <- paste0(md_bl$progression_group, md_bl$davies_based_risk)
md_bl <- md_bl |> dplyr::mutate(pfs_risk = dplyr::case_when(
    pfs_risk == "NPstandard_risk" ~ "NP.SR",
    pfs_risk == "RPstandard_risk" ~ "RP.SR",
    .default = pfs_risk
))
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(!!sym(cluster_group), isDoublet, subcluster_V03072023_compartment, lineage_group, lineage_order, cellID_short, cellID_long)
colnames(cluster_md)[1] <- "cluster"
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")
```

```{r}
all_cells <- new.env()

all_cells$counts <- counts |> dplyr::filter(lineage_group != "LQ")

all_cells$outs <- quickWilcoxAndPlots(all_cells$counts |> dplyr::filter(lineage_group != "LQ"), "pfs_risk", "RP.SR", "NP.SR")
all_cells$cluster_size <- all_cells$outs$counts_with_props |>
    group_by(cellID_short) |>
    dplyr::summarise(sum = sum(Counts))

all_cells$cluster_md <- cluster_md
all_cells$cluster_md$subcluster_V03072023 <- all_cells$cluster_md$cluster

all_cells$outs$wilcox_out <- dplyr::left_join(all_cells$outs$wilcox_out, all_cells$cluster_md, by = "cluster")
all_cells$outs$wilcox_out_select <- all_cells$outs$wilcox_out |> dplyr::select(cluster, mean_A, mean_B, fold_change, fold_change_for_display, p, padj, group1, group2)

all_cells$counts_plotting <- all_cells$outs$counts_with_props
all_cells$counts_plotting <- all_cells$counts_plotting |> dplyr::left_join(all_cells$cluster_size, "cellID_short")
all_cells$counts_plotting <- all_cells$counts_plotting |> dplyr::left_join(all_cells$outs$wilcox_out_select, "cluster")
```

# Wilcoxon Volcano
```{r}
wilcox_plot_env <- new.env()
wilcox_plot_env$wilcox_out <- all_cells$outs$wilcox_out
wilcox_plot_env$wilcox_out <- wilcox_plot_env$wilcox_out |> dplyr::mutate(fold_change = dplyr::case_when(
    fold_change > 2^2.5 ~ 2^2.5,
    fold_change < 2^-2.5 ~ 2^-2.5,
    .default = fold_change
))
x_max_range <- max(abs(wilcox_plot_env$wilcox_out$rank_biserial))

wilcox_plot_env$p <- ggplot(wilcox_plot_env$wilcox_out) +
    geom_point(aes(x = rank_biserial, y = -log10(p + 1e-200), fill = direction, color = direction, alpha = 0.5), show.legend = F) +
    geom_vline(xintercept = 0, color = "black", size = 1) +
    geom_hline(yintercept = -log10(0.05 + 1e-200), color = "black", size = 1) +
    geom_label_repel(aes(x = rank_biserial, y = -log10(p + 1e-200), label = cellID_short, fill = direction), max.overlaps = 15) +
    xlim(-x_max_range, x_max_range) +
    ggtitle("Differential Abundance - All Cells") +
    theme_prism()

wilcox_plot_env$annotations <- data.frame(
    xpos = c(-Inf, Inf),
    ypos = c(Inf, Inf),
    annotateText = c(
        "NP.SR Enriched",
        "RP.SR Enriched"
    ),
    hjustvar = c(-.05, 1),
    vjustvar = c(1, 1)
)

wilcox_plot_env$p <- wilcox_plot_env$p +
    scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP.SR*", "NP.SR", "RP.SR*", "RP.SR")) +
    scale_color_manual(values = floral_color_theme) +
    xlab("Rank-Biserial Correlation") +
    ylab("-log10 P Value") +
    ggtitle("Differential Abundance, All Clusters") +
    geom_text(data = wilcox_plot_env$annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)
```

# Generate the boxplots

```{r}
box_plots <- new.env()
box_plots$data <- all_cells$counts_plotting |> dplyr::filter(pfs_risk %in% c("NP.SR", "RP.SR"))

box_plots$data <- box_plots$data |> dplyr::arrange(lineage_group, desc(sum))

box_plots$panels <- unique(box_plots$data$cluster)

box_plots$plot_list <- list()
box_plots$plot_list_bar <- list()

box_plots$per_clust_max <- box_plots$data |>
    dplyr::group_by(cluster, pfs_risk) |>
    dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.5 * IQR_max, y.position.alt = upper_max + 0.2 * IQR_max)

box_plots$per_clust_max <- box_plots$per_clust_max |> dplyr::mutate(
    y.position = dplyr::case_when(
        y.position == 0 ~ 0.001,
        .default = y.position
    ),
    y.position.alt = dplyr::case_when(
        y.position.alt == 0 ~ 0.001,
        .default = y.position.alt
    )
)

box_plots$data <- box_plots$data |> dplyr::left_join(box_plots$per_clust_max, by = "cluster")


grouping_var <- "pfs_risk"
wrap_point <- 6

custom_lin_group_order <- c("CD8", "CD4", "Nk", "M", "B", "Other", "E", "P")
for (k in custom_lin_group_order) {
    box_plots$panels_select <- box_plots$data |>
        dplyr::filter(lineage_group == k) |>
        dplyr::distinct(cluster)

    box_plots$panels_select <- box_plots$panels_select[["cluster"]]

    box_plots$plot_list[[k]] <- list()

    for (i in box_plots$panels_select) {
        box_plots$tmp <- box_plots$data |> dplyr::filter(cluster == i)
        box_plots$tmp[[grouping_var]] <- factor(box_plots$tmp[[grouping_var]], levels = c("NP.SR", "RP.SR"))

        box_plots$tmp_distinct <- box_plots$tmp |> dplyr::distinct(cluster, p, y.position, y.position.alt, cluster_size, cellID_short, group1, group2)


        box_plots$plot_list[[k]][[i]] <- ggplot(box_plots$tmp) +
            geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
            geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
            stat_pvalue_manual(box_plots$tmp_distinct, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
            theme_prism() +
            ylim(c(0, box_plots$tmp_distinct$y.position)) +
            scale_fill_manual(values = fill_theme_binary) +
            scale_color_manual(values = color_theme_binary) +
            theme(axis.title.x = element_blank()) +
            ylab(bquote(atop("Proportion", "of all Cells"))) +
            NoLegend() +
            coord_cartesian(clip = "off") +
            theme(axis.title.x = element_blank()) +
            ggtitle(box_plots$tmp_distinct$cellID_short) +
            theme(plot.title = element_text(size = 14, face = "bold"))
    }
}

test <- list()
test_b <- list()
heights <- c()
heights_b <- list()
widths_b <- list()

heights_b_con <- c()
widths_b_con <- c()

for (i in 1:(4 * wrap_point)) {
    test_b[[i]] <- plot_spacer()
}
# Use the Design matrix to add gaps in the plot
for (k in custom_lin_group_order) {
    for (i in seq_along(box_plots$plot_list[[k]])) {
        if (mod(i - 1, wrap_point) != 0) {
            box_plots$plot_list[[k]][[i]] <- box_plots$plot_list[[k]][[i]] + theme(axis.title.y = element_blank())
        }
    }
    numPlots <- length(box_plots$plot_list[[k]])
    if (numPlots > 0) {
        if (mod(numPlots, wrap_point) != 0) {
            idx <- 1
            for (i in (mod(numPlots, wrap_point) + 1):wrap_point) {
                box_plots$plot_list[[k]][[numPlots + idx]] <- plot_spacer()
                idx <- idx + 1
            }
        }
        updated_size <- length(box_plots$plot_list[[k]])
        for (i in 1:6) {
            box_plots$plot_list[[k]][[updated_size + i]] <- plot_spacer()
        }
        widths_b[[k]] <- rep(1.0, length(box_plots$plot_list[[k]]))
        heights_b[[k]] <- rep(1.0, length(box_plots$plot_list[[k]]))

        widths_b_con <- c(widths_b_con, widths_b[[k]])
        heights_b_con <- c(heights_b_con, heights_b[[k]])

        test_b <- c(test_b, box_plots$plot_list[[k]])
        test[[k]] <- wrap_plots(box_plots$plot_list[[k]], ncol = wrap_point)
        heights[k] <- ceiling(length(box_plots$plot_list[[k]]) / wrap_point)
    }
}

patchwork_nrows <- ceiling(length(test_b) / wrap_point)
patchwork_area <- ceiling(sqrt(length(test_b))) # ?????????????????

# dark magic
design_mat <- patchwork::area()
base <- 0
for (i in seq_along(test_b)) {
    col <- mod(i - 1, wrap_point) + 1
    row <- ceiling(i / wrap_point)

    design_mat <- c(design_mat, area(t = base + row, b = base + row, l = col, r = col))
}

test_multi <- wrap_plots(test_b, ncol = wrap_point, heights = heights_b_con)
```

```{r}
patchwork_nrows <- ceiling(length(test_b) / wrap_point)
test_multi_with_wilcox <- test_multi + wilcox_plot_env$p + plot_layout(design = c(
    design_mat,
    area(t = 1, b = 3, l = 2, r = 5)
))

# test_multi <- wrap_plots(test, ncol = 1, heights = heights) + plot_annotation(tag_levels = list(names(test)))
# test_2 <- wilcox_plot_env$p / test_all
pdf_and_png(test_multi_with_wilcox, output_expr, "wrapped_cd8_test", pdfWidth = 3 * W_FULL, pdfHeight = H_FULL * patchwork_nrows / 2, scale = 3)

test_multi_with_wilcox_free <- test_multi + free(wilcox_plot_env$p) + plot_layout(design = c(
    design_mat,
    area(t = 1, b = 3, l = 2, r = 5)
))

pdf_and_png(test_multi_with_wilcox_free, output_expr, "wrapped_cd8_test_free_wilcox", pdfWidth = 3 * W_FULL, pdfHeight = H_FULL * sum(heights) / 2, scale = 3)
```

# Supplemental Figure 13

Generate box plots for various subcompartments, and compile into large figures

Supplemental Figure 13 - Non-Malignant

```{r}
baseline <- merged |> subset(VJ_INTERVAL == "Baseline")
Idents(baseline) <- baseline@meta.data[, cluster_group]
#  baseline$pfs_risk <- paste0(baseline$pfs_risk, "_", baseline$davies_based_risk)

counts <- table(baseline$public_id, baseline@meta.data[, cluster_group])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl <- baseline@meta.data |> dplyr::distinct(public_id, progression_group, Batch, siteXbatch, Study_Site, Batch, davies_based_risk, d_pt_race_1)

md_bl$pfs_risk <- paste0(md_bl$progression_group, md_bl$davies_based_risk)
md_bl <- md_bl |> dplyr::mutate(pfs_risk = dplyr::case_when(
    pfs_risk == "NPstandard_risk" ~ "NP.SR",
    pfs_risk == "RPstandard_risk" ~ "RP.SR",
    .default = pfs_risk
))
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(!!sym(cluster_group), isDoublet, subcluster_V03072023_compartment, lineage_group, lineage_order, cellID_short, cellID_long)
colnames(cluster_md)[1] <- "cluster"
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")

types_list <- list(
    "CD3+" = c("CD8", "CD4"),
    "Myeloid" = "M",
    "B" = "B",
    "CD8+" = "CD8",
    "CD4+" = "CD4",
    "Nk" = "Nk",
    "Immune" = c("CD8", "CD4", "M", "B", "Other", "Nk"),
    "Non-Malignant" = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    "All" = c("CD8", "CD4", "M", "B", "Other", "Nk", "P", "E")
)
file_name_list <- list(
    "CD3+" = "cd3_only",
    "Myeloid" = "myeloid_only",
    "B" = "b_only",
    "CD8+" = "cd8_only",
    "CD4+" = "cd4_only",
    "Nk" = "nk_only",
    "Immune" = "immune_only",
    "Non-Malignant" = "all_clusters_no_plasma",
    "All" = "all_cells"
)

for (idx_name in names(types_list)) {
    all_cells <- new.env()

    all_cells$counts <- counts |> dplyr::filter(lineage_group %in% types_list[[idx_name]])
    all_cells$outs <- quickWilcoxAndPlots(all_cells$counts, "pfs_risk", "RP.SR", "NP.SR")
    all_cells$cluster_size <- all_cells$outs$counts_with_props |>
        group_by(cellID_short) |>
        dplyr::summarise(sum = sum(Counts))

    all_cells$cluster_md <- cluster_md
    all_cells$cluster_md$subcluster_V03072023 <- all_cells$cluster_md$cluster

    all_cells$outs$wilcox_out <- dplyr::left_join(all_cells$outs$wilcox_out, all_cells$cluster_md, by = "cluster")
    all_cells$outs$wilcox_out_select <- all_cells$outs$wilcox_out |> dplyr::select(cluster, mean_A, mean_B, fold_change, fold_change_for_display, p, padj, group1, group2)

    all_cells$counts_plotting <- all_cells$outs$counts_with_props
    all_cells$counts_plotting <- all_cells$counts_plotting |> dplyr::left_join(all_cells$cluster_size, "cellID_short")
    all_cells$counts_plotting <- all_cells$counts_plotting |> dplyr::left_join(all_cells$outs$wilcox_out_select, "cluster")


    wilcox_plot_env <- new.env()
    wilcox_plot_env$wilcox_out <- all_cells$outs$wilcox_out
    wilcox_plot_env$wilcox_out <- wilcox_plot_env$wilcox_out |> dplyr::mutate(fold_change = dplyr::case_when(
        fold_change > 2^2.5 ~ 2^2.5,
        fold_change < 2^-2.5 ~ 2^-2.5,
        .default = fold_change
    ))
    wilcox_plot_env$wilcox_out$direction <- factor(wilcox_plot_env$wilcox_out$direction, levels = c("NP.SR (SIG)", "NP.SR", "RP.SR (SIG)", "RP.SR"))

    x_max_range <- max(abs(wilcox_plot_env$wilcox_out$rank_biserial))

    wilcox_plot_env$p <- ggplot(wilcox_plot_env$wilcox_out) +
        geom_point(aes(x = rank_biserial, y = -log10(p + 1e-200), fill = direction, color = direction, alpha = 0.5), show.legend = F) +
        geom_vline(xintercept = 0, color = "black", size = 1) +
        geom_hline(yintercept = -log10(0.05 + 1e-200), color = "black", size = 1) +
        geom_label_repel(aes(x = rank_biserial, y = -log10(p + 1e-200), label = cellID_short, fill = direction), color = "white", max.overlaps = 15, segment.colour = "black") +
        xlim(-x_max_range, x_max_range) +
        ggtitle(paste0("Differential Abundance, ", idx_name)) +
        theme_prism()

    wilcox_plot_env$annotations <- data.frame(
        xpos = c(-Inf, Inf),
        ypos = c(Inf, Inf),
        annotateText = c(
            "NP.SR Enriched",
            "RP.SR Enriched"
        ),
        hjustvar = c(-.05, 1),
        vjustvar = c(1, 1)
    )

    wilcox_plot_env$p <- wilcox_plot_env$p +
        scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP.SR*", "NP.SR", "RP.SR*", "RP.SR")) +
        scale_color_manual(values = floral_color_theme) +
        xlab("Rank-Biserial Correlation") +
        ylab("-log10 P Value") +
        ggtitle(paste0("Differential Abundance, ", idx_name)) +
        geom_text(data = wilcox_plot_env$annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

    box_plots <- new.env()
    box_plots$data <- all_cells$counts_plotting |> dplyr::filter(pfs_risk %in% c("NP.SR", "RP.SR"))

    box_plots$data <- box_plots$data |> dplyr::arrange(lineage_group, desc(sum))

    box_plots$panels <- unique(box_plots$data$cluster)

    box_plots$plot_list <- list()
    box_plots$plot_list_bar <- list()

    box_plots$per_clust_max <- box_plots$data |>
        dplyr::group_by(cluster, pfs_risk) |>
        dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
        dplyr::group_by(cluster) |>
        dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.5 * IQR_max, y.position.alt = upper_max + 0.2 * IQR_max)

    box_plots$per_clust_max <- box_plots$per_clust_max |> dplyr::mutate(
        y.position = dplyr::case_when(
            y.position == 0 ~ 0.001,
            .default = y.position
        ),
        y.position.alt = dplyr::case_when(
            y.position.alt == 0 ~ 0.001,
            .default = y.position.alt
        )
    )

    box_plots$data <- box_plots$data |> dplyr::left_join(box_plots$per_clust_max, by = "cluster")


    grouping_var <- "pfs_risk"
    wrap_point <- 6

    custom_lin_group_order <- unique(all_cells$counts$lineage_group)
    for (k in types_list[[idx_name]]) {
        box_plots$panels_select <- box_plots$data |>
            dplyr::filter(lineage_group == k) |>
            dplyr::distinct(cluster)

        box_plots$panels_select <- box_plots$panels_select[["cluster"]]

        box_plots$plot_list[[k]] <- list()

        for (i in box_plots$panels_select) {
            box_plots$tmp <- box_plots$data |> dplyr::filter(cluster == i)
            box_plots$tmp[[grouping_var]] <- factor(box_plots$tmp[[grouping_var]], levels = c("NP.SR", "RP.SR"))

            box_plots$tmp_distinct <- box_plots$tmp |> dplyr::distinct(cluster, p, y.position, y.position.alt, cluster_size, cellID_short, group1, group2)

            box_plots$plot_list[[k]][[i]] <- ggplot(box_plots$tmp) +
                geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
                geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
                stat_pvalue_manual(box_plots$tmp_distinct, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
                theme_prism() +
                ylim(c(0, box_plots$tmp_distinct$y.position)) +
                scale_fill_manual(values = fill_theme_binary) +
                scale_color_manual(values = color_theme_binary) +
                theme(axis.title.x = element_blank()) +
                ylab(bquote(atop("Proportion", "of " ~ .(idx_name) ~ " Cells"))) +
                NoLegend() +
                coord_cartesian(clip = "off") +
                theme(axis.title.x = element_blank()) +
                ggtitle(box_plots$tmp_distinct$cellID_short) +
                theme(plot.title = element_text(size = 14, face = "bold"))
        }
    }

    test <- list()
    test_b <- list()
    heights <- c()
    heights_b <- list()
    widths_b <- list()

    heights_b_con <- c()
    widths_b_con <- c()

    for (i in 1:(4 * wrap_point)) {
        test_b[[i]] <- plot_spacer()
    }
    # Use the Design matrix to add gaps in the plot
    for (k in types_list[[idx_name]]) {
        for (i in seq_along(box_plots$plot_list[[k]])) {
            if (mod(i - 1, wrap_point) != 0) {
                box_plots$plot_list[[k]][[i]] <- box_plots$plot_list[[k]][[i]] + theme(axis.title.y = element_blank())
            }
        }
        numPlots <- length(box_plots$plot_list[[k]])
        if (numPlots > 0) {
            if (mod(numPlots, wrap_point) != 0) {
                idx <- 1
                for (i in (mod(numPlots, wrap_point) + 1):wrap_point) {
                    box_plots$plot_list[[k]][[numPlots + idx]] <- plot_spacer()
                    idx <- idx + 1
                }
            }
            updated_size <- length(box_plots$plot_list[[k]])
            for (i in 1:6) {
                box_plots$plot_list[[k]][[updated_size + i]] <- plot_spacer()
            }
            widths_b[[k]] <- rep(1.0, length(box_plots$plot_list[[k]]))
            heights_b[[k]] <- rep(1.0, length(box_plots$plot_list[[k]]))

            widths_b_con <- c(widths_b_con, widths_b[[k]])
            heights_b_con <- c(heights_b_con, heights_b[[k]])

            test_b <- c(test_b, box_plots$plot_list[[k]])
            test[[k]] <- wrap_plots(box_plots$plot_list[[k]], ncol = wrap_point)
            heights[k] <- ceiling(length(box_plots$plot_list[[k]]) / wrap_point)
        }
    }

    patchwork_nrows <- ceiling(length(test_b) / wrap_point)
    patchwork_area <- ceiling(sqrt(length(test_b))) # ?????????????????

    # dark magic
    design_mat <- patchwork::area()
    base <- 0
    for (i in seq_along(test_b)) {
        col <- mod(i - 1, wrap_point) + 1
        row <- ceiling(i / wrap_point)

        design_mat <- c(design_mat, area(t = base + row, b = base + row, l = col, r = col))
    }

    test_multi <- wrap_plots(test_b, ncol = wrap_point, heights = heights_b_con)

    patchwork_nrows <- ceiling(length(test_b) / wrap_point)
    test_multi_with_wilcox <- test_multi + wilcox_plot_env$p + plot_layout(design = c(
        design_mat,
        area(t = 1, b = 3, l = 2, r = 5)
    ))

    # test_multi <- wrap_plots(test, ncol = 1, heights = heights) + plot_annotation(tag_levels = list(names(test)))
    # test_2 <- wilcox_plot_env$p / test_all
    pdf_and_png(test_multi_with_wilcox, output_expr, "wrapped_cd8_test", pdfWidth = 3 * W_FULL, pdfHeight = H_FULL * patchwork_nrows / 2, scale = 3)

    test_multi_with_wilcox_free <- test_multi + free(wilcox_plot_env$p) + plot_layout(design = c(
        design_mat,
        area(t = 1, b = 3, l = 2, r = 5)
    ))

    pdf_and_png(test_multi_with_wilcox_free, output_expr, paste0("wrapped_free_", file_name_list[[idx_name]], ".pdf"), pdfWidth = 3 * W_FULL, pdfHeight = H_FULL * row / 2, scale = 3)
}
```



### Major CellTypes

```{r}
Idents(baseline) <- baseline@meta.data[, "lineage_group"]
#  baseline$pfs_risk <- paste0(baseline$pfs_risk, "_", baseline$davies_based_risk)

counts <- table(baseline$public_id, baseline@meta.data[, "lineage_group"])

counts <- counts[, colSums(counts) != 0] # filter clusters with no cells... will show up because cluster vector is a factor
cluster_size <- colSums(counts)
counts <- counts |> as.data.frame()
colnames(counts) <- c("public_id", "cluster", "Counts")
md_bl <- baseline@meta.data |> dplyr::distinct(public_id, progression_group, Batch, siteXbatch, Study_Site, Batch, davies_based_risk, d_pt_race_1)
md_bl$pfs_risk <- paste0(md_bl$progression_group, md_bl$davies_based_risk)
md_bl <- md_bl |> dplyr::mutate(pfs_risk = dplyr::case_when(
    pfs_risk == "NPstandard_risk" ~ "NP.SR",
    pfs_risk == "RPstandard_risk" ~ "RP.SR",
    .default = pfs_risk
))
counts <- counts |> dplyr::left_join(md_bl, by = "public_id")

cluster_md <- baseline@meta.data |> dplyr::distinct(lineage_group)
cluster_md$cluster <- cluster_md$lineage_group
counts <- counts |> dplyr::left_join(cluster_md, by = "cluster")

types_list <- list(
    "CD3+ Cells" = c("CD8", "CD4"),
    "Immune" = c("CD8", "CD4", "M", "B", "Other", "Nk"),
    "Non-Malignant" = c("CD8", "CD4", "M", "B", "Other", "Nk", "E"),
    "All" = c("CD8", "CD4", "M", "B", "Other", "Nk", "P", "E")
)
file_name_list <- list(
    "CD3+ Cells" = "LINEAGE_GROUP_cd3_only",
    "Immune" = "LINEAGE_GROUP_immune_only",
    "Non-Malignant" = "LINEAGE_GROUP_All_NoPlasma",
    "All" = "LINEAGE_GROUP_all_cells"
)

for (idx_name in names(types_list)) {
    all_cells <- new.env()

    all_cells$counts <- counts |> dplyr::filter(lineage_group %in% types_list[[idx_name]])
    all_cells$outs <- quickWilcoxAndPlots(all_cells$counts, "pfs_risk", "RP.SR", "NP.SR")
    all_cells$cluster_size <- all_cells$outs$counts_with_props |>
        group_by(lineage_group) |>
        dplyr::summarise(sum = sum(Counts))

    all_cells$cluster_md <- cluster_md
    all_cells$cluster_md$subcluster_V03072023 <- all_cells$cluster_md$cluster

    all_cells$outs$wilcox_out <- dplyr::left_join(all_cells$outs$wilcox_out, all_cells$cluster_md, by = "cluster")
    all_cells$outs$wilcox_out_select <- all_cells$outs$wilcox_out |> dplyr::select(cluster, mean_A, mean_B, fold_change, fold_change_for_display, p, padj, group1, group2)

    all_cells$counts_plotting <- all_cells$outs$counts_with_props
    all_cells$counts_plotting <- all_cells$counts_plotting |> dplyr::left_join(all_cells$cluster_size, "lineage_group")
    all_cells$counts_plotting <- all_cells$counts_plotting |> dplyr::left_join(all_cells$outs$wilcox_out_select, "cluster")


    wilcox_plot_env <- new.env()
    wilcox_plot_env$wilcox_out <- all_cells$outs$wilcox_out
    wilcox_plot_env$wilcox_out <- wilcox_plot_env$wilcox_out |> dplyr::mutate(fold_change = dplyr::case_when(
        fold_change > 2^2.5 ~ 2^2.5,
        fold_change < 2^-2.5 ~ 2^-2.5,
        .default = fold_change
    ))
    wilcox_plot_env$wilcox_out$direction <- factor(wilcox_plot_env$wilcox_out$direction, levels = c("NP.SR (SIG)", "NP.SR", "RP.SR (SIG)", "RP.SR"))
    x_max_range <- max(abs(wilcox_plot_env$wilcox_out$rank_biserial))

    wilcox_plot_env$p <- ggplot(wilcox_plot_env$wilcox_out) +
        geom_point(aes(x = rank_biserial, y = -log10(p + 1e-200), fill = direction, color = direction, alpha = 0.5), show.legend = F) +
        geom_vline(xintercept = 0, color = "black", size = 1) +
        geom_hline(yintercept = -log10(0.05 + 1e-200), color = "black", size = 1) +
        geom_label_repel(aes(x = rank_biserial, y = -log10(p + 1e-200), label = lineage_group, fill = direction), color = "white", max.overlaps = 15, segment.colour = "black") +
        xlim(-x_max_range, x_max_range) +
        ggtitle(paste0("Differential Abundance, ", idx_name)) +
        theme_prism()

    wilcox_plot_env$annotations <- data.frame(
        xpos = c(-Inf, Inf),
        ypos = c(Inf, Inf),
        annotateText = c(
            "NP.SR Enriched",
            "RP.SR Enriched"
        ),
        hjustvar = c(-.05, 1),
        vjustvar = c(1, 1)
    )

    wilcox_plot_env$p <- wilcox_plot_env$p +
        scale_fill_manual(values = floral_fill_theme, breaks = names(floral_fill_theme), labels = c("NP.SR*", "NP.SR", "RP.SR*", "RP.SR")) +
        scale_color_manual(values = floral_color_theme) +
        xlab("Rank-Biserial Correlation") +
        ylab("-log10 P Value") +
        ggtitle(paste0("Differential Abundance, ", idx_name)) +
        geom_text(data = wilcox_plot_env$annotations, aes(x = xpos, y = ypos, hjust = hjustvar, vjust = vjustvar, label = annotateText, fontface = "bold"), size = 6)

    box_plots <- new.env()
    box_plots$data <- all_cells$counts_plotting |> dplyr::filter(pfs_risk %in% c("NP.SR", "RP.SR"))

    box_plots$data <- box_plots$data |> dplyr::arrange(lineage_group, desc(sum))

    box_plots$panels <- unique(box_plots$data$cluster)

    box_plots$plot_list <- list()
    box_plots$plot_list_bar <- list()

    box_plots$per_clust_max <- box_plots$data |>
        dplyr::group_by(cluster, pfs_risk) |>
        dplyr::summarise(Q1 = quantile(prop, 0.25), Q3 = quantile(prop, 0.75), IQR = Q3 - Q1, upper = min(max(prop), Q3 + 1.5 * IQR), lower = max(min(prop), Q1 - 1.5 * IQR)) |>
        dplyr::group_by(cluster) |>
        dplyr::summarise(lower_min = min(lower), upper_max = max(upper), IQR_max = max(IQR), y.position = upper_max + 0.5 * IQR_max, y.position.alt = upper_max + 0.2 * IQR_max)

    box_plots$per_clust_max <- box_plots$per_clust_max |> dplyr::mutate(
        y.position = dplyr::case_when(
            y.position == 0 ~ 0.001,
            .default = y.position
        ),
        y.position.alt = dplyr::case_when(
            y.position.alt == 0 ~ 0.001,
            .default = y.position.alt
        )
    )

    box_plots$data <- box_plots$data |> dplyr::left_join(box_plots$per_clust_max, by = "cluster")


    grouping_var <- "pfs_risk"
    wrap_point <- 6

    custom_lin_group_order <- c("ALL CLUSTERS") # unique(all_cells$counts$lineage_group)
    for (k in custom_lin_group_order) {
        box_plots$panels_select <- box_plots$data |>
            # dplyr::filter(lineage_group == k) |>
            dplyr::distinct(cluster)

        box_plots$panels_select <- box_plots$panels_select[["cluster"]]

        box_plots$plot_list[[k]] <- list()

        for (i in box_plots$panels_select) {
            box_plots$tmp <- box_plots$data |> dplyr::filter(cluster == i)
            box_plots$tmp[[grouping_var]] <- factor(box_plots$tmp[[grouping_var]], levels = c("NP.SR", "RP.SR"))

            box_plots$tmp_distinct <- box_plots$tmp |> dplyr::distinct(cluster, p, y.position, y.position.alt, cluster_size, lineage_group, group1, group2)


            box_plots$plot_list[[k]][[i]] <- ggplot(box_plots$tmp) +
                geom_boxplot(outlier.shape = NA, alpha = 1, width = 0.4, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
                geom_quasirandom(shape = 1, show.legend = F, alpha = 0.5, aes(x = !!sym(grouping_var), fill = !!sym(grouping_var), color = !!sym(grouping_var), group = !!sym(grouping_var), y = prop)) +
                stat_pvalue_manual(box_plots$tmp_distinct, label = "{scales::pvalue(p)}", size = 5, y.position = "y.position.alt", hide.ns = T) +
                theme_prism() +
                ylim(c(0, box_plots$tmp_distinct$y.position)) +
                scale_fill_manual(values = fill_theme_binary) +
                scale_color_manual(values = color_theme_binary) +
                theme(axis.title.x = element_blank()) +
                ylab(bquote(atop("Proportion", "of " ~ .(idx_name) ~ " Cells"))) +
                NoLegend() +
                coord_cartesian(clip = "off") +
                theme(axis.title.x = element_blank()) +
                ggtitle(box_plots$tmp_distinct$lineage_group) +
                theme(plot.title = element_text(size = 14, face = "bold"))
        }
    }

    test <- list()
    test_b <- list()
    heights <- c()
    heights_b <- list()
    widths_b <- list()

    heights_b_con <- c()
    widths_b_con <- c()

    for (i in 1:(4 * wrap_point)) {
        test_b[[i]] <- plot_spacer()
    }
    # Use the Design matrix to add gaps in the plot
    for (k in custom_lin_group_order) {
        for (i in seq_along(box_plots$plot_list[[k]])) {
            if (mod(i - 1, wrap_point) != 0) {
                box_plots$plot_list[[k]][[i]] <- box_plots$plot_list[[k]][[i]] + theme(axis.title.y = element_blank())
            }
        }
        numPlots <- length(box_plots$plot_list[[k]])
        if (numPlots > 0) {
            if (mod(numPlots, wrap_point) != 0) {
                idx <- 1
                for (i in (mod(numPlots, wrap_point) + 1):wrap_point) {
                    box_plots$plot_list[[k]][[numPlots + idx]] <- plot_spacer()
                    idx <- idx + 1
                }
            }
            updated_size <- length(box_plots$plot_list[[k]])
            for (i in 1:6) {
                box_plots$plot_list[[k]][[updated_size + i]] <- plot_spacer()
            }
            widths_b[[k]] <- rep(1.0, length(box_plots$plot_list[[k]]))
            heights_b[[k]] <- rep(1.0, length(box_plots$plot_list[[k]]))

            widths_b_con <- c(widths_b_con, widths_b[[k]])
            heights_b_con <- c(heights_b_con, heights_b[[k]])

            test_b <- c(test_b, box_plots$plot_list[[k]])
            test[[k]] <- wrap_plots(box_plots$plot_list[[k]], ncol = wrap_point)
            heights[k] <- ceiling(length(box_plots$plot_list[[k]]) / wrap_point)
        }
    }

    patchwork_nrows <- ceiling(length(test_b) / wrap_point)
    patchwork_area <- ceiling(sqrt(length(test_b))) # ?????????????????

    # dark magic
    design_mat <- patchwork::area()
    base <- 0
    for (i in seq_along(test_b)) {
        col <- mod(i - 1, wrap_point) + 1
        row <- ceiling(i / wrap_point)

        design_mat <- c(design_mat, area(t = base + row, b = base + row, l = col, r = col))
    }

    test_multi <- wrap_plots(test_b, ncol = wrap_point, heights = heights_b_con)

    patchwork_nrows <- ceiling(length(test_b) / wrap_point)
    test_multi_with_wilcox <- test_multi + wilcox_plot_env$p + plot_layout(design = c(
        design_mat,
        area(t = 1, b = 3, l = 2, r = 5)
    ))

    # test_multi <- wrap_plots(test, ncol = 1, heights = heights) + plot_annotation(tag_levels = list(names(test)))
    # test_2 <- wilcox_plot_env$p / test_all
    pdf_and_png(test_multi_with_wilcox, output_expr, "wrapped_cd8_test", pdfWidth = 3 * W_FULL, pdfHeight = H_FULL * patchwork_nrows / 2, scale = 3)

    test_multi_with_wilcox_free <- test_multi + free(wilcox_plot_env$p) + plot_layout(design = c(
        design_mat,
        area(t = 1, b = 3, l = 2, r = 5)
    ))

    pdf_and_png(test_multi_with_wilcox_free, output_expr, paste0("wrapped_free_", file_name_list[[idx_name]], ".pdf"), pdfWidth = 3 * W_FULL, pdfHeight = H_FULL * row / 2, scale = 3)
}
```


## Setup Categories 

# Flag CD45 Positive/Immune Populations
```{r}

```
