---
title: "Slingshot with CD8 T Cells"
author: "William Pilcher"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
---

For computing the CD8 Trajectory Object

Note: This file is a bit older - contains some code blocks/annotations which are not final. See visualization codes for the up-to-date code.

# Setup
```{r init}
library(kableExtra)
library(Seurat)
library(future)

source("source/Figure_3/Fig3_standard_load_data.R", local = T)


output <- file.path("/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation/output", "monocle3_cd8_LQ_included")
dir.create(output)
# dataPath <- "/opt/megaseq-data/wcpilch/MMRF-ImmuneAtlas-Analysis-Sync"

# objectPath <- "data/INTEGRATED_OBJECTS_MMRF/Integration/SeuratObj_rm38_44_in_361_samples_Human_Ref_SM_CB_LogNorm_PC25_Harmony_singleR_doublet.rds"
# aliquotMetadata <- "/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation/data/metadata/616_metadata_by_sampleid.rds"
# SKIP_SETUP <- T
# LOAD_PATH <- "/opt/localdata/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation/output/monocle3_cd8/slingshot_cd8_NEWMD_25/sce_slingshot_cd8_NEWMD_25.rds"

# clinicalMD <- read.csv(clinicalMetadataPath, header = T)
# rownames(clinicalMD) <- clinicalMD$aliquot_id
# patMetadata <- "/opt/localdata/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation/data/metadata/pat_metadata_5_3_23_WP_slim.rds"
# visitMetadata <- "/opt/localdata/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation/data/metadata/visit_metadata_5_1_23_WP.rds"
# aliquotMetadata <- "/opt/localdata/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation/data/metadata/aliquot_metadata_3_10_23_WP.rds"

# clusterMetadata <- "/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation/cluster_metadata_WP_2_10_23.rds"

# allDimReducs <- "/opt/pmedseq-data/wcpilch/Projects/MMRF-ImmuneAtlas-Annotation/data/metadata/subumap_dimreduc.rds"
```

# Load 
```{r load}
if (!SKIP_SETUP) {
    # merged <- readRDS(file.path(dataPath, objectPath))
    # merged[["umap.sub"]] <- readRDS(allDimReducs) # Adds umap embeddings for all subclusters for display
    # # patMD <- readRDS(patMetadata)
    # # visitMD <- readRDS(visitMetadata)
    # aliquotMD <- readRDS(aliquotMetadata)
    # clusterMD <- readRDS(clusterMetadata)


    # merged_md_bckup <- merged@meta.data
    # aliquotMD_orig <- aliquotMD
    # patMD_orig <- patMD
    # visitMD_orig <- visitMD
}
```

# Merging metadata objects 

See metadata file
```{r sample_id_rename}
if (!SKIP_SETUP) {
    # merged@meta.data <- merged_md_bckup
    # aliquotMD <- aliquotMD_orig
    # # patMD <- patMD_orig
    # # visitMD <- visitMD_orig
    # merged@meta.data <- dplyr::select(merged@meta.data, -c("public_id"))

    # rownames(aliquotMD) <- aliquotMD$aliquot_id
    # aliquotMD <- dplyr::select(aliquotMD, -X, -sample_id, -site, -batch_num, -siteXbatch)

    # merged <- AddMetaData(merged, clusterMD)

    # merged@meta.data$aliquot_id <- paste0("MMRFA-", vapply(strsplit(merged@meta.data$sample_id, "_", fixed = TRUE), "[", "", 2))
    # merged@meta.data <- rownames_to_column(merged@meta.data, "cellname")


    # # visitMD$immune_analysis_id <- visitMD$visit_id
    # merged@meta.data <- left_join(merged@meta.data, aliquotMD, by = "aliquot_id")
    # # merged@meta.data <- left_join(merged@meta.data, patMD, by = "public_id")
    # # merged@meta.data <- left_join(merged@meta.data, visitMD, by = "immune_analysis_id")
    # rownames(merged@meta.data) <- merged@meta.data$cellname
}
```

# Setup Variables of Interest

## Clinical group variable, cluster group variable, and triplet/doublet variable.
```{r doubletFilter}
clinical_group <- "progression_group"

cluster_group <- "subcluster_V03072023" #
if (!SKIP_SETUP) {
    merged@meta.data$seurat_cluster_subcluster <- merged@meta.data[, cluster_group]
    merged@meta.data <- merged@meta.data %>% mutate(treatment_simple = dplyr::case_when(
        d_tx_induction_cat %in% c("imid_pi_steroid", "chemo_imid_pi_steroid") ~ "Triplet",
        .default = "Doublet"
    ))
}
```

## Flag Doublets 
```{r doublet-flag}
doublet_clusters <- c("NkT.7", "NkT.15", "NkT.16", "Myeloid.7", "Myeloid.9", "Myeloid.14", "BEry.5", "BEry.8", "BEry.13", "BEry.14", "Ery.7", "Ery.9", "Ery.10", "Ery.11", "Plasma.3", "Plasma.9", "Plasma.10", "Plasma.11")
if (!SKIP_SETUP) {
    merged@meta.data <- merged@meta.data %>% mutate(isDoublet = dplyr::case_when(
        .data[[cluster_group]] %in% doublet_clusters ~ "Doublet",
        .default = "Singlet"
    ))
}
```



## Flag Major Populations
```{r cell-type-flag}
if (!SKIP_SETUP) {
    labelstmp <- c(
        `BEry.0` = "Erythroblast",
        `BEry.1` = "Naive B",
        `BEry.2` = "Pre B",
        `BEry.3` = "Memory B",
        `BEry.4` = "Erythroblast",
        `BEry.5` = "CMP/E-M Doublet",
        `BEry.6` = "Naive B",
        `BEry.7` = "CD34 Progen",
        `BEry.8` = "T-B Doublet",
        `BEry.9` = "Memory B",
        `BEry.10` = "Pro B",
        `BEry.11` = "pDC",
        `BEry.12` = "Pro B",
        `BEry.13` = "Plasma-E Doublet",
        `BEry.14` = "T-Ery Doublet",
        `BEry.15.0` = "Platelet/MK",
        `BEry.15.1` = "Mast Cell",
        `BEry.16` = "Pro B",
        `BEry.17` = "Erythrocyte",
        `Ery.0` = "Erythrocyte",
        `Ery.1` = "Erythrocyte",
        `Ery.2` = "Erythrocyte",
        `Ery.3` = "Erythrocyte",
        `Ery.4` = "Erythrocyte",
        `Ery.5` = "Erythrocyte",
        `Ery.6` = "Erythrocyte",
        `Ery.7` = "Erythrocyte",
        `Ery.8` = "Erythrocyte",
        `Ery.9` = "Erythrocyte",
        `Ery.10` = "Erythrocyte",
        `Ery.11` = "Erythrocyte",
        `Full.23` = "Stromal Cell",
        `Myeloid.0` = "CD14 Monocyte",
        `Myeloid.1` = "CD14 Monocyte (CD68)",
        `Myeloid.2` = "CD14 Monocyte ISG (CD68)",
        `Myeloid.3` = "CD14 Monocyte Stressed (CD68)",
        `Myeloid.4` = "CD16 Monocyte",
        `Myeloid.5` = "GMP (Proliferating)",
        `Myeloid.6` = "Neutrophil",
        `Myeloid.7` = "M-T Doublet",
        `Myeloid.8` = "CD14+ Monocyte Active (TNF)",
        `Myeloid.9` = "M-Plasma/Ery Doublet",
        `Myeloid.10` = "Neutrophil",
        `Myeloid.11.0` = "CLEC10A DC",
        `Myeloid.11.1` = "CLEC9A DC",
        `Myeloid.11.2` = "DC-T Misc",
        `Myeloid.12` = "CD14 Monocyte, IL2RA",
        `Myeloid.13` = "MDSC",
        `Myeloid.14` = "M-B Doublet",
        `Myeloid.15` = "M2 Macrophage",
        `NkT.0` = "CD4 Naive",
        `NkT.1.0` = "CD4 Memory (Naive/CM)",
        `NkT.1.1` = "CD4 Memory (Effector)",
        `NkT.1.2` = "CD4 Memory (TNFAIP3/CXCR3)",
        `NkT.1.3` = "CD8 Central Memory (Naive)",
        `NkT.1.4` = "CD4 Memory ISG",
        `NkT.1.5` = "CD4 Memory",
        `NkT.2.0` = "CD8 Terminal Effector",
        `NkT.2.1` = "CD8 Effector",
        `NkT.2.2` = "CD4 Effector",
        `NkT.2.3` = "CD8 Terminal Effector",
        `NkT.2.4` = "CD8 Terminal Effector",
        `NkT.3.0` = "CD8 Effector Memory",
        `NkT.3.1` = "CD8 Central Memory",
        `NkT.3.2` = "CD8 CXCR4 EM",
        `NkT.4` = "CD56 Dim Nk",
        `NkT.5.0` = "NkT",
        `NkT.5.1` = "CD56 Dim Nk",
        `NkT.5.2` = "CD56 Dim Nk",
        `NkT.6` = "CD8 Naive",
        `NkT.7` = "CD4 LQ",
        `NkT.8` = "CD4 Regulatory",
        `NkT.9.0` = "CD56 Bright Nk (SELL+)",
        `NkT.9.1` = "CD56 Bright Nk (CD160+)",
        `NkT.10.0` = "CD8 ISG TNF",
        `NkT.10.1` = "CD4 ISG TNF",
        `NkT.11` = "CD8 LQ",
        `NkT.12` = "CD4 Naive ISG",
        `NkT.13` = "CD8 Gamma-Delta",
        `NkT.14` = "CD8 Effector ISG",
        `NkT.15` = "CD56 Dim Nk",
        `NkT.16` = "CD4 LQ",
        `Plasma.0` = "Plasma",
        `Plasma.1` = "Plasma",
        `Plasma.2` = "Plasma",
        `Plasma.3` = "Plasma",
        `Plasma.4` = "Plasma",
        `Plasma.5` = "Plasma",
        `Plasma.6` = "Plasma",
        `Plasma.7` = "Plasma",
        `Plasma.8` = "Plasma",
        `Plasma.9` = "Plasma",
        `Plasma.10` = "Plasma",
        `Plasma.11` = "Plasma",
        `Plasma.12` = "Plasma",
        `Plasma.13` = "Plasma",
        `Plasma.14` = "Plasma",
        `Plasma.15` = "Plasma",
        `Plasma.16` = "Plasma",
        `Plasma.17` = "Plasma",
        `Plasma.18` = "Plasma",
        `Plasma.19` = "Plasma",
        `Plasma.20` = "Plasma",
        `Plasma.21` = "Plasma",
        `Plasma.22` = "Plasma",
        `Plasma.23` = "Plasma"
    )
    # merged@meta.data <- merged@meta.data %>% mutate(cellID = labelstmp[[.data[[cluster_group]]]])
    merged@meta.data$cellID <- labelstmp[merged@meta.data[[cluster_group]]]



    labels_comp <- c(
        `BEry.0` = "Ery",
        `BEry.1` = "B",
        `BEry.2` = "B",
        `BEry.3` = "B",
        `BEry.4` = "Ery",
        `BEry.5` = "Doublet",
        `BEry.6` = "B",
        `BEry.7` = "CD34 Progen",
        `BEry.8` = "Doublet",
        `BEry.9` = "B",
        `BEry.10` = "B",
        `BEry.11` = "pDC",
        `BEry.12` = "B",
        `BEry.13` = "Doublet",
        `BEry.14` = "Doublet",
        `BEry.15.0` = "Ery",
        `BEry.15.1` = "Ery",
        `BEry.16` = "B",
        `BEry.17` = "Ery",
        `Ery.0` = "Ery",
        `Ery.1` = "Ery",
        `Ery.2` = "Ery",
        `Ery.3` = "Ery",
        `Ery.4` = "Ery",
        `Ery.5` = "Ery",
        `Ery.6` = "Ery",
        `Ery.7` = "Ery",
        `Ery.8` = "Ery",
        `Ery.9` = "Ery",
        `Ery.10` = "Ery",
        `Ery.11` = "Ery",
        `Full.23` = "Stromal Cell",
        `Myeloid.0` = "Monocyte",
        `Myeloid.1` = "Monocyte",
        `Myeloid.2` = "Monocyte",
        `Myeloid.3` = "Monocyte",
        `Myeloid.4` = "Monocyte",
        `Myeloid.5` = "GMP",
        `Myeloid.6` = "Neutrophil",
        `Myeloid.7` = "Doublet",
        `Myeloid.8` = "Monocyte",
        `Myeloid.9` = "Doublet",
        `Myeloid.10` = "Neutrophil",
        `Myeloid.11.0` = "DC",
        `Myeloid.11.1` = "DC",
        `Myeloid.11.2` = "DC",
        `Myeloid.12` = "Monocyte",
        `Myeloid.13` = "Monocyte",
        `Myeloid.14` = "Doublet",
        `Myeloid.15` = "Monocyte",
        `NkT.0` = "CD4",
        `NkT.1.0` = "CD4",
        `NkT.1.1` = "CD4",
        `NkT.1.2` = "CD4",
        `NkT.1.3` = "CD8",
        `NkT.1.4` = "CD4",
        `NkT.1.5` = "CD4",
        `NkT.2.0` = "CD8",
        `NkT.2.1` = "CD8",
        `NkT.2.2` = "CD4",
        `NkT.2.3` = "CD8",
        `NkT.2.4` = "CD8",
        `NkT.3.0` = "CD8",
        `NkT.3.1` = "CD8",
        `NkT.3.2` = "CD8",
        `NkT.4` = "Nk",
        `NkT.5.0` = "NkT",
        `NkT.5.1` = "Nk",
        `NkT.5.2` = "Nk",
        `NkT.6` = "CD8",
        `NkT.7` = "LQ",
        `NkT.8` = "CD4",
        `NkT.9.0` = "Nk",
        `NkT.9.1` = "Nk",
        `NkT.10.0` = "CD8",
        `NkT.10.1` = "CD4",
        `NkT.11` = "CD8",
        `NkT.12` = "CD4",
        `NkT.13` = "CD8",
        `NkT.14` = "CD8",
        `NkT.15` = "Nk",
        `NkT.16` = "CD4",
        `Plasma.0` = "Plasma",
        `Plasma.1` = "Plasma",
        `Plasma.2` = "Plasma",
        `Plasma.3` = "Plasma",
        `Plasma.4` = "Plasma",
        `Plasma.5` = "Plasma",
        `Plasma.6` = "Plasma",
        `Plasma.7` = "Plasma",
        `Plasma.8` = "Plasma",
        `Plasma.9` = "Plasma",
        `Plasma.10` = "Plasma",
        `Plasma.11` = "Plasma",
        `Plasma.12` = "Plasma",
        `Plasma.13` = "Plasma",
        `Plasma.14` = "Plasma",
        `Plasma.15` = "Plasma",
        `Plasma.16` = "Plasma",
        `Plasma.17` = "Plasma",
        `Plasma.18` = "Plasma",
        `Plasma.19` = "Plasma",
        `Plasma.20` = "Plasma",
        `Plasma.21` = "Plasma",
        `Plasma.22` = "Plasma",
        `Plasma.23` = "Plasma"
    )
    merged@meta.data$cellID_comp <- labels_comp[merged@meta.data[[cluster_group]]]
}
```

## Flag CD45+, plasma
Flagging CD45+/- populations to analyze proportions wrt immune populations
```{r, fig.size=50,fig.height=6}
if (!SKIP_SETUP) {
    # Not flagging CD45- plasma... just erythroid
    VlnPlot(merged, features = c("PTPRC"), pt.size = 0, group.by = cluster_group)

    cd45neg_notplasma <- c("BEry.0", "BEry.4", "BEry.13", "BEry.17", "Ery.0", "Ery.1", "Ery.2", "Ery.3", "Ery.4", "Ery.5", "Ery.6", "Ery.8", "Ery.9", "Full.23")

    merged@meta.data <- merged@meta.data %>% mutate(CD45_flag = dplyr::case_when(
        .data[[cluster_group]] %in% cd45neg_notplasma ~ "Negative",
        .default = "Positive"
    ))
}
```


## Flag CD8
```{r}
if (!SKIP_SETUP) {
    # cd8_clusters <- c("NkT.1.3", "NkT.2.0", "NkT.2.1", "NkT.2.3", "NkT.2.4", "NkT.3.0", "NkT.3.1", "NkT.3.2", "NkT.6", "NkT.10.0", "NkT.14", "NkT.5.0")

    # no lq
    cd8_clusters <- c("NkT.1.3", "NkT.2.0", "NkT.2.1", "NkT.3.0", "NkT.3.1", "NkT.3.2", "NkT.6", "NkT.10.0", "NkT.14", "NkT.5.0", "NkT.5.2", "NkT.13", "NkT.2.3", "NkT.2.4", "NkT.11")


    merged@meta.data <- merged@meta.data %>% mutate(isCD8 = dplyr::case_when(
        .data[[cluster_group]] %in% cd8_clusters ~ "CD8",
        .default = "Other"
    ))
}
```


# Create Baseline CD8
Multiple baseline samples are possible, verify
```{r baseline_objects}
if (!SKIP_SETUP) {
    gc()
    cd8 <- subset(merged, subset = (
        isDoublet %in% c("Singlet") &
            isCD8 %in% c("CD8")))

    rm(merged)
    rm(merged_md_bckup)
    rm(clusterMD)
    gc()
    pcs <- c(25) # , 20, 25, 30)
    cd8_list <- list()
    for (i in pcs) {
        cd8_list[[i]] <- subset_and_markers(cd8, name = paste0("cd8_all_pcs_NEWMD_", i), output = output, SKIP_MARKERS = T, DIMPLOT_ADDITIONAL_DISPLAY = "subcluster_V03072023", res = NULL, rebuildObject = F, dims = i)
        cd8_list[[i]][["umapold"]] <- cd8_list[[i]][["umap.sub"]]
        cd8_list[[i]][["umap"]] <- cd8_list[[i]][["umapsub"]]
    }

    # DimPlot(cd8_list[[15]], group.by = cluster_group, label = T, repel = T, reduction = "umapsub")
}
```

```{r}
# DimPlot(cd8_list[[30]], group.by = cluster_group, label = T, repel = T)
```

# SlingShot 
```{r}
if (!SKIP_SETUP) {
    library(slingshot)
    library(SingleCellExperiment)
    library(Polychrome)
    library(BUSpaRse)
    library(tidyverse)
    library(tidymodels)
    output_list <- list()

    # Idents(cd8_list[[15]]) <- cd8_list[[15]]@meta.data[, cluster_group]
    # test <- cd8_list[[15]] |> subset(downsample = 500)
    # testout <- runSlingshot(test,
    #     groupingVar = "seurat_clusters_compartment_1_20_23_fullname",
    #     output = output,
    #     nameMod = paste0("slingshot_test"),
    #     slingshotDim = "HARMONY",
    #     dims = 1:5,
    #     visualizationDim = "UMAP",
    #     startCluster = "NkT.6",
    #     CREATE_SUBFOLDER_WITH_NAMEMOD = T,
    #     REBUILD_TRAJECTORY = T
    # )

    # outputs format: TrajectoryObject, TrajectoryLines, numLineages, AllGenes, TopGenes, validation results
    for (i in pcs) {
        obj <- cd8_list[[i]]
        output_list[[i]] <- runSlingshot(obj,
            groupingVar = "subcluster_V03072023",
            output = output, nameMod = paste0("slingshot_cd8_NEWMD_", i),
            slingshotDim = "HARMONY",
            dims = 1:i,
            visualizationDim = "UMAP",
            startCluster = "NkT.6",
            CREATE_SUBFOLDER_WITH_NAMEMOD = T,
            REBUILD_TRAJECTORY = F
        )
    }
} else {
    sce <- readRDS(LOAD_PATH)
}
# Then, do downstream stuff
# See: https://bioconductor.org/packages/release/bioc/vignettes/tradeSeq/inst/doc/tradeSeq.html


# cd8_reclustered <- subset_and_markers(cd8, name = "cd8-only-alltimepoints-reclustered_no_nkt12", res = 0.8, output = output, SKIP_MARKERS = T, rebuildObject = F)
# outputs_reclustered <- runSlingshot(cd8_reclustered, groupingVar = "seurat_clusters_0.8", output = output, nameMod = "slingshot_cd8_alltimepoints_0.8res_startclus2_no_nkt12", slingshotDim = "HARMONY", visualizationDim = "UMAPSUB", startCluster = "2", CREATE_SUBFOLDER_WITH_NAMEMOD = T, REBUILD_TRAJECTORY = F)

# # https://kstreet13.github.io/bioc2020trajectories/articles/workshopTrajectories.html - selected an external parameter with only two clusters for groupingVar... this is probably not valid
# outputs_oneaxis <- runSlingshot(cd8, groupingVar = "progression_group", output = output, nameMod = "slingshot_cd8_alltimepoints", slingshotDim = "HARMONY", visualizationDim = "UMAPSUB", startCluster = "NkT.6", CREATE_SUBFOLDER_WITH_NAMEMOD = T, REBUILD_TRAJECTORY = F)

# cd8_nogdt <- subset_and_markers(cd8 |> subset(subset = (seurat_clusters_compartment_1_20_23_fullname %ni% c("NkT.13"))), name = "cd8-only-nogdt", output = output, SKIP_MARKERS = T, DIMPLOT_ADDITIONAL_DISPLAY = "seurat_clusters_compartment_1_20_23_fullname", res = NULL, rebuildObject = F)
# outputs_nogdt <- runSlingshot(cd8_nogdt, groupingVar = "seurat_clusters_0.8", output = output, nameMod = "slingshot_cd8_alltimepoints_0.8res", slingshotDim = "HARMONY", visualizationDim = "UMAPSUB", startCluster = "2", CREATE_SUBFOLDER_WITH_NAMEMOD = T, REBUILD_TRAJECTORY = F)
```
